\input pdfToolbox

\setlayout{horizontal margin=2cm, vertical margin=2cm}
\parindent=0pt
\parskip=3pt plus 2pt minus 2pt

\input preamble

\footline={}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\headline={\pageborder{rgb{.5 .5 1}}{rgb{0 0 .4}}{5}}

\color rgb{.1 0 .8}

{\def\boxshadowcolor{rgb{.3 .3 .8}}
\bppbox{rgb{.6 .6 1}}{rgb{0 0 .4}}{rgb{.1 0 .4}}

    \centerline{\setfontandscale{bf}{20pt}Algebraic Topology I}
    \smallskip
    \centerline{\setfont{it}Lectures by Tahl Nowik}
    \centerline{\setfont{it}Summary by Ari Feiglin \setfont{rm}({\tt ari.feiglin@gmail.com})}

\eppbox

\bigskip

\bppbox{rgb{.6 .6 1}}{rgb{0 0 .4}}{rgb{.1 0 .4}}
    \section*{Contents}
    
    \tableofcontents
\eppbox

}

\vfill\break

\color{black}

\pageno=1
\newif\ifpageodd
\pageoddtrue
\headline={%
    \hbox to \hsize{\color{black}%
        \ifpageodd\hfil{\it\currsubsection\quad\bf\folio}\global\pageoddfalse%
        \else{\bf\folio\quad\it\currsubsection}\hfil\global\pageoddtrue\fi%
    }%
}

\section{Categories}

\bdefn

    A {\emphcolor category} ${\cal C}$ is a mathematical object which contains the following
    \benum
        \item a class of objects $\obof{\cal C}$ (the objects need not be sets),
        \item for every two objects $A,B\in\obof{\cal C}$ a class of {\emphcolor morphisms} $\Morof{A,B}$,
        \item an operation on morphisms $\circ$, where for every $f\in\Morof{A,B}$ and $g\in\Morof{B,C}$, $g\circ f\in\Morof{A,C}$,
        \item for every object $A\in\obof{\cal C}$ there exists an identity morphism $1_A\in\Morof{A,A}$ where for every $A,B\in\obof{\cal C}$ and $f\in\Morof{A,B}$, $f\circ 1_A=1_B\circ f=f$,
        \item for every $A,B,C,D\in\obof{\cal C}$ and $f\in\Morof{A,B},g\in\Morof{B,C},h\in\Morof{C,D}$, there is associativity: $(h\circ g)\circ f=h\circ (g\circ f)$.
    \eenum

\edefn

Although morphisms are not necessarily functions, we use similar notation: both $f\colon A\longto B$ and $A\xvarrightarrow{\,f\,}B$ are to be understood to mean $f\in\Morof{A,B}$.
And we write $A\in{\cal C}$ to mean $A\in\obof{\cal C}$.

Notice that for every $A\in{\cal C}$, $1_A$ is unique: suppose $1_A$ and $1_A'$ are both identity morphisms then $1_A\circ1_A'=1_A$ since $1_A'$ is an identity, but $1_A\circ1_A'=1_A'$ since $1_A$ is an
identity so $1_A=1_A'$.

\bdefn

    Suppose ${\cal C}$ and ${\cal D}$ are two categories, a {\emphcolor functor} $F$ from ${\cal C}$ to ${\cal D}$ is a correspondence where for every $A\in{\cal C}$ there is defined a single
    $F(A)\in{\cal D}$, and for every $f\in\Morof{A,B}$ there exists a unique $F(f)\in\Morof{F(A),F(B)}$ such that for all $A,B,C\in{\cal C}$ and $f\in\Morof{A,B}$ and $g\in\Morof{B,C}$ we have that
    $F(g\circ f)=F(g)\circ F(f)$ and $F(1_A)=1_{F(A)}$.

\edefn

\bexam

    The following are examples of categories:
    \benum
        \item The category of all groups, morphisms are taken to be homomorphisms between groups;
        \item The category of all topological spaces, morphisms are taken to be homeomorphisms;
        \item The category of all sets, the morphisms are taken to be set functions;
        \item The category of pairs of topological spaces: the objects are of the form $(X,A)$ where $X$ is a topological space and $A\subseteq X$.
            Morphisms between $(X,A)$ and $(Y,B)$ of this category are continuous functions $f$ between $X$ and $Y$ such that $f(A)\subseteq B$.
        \item The category of pointed topological spaces: the objects are $(X,a)$ where $X$ is a topological space and $a\in X$ and the morphisms between $(X,a)$ and $(Y,b)$ are continuous functions between
            $X$ and $Y$ such that $a\mapsto b$.
    \eenum
    An example of a functor is the so-called {\it forgetful functor} from the category of topological spaces to the category of sets: map a topological to itself as a pure set.

\eexam

This course will focus on a specific functor between the category of pointed topological spaces to the category of groups.

\bdefn

    Let ${\cal C}$ be a category, and $A,B\in{\cal C}$.
    A morphism $f\colon A\longto B$ is an {\emphcolor isomorphism} if there exists a morphism $g\colon B\longto A$ such that $g\circ f=1_A$ and $f\circ g=1_B$.
    Such a $g$ is called the {\emphcolor inverse} of $f$ and is denoted $f^{-1}$ (notice that by symmetry the inverse is also an isomorphism).
    If there exists an isomorphism between $A$ and $B$, we denote this by $A\cong B$ and $A$ and $B$ are called {\emphcolor isomorphic}.

\edefn

Inverses are unique: if $g_1$ and $g_2$ are inverses of $f$ then $(g_1\circ f)\circ g_2=1_A\circ g_2=g_2$ but $g_1\circ(f\circ g_2)=g_1\circ 1_B=g_1$ and by associativity these are equal.
Furthermore the composition of isomorphisms is an isomorphism: it is easily verified that $(f\circ g)^{-1}=g^{-1}\circ f^{-1}$.
Notice that $1_A$ is an isomorphism and it is its own inverse.

\bprop

    A functor maps isomorphisms to isomorphisms, in particular $F(f^{-1})=F(f)^{-1}$ if $f\colon A\longto B$ is an isomorphism.

\eprop

\Proof notice that $F(f)\circ F(f^{-1})=F(f\circ f^{-1})=F(1_B)=1_{F(B)}$ and $F(f^{-1})\circ F(f)=F(f^{-1}\circ f)=F(1_A)=1_{F(A)}$.
So $F(f^{-1})$ is indeed the inverse of $F(f)$.
\qed

\subsection{Homotopies}

\bdefn

    Let $X$ and $Y$ are topological spaces and $f,g\colon X\longto Y$ (meaning they are morphisms, thus continuous).
    We say that $f$ is homotopic to $g$, denoted $f\sim g$, if there exists an $H\colon X\times I\longto Y$ ($I=[0,1]$, $X\times I$ is the product topology) such that $H(x,0)=f(x)$ and
    $H(x,1)=g(x)$ for all $x\in X$.
    We denote $h_t(x)\coloneqq H(x,t)$, and $H$ is called a {\emphcolor homotopy} from $f$ to $g$.

\edefn

A homotopy is essentially a smooth mapping from one morphism $f$ to another $g$.
Homotopy is indeed an equivalence relation: firstly $f\sim f$ as we can define $H(x,t)=f(x)$ which is continuous as the composition of continuous functions ($H=f\circ\pi_1$), if $f\sim g$ then
define $H'(x,t)=H(x,1-t)$ which is also continuous (since $(x,t)\mapsto(x,1-t)$ is continuous since its components are) and $H'(x,0)=g(x)$ and $H'(x,1)=f(x)$ so $g\sim f$, and if $H_1$ is a homotopy from $f$
to $g$ and $H_2$ is a homotopy from $g$ to $h$, define
$$ H(x,t) = \cases{H_1(x,2t) & $0\leq t\leq\frac12$\cr H_2(x,2t-1) & $\frac12\leq t\leq1$} $$
$X\times[0,1/2]$ and $X\times[1/2,1]$ are closed (since $X\times[0,1/2]$ is the preimage of $[0,1/2]$ in the mapping $(x,t)\mapsto t$) and $H(x,t)$ is continuous on both of these (since $H_1(x,2t)$ and
$H_2(x,2t-1)$ are continuous), so $H(x,t)$ is continuous.

\bprop

    For every topological space $X$ and every two morphisms $f,g\colon X\longto{\bb R}^n$, $f$ and $g$ are homotopic.

\eprop

\Proof define $H(x,t)=(1-t)f(x)+tg(x)$ (addition and scalar multiplication are continuous).
\qed

\bdefn

    A topological space $X$ is {\emphcolor contractible} if the identity map ${\rm id}_X$ is homotopic to some constant map.

\edefn

Notice that all two constant maps are homotopic if and only if the space is path connected.
If all two constant maps are homotopic, for $x_1,x_2\in X$ let $H(x,t)$ be a homotopy from $x_1$ to $x_2$ and define $\gamma(t)=H(x_0,t)$ for any $x_0\in X$, this is a continuous path from $x_1$ to $x_2$.
And if $X$ is path connected, for $x_1$ and $x_2$ and $\gamma$ connecting them, define $H(x,t)=\gamma(t)$.

\bprop

    Let $X,Y,Z$ be topological spaces, $f,f'\colon X\longto Y$ and $g,g'\colon Y\to Z$ such that $f\sim f'$ and $g\sim g'$, then $g\circ f\sim g'\circ f'$.

\eprop

\Proof let $H$ be a homotopy from $f$ to $f'$ and $K$ a homotopy from $g$ to $g'$.
Then define $J(x,t)=K(H(x,t),t)$ which is a composition of continuous functions (map $(x,t)$ to $((x,t),t)$ to $(H(x,t),t)$ to $K(H(x,t),t)$).
\qed

We call the equivalence classes of morphisms under $\sim$ {\it homotopy classes}, and the homotopy class of a morphism $f$ is denoted $[f]$.
So by above, $[f]\circ[g]\coloneqq[f\circ g]$ is a well-defined operation.
This gives us a new category whose objects are topological spaces and morphisms are homotopy classes.
What are the isomorphisms in this category?
Well the identities are obviously $[1_X]$ since $[f]\circ[1_X]=[f\circ1_X]=[f]$ and $[1_X]\circ[g]=[1_X\circ g]=[g]$.
So an isomorphism $X\xvarrightarrow{\,[f]\,}Y$ is a homotopy class such that there exists a $Y\xvarrightarrow{\,[g]\,}X$ such that $[f]\circ[g]=[f\circ g]=[1_X]$ and $[g\circ f]=[1_Y]$.
We give these isomorphisms a different name:

\bdefn

    Let $X$ and $Y$ be topological spaces, then $f\colon X\longto Y$ is a {\emphcolor homotopic equivalence} if there exists a $g\colon Y\longto X$ such that $g\circ f\sim{\rm id}_X$ and
    $f\circ g\sim{\rm id}_Y$.
    If a homotopy equivalence exists between $X$ and $Y$, then $X$ and $Y$ are said to be {\emphcolor homotopy equivalent}, denoted $X\simeq Y$.

\edefn

Notice that homeomorphisms are homotopic equivalences, since $f^{-1}\circ f={\rm id}_X$ and $f\circ f^{-1}={\rm id}_Y$.

\bdefn

    Let $X$ and $Y$ be topological spaces, $A\subseteq X$, and $f,g\colon X\longto Y$.
    We say that $f$ and $g$ are homotopic relative to $A$, denoted $f\buildrel A\over\sim g$, if there exists a homotopy $H$ from $f$ to $g$ such that $H(a,t)=f(a)$ for all $a\in A$ and $t\in I$.
    In such a case we must have $f\bigr|_A=g\bigr|_A$.

\edefn

It is not enough for $f\sim g$ and $f\bigr|_A=g\bigr|_A$ for $f$ and $g$ to be homotopic relative to $A$.
For example take $I$ and $S^1$ and the points $0$ and $1$ on $I$.
Then we can continuously deform $I$ so that it maps onto the bottom or top of the circle.
These are two continuous mappings which are homotopic, but no homotopy between them which keeps the image of $0$ and $1$ constant.

Notice that $\buildrel A\over\sim$ is an equivalence relation, the proof of this is analogous to the proof that homotpy is an equivalence relation.
It also preserves composition, if $f,f'\colon(X,A)\longto(Y,B)$ (meaning they are morphisms from $X$ to $Y$ and $f(A),f'(A)\subseteq B$) and $g,g'\colon(Y,B)\longto(Z,C)$ such that $f\buildrel A\over\sim f'$
and $g\buildrel B\over\sim g'$, then $g\circ f\buildrel A\over\sim g'\circ f'$.

\bdefn

    Let $X$ be a topological space.
    $A\subseteq X$ is called a {\emphcolor retract} if there exists an $r\colon X\longto A$ such that $r\circ\iota={\rm id}_A$ where $\iota\colon A\longto X$ is the inclusion map.
    In other words $r(a)=a$ for all $a\in A$.
    $r$ is called a {\emphcolor retraction}.

\edefn

For example $\partial I=\set{0,1}$ is not a retraction of $I$ since every continuous image of $I$ must be connected, and $\partial I$ is not.
But if we take $X$ to be an eight shape, and $A$ its bottom circle, then we can map the top circle to the middle point and $A$ to itself and this is a retraction.

\bdefn

    $A\subseteq X$ is called a {\emphcolor deformation retract} if there exists a retraction $r$ such that $\iota\circ r\buildrel A\over\sim{\rm id}_X$.

\edefn

Instead of requiring $r$ be a retraction, we can require only that $r(X)\subseteq A$.
Since then if $\iota\circ r\buildrel A\over\sim{\rm id}_X$, this means that $r(a)={\rm id}_X(a)=a$ for all $a\in A$ so it is already a retraction.
Explicitly, this is equivalent to saying that there exists a homotopy $H\colon X\times I\longto X$ such that $H(x,0)=x$ for all $x\in X$, $H(a,t)=a$ for all $a\in A,t\in I$, $H(x,1)\in A$ for all $x\in X$.

Notice that if $A\subseteq X$ is a deformation retract then $\iota\colon A\longto X$ is a homotopy equivalence, since $r\circ\iota={\rm id}_A$ and $\iota\circ r\sim{\rm id}_X$.

\bexam

    Let $X={\bb R}^n\setminus\set0$ and $A=S^{n-1}\subseteq{\bb R}^n\setminus\set0$.
    Then $r(x)\coloneqq\frac x{\norm x}$ is a retraction with the homotopy $H(x,t)=(1-t)x+t\frac x{\norm x}$.
    This is the homotopy we used to show that all morphisms to ${\bb R}^n$ are homotopic.

\eexam

A morphism $f$ is called {\it null-homotopic} if it is homotopic to a constant morphism.

\bprop

    Let $X$ be a topological space and $f\colon S^1\longto X$, then the following are equivalent
    \benum
        \item $f$ is null-homotopic,
        \item $f$ is null-homotopic relative to any point on $S^1$,
        \item $f$ can be expanded to a morphism on $D^2$ (the disk in ${\bb R}^2$), meaning there exists an $F\colon D^2\longto X$ such that $F\bigr|_{S'}=f$.
    \eenum

\eprop

$(2)\implies(1)$ is trivial since a null-homotopy relative to a point is still a null-homotopy.
$(3)\implies(2)$: let $\iota\colon S^1\longto D^2$ be the inclusion map, and let $a\in S^1$, define the homotopy $H\colon S^1\longto I\longto D^2$ by $H(x,t)=(1-t)\iota(x)+ta$, which is a homotopy from
$\iota$ to the constant map $K_a$.
Then $F\circ H$ is a null-homotopy between $f$ and $K_{f(a)}$ (since $F\circ H(x,0)=F(x)=f(x)$ and $F\circ H(x,1)=F(a)$) relative to $a$ since $F\circ H(a,t)=F(a)$.
$(1)\implies(3)$: so there exists a homotopy $H\colon S^1\times I\longto X$ such that $H(x,0)=f(x)$ for every $x\in S^1$ and there exists a $p\in X$ such that $H(x,1)=p$ for all $x\in S^1$.
Let us define $\rho\colon S^1\times I\longto D^2$ by $\rho(x,t)=(1-t)x$, this is a continuous map from a compact (since $S^1$ and $I$ are compact and therefore so is their product) to a Hausdorff space, and
so it is closed.
And it is surjective, so it is a quotient map.
So $D^2$ is the quotient space of $S^1\times I$ with respect to $\rho$, and $H$ respects $\rho$, since $\rho(x,t)=\rho(y,s)$ implies $(1-t)x=(1-s)y$ and this means that either $(x,t)=(y,s)$ or $t=s=1$.
But in both cases $H(x,t)=H(y,s)$, and so there exists an $F\colon D^2\longto X$ which is continuous such that $H=F\circ\rho$, meaning $F(x)=H(x,0)=f(x)$ as required.
\qed

This proof uses the fact that if $\rho$ is a quotient map, and $f\colon X\longto Y$ is continuous then there exists a $F\colon\overline X\longto Y$ such that $f=F\circ\rho$ if and only if $\rho(a)=\rho(b)$
implies $f(a)=f(b)$.

\bdefn

    Let $X$ be a topological space, and for every $a,b\in X$ define $\Gamma_{ab}$ to be the set of all paths from $a$ to $b$, which are continuous maps $I\longto X$.
    On $\Gamma_{ab}$ we take the equivalence relation of homotopy relative to $\partial I=\set{0,1}$.
    Take $\hat\Gamma_{ab}$ to be the partition defined by this relation, ie. $\hat\Gamma_{ab}=\slfrac{\Gamma_{ab}}{\buildrel\partial I\over\sim}$.

    If $[\gamma]\in\hat\Gamma_{ab}$ and $[\delta]\in\hat\Gamma_{bc}$ then we define $[\gamma][\delta]\coloneqq[\gamma*\delta]$ (their concatenation).

\edefn

We must show that this is well-defined, meaning we must show that if $\gamma\hdi\gamma'$ and $\delta\hdi\delta'$ then $\gamma*\delta\hdi\gamma'*\delta'$.
So let $H\colon I\times I\longto X$ be a homotopy relative to $\partial I$ between $\gamma$ and $\gamma'$, and $G\colon I\times I\longto X$ between $\delta$ and $\delta'$.
Then define
$$ K(s,t) \coloneqq \cases{H(2s,t) & $0\leq s\leq\frac12$\cr G(2s-1,t) & $\frac12\leq s\leq1$} $$
this is continuous, $K(0,t)=H(0,t)=0$ and $K(1,t)=G(1,t)=1$ so it is a homotopy between the concatenations relative to $\partial I$.

Notice that concatenation is not necessarily associative, since in $(\gamma*\delta)*\epsilon$, the speed of $\gamma$ and $\delta$ is quadrupled while in $\gamma*(\delta*\epsilon)$, $\gamma$'s speed is only
doubled.
But it is the case that $[\gamma]\bigl([\delta][\epsilon]\bigr)=\bigl([\gamma][\delta]\bigr)[\epsilon]$, so in homotopy concatenation is associative.
So we need to prove $\gamma(\delta\epsilon)\hdi(\gamma\delta)\epsilon$, the idea behind this is that for every $x$ and $y$ where $\gamma(\delta\epsilon)(x)=(\gamma\delta)\epsilon(y)$, define in $I\times I$
the line between $(x,0)$ and $(y,1)$.
These lines cover $I\times I$ and for every point $(t,s)$ which is on the line from $(x,0)$ map it to $\gamma(\delta\epsilon)(x)$.

We can prove in a similar manner that for $\gamma\in\Gamma_{ab}$, $[K_a][\gamma]=[\gamma][K_b]=[\gamma]$.

And so we have defined a category.
The objects of this category are the points $a\in X$ and the morphisms between $a$ and $b$ are $\hat\Gamma_{ab}$ (notice that $[\gamma]\in\hat\Gamma_{ab}$ can be composed with elements from
$\hat\Gamma_{bc}$, so the order of composition is reversed).
Here the identity morphisms are $[K_a]$.

Notice that every morphism in this category is an isomorphism.
This is since for every $\gamma\in\Gamma_{ab}$ we defined its reverse $\overline\gamma\in\Gamma_{ba}$ by $\overline\gamma(t)\coloneqq\gamma(1-t)$.

\bprop

    $[\gamma][\overline\gamma]=[K_a]$ and $[\overline\gamma][\gamma]=[K_b]$.

\eprop

The idea is that at time $t$ we take the path $\gamma$ but not all the way, then wait, then take the reverse path $\overline\gamma$.
So
$$ H(x,t) = \cases{\gamma(2x) & $0\leq x\leq\frac{1-t}2$\cr \gamma(1-t) & $\frac{1-t}2\leq x\leq\frac{1+t}2$\cr \gamma(2-2x) & $\frac{1+t}2\leq x\leq1$} $$
is a homotopy from $\gamma*\overline\gamma$ to $K_a$.
\qed

\bdefn

    A {\emphcolor groupoid} is a small category (a category whose objects form a set, not a pure class) such that every morphism is an isomorphism.
    If ${\cal C}$ is a groupoid, then $\Mor(A,A)$ is then a group for every $A\in{\cal C}$.

\edefn

\bdefn

    Given a pointed topological space $(X,a)$ (call $a$ the basis point), define the {\emphcolor first homotopy group} (or the {\emphcolor fundamental group}) $\pi_1(X,a)\coloneqq\hat\Gamma_{aa}$.
    And given a morphism $f\colon(X,a)\longto(Y,b)$ (meaning $f$ is continuous and $f(a)=b$), then we define a group homomorphism $f_*\colon\pi_1(X,a)\longto\pi_1(Y,b)$ by $f_*([\gamma])=[f\circ\gamma]$.
    The correspondence $(X,a)\mapsto\pi_1(X,a)$ and $f\mapsto f_*$ is a functor.

\edefn

We need to show that $f_*$ is well-defined and also a group homomorphism.
To show that it is well-defined, suppose $\gamma\hdi\delta$, then we must show $f\circ\gamma\hdi f\circ\delta$.
Now we showed that if $f\htopic Af'$ and $g\htopic Bg'$ such that $f(A),f'(A)\subseteq B$ then $g\circ f\htopic Ag'\circ f'$.
And we have that $\gamma\hdi\delta$ and $f\htopic{\set a}f$ so $f\circ\gamma\hdi f\circ\delta$ as required.
Now we must show that $f_*$ is a homomorphism, ie.
$$ f_*([\gamma][\delta]) = [f\circ(\gamma*\delta)] = [(f\circ\gamma)*(f\circ\delta)] = f_*([\gamma])f_*([\delta]) $$
Actually a stronger result holds, $f\circ*(\gamma*\delta)=(f\circ\gamma)*(f\circ\delta)$, as both are given by
$$ \cases{f\circ\gamma(2t) & $0\leq t\leq\frac12$\cr f\circ\delta(2t-1) & $\frac12\leq t\leq1$} $$

To finish the proof that the correspondence is a functor, we need to show that $(g\circ f)_*=g_*\circ f_*$ and $({\rm id}_X)_*={\rm id}_{\pi_1(X,a)}$.
We do so directly:
$$ (g\circ f)_*([\gamma]) = [(g\circ f)\circ\gamma] = [g\circ(f\circ\gamma)] = g_*([f\circ\gamma]) = g_*\circ f_*([\gamma]) $$
and
$$ ({\rm id}_X)_*([\phi]) = [{\rm id}_X\phi] = [\phi] $$
so $({\rm id}_X)_*={\rm id}_{\pi_1(X,a)}$ as required.
Thus we have defined a functor from the category of pointed topological spaces to the category of groups.

\bprop

    Let $X$ be a topological space, $a\in X$, and $A$ be $a$'s connected component.
    Let $\iota\colon A\longto X$ be the inclusion map, then $\iota_*\colon\pi_1(A,a)\longto\pi_1(X,a)$ is an isomorphism.

\eprop

\Proof $\iota_*$ is injective: $\iota_*([\gamma])=[K_a]$ if and only if $[\iota\circ\gamma]=[K_a]$, which means $\iota\circ\gamma\hdi K_a$, let $H$ be such a homotopy.
Then $H\colon I\times I\longto X$, but we claim that $H$'s image is contained in $A$ so that it is also a homotopy $\gamma\hdi K_a$, meaning $[\gamma]=[K_a]$.
Suppose not, that $H(t_0,s_0)\notin A$, then define $\delta(t)=H(t_0\cdot t,s_0)$.
Then $\delta(0)=H(0,s_0)=a$ since $H$ is a homotopy relative to $\partial I$, and $\delta(1)=H(t_0,s_0)\notin A$, so $a$ is connected to a value not in $A$, in contradiction.
So $\iota_*$ is indeed injective.

Now suppose $[\gamma]\in\pi_1(X,a)$, meaning $\gamma$ is a path connecting $a$ to itself in $X$.
But every point in $\gamma$'s image must also be connected to $a$, meaning $\gamma$ is a path connecting $a$ to itself contained within $A$.
So there exists a $\gamma'\in\Gamma_{aa}^A$ such that $\gamma=\iota\circ\gamma'$ and in particular $\iota_*([\gamma'])=[\gamma]$ as required.
So $\iota_*$ is a bijective homomorphism, an isomorphism.
\qed

Suppose $a,b\in X$ such that there exists a path between them, $\gamma\colon I\longto X,\gamma(0)=a,\gamma(1)=b$.
Let us define
$$ F_\gamma\colon\pi_1(X,a)\longto\pi_1(X,b),\qquad F_\gamma[\phi] = [\overline\gamma*\phi*\gamma] =[\overline\gamma][\phi][\gamma] $$
so $F_\gamma[\phi]$ is the class of curves homotopic to the curve obtained by walking from $b$ along $\gamma$ backward to $a$, then going back along $\gamma$ to $b$.
Notice that $F_\gamma[\phi]=[\gamma]^{-1}[\phi][\gamma]$, so $F_\gamma$ is simply conjugation by $[\gamma]$.

In general, suppose ${\cal G}$ is a groupoid and let $A,B\in{\cal G}$ such that $\Mor(A,B)\neq\varnothing$.
Then $\Mor(A,A)$ and $\Mor(B,B)$ are isomorphic groups.
Let $\phi\in\Mor(A,B)$ and define $F_\phi\colon\Mor(A,A)\longto\Mor(B,B)$ by $F_\phi(\varkappa)=\phi\circ\varkappa\circ\phi^{-1}$.
This is a group homomorphism:
$$ F_\phi(\varkappa_1)\circ F_\phi(\varkappa_2) = \phi\circ\varkappa_1\circ\phi^{-1}\circ\phi\circ\varkappa_2\circ\phi^{-1} = \phi\circ(\varkappa_1\circ\varkappa_2)\circ\phi^{-1} = F_\phi(\varkappa_1\circ\varkappa_2) $$
and it is injective:
$$ F_\phi(\varkappa) = 1_B \iff \phi\circ\varkappa\circ\phi^{-1} = 1_B \iff \phi\circ\varkappa=\phi \iff \varkappa = \phi^{-1}\circ\phi = 1_A $$
and it is surjective: let $\varkappa'\in\Mor(B,B)$ and define $\varkappa\coloneqq\phi^{-1}\circ\varkappa'\circ\phi$, it is clear $F_\phi(\varkappa)=\varkappa'$.
It is clear that $F_\phi^{-1}=F_{\phi^{-1}}$ by this.

Our $F_\gamma$ above is precisely this $F_{[\gamma]}$ defined in the groupoid of first homotopy groups above $X$, meaning it is an isomorphism between $\pi_1(X,a)$ and $\pi_1(X,b)$.
And so $F_\gamma^{-1}=F_{\overline\gamma}$.
Let us summarize this:

\bprop

    Let $a,b\in X$ be two points in $X$ connected by a path $\gamma$.
    Then $\pi_1(X,a)$ and $\pi_1(X,b)$ are isomorphic.

\eprop

\bprop

    Suppose $H\colon I\times I\longto X$ is a homotopy from the closed loop $\phi$ to the closed loop $\psi$ such that for all $t\in I$, $H(0,t)=H(1,t)$.
    Define the path $\gamma(t)=H(0,t)=H(1,t)$, then
    $$ [\psi] = [\overline\gamma][\phi][\gamma] $$

\eprop

\Proof this is equivalent to saying
$$ [\overline\psi][\overline\gamma][\phi][\gamma] = [K_p] \iff \overline\psi*\overline\gamma*\phi*\gamma \hdi K_p $$
Now, $\overline\psi*\overline\gamma*\phi*\gamma$ is a curve $I\longto X$ whose endpoints are the same, so it can be viewed as a curve $S^1\longto X$.
And it can be extended to the curve $H$ on $D^2$ (since the unit disc and unit square are homeomorphic), which we showed above is equivalent to $\overline\psi*\overline\gamma*\phi*\gamma$ being
null-homotopic relative to any point on $S^1$, so we can choose the point which corresponds to $0$ and $1$.
\qed

\bthrm

    Let $f,g\colon X\longto Y$ be homotopic with homotopy $H$.
    Define $\gamma(t)\coloneqq H(a,t)$, so $\gamma(0)=f(a)$ and $\gamma(1)=g(a)$.
    Then $g_*=F_\gamma\circ f_*$ (recall that $f_*\colon\pi_1(X,a)\longto\pi_1(Y,f(a))$ and $g_*\colon\pi_1(X,a)\longto\pi_1(Y,g(a))$).

\ethrm

\Proof let $[\phi]\in\pi_1(X,a)$ then
$$ F_\gamma(f_*[\phi]) = [\overline\gamma][f\circ\gamma][\gamma],\qquad g_*[\phi] = [g\circ\phi] $$
Define $K(s,t)\coloneqq H(\phi(s),t)$ which is continuous and
$$ K(s,0) = H(\phi(s),0) = f\circ\phi(s),\quad K(s,1) = H(\phi(s),1) = g\circ\phi(s),\quad K(0,t) = H(\phi(0),t) = H(a,t) = K(1,t) $$
so by the above proposition, since $K$ is a homotopy from the closed loop $f\circ\phi$ to the closed loop $g\circ\phi$,
$$ [g\circ\phi] = [\overline\gamma][f\circ\phi][\gamma] $$
as required.
\qed

Notice then that $g_*=f_*$ if and only if $F_\gamma={\rm id}$ (requiring $f(a)=g(a)=b$).
This happens when $[\gamma]=[K_b]$ for example, which can happen when $\gamma=K_b$.
Ie. if $H(a,t)=b$ for all $t\in I$, then $f_*=g_*$.
But notice that this is simply the condition for $H$ to be a homotopy relative to $\set a$, so

\bprop

    If $f\htopic{\set a}g$ then $f_*=g_*$.

\eprop

\bthrm

    If $f\colon X\longto Y$ is a homotopy equivalence, then $f_*\colon\pi_1(X,a)\longto\pi_1(Y,f(a))$ is an isomorphism of groups.

\ethrm

\Proof there exists a $g\colon Y\longto X$ such that $g\circ f\sim{\rm id}_X$ and $g\circ f\sim{\rm id}_Y$.
Now since $g\circ f\sim{\rm id}_X$, then by above $(g\circ f)_*=F_\gamma\circ({\rm id}_X)_*=F_\gamma$, since ${(\cdot)}_*$ is a functor, $g_*\circ f_*=F_\gamma$.
Now recall that $F_\gamma=g_*\circ f_*\colon\pi_1(X,a)\longto\pi_1(X,g(f(a)))$ is an isomorphism.
And similarly $f\circ g\sim{\rm id}_Y$, so there exists an $F_\delta$ such that $(f\circ g)_*=f_*\circ g_*=F_\delta\colon\pi_1(Y,f(a))\longto\pi_1(X,f(g(f(a))))$ is an isomorphism.

Recall that if $f\circ g$ is injective, then $g$ is injective, and if $f\circ g$ is surjective, $f$ is surjective.
Thus $f_*$ is injective and surjective, meaning it is a bijective homomorphism, an isomorphism.
\qed

Thus if $X$ is contractible, then it is homotopic to the singleton space (an exercise in homework), $\set b$.
Thus $\pi_1(X,a)\cong\pi_1(\set b,b)$, and $\pi_1(\set b,b)$ has a single point, meaning $\pi_1(X,a)$ is the trivial group.

\bcoro

    If $X$ is contractible, then $\pi_1(X,a)$ is trivial.

\ecoro

\bdefn

    Let $E,B$ be topological spaces, then a map $p\colon E\longto B$ is a {\emphcolor covering map} (or just a {\it covering}) if
    \benum
        \item $p$ is surjective,
        \item for every $x\in B$, there exists a neighborhood $x\in{\cal U}$ such that $p^{-1}({\cal U})$ is the disjoint union of open sets $\set{{\cal V}_\alpha}_{\alpha\in I}$ such that for every
            $\alpha\in I$, $p\bigl|_{{\cal V}_\alpha}\colon{\cal V}_\alpha\longto{\cal U}$ is a homeomorphism.
    \eenum

\edefn

For example, take $B=S^1$ and $E={\bb R}$.
Define $p(t)\coloneqq(\cos(2\pi t),\sin(2\pi t))$, or if we identify $S^1$ with $S^1=\set{z\in{\bb C}}[\abs z=1]$, then $p(t)=e^{2\pi it}$.
This is like spiraling the real line so that all the integers are all on a vertical line (they all map to $1\in S^1$), and projecting this spiral onto a plane to form the circle $S^1$.
$p$ is obviously a surjective continuous map.
For every $e^{2\pi it}$, let $\epsilon<2\pi$ then the preimage of the open set ${\cal U}_t\coloneqq\set{e^{2\pi is}}[t-\epsilon<s<t+\epsilon]$ is
$\bigdcup_{n\in{\bb Z}}\bigl(n+(t-\epsilon,t+\epsilon)\bigr)$ and every $n+(t-\epsilon,t+\epsilon)$ is homeomorphic (via $p$) to this ${\cal U}_t$.

\bthrm

    Suppose $p\colon E\longto B$ is a covering, $b\in B$ and $e\in E$ such that $p(e)=b$.
    Now suppose $\gamma\colon I\longto B$ is a curve starting at $b$, $\gamma(0)=b$.
    Then there exists a unique curve $\delta\colon I\longto E$ starting at $e$ such that $p\circ\delta=\gamma$.

\ethrm

\Proof let $\set{{\cal U}_\beta}_{\beta\in J}$ be an open covering of $B$, then $\set{\gamma^{-1}({\cal U}_\beta)}_{\beta\in J}$ is an open covering of $I$.
Since $I$ is a compact metric space, it has a Lebesgue number, and so we can take $n$ small enough such that for every $k$, $I_k\coloneqq\bracks{\frac{k-1}n,\frac kn}$ is contained within a single
$\gamma^{-1}({\cal U}_\beta)$.
Meaning for every $1\leq k\leq n$ there exists a $\beta_k\in J$ such that $I_k\subseteq\gamma^{-1}({\cal U}_{\beta_k})$, so $\gamma(I_k)\subseteq{\cal U}_{\beta_k}$.

Now since $p$ is a covering, there exists a neighborhood $b\in{\cal U}$ such that $p^{-1}({\cal U})=\bigdcup_{\alpha\in I}{\cal V}_\alpha$.
Since this is a disjoint union and $p(e)\in{\cal U}$, there exists a single $\alpha\in I$ such that $e\in{\cal V}_\alpha$, and let us denote this neighborhood by ${\cal V}$.
And so $p\bigl|_{\cal V}$ is a homeomorphism from ${\cal V}$ to ${\cal U}$.
So define
$$ \delta\bigl|_{I_1}\coloneqq \parens{p\bigl|_{\cal V}}^{-1}\circ\gamma\bigl|_{I_1} $$
We have then that $p\circ\parens{\delta\bigl|_{I_1}}=\gamma\bigl|_{I_1}$ and $\delta(0)=\parens{p\bigl|_{\cal V}}^{-1}\circ\gamma(0)=e$.
This is the only such transform of $\gamma\bigl|_{I_1}$ which satisfies $\delta\bigl|_{I_1}(0)=e$ since every transform $\delta\colon I_1\longto E$ where $p\circ\delta=\gamma$,
$\delta(I_1)=p^{-1}\gamma(I_1)\subseteq p^{-1}({\cal U})=\bigdcup_{\alpha\in I}{\cal V}_\alpha$.
Since $\delta(I_1)$ is connected, it must be contained within a single ${\cal V}_\alpha$, this being ${\cal V}$ since it contains $e$.

We continue with $\gamma\bigl|_{I_2}$ which starts at $\frac1n$ to get $\delta\bigl|_{I_2}$, and so on.
Together these form $\delta$, which is continuous as $\set{I_k}_{1\leq k\leq n}$ is a finite partition of $I$ into closed sets and $\delta$ is continuous over each of these closed sets.
\qed

Since this curve is unique, we can denote it by $\hat\gamma^e$ (since there is a unique curve for every $e\in p^{-1}(b)$).

\bprop

    Let $p\colon E\longto B$ be a covering, $\gamma,\delta\colon I\longto B$ are curves such that $\gamma\hdi\delta$.
    Let $a=\gamma(0)=\delta(0)$, and let $e\in E$ such that $p(e)=a$.
    Then $\hat\gamma^e\hdi\hat\delta^e$, and in particular $\hat\gamma^e(1)=\hat\delta^e(1)$; both of the curves finish on the same point.

\eprop

\Proof let $H\colon I\times I\longto B$ be a homotopy relative to $\partial I$ from $\gamma$ to $\delta$.
Let $\set{\U_\alpha}_{\alpha\in I}$ be an open cover of $B$ by good open sets, then $\set{H^{-1}(\U_\alpha)}_{\alpha\in I}$ is an open cover of $I\times I$.
Since $I\times I$ is a compact metric space, it has a Lebesgue number and so there is an $n$ large enough so that we can partition $I\times I$ into squares
$S_{ij}=\bracks{\frac in,\frac{i+1}n}\times\bracks{\frac jn,\frac{j+1}n}$ such that each $S_{ij}$ is contained within some $H^{-1}(\U_\alpha)$.

Continuing similar to the previous proof, we get a homotopy $\hat H\colon I\times I\longto E$ such that $p\circ\hat H=H$ and $\hat H(0,0)=e$.
So $p\circ\hat H(t,0)=H(t,0)=\gamma(t)$ and since $\hat\gamma^e$ is unique, this means $\hat H(t,0)=\hat\gamma^e(t)$.
And similarly $p\circ\hat H(t,1)=H(t,1)=\delta(t)$ so $\hat H(t,1)=\hat\delta^e(t)$.
And $p\circ\hat H(0,s)=H(0,s)=a$, we can view this as the curve $K_a$ from $a$ to $a$, and so by uniqueness we have again that $\hat H(0,s)=\hat K_a^e(s)=e$.
Similar for $\hat H(1,s)$.
Thus $\hat H$ is a homotopy from $\hat\gamma^e$ to $\hat\delta^e$ relative to $\partial I$.
\qed

Let $p\colon E\longto B$ be a covering, $a\in B$, and $p(e)=a$.
Then we define a function $F\colon\pi_1(B,a)\longto p^{-1}(a)$ by $F([\gamma])\coloneqq\hat\gamma^e(1)$.
This is well-defined by the previous proposition (it is in $p^{-1}(a)$ since $p\circ\hat\gamma^e=\gamma$, so $p\circ\hat\gamma^e(1)=a$).

\bprop

    \benum
        \item If $E$ is path connected, then $F$ is surjective.
        \item If $E$ is simply connected (see homework $2+3$, for every two $a,b\in E$ every two paths between them are homotopic relative to $\partial I$), then it is also injective (it is bijective).
    \eenum

\eprop

\Proof
\benum
    \item Let $p(x)=a$ then $x$ and $e$ are connected since $E$ is path connected, so let $\delta$ be a path $e$ to $x$, then $\widehat{p\circ\delta}^e=\delta$ by uniqueness.
        And so $F[p\circ\delta]=\delta(1)=x$.
        So $F$ is surjective.
    \item Suppose $F[\gamma]=F[\delta]$, then $\hat\gamma^e(1)=\hat\delta^e(1)$.
        Of course $\hat\gamma^e(0)=\hat\delta^e(0)=e$, and so by simple connectivity, $\hat\gamma^e\hdi\hat\delta^e$ since they begin at the same point and end at the same point.
        Now recall that $\gamma=p\circ\hat\gamma^e$ and since homotopy respects the composition of homotopic functions, $\gamma\hdi\delta$.
        \qed
\eenum

So for example, the covering $p\colon{\bb R}\longto S^1$ by $t\mapsto e^{2\pi it}$ is a covering from a simply connected space (since contractible implies simply connected, homework $2$) to $S^1$, this
means that $F\colon\pi_1(S^1,e^{2\pi it})\longto\set{t+n}[n\in{\bb Z}]\cong{\bb Z}$ is a bijection.
Since the fundamental group of a simply connected space is trivial, $S^1$ is not simply connected.

Notice that $F^{-1}$ can be viewed as mapping a curve $\gamma$ which starts at $e$ and ends at $b\in p^{-1}(a)$ to $[p\circ\gamma]$.
So for example, $\pi_1(S^1,1)$ is $\set{[\hbox{the curve which winds around the circle $n$ times}]}[n\in{\bb Z}]$.
This is since a curve from $0$ to $n$ in ${\bb R}$ is mapped to a curve which winds around the circle $n$ times (for negative $n$ this winds in the opposite direction).
Notice that $F$ is a group isomorphism here, since concatenating two curves which wind around the circle $n$ and $m$ times gives a curve which winds around the circle $n+m$ times.
So $\pi_1(S^1)\cong{\bb Z}$.

Now, since $S^1$ is a deformation retract of ${\bb R}^2\setminus\set0$, so the inclusion map is a homotopic equivalence and thus defines an isomorphism of their fundamental groups.
So $\pi_1({\bb R}^2\setminus\set0)\cong\pi_1(S^1)\cong{\bb Z}$ as well.

Now suppose $A\subseteq X$ is a retract, meaning there exists a retraction $r\colon X\longto A$ such that $r\circ\iota={\rm id}_A$.
So $\iota_*\colon\pi_1(A,a)\longto\pi_1(X,a)$ and $r_*\colon\pi_1(X,a)\longto\pi_1(A,a)$ and since this is a functor, $r_*\circ\iota_*=(r\circ\iota)_*=({\rm id}_A)_*={\rm id}_{\pi_1(A,a)}$.
So $\iota_*$ is injective.
Let us summarize this:

\bprop

    If $A\subseteq X$ is a retract, then $\iota_*\colon\pi_1(A,a)\longto\pi_1(X,a)$ is a monomorphism (injective).

\eprop

Now, for example $S^1=\partial D^2\subseteq D^2$ is not a retract: if it were then $\iota_*$ would be a monomorphism $\pi_1(S^1,1)\longto\pi_1(D^2,1)$.
But $\pi_1(S^1,1)\cong{\bb Z}$ and $\pi_1(D^2,1)=1$ since $D^2$ is contractible ($H(x,t)=(1-t)x$).
And there is no monomorphism ${\bb Z}\longto1$.

\bthrm[title=Brouwer Fixed-Point Theorem (for $D^2$), name=fpthrm]

    If $f\colon D^2\longto D^2$ is continuous, then it has a fixed point: a point $x\in D^2$ such that $f(x)=x$.

\ethrm

\Proof suppose not, then $f(x)\neq x$ for all $x$.
Using this $f$ we will construct an $r\colon D^2\longto S^1$ such that $r\circ\iota={\rm id}_{S^1}$, meaning then $S^1$ would be a retract of $D^2$, which we just showed it is not.
For $x\in D^2$, look at the segment which begins at $f(x)$ to $x$, and set $r(x)$ to be the intersection of this line with the segment.
The segment is $f(x)+t(x-f(x))$ and so we want to solve (setting $f(x)=(f_1,f_2)$ and $x=(x_1,x_2)$)
$$ \bigl(f_1+t(x_1-f_1)\bigr)^2 + \bigl(f_2+t(x_2-f_2)\bigr)^2 = 1 $$
which can be solved, and gives a $t$ which is continuous in $f$ and $x$.
Notice that if $x\in S^1$ then by definition $r(x)=x$ since the intersection of the segment with the boundary of the circle is $x$.
So $r$ is indeed a retract, in contradiction.
\qed

\bye

