\input pdfToolbox

\setlayout{horizontal margin=2cm, vertical margin=2cm}
\parindent=0pt
\parskip=4pt plus 2pt minus 2pt

\input ../preamble

\footline={}

\setcounter{section}{2}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\def\printmcount{\the\counter{section}.\the\counter{math counter}}

{\bppbox{rgb{.5 1 .5}}{rgb{0 .4 0}}{rgb{.1 .4 0}}

    \centerline{\setfontandscale{bf}{20pt}Advanced Algorithms}
    \smallskip
    \centerline{\setfont{it}Homework \the\counter{section}}
    \centerline{\setfont{it}Ari Feiglin}

\eppbox}

\bexerc

    Consider the following metric TSP problem:
    \blist
        \item Construct an undirected graph $G=(V,E,w)$ which is a path of $n=2k+1$ vertices, such that for every $1\leq i<n$, $(v_i,v_{i+1})\in E$, $w(v_i,v_{i+1})=1$.
        \item Alter $G$ such that for vertices two steps apart, $w(v_i,v_{i+2})=1+\epsilon$ for some $\epsilon>0$.
        \item Define $G'=(V,E)$ with metric $d$ defined to be $d(u,v)=\delta_G(u,v)$.
    \elist
    \benum
        \item Show and compare the results of the twice-MST and Christofide's algorithms to the optimal solution's cost.
        \item Describe a function $f(n)$ such that when $\epsilon=f(n)$, the analysis of Christofide's algorithm becomes tight as $n$ grows.
    \eenum

\eexerc

\benum
    \item An MST of $G'$ must connect $n$ vertices, so it must have $n-1$ edges, and thus an MST of $G'$ is the original path.
    So our MST is $v_1,v_2,\dots,v_n$ and so an Euler tour on the doubled MST after finding shortcuts is $v_1\to v_2\to\cdots\to v_n\to v_1$.
    This has a weight of
    $$ \sum_i d(v_i,v_{i+1}) + d(v_n,v_1) = n-1 + d(v_n,v_1) $$
    The lightest path from $v_n$ to $v_1$ in $G$ is $v_{2k+1}\to v_{2k-1}\to\cdots\to v_3\to v_1$ which has a weight of $k(1+\epsilon)$.
    Thus the weight of the lightest path, and thus $d(v_n,v_1)$, is 
    $$ 2k+1-1+k(1+\epsilon)=k(3+\epsilon)=\frac{n-1}2(3+\epsilon) = n\cdot\frac{3+\epsilon}2 - \frac{3+\epsilon}2 $$

    For Christofide's, the only vertices in the MST with odd degree are $v_1,v_n$.
    The minimum distance between them is $k(1+\epsilon)$, as proven before.
    And an Euler tour in $T+M$ is just $v_1\to\cdots\to v_n\to v_1$, which has weight $k(3+\epsilon)$.
    So twice-MST and Christofide's have the same approximation here.

    The optimal solution only uses edges of size $1$ and $1+\epsilon$, as otherwise we would have skipped a vertex for nothing.
    The optimal cycle is then
    $$ (v_1 \to v_3 \to \cdots \to v_n) \to (v_{n-1} \to v_{n-2} \to \cdots \to v_2) \to v_1 $$
    Which has a weight of
    $$ k(1+\epsilon) + 2 + (k-2)(1+\epsilon) = (1+\epsilon)(2k-2) + 2 = (1+\epsilon)(n-3) + 2 = n(1+\epsilon) - 1 - 3\epsilon $$

    \item We want
    $$ n\frac{3+\epsilon}2 - \frac{3+\epsilon}2 = \frac32\bigl(n(1+\epsilon) - 1 - 3\epsilon\bigr) $$
    $$ \epsilon n = 4\epsilon $$
    So there is no such $f(n)$.
\eenum

\bexerc

    Consider the greedy algorithm we discussed in class for Set-Cover.
    Describe a family of instances (for arbitrarily large $k$) such that the analysis is tight.

\eexerc

Let us take $k$ to be a power of $2$, then we define $X=[1,2k]^2$ (the set of all integer points $(x_1,x_2)$ where $x_i\in[1,2k]$).
Then let us define the following the sets
$$ S_1 = [1,2k] \times [1,k],\quad S_2 = [1,2k] \times [k,2k] $$
Then let us define a sequence of values $a_0=0,a_1,\dots$ and sets
$$ A_n = [a_{n-1},a_n] \times [1,2k] $$
The optimal solution is $2$, but we want at each step to choose a set $A_n$.
Let
$$ U_n=X\setminus\bigcup_{i<n}A_n X \setminus [1,a_{n-1}] \times [1,2k] = [a_{n-1}+1,2k] \times [1,2k] $$
then we want $A_n$ to maximize $S\cap U_n$.
Note that the size of $S_i\cap U_n$ is $(2k-a_{n-1}-1)k$, and the size of $A_n\cap U_n$ is $(a_n-a_{n-1}-1)2k$, so we want
$$ 2k^2 - a_{n-1}k - k \leq 2a_nk - 2a_{n-1}k - 2k \iff 2k + a_{n-1} + 1 \leq 2a_n \iff a_n \geq k + \frac{a_{n-1}}2 + \frac12 $$
So let us define $a_n=k+a_{n-1}/2+1/2$, and solving this gives
$$ a_n = 2k+1 - \frac{2k+1}{2^n} $$
In order for $a_n\leq2k$ we must have
$$ 1 \leq \frac{2k+1}{2^n} \iff n \leq \log_2(2k+1) $$
So we have that the optimal solution is $\log_2(2k+1)$ as required.

\bexerc

    Consider the following bin packing algorithm: given an item, pack it into the last opened bin.
    If this is impossible, open a new bin.
    Show that $\abs{\rm alg}\leq2\abs{\rm opt}-1$.

\eexerc

Let $U=\set{\mu_1,\dots,\mu_n}$ be our objects, with corresponding weights $\s s(\mu_1),\dots,\s s(\mu_n)$.
Recall that
$$ \sum_{i=1}^n\s s(\mu_i) \leq \abs{\rm opt} $$
let $B_1^A,\dots,B_m^A$ be the bins used by any algorithm $A$, then
$$ \sum_{i=1}^n\s s(\mu_i) = \sum_{i=1}^m\s s(B_i^A) \leq \sum_{i=1}^m1 = m $$
so in particular, $\sum_{i=1}^n\s s(\mu_i)\leq\abs{\rm opt}$.
Now let $B_1,\dots,B_m$ be the bins used by our algorithm, notice that $\s s(B_i)+\s s(B_{i+1})>1$ as otherwise the algorithm would've just combined the two bins.
So
$$ 2\sum_{i=1}^m\s s(B_i) = \sum_{i=1}^{m-1}\bigl(\s s(B_i)+\s s(B_{i+1})\bigr) + \s s(B_1) + \s s(B_n) > m - 1 + \s s(B_1) + \s s(B_n) $$
So we have that
$$ m - 1 + \s s(B_1) + \s s(B_n) < 2\abs{\rm opt} \implies m + \epsilon < 2\abs{\rm opt} + 1 \implies m + \epsilon \leq 2\abs{\rm opt} \implies m \leq 2\abs{\rm opt} - 1 $$
As required.

\def\opt{{\rm opt}}
\bexerc

    Consider an alternative algorithm for the bin-packing algorithm where all the items have weight at least $\delta$.
    In this algorithm, we convert our set of items $U$ to a new set $U'$ where for every $u\in U$ we add an item $u'$ to $U'$ where $w_{u'}$ is $w_u$ rounded up to the nearest multiple of $\frac\delta k$.
    I.e. $w_{u'}=\frac\delta k\cdot\ceil{\frac k\delta w_u}$.
    \benum
        \item Show that $\abs{\opt(U)}\leq\abs{\opt(U')}$.
        \item Show that $\abs{\opt(U')}\leq t\abs{\opt(U)}+O(1)$ for as small of a $t$ as possible.
    \eenum

\eexerc

\benum
    \item Since $w_u\leq w_{u'}$ if $B_1,\dots,B_m$ is a valid allocation of $U'$, it is also a valid allocation of $U$.
    Thus the optimal solution to $U'$ is a solution to $U$.
    \item Notice that
    $$ w_{u'} \leq \frac\delta k\cdot\parens{\frac k\delta w_u+1} = w_u + \frac\delta k $$
    and so now suppose that $\opt$ gives the bins $B_1,\dots,B_m$, we'd have that
    $$ s(B_i') \leq s(B_i) + \frac\delta k\cdot\abs{B_i} \leq 1 + \frac\delta k\cdot\abs{B_i} $$
    and so if we require that $k\geq\abs{B_i}$ for $1\leq i\leq m$ we have that $s(B_i')\leq 1+\delta$.
    Now for each $B_i$ such that $s(B_i')>1$ we can remove the minimal element (which is greater than $\delta$), and this will reduce the bin's weight to at most $1$.
    This minimal element's original weight is at most $\frac12$ (since otherwise, it must be equal to $1$, and we can just require that $\frac k\delta\in{\bb N}$ so $w'=w=1$, and so $s(B_i)=1$).
    If it is $\frac12$, then this means that there is one other element in the bin which is also $\frac12$, but then $w'=\frac\delta k\ceil{\frac k{2\delta}}$ and we can assume $\frac k{2\delta}\in{\bb N}$.
    So $w'=\frac12$ and so $s(B_i)=1$.
    Thus we have that $w<\frac12$ and $w'\leq w+\frac\delta k$, and we want $w'\leq\frac12$, so we require $w+\frac\delta k\leq\frac12$, so $k\geq2\delta\frac1{1-2w}$.
    Thus we also require
    $$ k \geq 2\delta\max_{w_i<\frac12}\frac1{1-2w_i} $$
    and then we have that $w'\leq\frac12$ for all of the weights which we removed, and thus we can pack two weights into a single box.
    So for every two boxes, we create one extra box, thus
    $$ \abs{\rm opt'} \leq 1.5\abs{\rm opt} + O(1) $$
\eenum

\bye

