\input pdfToolbox

\setlayout{horizontal margin=2cm, vertical margin=2cm}
\parindent=0pt
\parskip=3pt plus 2pt minus 2pt

\input preamble

\footline={}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\headline={\pageborder{rgb{1 .5 1}}{rgb{.6 .2 .2}}{5}}

\color rgb{.8 .1 .1}

{\def\boxshadowcolor{rgb{.6 .4 .6}}
\bppbox{rgb{1 .6 1}}{rgb{.6 .1 .1}}{rgb{.4 .1 .1}}

    \centerline{\setfontandscale{bf}{20pt}Fields and Galois Theory}
    \smallskip
    \centerline{\setfont{it}Lectures by Uzi Vishne}
    \centerline{\setfont{it}Summary by Ari Feiglin \setfont{rm}({\tt ari.feiglin@gmail.com})}

\eppbox

\bigskip

\bppbox{rgb{1 .6 1}}{rgb{.6 .1 .1}}{rgb{.4 .1 .1}}
    \section*{Contents}
    
    \tableofcontents
\eppbox

}

\vfill\break

\color{black}

\pageno=1
\newif\ifpageodd
\pageoddtrue
\headline={%
    \hbox to \hsize{\color{black}%
        \ifpageodd\hfil{\it\currsubsection\quad\bf\folio}\global\pageoddfalse%
        \else{\bf\folio\quad\it\currsubsection}\hfil\global\pageoddtrue\fi%
    }%
}

\section{Field Extensions}

Suppose $F\subseteq K$ are fields, then $K$ is certainly also an $F$-vector space and therefore has a dimension and we denote it $[K:F]\coloneqq\dim_FK$.

\bthrm

    Suppose $F\subseteq K$ and $V$ is a $K$-vector space, then $V$ is also a vector space over $F$ as well, and $\dim_FV=[K:F]\dim_KV$.

\ethrm

\Proof Let $B_1\subseteq V$ be a basis for $V$ over $K$ and $B_2\subseteq K$ be a basis for $K$ over $F$, then define $B=\set{\alpha v}[\alpha\in B_2,v\in B_1]$.
This is a basis for $V$ in $F$, it is linearly independent since if $\alpha_1v_1,\dots,\alpha_nv_n\in B$ and $\beta_1,\dots,\beta_n\in F$ then $\sum_{i=1}^n\beta_i\alpha_iv_i=0$ implies $\beta_i\alpha_i=0$
for all $i$ since $B_1$ is a basis, and this means that $\beta_i$ or $\alpha_i$ is zero, but $\alpha_iv_i\in B$ so $\beta_i=0$ as required.
$B$ spans $V$ since for $v\in B$ there exist $v_1,\dots,v_n\in B_1$ and $\alpha_1,\dots,\alpha_n\in K$ such that $v=\sum_{i=1}^n\alpha_iv_i$ and $\alpha_i$ can be written as the linear combination of
elements in $B_2$ by elements of $F$ which gives a linear combination of elements in $B$ of $F$.
So $B$ is indeed a basis for $V$ over $F$.
Finally $B\cong B_2\times B_1$ since $(\alpha,v)\mapsto\alpha v$ is a bijection: it is obviously surjective and $\alpha_1v_1=\alpha_2v_2$ implies $\alpha_1=\alpha_2,v_1=v_2$ since $v_1,v_2$ are independent.
Thus we have
$$ \dim_FV = \abs B = \abs{B_2\times B_1} = [K:F]\dim_K V $$
\qed

In particular if $F\subseteq K\subseteq E$ are fields then $[E:F]=[E:K]\cdot[K:F]$.

The following are methods of constructing fields:
\benum
    \item If $R$ is a commutative ring and $M\triangleleft R$ is a maximal ideal then $\slfrac RM$ is a field.
        Specifically if $R=F[x]$ and $p$ is an irreducible polynomial, $\gen p$ is maximal and $\slfrac{F[x]}{\gen p}$ is a field.
    \item If $F$ is a field, then the set of rational functions is also a field:
        $$ F \subseteq F(x) \coloneqq \set{\frac{f(x)}{g(x)}}[f,g\in F[x],\,g(x)\neq0] $$
        In general if $R$ is an integral domain then its field of fractions/quotients $q(R)\coloneqq\set{\frac ab}[a,b\in R,\,b\neq0]$ is a field.
        And $F(x)$ is the quotient field of $F[x]$.
    \item If $F_0\subseteq F_1\subseteq F_2\subseteq\cdots$ is a chain of fields then so is $\bigcup F_n$ (the theory of fields is inductive, this holds for arbitrary chains, not just inductive ones).
        So for example $F(\lambda_1,\lambda_2,\dots)$ is a field since we can define $F_n=F(\lambda_1,\dots,\lambda_n)$ (the quotient field of $F[\lambda_1,\dots,\lambda_n]$) and the union of this chain
        is $F(\lambda_1,\lambda_2,\dots)$.
\eenum

Let $F$ be a field and $F\subseteq K$ a ring with $a\in K$, we define a homomorphism $F[\lambda]\xvarrightarrow{\,\psi_a\,}K$ defined by $\alpha\mapsto\alpha$ for $\alpha\in F$
and $\lambda\mapsto a$, meaning
$$ \psi_a\parens{\sum\alpha_i\lambda^i} = \sum\alpha_i a^i \qquad (\psi_a(f)=f(a)) $$
In particular $\psi_a$ is a linear transformation from $F$ to $K$, and is called the {\it evaluation homomorphism} at $a$.
The kernel of the homomorphism is
$$ \ker\psi_a = \set{f\in F[\lambda]}[f(a)=0] \triangleleft F[\lambda] $$

\bdefn

    $a\in K$ is {\emphcolor algebraic} if $\ker\psi_a\neq0$ and {\emphcolor transcendental} if the kernel is trivial.

\edefn

If $a$ is transcendental then $\ker\psi_a$ and so $\Im\psi_a=\set{f(a)}[{f\in F[\lambda]}]=F[a]\cong F[\lambda]$.
In fact we get
$$ \vcenter{\ialign{&${}\m@th#{}$\cr
    F &\subseteq& F[a] &\subseteq& F(a) &\subseteq& K\cr 
      &         & \cong &        & \cong\cr
      &         & F[x]  &        & F(x)\cr
    }}
$$

Now if $a$ is algebraic, since $F[x]$ is a euclidean domain and therefore a PID, the kernel has a generator $\ker\psi_a=\gen h=h\cdot F[\lambda]$.
So $h(a)=0$ and $f(a)=0\implies h\divides f$, and $h$ is called the {\it minimal polynomial} of $a$.
And so
$$ \slfrac{F[\lambda]}{\gen h} = \slfrac{F[\lambda]}{\ker\psi_a} \cong \Im\psi_a = \set{f(a)}[{f\in F[\lambda]}] = F[a] = \lspanof{1,a,\dots,a^{n-1}} \subseteq K $$
where $n=\deg h$, since $f(x)=q(x)h(x)+r(x)$ where $\deg r<\deg h=n$ and so $f(a)=r(a)$.
$\set{1,\dots,a^{n-1}}$ is a basis due to $h$ being minimal, a zeroing linear combination would give a zeroing polynomial of $a$ of degree less than $h$.
This means that the dimension of $F[a]$ as an $F$-vector space is $n$, ie. $\bigl[F[a]:F\bigr]=n$.

Since $K$ is an integral domain and therefore so too is $F[a]$ and this means that $\gen h$ is a prime ideal (since $\slfrac RI$ is an integral domain if and only if $I$ is prime), this means
that $h$ is a prime (irreducible) polynomial.
And since $F[a]$ is a PID, prime and maximal ideals are one and the same, so $\gen h$ is maximal and therefore $\slfrac{F[\lambda]}{\gen h}\cong F[a]$ is a field.
Let us summarize this:

\bprop

    Let $F\subseteq K$ where $K$ is an integral domain and $a\in K$ is algebraic in $F$, let $h_a$ be its minimal polynomial.
    Then $(1)$ $h_a$ is irreducible, $(2)$ $F[a]$ is a field, $(3)$ $\bigl[F[a]:F\bigr]=\deg h_a$.

\eprop

So for example let $a\in K\setminus F$ be algebraic then $F\subseteq F[a]\subseteq K$ and suppose $[K:F]=p$ is prime.
Then $p=[K:F]=[K:F[a]]\cdot[F[a]:F]$, and since $a\in F[a]\setminus F$ this means $[F[a]:F]>1$ so $[F[a]:F]=p$ and $[K:F[a]]=1$ since $p$ is prime so $F[a]=K$.

\bcoro

    Suppose $F$ is a field and $F\subseteq K$ is an integral domain with finite dimension.
    Then every element of $K$ is algebraic and $K$ is a field.

\ecoro

\Proof Let $a\in K$ then $[K:F]=[K:F[a]]\cdot[F[a]:F]$ so $[F[a]:F]$ is finite.
If $a$ were transcendental then $F[a]\cong F[x]$ and $F[x]$ has infinite dimension over $F$.
$K$ is a field since every $a\in K$ must have a multiplicative inverse, since $F[a]$ is a field.
\qed

Notice that $[F[a,b]:F[a]]\leq [F[b]:F]$ since if $h_b$ is $b$'s minimal polynomial in $F$ then it is also a zeroing polynomial in $F[a]$.
This means that
$$ [F[a,b]:F] = [F[a,b]:F[a]]\cdot[F[a]:F] \leq [F[b]:F]\cdot[F[a]:F] $$

\bcoro

    Let $F$ be a field and $K$ a field extension, define
    $$ {\rm Alg}_F(K) \coloneqq \set{a\in K}[\hbox{$a$ is algebraic over $F$}] . $$
    This is a field.
    Furthermore $F\subseteq{\rm Alg}_F(K)$ is an algebraic extension (all elements of ${\rm Alg}_F(K)$ are algebraic in $F$), and ${\rm Alg}_F(K)\subseteq K$ is a purely transcendental extension (all
    elements in $K\setminus{\rm Alg}_F(K)$ are transcendental in ${\rm Alg}_F(K)$).

\ecoro

\Proof Notice that $F[a\cdot b],F[a+b]\subseteq F[a,b]$ and so $[F[a,b]:F]\leq[F[b]:F]\cdot[F[a]:F]<\infty$, so ${\rm Alg}_F(K)$ is closed under addition and multiplication (and obviously additive inverses).
For $a$ algebraic, $F[a]$ is a field so $a^{-1}\in F[a]$ and so $F[a^{-1}]\subseteq F[a]$ and therefore $[F[a^{-1}]:F]<\infty$ so $a^{-1}$ is algebraic as well (and so by symmetry $F[a]=F[a^{-1}]$).
So ${\rm Alg}_F(K)$ is indeed a field.

To show that ${\rm Alg}_F(K)\subseteq K$ is a pure transcendental extension, notice that if $F_1\subseteq F_2\subseteq F_3$ where $F_1\subseteq F_2$ is algebraic, if $a\in F_3$ is algebraic in $F_2$
it is also algebraic in $F_1$.
Indeed if $f\in F_2[x]$ such that $f(a)=0$, let its coefficients be $b_i$ then $a$ is algebraic in $F_1[b_0,\dots,b_n]$ and so
$$ [F_1[b_0,\dots,b_n,a]:F_1[b_0,\dots,b_n]] = [F_1[b_0,\dots,b_n,a]:F_1[b_0,\dots,b_n]]\cdot[F_1[b_0,\dots,b_n]:F_1] $$
and this is finite since $b_0,\dots,b_n$ are algebraic in $F_1$ as they are in $F_2$, so both terms are finite.
So if $K$ had any algebraic numbers not in ${\rm Alg}_F(K)$, they would be algebraic in $F$ and thus in ${\rm Alg}_F(K)$ in contradiction.
\qed

\bprop

    Let $F$ be a field and $f\in F[\lambda]$ be irreducible, then there exists a field extension $F\subseteq K$ such that $f$ has a root in $K$, and $[K:F]=\deg f$.

\eprop

\Proof since $f$ is irreducible, $\gen f$ is prime and $F[\lambda]$ is a PID so it is maximal.
So $K\coloneqq\slfrac{F[\lambda]}{\gen f}$ is a field, and its dimension is $\deg f$, since it can be generated by $\set{1,x,\dots,x^{\deg f-1}}$.
Now recall that by the second isomorphism theorem, $\slfrac F{F\cap\gen f}\cong\slfrac{F+\gen f}{\gen f}\subseteq\slfrac{F[\lambda]}{\gen f}=K$.
But since elements of $\gen f$ are multiples of $f$, which is disjoint from $F$, so $F\cap\gen f=(0)$ so $\slfrac F{F\cap\gen f}\cong F$, and so $F$ can be embedded into $K$ and is thus for all intents and
purposes, a subfield of $K$.
Now define $\alpha\coloneqq\lambda+\gen f$, and suppose $f(\lambda)=\sum_{i=0}^n a_i\lambda^i$ where $a_i\in F$ (viewing $f$ as a polynomial over $K$, $a_i$ is actually $a_i+\gen f$).
Then
$$ f(\alpha) = \sum_{i=0}^na_i(\lambda + \gen f)^i = \sum_{i=0}^na_i(\lambda^i+\gen f) = \sum_{i=0}^na_i\lambda^i + \gen f = f + \gen f = \gen f = 0_K $$
so $\alpha$ is indeed a root of $f(\lambda)$, as required.
\qed

\bcoro

    Let $F$ be a field and $f\in F[\lambda]$ any polynomial.
    Then there exists a field extension $F\subseteq K$ such that $f$ has a root in $K$ and $[K:F]\leq\deg f$.

\ecoro

\Proof find $f$'s irreducible factorization $f=f_1\cdots f_t$, then extend $F$ to a field $K$ such that $f_1$ has a root in $K$, and by above $[K:F]=\deg f_1\leq\deg f$.
\qed

\bdefn

    Let $F$ be a field, and $f$ a polynomial over $F$.
    A field $F\subseteq K$ {\emphcolor splits $f$} if there exist $\alpha_1,\dots,\alpha_n\in K$ such that $f(\lambda)=(\lambda-\alpha_1)\cdots(\lambda-\alpha_n)$.

\edefn

\bthrm

    Every polynomial $f$ over a field $F$ has a field $K$ which splits it, such that $[K:F]\leq(\deg f)!$.

\ethrm

\Proof by induction on $n=\deg f$.
For $n=1$ then $f$ already has a root, and so take $F=K$ and $[K:F]=1=(\deg f)!$.
Now suppose $\deg f=n+1$, then by above there exists a field extension $F\subseteq K_0$ such that there exists an $\alpha_1\in K_0$ such that $f(\alpha_1)=0$ and $[K_0:F]\leq\deg f=n+1$.
And so $(\lambda-\alpha_1)\divides f(\lambda)$, so $f(\lambda)=(\lambda-\alpha_1)g(\lambda)$.
Then $\deg g=n$, and $g$ is a polynomial over $K_0$, so there exists a field extension $F\subseteq K_0\subseteq K$ such that $g(\lambda)=(\lambda-\alpha_2)\cdots(\lambda-\alpha_{n+1})$ for $\alpha_i\in K$
and $[K:K_0]\leq n!$.
Then $f(\lambda)=(\lambda-\alpha_1)\cdots(\lambda-\alpha_{n+1})$ for $\alpha_i\in K$ and $[K:F]=[K:K_0][K_0:F]\leq (n+1)n!=(n+1)!$.
\qed

Notice the following
\benum
    \item the split of a polynomial over any field into its roots is unique,
    \item the number of roots is $\leq\deg f$.
\eenum

Recall that a field $F$ is {\it algebraically closed} if it splits every polynomial in $F[\lambda]$.

\bdefn

    Let $F$ be a field, then $F\subseteq\overline F$ is an {\emphcolor algebraic closure} of $F$ if $\overline F$ is algebraically closed.

\edefn

\bnote

    Every field has a unique (up to isomorphism) algebraic closure.

\enote

So let $f(\lambda)\in F[\lambda]$, then $f(\lambda)\in\overline F[\lambda]$ and so $f=(\lambda-\alpha_1)\cdots(\lambda-\alpha_n)$ for $\alpha_i\in\overline F$.
Then take $F\subseteq K=F[\alpha_1,\dots,\alpha_n]\subseteq\overline F$, it can be shown that $[K:F]\leq(\deg f)!$.

Now suppose $F\subseteq K$ are fields, and $E$ is a field which $F$ is embeddable into, suppose $\phi\colon F\longembeds E$ is an embedding.
An embedding $\phi'\colon K\longembeds E$ is an {\it extension} of $\phi$ if $\phi'\bigr|_F=\phi$.
Denote
$$ \eta_{F\subseteq K}^E \coloneqq \#\set{\hbox{$\phi'$ is an extension of $\phi$}} $$
where $\phi$ is held constant and understood.
Then

\bprop

    Suppose $K=F[\alpha]$, then $\eta_{F\subseteq K}^E$ is equal to the number of roots the minimal polynomial of $\alpha$ in $F$ has in $E$.

\eprop

\Proof since $\alpha$ generates $K$ over $F$, every extension of $\phi$ is defined by its image on $\alpha$.
Let $h$ be the minimal polynomial of $\alpha$ over $F$.
Denote $\hat b\coloneqq\phi(b)$ for all $b\in F$, and this definition extends to polynomials, $\varwidehat{\sum_{i=0}^n b_ix^i}=\sum_{i=0}\hat b_ix^i$.
Then if $\phi'$ is an extension of $\phi$,
$$ \hat h(\phi'(\alpha)) = \phi'(h(\alpha)) = \phi'(0) = 0 $$
this is since if $h(\lambda)=\sum_{i=0}^n a_i\lambda^i$, then $\hat h(\lambda)=\sum_{i=0}^n\hat a_i\lambda^i$, so
$$ \hat h(\phi'(\alpha)) = \sum_{i=0}^n\hat a_i\phi'(\alpha)^i = \sum_{i=0}^n\phi(a_i)\phi'(\alpha)^i = \sum_{i=0}^n\phi'(a_i)\phi'(\alpha)^i = \phi'\parens{\sum_{i=0}^n a_i\alpha^i} = \phi'(h(\alpha)) $$
so $\phi'(\alpha)$ must be one of $\hat h$'s roots, precisely as stated.
\qed

\bdefn

    A polynomial $f$ which splits over $E$ is called {\emphcolor separable} in $E$ if its linear factors are distinct (ie. all of its roots in $E$ are distinct).

\edefn

\bthrm

    Let $F\subseteq K$ be a finite extension (meaning $[K:F]<\infty$), and let $\phi\colon F\longembeds E$ be a given embedding.
    Then
    \benum
        \item $\eta_{F\subseteq K}^E\leq[K:F]$,
        \item if $K$ is generated by the roots of $f$, assuming that $E$ splits $f$, then $1\leq\eta_{F\subseteq K}^E$,
        \item if $f$ is separable over $E$, then $\eta_{F\subseteq K}^E=[K:F]$.
    \eenum

\ethrm

\Proof suppose $K=F[\alpha_1,\dots,\alpha_n]$ (the generators of $K$ can be taken to be the basis of $K$ as an $F$-vector space).
We prove this by induction on $n$, for $n=1$ this is given by the previous proposition, since $\eta_{F\subseteq K}^E$ is the number of roots $h$ has in $E$, and $[K:F]=\deg h$ which is at least this.
Define $F_1\coloneqq F[\alpha_1]$, then
$$ \eqalign{
    \eta_{F\subseteq K}^E &= \#\set{\hbox{$\phi''\colon K\longto E$ is an extension of $\phi$}} \cr
    &= \#\bigcup\set{{\hbox{$\phi''\colon F_1\longto E$ is an extension of $\phi'$}}}[\hbox{$\phi'\colon F_1\longto E$ is an extension of $\phi$}]\cr
    &= \sum_{\phi'}\eta_{F_1\subseteq K}^E = \eta_{F\subseteq F_1}^E\cdot\eta_{F_1\subseteq K}^E \subseteq [F_1:F]\cdot[K:F_1] = [K:F]
} $$

For $(2)$, by the assumption there is an extension of $F\longembeds E$ to $F_1\longembeds E$, and continue inductively.
For $(3)$, since $f$ is separable, makes the bound an equality.
\qed

\bdefn

    Let $f$ be a polynomial over $F$, a field $F\subseteq K$ is a {\emphcolor splitting field} if it is the smallest field in which the polynomial splits.

\edefn

Notice that if $K$ is a splitting field, it is of the form $K=F[\alpha_1,\dots,\alpha_n]$ where $\alpha_i$ are roots of the polynomial, so they are algebraic.
This means that $[K:F]\leq\prod_i[F:\alpha_i]<\infty$.

Furthermore, if $K$ is a splitting field of $f$, then it is generated by the roots of $f$: $K=F[\alpha_1,\dots,\alpha_n]$, then if $E$ is any field which splits $f$, we have $\eta^E_{F\subseteq K}\geq1$,
meaning there exists an embedding $K\embeds E$ which extends the embedding $F\embeds E$.
And in particular if $K,K'$ are two splitting fields of $f$, there exists two embeddings $K\embeds K'$ and $K'\embeds K$, which means $[K:F]=[K':F]$ and so $K$ and $K'$ are isomorphic as $F$-vector spaces.
And so $K\cong K'$ as fields.

Recall that there exists a unique ring homomorphism $f\colon{\bb Z}\longto F$, and $\slfrac{{\bb Z}}{\ker f}\cong\Im f\subseteq F$.
Since $\Im f$ is a subring of $F$, it is an integral domain and so $\ker f$ is a prime ideal.
Thus $\ker f=p{\bb Z}$ for $p$ prime or $0$, and this $p$ is called {\it $F$'s characteristic}.
In other words $F$ has characteristic $p$ if and only if $1+\cdots+1=0$ ($p$ times) since then $p\in\ker f$ and so $(p)\subseteq\ker f$, but ${\bb Z}$ is a PID and so $(p)$ is maximal.
And $F$ has characteristic $0$ if $1+\cdots+1$ is never zero.

If $F$ has characteristic $0$, then $f$ is an embedding into $F$, so ${\bb Z}\subseteq F$ and since it is a field ${\bb Q}\subseteq F$, up to embedding.
And for characteristic $p$, $\slfrac{{\bb Z}}{p{\bb Z}}={\bb F}_p\subseteq F$.

Notice that in characteristic $p$, $\binom pk=\frac{p!}{k!(p-k)!}$ is zero for $k\neq0,p$.
$$ (a+b)^p = \sum_{k=0}^p\binom pka^kb^{p-k} = a^p + b^p $$
And so $e(x)=x^p$ is a field homomorphism $F\longto F^p=\set{x^p}[x\in F]$, and it has a trivial kernel, and so $F\cong\slfrac F{\ker f}\cong F^p$.

\bdefn

    We define the {\emphcolor derivative} over a field $F$ to be the function $F[\lambda]\longto F[\lambda]$ defined by
    $$ \parens{\sum_{i=0}^n\alpha_i\lambda^i}' = \sum_{i=1}^n\alpha_i\cdot i\lambda^{i-1} $$

\edefn

It is trivial to show that $(f+g)'=f'+g'$ and $(fg)'=fg+f'g$, meaning that $(f^2g)'=f^2g'+2ff'g$.
This means that if $f^2\divides h$ then $f\divides h'$.
In particular if $f$ is not separable, then there exists some $(\lambda-\alpha)^2$ which divides $f$ over a field which splits it, then $\lambda-\alpha$ divides $f'$, meaning $f'(\alpha)=0$.
But this means that $f'=0$, so $\alpha_ii=0$ for all $i$, and so if $p$ doesn't divide $i$ this means $i\neq0$ so $\alpha_i=0$.
Thus
$$ f(\lambda) = \sum_{p\divides i}\alpha_i\lambda^i = \sum_j\alpha_{pj}(\lambda^p)^j $$
So we get that

\bprop

    Let $f$ be irreducible over a field of characteristic $p>0$, then $f$ is not separable if and only if $f'=0$ if and only if $f(\lambda)=g(\lambda^p)$ for some polynomial $g$.

\eprop

\bexam

    Let $\lambda^p-a$ be a polynomial over $F$ of characteristic $p$, and $\alpha$ a root in a field which splits it.
    Then
    $$ \lambda^p-a = \lambda^p-\alpha^p = (\lambda-\alpha)^p $$
    so $\lambda^p-a$ is not separable (which we can see since it is $g(\lambda^p)$ for $g(\lambda)=\lambda-a$).

\eexam

\bdefn

    Let $K/F$ be a field extension (meaning $F\subseteq K$), then an automorphism of $K$ over $F$ is an automorphism $\sigma\colon K\longto K$ which holds $F$ constant: $\sigma(a)=a$ for all
    $a\in F$.

\edefn

Notice that all field homomorphisms are either injective or trivial, since the kernel is an ideal and fields only have trivial ideals, so if $\sigma$ is a field homomorphism there is no need
to check injectivity.
And $\sigma(ax)=\sigma(a)\sigma(x)=a\sigma(x)$ for $a\in F$ and $x\in K$ so $\sigma$ is an $F$-linear transformation, so if $[K:F]$ is finite $\sigma$ must be surjective.
Thus in the case that $K/F$ is a finite field extension, all monomorphisms of $K$ over $F$ are automorphisms.

\bdefn

    Let $K/F$ be a field extension, then define its {\emphcolor Galois group} to be
    $$ \Galof{K/F} \coloneqq \set{\sigma}[\hbox{$\sigma$ is an automorphism of $K$ over $F$}] $$
    and this is indeed a group relative to composition.

\edefn

Notice that if $K/F$ is a field extension and $\alpha\in K$ algebraic.
Let $h$ be its minimal polynomial and $\sigma\in\Galof{K/F}$, then
$$ h(\sigma(\alpha)) = \sigma(h(\alpha)) = \sigma(0) = 0 $$
This is since $\sigma\parens{\sum_ia_i\alpha^i}=\sum_i\sigma(a_i)\sigma(\alpha)^i=\sum_ia_i\sigma(\alpha)^i=h(\sigma(\alpha))$.
So permutations in Galois groups map roots to roots of polynomials.

So for example, let $G=\Galof{{\bb Q}[\sqrt3]/{\bb Q}}$ and $\lambda^2-3=(\lambda-\sqrt3)(\lambda+\sqrt3)$ and so $\sigma$ must map $\sqrt3$ to $\pm\sqrt3$.
And since all automorphisms of ${\bb Q}[\sqrt3]$ over ${\bb Q}$ are defined by $\sqrt3$'s image,
$$ G = \set{1,\sqrt3\xvarmapsto{\sigma}-\sqrt3} \cong {\bb Z}_2 $$
And similarly let $G=\Galof{{\bb Q}[\sqrt3,\sqrt2]/{\bb Q}}$, $\sqrt3$ must be mapped to $\pm\sqrt3$ (due to $\lambda^2-3$) and $\sqrt2$ must be mapped to $\pm\sqrt3$, so
$$ G = \set{1,\,\stackmath{\sqrt2\mapsto\sqrt2\cr\sqrt3\mapsto-\sqrt3},\,\stackmath{\sqrt2\mapsto-\sqrt2\cr\sqrt3\mapsto\sqrt3},\,\stackmath{\sqrt2\mapsto-\sqrt2\cr\sqrt3\mapsto-\sqrt3}} \cong
{\bb Z}_2\times{\bb Z}_2 $$

Notice that if $K$ has characteristic $p$, every automorphism must keep elements of ${\bb F}_p$ constant (since $\sigma(1)=1$).
And if $K$ has characteristic $0$, every automorphism must keep elements of ${\bb Q}$ constant (since $\sigma(a/b)=\sigma(a)/\sigma(b)=a/b$).
So let $F_0$ be the characteristic field of $K$ (either ${\bb F}_p$ or ${\bb Q}$), so
$$ \Aut(k) = \Galof{K/F_0} $$

\bdefn

    Let $K$ be a field, then for every subfield $G\leq\Autof K$, define the {\emphcolor fixed-point field},
    $$ K^G \coloneqq \set{a\in K}[\forall\sigma\in G\colon \sigma(a)=a] $$
    This is indeed a field.

\edefn

Notice that if $F\subseteq K$ is a subfield, then $\Galof{K/F}$ is a subgroup of $\Aut(K)$.
And if $G\leq\Aut(K)$ is a subgroup, then $K^G$ is a subfield of $K$.
So we have the following correspondences:

\penalty-100\medskip
{\tabskip=0pt plus1fil
\offinterlineskip\halign to\hsize{$#$\hfil\tabskip=.25cm&\hfil$#$\hfil\tabskip=.25cm&$#$\hfil\tabskip=0pt plus1fil\cr
& \xvarmapsto{\mathtowd{3cm}{\Galof{K,\bullet}}}\cr
\set{\hbox{Subgroups of $\Autof K$}} & &\set{\hbox{Subfields of $K$}}\cr
& \xvarmapsfrom{\mathtowd{3cm}{K^{\textstyle\bullet}}}\cr
}}
\medskip

And if $F\subseteq K$ is a subfield, and $F\subseteq L\subseteq K$ is a field between them, $\Galof{K,L}$ is a subgroup of $\Galof{K/F}$ (since $\sigma\in\Galof{K/L}$ keeps elements of $L$, and thus $F$
constant).
And if $G\leq\Galof{K/F}$, then $K^G$ is a field between $F$ and $K$.
So we have

\medskip
{\tabskip=0pt plus1fil
\offinterlineskip\halign to\hsize{$#$\hfil\tabskip=.25cm&\hfil$#$\hfil\tabskip=.25cm&$#$\hfil\tabskip=0pt plus1fil\cr
& \xvarmapsto{\mathtowd{3cm}{\Galof{K,\bullet}}}\cr
\set{\hbox{Subgroups of $\Galof{K/F}$}} & &\set{\hbox{Fields between $F$ and $K$}}\cr
& \xvarmapsfrom{\mathtowd{3cm}{K^{\textstyle\bullet}}}\cr
}}
\medskip

Some properties:
\benum
    \item If $L_2\subseteq L_1$ then $\Galof{K/L_2}\supseteq\Galof{K/L_1}$ since an automorphism which keeps elements of $L_1$ constant keeps elements of $L_1$ constant.
    \item If $H_2\subseteq H_1$ then $K^{H_2}\supseteq K^{H_1}$ since if $a$ is held constant by every $\sigma\in H_1$, it is held constant by every $\sigma\in H_2$.
    \item For every $L$, $L\subseteq K^{\Galof{K/L}}$ since $K^{\Galof{K/L}}$ are elements held constant by every automorphism in $\Galof{K/L}$, which includes all elements of $L$ by definition.
    \item For every $H$, $H\subseteq\Galof{K/K^H}$ since for $\sigma\in H$ every element of $K^H$ is held constant.
\eenum

\bdefn

    Let $X,Y$ be posets (partially ordered sets), then a pair of functions $\alpha\colon X\longto Y$ and $\beta\colon Y\longto X$ is an {\emphcolor Galois correspondence} if
    \benum
        \item $\alpha$ and $\beta$ reverse order, meaning if $x_1\leq x_2$ then $\alpha(x_2)\leq\alpha(x_1)$ and similar for $\beta$,
        \item for every $x\in X$ and $y\in Y$, $x\leq\beta(\alpha(x))$ and $y\leq\alpha(\beta(y))$.
    \eenum
\edefn

For example, let $X$ and $Y$ both be the lattice of subgroups of a group $G$, $\alpha=\beta\colon H\mapsto C_G(H)$.
But our important example is $\alpha\colon F\mapsto\Galof{K/F}$ and $\beta\colon H\mapsto K^H$.

\blemm

    $\alpha,\beta$ form a Galois correspondence if and only if for all $x\in X$ and $y\in Y$ $y\leq\alpha(x)\iff x\leq\beta(y)$.

\elemm

\Proof suppose $\alpha,\beta$ form a Galois correspondence.
If $x\leq\beta(y)$ then $y\leq\alpha(\beta(y))\leq\alpha(x)$ and similar for $\beta$, so we get the desired result.
Now suppose $y\leq\alpha(x)\iff x\leq\beta(y)$.
Since $\beta(y)\leq\beta(y)$, we get $y\leq\alpha(\beta(y))$, similar for $\beta(\alpha(x))$.
And if $x\leq x'$ then $x\leq x'\leq\beta(\alpha(x'))=\beta(y)$ which is equivalent to $\alpha(x')=y\leq\alpha(x)$.
Similar for $\beta$.
\qed

\bprop

    Let $\alpha,\beta$ be a Galois correspondence.
    Then
    \benum
        \item $\alpha\circ\beta\circ\alpha=\alpha$ and $\beta\circ\alpha\circ\beta=\beta$.
        \item $\beta(\alpha(x))=x$ if and only if $x\in\beta(Y)$ and $\alpha(\beta(y))=y$ if and only if $y\in\alpha(X)$.
        \item $\alpha$ and $\beta$ are inverses as functions between $\beta(Y)$ and $\alpha(X)$.
    \eenum

\eprop

\Proof 
\benum
    \item Since $x\leq\beta(\alpha(x))$, we get $\alpha\beta\alpha(x)\leq\alpha x$.
    On the other hand let $y=\alpha(x)$ then $y\leq\alpha\beta y=\alpha\beta\alpha(x)$, so we have equality.
    \item This is direct from $(1)$, since if $\alpha\beta(y)=y$ then trivially $y\in\alpha(X)$, and if $y\in\alpha(X)$ then $y=\alpha(x)$ so $\alpha\beta(y)=\alpha\beta\alpha(x)=\alpha(x)=y$.
    \item This is direct from $(2)$.
    \qed
\eenum

In particular $\alpha(X)$ is isomorphic to the reverse order of $\beta(X)$, $\alpha(X)\cong\beta(X)^{\rm op}$.

\bdefn

    A field extension $K/F$ is a {\emphcolor separable extension} if the minimal polynomial of every $a\in K$ over $F$ is separable (meaning $f$ splits into distinct linear factors over its splitting field).
    And it is a {\emphcolor normal extension} if the minimal polynomial of every $a\in K$ over $F$ splits over $K$.
    Equivalently for every irreducible polynomial $f$ over $F$, if $f$ has a root in $K$ then $f$ splits in $K$.
    If it is both a normal and separable extension then it is called a {\emphcolor Galois extension}.

\edefn

\bthrm

    Let $K/F$ be a finite field extension, then the following are equivalent:
    \benum
        \item $K/F$ is a Galois extension,
        \item $K$ is the splitting field of a separable polynomial over $F$,
        \item $F=K^G$ for some $G\leq\Autof K$,
        \item $F=K^{\Galof{K/F}}$,
        \item $\abs{\Galof{K/F}}=[K:F]$.
    \eenum

\ethrm

\Proof $(1)\implies(2)$: suppose $K=F[a_1,\dots,a_n]$, then since $K/F$ is separable the minimal polynomial $f_i$ of every $a_i$ is separable (meaning its linear factors are distinct in its splitting field).
By normality, since $f_i$ has a root in $K$ it splits, and the factors must be distinct.
Define $f=\prod f_i$, which is separable and splits over $K$.
$K$ must be the splitting field of $f$ since $f$ splits into distinct linear terms over $K$ and $K$ is generated from its roots.

$(2)\implies(5)$: we showed that if $K$ is generated by the roots of $f$ which has a splitting field $E$, then $\eta_{F\subseteq K}^E=[K:F]$.
Take $E=K$ so $\eta_{F\subseteq K}^K=[K:F]$.
Extensions of $F\embeds K$ to $K\embeds K$ are simply automorphisms which hold $F$ constant (since field homomorphisms are either injective or trivial, and if it holds $F$ constant it cannot be trivial).
Thus $\eta_{F\subseteq K}^K=\abs{\Galof{K/F}}$, so we have $\abs{\Galof{K/F}}=[K:F]$.

$(2)\implies(4)$: let $F'=K^{\Galof{K/F}}$ then by the Galois correspondence, $F\subseteq F'$.
We now that $(2)\implies(5)$ so $\abs{\Galof{K/F'}}=[K:F']$ and $\abs{\Galof{K/F}}=[K:F]$
Since $\Galof{K/F'}=\alpha\beta\alpha(F)$ we know that $\Galof{K/F'}=\Galof{K/F}$ so $[K:F]=[K:F']$ and $F\subseteq F'$ so $F=F'$.

$(4)\implies(3)$ is trivial.

$(3)\implies(1)$: let $a\in K$ and let $g$ be its minimal polynomial in $F$.
Let $a_1,\dots,a_k$ be its roots in $K$, then let $h=\prod(\lambda-a_i)\in K[\lambda]$, so $h\divides g$ in $K[\lambda]$.
Now let $\sigma\in G$, this will permute a root of $g$ to another root of $g$ which is in $K$, we have $h\in K^G[\lambda]=F[\lambda]$ (since $a_i$ is mapped to $a_j$), we then get that $h\divides h$ in $F$.
But $g$ is the minimal polynomial so $h=g$, meaning the minimal polynomial of every $a\in K$ splits into distinct linear terms.

$(5)\implies(4)$: let $G=\Galof{K/F}$ and $F'=K^G$, then it satisfies the condition for $(3)$, which implies $(1)$ which implies $(5)$, so we get $\abs{\Galof{K/F'}}=[K:F']$.
Again since $\Galof{K/F'}=\alpha\beta\alpha(F)=\Galof{K/F}$, we get $[K:F]=[K:F']$ and $F\subseteq F'$ so $F=F'$.
\qed

Notice that if $F\subseteq L\subseteq K$ are fields such that $K/F$ is a Galois extension, then $K/L$ is also a Galois extension, since if the minimal polynomial of $a\in K$ over $F$ splits, then since the
minimal polynomial of $a$ in $L$ divides it, it must also split.
Thus
$$ K^{\Galof{K/L}} = L $$
So if we look at our previous diagram

\penalty-100\medskip
{\tabskip=0pt plus1fil
\offinterlineskip\halign to\hsize{$#$\hfil\tabskip=.25cm&\hfil$#$\hfil\tabskip=.25cm&$#$\hfil\tabskip=0pt plus1fil\cr
& \xvarmapsto{\mathtowd{3cm}{\alpha=\Galof{K,\bullet}}}\cr
\set{\hbox{Subgroups of $\Galof{K/F}$}} & &\set{\hbox{Fields between $F\subseteq K$}}\cr
& \xvarmapsfrom{\mathtowd{3cm}{\beta=K^{\textstyle\bullet}}}\cr
}}
\medskip

We have that $\beta\alpha=1$, meaning we know that for every $F\subseteq L\subseteq K$ there exists a subgroup of $\Galof{K/F}$ such that $K^G=L$.
But for which subgroups $H\leq G$ is there a field $F\subseteq L\subseteq K$ such that $\Galof{K/L}=G$?

\blemm[title=Artin's Lemma, name=artinlemma]

    Let $H\leq\Autof K$ be a finite subgroup, then $[K:K^H]\leq\abs H$

\elemm

\Proof let $H=\set{\sigma_1=1,\sigma_2,\dots,\sigma_n}$, $n<m$, and $x_1,\dots,x_m\in K$.
We need to show that $x_1,\dots,x_m$ are linearly dependent in $K$ over $K^H$.
So we'd like to find $a_1,\dots,a_m\in K^H$ such that $\sum_ia_ix_i=0$.
Applying $\sigma_i\in H$, since $a_j\in K^H$ we have that by definition $\sigma_ia_j=a_j$ so
$$ \sigma_i\parens{\sum_ja_jx_j} = \sum_ja_j\sigma_i(x_j) = 0 $$
Let $X$ be an $n\times m$ matrix defined by $X=(\sigma_i(x_j))_{ij}$ and $\vec a=(a_1,\dots,a_m)^\top$.
Then we need to solve for $\vec a$ in
$$ X\vec a = 0 $$
But $X$ is an $M_{n\times m}(K)$ matrix, so it must have a nontrivial nullspace, meaning there exists a solution $\vec a\neq0$ in $K$.
Recall that our goal is to find such a $\vec a$ in $K^H$.

Let us choose one non-trivial solution $\vec a$ such that its number of zeroes is minimal (meaning we choose $\vec a$ such that $\#\set{1\leq i\leq m}[a_i=0]$ is minimal).
We can reorder the solutions to assume that $a_1\neq0$, and since $a_1^{-1}\vec a$ is also a solution, we can assume $a_1=1$.
Now we claim that $a_i\in K^H$ for all $1\leq i\leq m$, and so once we prove this we have finished.
Suppose that $a_2\notin K^H$, then there exists a $\sigma_k\in H$ such that $\sigma_k(a_2)\neq a_2$.
Then we know that $\sum_ja_j\sigma_i(x_j)=0$ for all $i$, so compose this with $\sigma_k$ to get
$$ \sum_j\sigma_k(a_j)\sigma_k(\sigma_i(x_j)) = 0 $$
since $\sigma_k\sigma_i\in H$, this is just a permutation of the indexing of $i$, so we still have
$$ \sum_j\sigma_k(a_j)\sigma_k(x_j) = 0 $$
meaning $(1,\sigma_k(a_2),\dots,\sigma_k(a_m))$ is another solution to the system of equations, and since the set of solutions form a vector space, this means that
$(0,a_2-\sigma_k(a_2),\dots,a_m-\sigma_k(a_m))$ is a solution.
This is a non-trivial solution, but it has fewer zeroes than $\vec a$ since now the first coefficient is zero (and all zero coefficients in $\vec a$ remain zero here), in contradiction.
\qed

Since $K^H$ is Galois by the above theorem, we have $[K:K^H]=\abs{\Galof{K/K^H}}$.
Now, $H\subseteq\Galof{K/K^H}$ by the definition of a Galois correspondence, and so we have $\abs H\leq\abs{\Galof{K/K^H}}=[K:K^H]\leq\abs H$.
Thus $H=\Galof{K/K^H}$.

And so we have shown that $H\mapsto K^J$ and $L\mapsto\Galof{K/L}$ are inverse functions:

\bthrm[title=The Fundamental Theorem of Galois Theory, name=ftgalois]

    Let $K/F$ be a finite-dimensional field extension.
    Then the maps $H\mapsto K^J$ and $L\mapsto\Galof{K/L}$ are inverse functions which invert order between fields in between $F$ and $K$ and subgroups of $\Galof{K/F}$.

\ethrm

\bcoro

    For every finite-dimensional Galois field extension, there is a finite number of in-between fields.

\ecoro

\Proof since the number of in-between fields is equal to the number of subgroups of $\Galof{K/F}$, we need simply to show that $\Galof{K/F}$ is finite.
This is since $\abs{\Galof{K/F}}=[K:F]$ since $K/F$ is Galois, and by assumption this is finite.
\qed

\bcoro

    Let $G=\Galof{K/F}$, then $H$ is normal in $G$ if and only if $\sigma(K^H)=K^H$ for every $\sigma\in G$.

\ecoro

\Proof if $H$ is normal in $G$, then let us consider the map $\sigma\mapsto\sigma\bigl|_{K^H}$ from $\Galof{K/F}$ to $\Galof{K^H/F}$.
The kernel of this is the set of all permutations which hold $K^H$ constant, $\Galof{K/K^H}$ which is just $H$.
So $H$ is therefore normal.
Now notice that
\multlines{
    K^{\sigma H\sigma^{-1}} = \set{x\in K}[\forall h\in H\colon \sigma h\sigma^{-1}(x)=x] = \set{\sigma(y)}[\forall h\in H\colon \sigma h(y)=\sigma(y)]\cr
    &= \sigma\set{y\in K}[\forall h\in H\colon h(y)=y] = \sigma(K^H)
}
so if $\sigma(K^H)=K^H$ then $K^{\sigma H\sigma^{-1}}=K^H$, meaning $H=\sigma H\sigma^{-1}$ (the $H\mapsto K^H$ map is injective; it has an inverse).
\qed

Notice that we defined a homomorphism $\Galof{K/F}\longto\Galof{K^H/F}$ whose kernel is $\Galof{K/K^H}$, thus
$$ \Galof{K^H/F} \cong \slfrac{\Galof{K/F}}{\Galof{K/K^H}} $$

\bprop

    If $K/F$ is Galois and $G=\Galof{K/F}$, then $H\leq G$ is normal if and only if $K^H/F$ is a normal extension.

\eprop

\Proof let $L=K^H$, $a\in L$, and $h$ be its minimum polynomial over $F$.
Thus it splits in $K$, meaning $h(\lambda)=\prod(\lambda-a_i)$ for $a_i\in K$.
We showed previously that $\sigma\in G$ permutes the roots of $h$.
If $H$ is normal, then $\sigma(K^H)=K^H$ meaning $\sigma(a)\in L$.
So if there is an $a_i$ not in $L$, then we can define a permutation $a\mapsto a_i$, but then $\sigma(a)\notin L$ in contradiction.
So all $a_i$ are in $L$, meaning $h$ splits in $L$, so $L/F$ is normal.

And if $L/F$ is normal, then $a_i\in L$ so $\sigma(a)=a_i$ for some $i$, since again $G$ permutes the roots, and so $\sigma(a)\in L$.
Thus $\sigma(L)=L$, meaning $H$ is normal.
\qed

\bprop

    Let $F\subseteq L\subseteq K$ be field extensions such that $K/F$ is Galois.
    Then $L/F$ is Galois if and only if $\Galof{K/L}$ is normal in $\Galof{K/F}$.
    In such a case,
    $$ \Galof{L/F} \cong \slfrac{\textstyle\Galof{K/F}}{\textstyle\Galof{K/L}} $$

\eprop

\Proof by the above proposition, $\Galof{K/L}$ is normal in $\Galof{K/F}$ if and only if $K^{\Galof{K/L}}/F$ is normal.
Now, since $K/F$ is Galois, so is $K/L$ since for $a\in K$ the minimal polynomial over $L$ divides the minimal polynomial over $F$ which splits into distinct linear factors, which means that so too must the
minimal polynomial over $L$.
Thus $K^{\Galof{K/L}}=L$, so $\Galof{K/L}$ is normal if and only if $L/F$ is normal.
And since $K/F$ is separable, so is $L/F$ since for $a\in L$ the minimal polynomial of $a$ in $F$ splits into distinct linear factors over its splitting field since $K$ is separable.
Thus $L/F$ is normal if and only if it is Galois, as required.

As noted before, 
$$ \Galof{K^H/F} \cong \slfrac{\Galof{K/F}}{\Galof{K/K^H}} $$
for $H=\Galof{K/L}$, and as shown above $K^H=L$ as required.
\qed

\bdefn

    Let $L/F$ be a finite-dimensional separable extension.
    Then $E\supseteq L$ is a {\emphcolor Galois closure} of $L/F$ if $E$ is the smallest field containing $L$ such that $E/F$ is Galois.

\edefn

The Galois closure can be found by taking the splitting field of the product of the minimal polynomials of the generators of $L/F$.
The Galois closure is unique up to isomorphism.

\bthrm

    Let $K/F$ be separable, then there exist only finitely many in-between fields.

\ethrm

\Proof let us look at the Galois closure of $K/F$, $E$.
We showed already that $E/F$ has finitely-many in-between fields, and therefore so does $K/F$.
\qed

\bthrm[title=Steinitz's Theorem, name=steinitz]

    Every finite-dimension separable field extension $K/F$ is generated by a single element.

\ethrm

\Proof we prove for the case that the fields are infinite, and we induct on the number of generators (which is finite as the extension is finite).
It is sufficient to prove this for the case $K=F[x,y]$ as we can go from $F[x_1,\dots,x_n]$ to $F[x_1,\dots,x_{n-1}]$ and continue inductively.
Let us focus on elements of the form $x+\alpha y$ for $\alpha\in F$, and so we get infinitely many (with repetitions) subfields $F[x+\alpha y]$.
Now, using the Galois closure of $K/F$ we can see that there are only finitely many in-between fields of $K/F$.
Thus there are $\alpha\neq\beta\in F$ such that $L=F[x+\alpha y]=F[x+\beta y]$.
And so $x+\alpha y-(x+\beta y)=(\alpha-\beta)y\in L$, meaning $y\in L$.
And similarly $x\in L$ so $L=F[x,y]$ and is generated by a single element.
\qed

\bdefn

    Let $n>1$ be a natural number, then a {\emphcolor root of unity} of order $n$ is an element in the field $\rho$ such that $\rho^n=1$.
    Denote $\mu_n(F)=\set{\rho\in F}[\rho^n=1]$ to be the set of all roots of unity in $F$ of order $n$.

\edefn

Obviously $\mu_n(F)$ is a subgroup of the multiplicative group of $F$: $1\in\mu_n(F)$ and if $\rho_1^n=\rho_2^n=1$ then $(\rho_1\rho_2)^n=\rho_1^n\rho_2^n=1$.

\bprop

    Every finite subgroup of the multiplicative group of a field is cyclic.

\eprop

\Proof let $A\leq F^\times$, then recall $e\coloneqq\exp A\coloneqq\minof[m>1]{\forall a\in A\colon a^m=1}={\rm lcm}\set{o(a)}[a\in A]$.
Then let us focus on the polynomial $\lambda^e-1$, for which by definition every element of $A$ is a root.
The degree of the polynomial is $e$ and so there are at most $e$ roots, so $\abs A\leq e$.
Thus $\abs A=e$ since $e\leq\abs A$ as well in general.
From this it follows that $A$ is cyclic.
\qed

From this it follows that $\mu_n(F)$ is cyclic as it is finite (since elements of it are roots of $\lambda^n-1$, so it has at most $n$ elements).
Denote its generator by $\rho_n$, called a {\it primitive root of unity of order $n$}.

In ${\bb C}$ the roots of unity of order $n$ are of the form $e^{2\pi i m/n}$ for $1\leq m\leq n$.
So let us focus on the polynomial $\lambda^n-1$, the polynomial which defines the roots of unity of order $n$, in ${\bb C}$.
Define $\rho=e^{2\pi i/n}$ so that the roots of the polynomial are $\rho^i$ for $1\leq i\leq n$.
So the splitting field of this polynomial over ${\bb Q}$ is
$$ {\bb Q}[1,\rho,\rho^2,\dots,\rho^{n-1}] = {\bb Q}[\rho] $$
Since $\lambda^n-1$ splits into distinct linear factors, and ${\bb Q}[\rho]$ is the splitting field of this separable polynomial, we get that ${\bb Q}[\rho]/{\bb Q}$ is a Galois extension.

So what is the minimal polynomial of these roots?
For $n=3$, $\lambda^3-1=(\lambda-1)(\lambda^2+\lambda+1)$ and both these polynomials are irreducible (since the first is linear, the second has no rational roots) so they are the minimal polynomials of
$1$ and $\rho,\rho^2$ respectively.
For $n=4$, $\lambda^4-1=(\lambda-1)(\lambda+1)(\lambda^2+1)$, so the minimal polynomial of $\rho$ is $\lambda^2+1$.

\bdefn

    The {\emphcolor cyclotomic polynomial} of degree $n$ is defined to be
    $$ \Phi_n(\lambda) \coloneqq \prod_{(j,n)=1}(\lambda-\rho_n^j) \in {\bb Q}[\rho_n][\lambda] $$

\edefn

The degree of $\Phi_n$ is the number of $1\leq j\leq n$ such that $(j,n)=1$, meaning $\deg\Phi_n=\phi(n)$ where $\phi$ is the Euler totient function.
For example,
$$ \Phi_3(\lambda) = (\lambda-\rho)(\lambda-\rho^2) = \lambda^2 - (\rho+\rho^2)\lambda + \rho^3 = \lambda^2 + \lambda + 1 $$
so we see that $\Phi_3$ is the minimal polynomial of $\rho$.

Recall that $o(g^k)=\frac{o(g)}{(o(g),k)}$ for $g\in G$.
So $o(\rho_n^k) = \frac{n}{(n,k)}$, meaning if $n$ and $k$ are coprime then $\rho_n^k$ is also a primitive root of unity.

Further notice that
$$ \prod_{d\divides n}\Phi_d(\lambda) = \prod_{d\divides n}\prod_{(j,d)=1}(\lambda-\rho_d^j) = \prod_{d\divides n}\prod_{(j,d)=1}(\lambda-\rho_n^{jn/d}) =
\prod_{d\divides n}\prod_{(j\cdot n/d,n)=n/d}(\lambda-\rho_n^{jn/d}) = \prod_{d\divides n}\prod_{(k,n)=n/d}(\lambda-\rho_n^k) $$
Since every $0\leq k\leq n-1$ has a unique $d$ such that $(k,n)=n/d$, this is just equal to
$$ = \prod_{k=0}^{n-1}(\lambda-\rho_n^k) = \lambda^n - 1 $$

\bprop

    $\Phi_n(\lambda)\in{\bb Z}[\lambda]$

\eprop

\Proof by induction on $n$.
For $n=1$, $\Phi_1(\lambda)=\lambda-1$.
Then by the above equality
$$ \Phi_n(\lambda) = \frac{\lambda^n-1}{\displaystyle\prod_{d\divides n,d<n}\Phi_d(n)} $$
So by our inductive hypothesis, this is in ${\bb Q}(\lambda)$ (the field of rational rational functions).
Thus $\Phi_n(\lambda)\in{\bb C}[\lambda]\cap{\bb Q}(\lambda)={\bb Q}[\lambda]$ ($K[\lambda]\cap F(\lambda)=F[\lambda]$ was a homework question).
So by Gauss's lemma, $\Phi_n(\lambda)\in{\bb Z}[\lambda]$ as required.
\qed

\bthrm

    $\Phi_n(\lambda)$ is irreducible over ${\bb Q}$.

\ethrm

\Proof suppose not, so there exists a factorization $\Phi_n(\lambda)=g(\lambda)h(\lambda)$ where $g$ is irreducible and $g,h\in{\bb Z}[\lambda]$ (by Gauss's lemma such a factorization exists).
Let $\rho$ be a root of $g$, then there must exist a prime $p>0$ such that $\rho^p$ is a root of $h$ (otherwise all of $\Phi_n$'s roots are roots of $g$).
Let us focus on $h(\lambda^p)$.
Since $\rho$ is a root of $h(\lambda^p)$ and $g(\lambda)$ is its minimal polynomial, $g(\lambda)\divides h(\lambda^p)$.
Since $h(\lambda^p)\equiv h(\lambda)^p\pmod p$, we have that $g(\lambda)\divides h(\lambda)^p)$ in ${\bb Z}/p{\bb Z}$.
But modulo $p$, $\lambda^n-1$ is separable since the gcd of $\lambda^n-1$ and its derivative $n\lambda^{n-1}$ is $1$.
Therefore $g$ is also separable modulo $p$, so $g(\lambda)\divides h(\lambda)$ modulo $p$, so $g^2\divides gh=\Phi_n(\lambda)\divides\lambda^n-1$ which is a contradiction to $\lambda^n-1$'s separability.
\qed

Therefore $\Phi_n(\lambda)$ is the minimal polynomial of $\rho_n$.
Thus $[{\bb Q}[\rho_n]:{\bb Q}]=\deg\Phi_n=\phi(n)$.

\bprop

    $\Galof{{\bb Q}[\rho_n],{\bb Q}}\cong{\cal U}_n$ (where ${\cal U}_n$ is the Euler group over $n$ elements).

\eprop

\Proof define $\psi\colon{\cal U}_n\longto\Galof{{\bb Q}[\rho_n],{\bb Q}}$ by $k\mapsto\sigma_k$ where $\sigma_k(\rho_n)=\rho_n^k$.
This is well-defined since ${\bb Q}[\rho_n]$ is generated by $\rho_n$ and $\rho_n^k$ is a root of its minimal polynomial (since automorphisms over $F$ must preserve roots of minimal polynomials).
This is surjective since every automorphism maps $\rho_n$ to some $\rho_n^k$.
And obviously $\sigma_{kk'}=\sigma_k\sigma_{k'}$ so $\psi$ is a homomorphism.
\qed

\bcoro

    Every subfield of finite-dimension of ${\bb Q}_{ab}\coloneqq\bigcup_n{\bb Q}[\rho_n]$ is Galois over ${\bb Q}$, and its Galois group is Abelian.

\ecoro

\Proof this is since every subfield generated by $\rho_{n_1},\dots,\rho_{n_k}$ is contained in some ${\bb Q}[\rho_n]$ (take for example their product).
This is Galois as proven before, and since $\Galof{{\bb Q}[\rho_{n_1},\dots,\rho_{n_k}],{\bb Q}}$ is normal in $\Galof{{\bb Q}[\rho_n],{\bb Q}}$ (as it is cyclic), it also forms a Galois extension.
\qed

\bthrm[title=Kronecker-Weber Theorem]

    Every Galois extension of ${\bb Q}$ is a subfield of ${\bb Q}_{ab}$.

\ethrm

Let $K/F$ be a Galois extension, then we can define the {\it trace} of the extension, $T\colon K\longto F$ by
$$ T(a) = \sum_{\sigma\in G}\sigma(a) $$
for $G=\Galof{K/F}$ (recall that this means $F=K^G$).
This is well-defined since ${\rm Im}T$ contains only fixed points:
$$ \tau T(a) = \sum_{\sigma\in G}\tau\sigma(a) = \sum_{\sigma\in G}\sigma(a) = T(a) $$
and so ${\rm Im}T\subseteq K^G$.
On the other hand, for $a\in F$, $T(a/\abs G)=a$ so ${\rm Im}T=F$.
The trace is a linear functional (viewing $K$ as a linear space over $F$).

We also define the {\it norm} of the extension similarly to be $N\colon K^\times\longto F^\times$:
$$ N(a) = \prod_{\sigma\in F}\sigma(a) $$
This is a group homomorphism.

For example, ${\bb C}/{\bb R}$ then $G=\set{1,z\mapsto\overline z}$.
So
$$ T(x+iy) = (x+iy) + (x-iy) = 2x $$
and
$$ N(x+iy) = (x + iy)\cdot(x-iy) = x^2 + y^2 = \abs{x+iy}^2 $$
Here $\ker T=i{\bb R}$ and $\ker N=S^1$.

\bdefn

    Call a Galois extension $K/F$ {\emphcolor cyclic} if $\Galof{K/F}$ is cyclic.

\edefn

If $K/F$ is cyclic and $\Galof{K/F}=\gen\sigma$ then define $D\colon K\longto K$ by $D(a)=a-\sigma(a)$.
Then its kernel is $K^G=F$, and
$$ T\circ D(a) = \sum_{i=0}^n\sigma^i(a) - \sum_{i=0}^n\sigma^{i+1}(a) = 0 $$
so ${\rm Im}D\subseteq\ker T$.
We can compare dimensions and conclude that $\ker T={\rm Im}D=\set{a-\sigma(a)}[a\in K]$.

A cyclic field extension must be generated by a single element, otherwise for $F[\alpha,\beta]$ we could map $\alpha$ to $-\alpha$ and have another map $\beta$ to $-\beta$ and neither are powers of the
other.
So suppose
$$ K = F[\alpha] = \set{a_0+a_1\alpha+\cdots+a_n\alpha^n}[a_i\in F] $$

\bthrm[title=Hilbert's Theorem 90, name=hilbert90]

    Suppose $K/F$ is a cyclic Galois extension whose group is generated by $\sigma$.
    Then
    $$ \ker N = \set{\frac a{\sigma(a)}}[a\in K^\times] $$

\ethrm

\Proof in one direction,
$$ N\parens{\frac a{\sigma(a)}} = \frac{N(a)}{N(\sigma(a))} = \frac{N(a)}{\sigma(N(a))} = \frac{N(a)}{N(a)} = 1 $$
And in the other, if $N(b)=1$ then we want an $a$ such that $\sigma(a)=b^{-1}a$.
Let us write $a = a_0 + \cdots + a_n$, where $n$ is the order of $\gen\sigma$, and so $\sigma(a)=\sigma(a_1)+\cdots+\sigma(a_{n-1})+\sigma(a_0)$.
Now let us require that $\sigma(a_i)=b^{-1}a_{i-1}$ (where $i-1$ is modulo $n$), so that we have
$$ \sigma(a) = \sigma(a_1) + \cdots + \sigma(a_{n-1}) + \sigma(a_0) = b^{-1}a_0 + \cdots + b^{-1}a_{n-2} + b^{-1}a_{n-1} = b^{-1}a $$
So let us set $a_0$, then we have that
$$ a_1=\sigma^{-1}(b^{-1}a_0)=\sigma^{-1}(b)^{-1}\sigma^{-1}(a_0),\qquad a_2=\sigma^{-1}(b)^{-1}\sigma^{-1}(a_1)=\sigma^{-1}(b)^{-1}\sigma^{-2}(b)^{-1}\sigma^{-2}(a_0)$$
And so on, until we get
$$ a_i = \bigl(\sigma^{-1}(b)\cdots\sigma^{-i}(b)\bigr)^{-1}\sigma^{-i}(a_0) $$
This holds for $1\leq i\leq n-1$, so now let us check for $i=0$ (equivalently $n$):
$$ \sigma(a_0) = \bigl(b\sigma^{-1}(b)\cdots\sigma^{-(n-1)}(b)\bigr)^{-1}\sigma^{1-n}(a_0) = N(b)^{-1}\sigma(a_0) $$
Since $N(b)=1$, this means that every choice of $a_0$ defines $a_1,\dots,a_{n-1}$ so that this system has a solution.
\qed

Suppose $K/F$ is a Galois extension such that $a\in K$ generates the extension.
Then there exists an injection $K\longembeds{\rm End}(K)\cong M_n(F)$ (where $n=[K:F]$) by $a\mapsto\ell_a$ (left multiplication by $a$, $\ell_a\colon b\mapsto ab$).
Notice that if we investigate $\ell_a$ with respect to the basis $B_a=\set{1,a,\dots,a^{n-1}}$ then
$$ [\ell_a]_{B_a} = \pmatrix{0 & 0 & \cdots & 0 & -\beta_0\cr 1 & 0 & \cdots & 0 & -\beta_1\cr 0 & 1 & \cdots & 0 & -\beta_2\cr \vdots & \vdots & \ddots & \vdots \cr 0 & 0 & \cdots & 1 & -\beta_n} $$
Where $h(\lambda)=\lambda^n+\sum_{i=0}^{n-1}\beta_i\lambda^i$ is the minimal polynomial of $a$.
This is called the {\it companion matrix} of $a$ denoted $C(a)$, and it can be shown (without too much trouble) that its characteristic polynomial is the minimal polynomial $h(\lambda)$.
Since permutations in the Galois group map roots of $h(\lambda)$ to roots and this is a Galois extension, we have that $h(\lambda)=\prod_{\sigma\in\Galof{K/F}}(\lambda-\sigma(a))$.
And thus
$$ h(\lambda) = \prod_{\sigma\in\Galof{K/F}}(\lambda-\sigma(a)) = \lambda^n - T(a)\lambda^{n-1} \pm \cdots \pm N(a) $$
And so $T(a)$ is the trace of $C(a)$ and $N(a)$ is the determinant of $C(a)$.

\bdefn

    Let $K/F$ be an $n$-dimensional field extension.
    It is called a {\emphcolor radical extension} if there exists $\alpha\in K$ such that $K=F[\alpha]$ and $a=\alpha^n$ is an element of $F$.
    In such a case, we write $K=F\bigl[\root n\of a\bigr]$.

\edefn

In such a case, $\lambda^n-a$ is the minimal polynomial of $\alpha$ (since it has degree $n$).

\bprop

    If $\rho\in F$ (meaning $F$ has a primitive root of unity of order $n$), then a radical extension is cyclic.

\eprop

\Proof suppose $K=F[\alpha]$ and $\alpha^n=a\in F$.
Then
$$ \lambda^n - a = \prod_{i=0}^{n-1}(\lambda - \rho^i\alpha) $$
Meaning that the splitting field of $\lambda^n-a$ is $F[\alpha,\rho\alpha,\dots,\rho^{n-1}\alpha]=F[\rho,\alpha]=F[\alpha]$ since $\rho$ is already in $F$.
Thus $K/F$ is Galois as the splitting field of an irreducible polynomial.
$\sigma(\alpha)=\rho\alpha$ defines an automorphism of $K/F$ since this maps $\alpha$ to another root of the irreducible polynomial.
Thus $\sigma^i(\alpha)=\rho^i\alpha$ and so the order of $\sigma$ is at least $n$, but $\abs{\Galof{K/F}}=[K:F]=n$ so this means that $\Galof{K/F}$ is generated by $\sigma$, as required.
\qed

\bthrm[title=Kummer's Theorem, name=kummer]

    Suppose $F$ has a primitive root of unity of degree $n$.
    Then every cyclic field extension of dimension $n$ is radical.

\ethrm

\Proof suppose $K/F$ is cyclic of degree $n$, then $N(\rho)=\rho^n=1$ (since for $a\in F$, $N(a)=\prod_\sigma \sigma(a)=\prod_\sigma a=a^{\abs{\Galof{K/F}}}$).
Thus $\rho\in\ker N$, meaning $\rho=\frac\alpha{\sigma\alpha}$ (where $\sigma$ generates $\Galof{K/F}$).
Thus
$$ 1 = \rho^n = \frac{\sigma(\alpha)^n}{\alpha^n} = \frac{\sigma(\alpha^n)}{\alpha^n} $$
This means that $\sigma(\alpha^n)=\alpha^n$, but since $\sigma$ generates $\Galof{K/F}$ a fixed point of $\sigma$ is in $K^{\Galof{K/F}}=F$.
And so let us define $a=\alpha^n$, so $a\in F$.
But $K/F$ is of dimension $n$ and so $F[\alpha]=K$ since $\lambda^n-a$ is a zeroing polynomial of $\alpha$, meaning $\alpha$ must generate the extension.
Thus $F[\root n\of a]=K$, as required.
\qed

\bdefn

    We say that a polynomial $f\in F[\lambda]$ is {\emphcolor solvable by radicals} if there exists a chain of radical extensions: $F\subseteq F_1\subseteq F_2\subseteq\cdots\subseteq F_n$ where
    $F_{i+1}/F_i$ is a radical extension; such that $F_n$ has a root of $f$.

\edefn

\bye

