\section{Induction}

\subsection{Linear categories and functors}

We will define here linear category theory.
In the language of category theory, this is simply the theory of categories enriched over the monoidal category of vector spaces (over a given field).
But we will ignore this abstraction, and instead define everything from the ground-up.

\bnote

    I have grown weary of denoting fields using blackboard bold (e.g.\ $\bF$).
    Instead a typical field will be denoted by $k$.
    Blackboard bold seems to give more significance to the field than necessary in my opinion.

\enote

\bdefn

    A {\emphcolor $k$-linear category} (or just a category) $\C$ consists of:
    \benum
        \item a class of objects (we abuse notation and denote this class also by $\C$);
        \item for every two objects $M,N\in\C$, a $k$-vector space $\hom_\C(M,N)$, called a {\emphcolor hom-space};
        \item for every three objects $M,N,L\in\C$, a $k$-bilinear map
        $$ {\circ}\colon\hom(N,L)\times\hom(M,N)\to\hom(M,L) $$
        (Equivalently, this is a $k$-linear map from the tensor product of the left hom-spaces.)
        Such that for every $M\in\C$ there exists a ${\rm id}_M\in\hom(M,M)$ which is the identity under composition (left/right according to where composition is defined), and $\circ$ is associative
        when defined.
    \eenum

\edefn

When declaring $\alpha\in\hom_\C(M,N)$ we may instead write $\alpha\colon M\to N$.

For example, given a group $G$ and field $k$, we can form a category $\Rep(G)$ whose objects are $G$-representations over $k$, and morphisms are $G$-equivariant morphisms.
In particular, the hom-spaces are simply $\hom_G(M,N)$.

Similarly, for a $k$-algebra $A$, we can consider the category ${}_A\Mod$ of left $A$-modules, and $\Mod_A$ of right $A$-modules.

\bdefn

    Let $\C$ and $\D$ be $k$-linear categories.
    A {\emphcolor functor} from $\C$ to $\D$, denoted $F\colon\C\to\D$, associates for every object $M\in\C$ an object $F(M)\in\D$, and for every $M,N\in\C$ a $k$-linear map
    $$ F\colon\hom_\C(M,N) \to \hom_\D(FM,FN) $$
    such that $F{\rm id}_M={\rm id}_{FM}$ and $F(\alpha\circ\beta)=F(\alpha)\circ F(\beta)$.

\edefn

\bdefn

    Let $F,G\colon\C\to\D$ be functors.
    A {\emphcolor natural transformation} $\sigma\colon F\to G$ consists of associating to every $M\in\C$ a map $\sigma_M\colon FM\to GM$ such that for every $\alpha\colon M\to N$ in $\C$,
    we have
    $$ G(\alpha)\circ\sigma_M = \sigma_N\circ F(\alpha) $$
    (Write out the commutative square for clarity.)

\edefn

A natural transformation for which every component ($\sigma_M$) is an isomorphism (in the usual category-theoretic sense) is called a {\it natural isomorphism}.
Two functors are {\it naturally isomorphic} if there exists a natural isomorphism between them.

Note that we can easily define the composition of functors, the product of categories and functors, opposite categories, in the standard way.
In particular we can define the {\it hom-functor}:
$$ \hom\colon\C^\op\times\C\to{\rm Vec}_k $$
which maps pairs of objects $(M,N)$ to $\hom_\C(M,N)$ and pairs of morphisms $f\colon M\to N$ (in $\C$, so $N\to M$ in $\C^\op$) and $g\colon L\to K$ to
$$ \hom(f,g)\colon\hom(N,L)\to\hom(M,K),\qquad \hom(f,g)(\alpha) = g\circ\alpha\circ f $$
For a set object $M\in\C$, this restricts to give two functors:
$$ \hom(M,\bul)\colon\C\to{\rm Vec}_k,\qquad\hom(\bul,M)\colon\C\to{\rm Vec}_k $$
where $\hom(M,f)$ is post-composition with $f$ (i.e.\ $\hom(M,f)\alpha=f\alpha$) and $\hom(f,M)$ is pre-composition with $f$ ($\hom(f,M)\alpha=\alpha f$).

We now define adjoint pairs;

\bdefn

    Let $F\colon\C\to\D$ and $G\colon\D\to\C$ be two functors.
    They form an {\emphcolor adjoint pair} (or an {\emphcolor adjunction}) if the functors $\hom_\D(F\bul,\bul)$ and $\hom_\C(\bul,G\bul)\colon\C^\op\times\D\to{\rm Vec}_k$ are naturally isomorphic.
    In such a case, we write $F\dashv G$, and call $F$ the {\emphcolor left adjoint} and $G$ the {\emphcolor right adjoint}.

\edefn

The naturality condition is equivalent to saying that for every $M\in\C$ and $U\in\D$ there exists a $k$-linear isomorphism $\gamma_{MU}\colon\hom_\D(FM,U)\to\hom_\C(M,GU)$ such that
for every $\alpha\colon M_1\to M_2$ in $\C$ the following diagram commutes:

\centerline{\drawdiagram{
    $\hom_\D(FM_2,U)$&$\hom_\C(M_2,GU)$\cr
    $\hom_\D(FM_1,U)$&$\hom_\C(M_1,GU)$
}{
    \diagarrow{from={1,1}, to={1,2}, text=$\gamma_{M_2U}$, y distance=.25cm}
    \diagarrow{from={2,1}, to={2,2}, text=$\gamma_{M_1U}$, y distance=.25cm}
    \diagarrow{from={1,1}, to={2,1}, text=$\bul\circ F\alpha$, x distance=-.5cm}
    \diagarrow{from={1,2}, to={2,2}, text=$\bul\circ \alpha$, x distance=.5cm}
}}

as well as for every $\beta\colon U_1\to U_2$ in $\D$ an analogous diagram commute.

An equivalent definition of adjunctions are via unit-counit pairs.

\bdefn

    Let $F\colon\C\to\D$ and $G\colon\D\to\C$ be two functors.
    A {\emphcolor unit-counit pair} for $(F,G)$ (note the order) are two natural transformations $\eta\colon{\rm Id}_\C\to GF$ (the unit) and $\epsilon\colon FG\to{\rm Id}_\D$ such that the compositions
    $$ F\xvarrightarrow{F\eta}FGF\xvarrightarrow{\epsilon F}F,\qquad G\xvarrightarrow{\eta G}GFG\xvarrightarrow{G\epsilon}G $$
    are the respective identity natural transformations.

\edefn

In the above definition, we use the concept of {\it whiskering}.
For a natural transformation $\sigma\colon F\to G$ and another functor $H$, we define the natural transformation $H\sigma\colon HF\to HG$ (if such composition is well-defined) by $(H\sigma)_X=H(\sigma_X)$
and $\sigma H\colon FH\to GH$ by $(\sigma H)_X=\sigma_{H(X)}$.
So the unit-counit conditions say that
$$ {\rm id}_{FX} = \epsilon_{FX}\circ F(\eta_X),\qquad {\rm id}_{GY} = G(\epsilon_Y)\circ\eta_{GY} $$
for all $X\in\C$ and $Y\in\D$.

\bthrm

    $F\dashv G$ form an adjunction if and only if there exists a unit-counit pair $(\eta,\epsilon)$ for $(F,G)$.

\ethrm

\bproof

    Given an adjunction $F\dashv G$, let
    $$ \eta_X=\gamma_{X,FX}({\rm id}_{FX})\in\hom_\D(FX,FX),\quad\hbox{and}\quad\epsilon_Y=\gamma^{-1}_{GY,Y}({\rm id}_{GY})\in\hom_\C(GY,GY) . $$
    This forms the unit-counit pair.

    Conversely, given a unit-counit pair $(\eta,\epsilon)$, define $\gamma\colon\hom_\D(F\bul,\bul)\to\hom_C(\bul,G\bul)$ by $\gamma_{XY}(\alpha)=G(\alpha)\circ\eta_X$ for $f\in\hom_\D(FX,Y)$.
    Its inverse is $\gamma^{-1}_{YX}(\beta)=\epsilon_Y\circ F(\beta)$.
    \qed

\eproof

Let's construct an example of an adjunction.
Let $\iota\colon B\to A$ be a morphism of $k$-algebras.
Then every $A$-module can be viewed as a $B$-module; given an $A$-module $M$, give it a $B$-module structure by $b\cdot m=\iota(b)m$ for $b\in B,m\in M$ (this is just composition of $\iota$ with the $A$-module
structure).
Note that with such a view, every $A$-module morphism is a $B$-module morphism: $f(bm)=f(\iota(b)m)=\iota(b)f(m)=bf(m)$.
Thus we can define a functor
$$ \res^A_B\colon\Mod(A)\to\Mod(B) $$
which maps an $A$-module $M$ to its induced $B$-module structure, and an $A$-module morphism to itself.
$\res^A_B$ is called the {\it restriction functor} (or the {\it forgetful functor}).

Now, we ask ourselves if the forgetful functor has a left adjoint.
Consider the functor
$$ A\otimes_B\bul\colon\Mod(B)\to\Mod(A) $$
which maps a $B$-module $M$ to the $A$-module $A\otimes_BM$ (this is well-defined, $A$ is a right $B$-module by $a\cdot b=a\iota(b)$).
And it maps a morphism of $B$ modules $f\colon M_1\to M_2$ to $A\otimes f\colon A\otimes_BM_1\to A\otimes_BM_2$ which maps $a\otimes m$ to $a\otimes f(m)$.
Such a morphism exists and is unique.

We claim that $A\otimes_B\bul$ and $\res^A_B$ forms an adjunction.
Given $M\in\Mod(B)$ and $N\in\Mod(A)$ we define an isomorphism
$$ \gamma_{MN}\colon\hom_{\Mod(A)}(A\otimes_BM,N) \to \hom_{\Mod(B)}(M,\res^A_BN) $$
by mapping $\alpha\colon A\otimes_BM\to N$ to $\beta\colon M\to\res^A_BN$ defined by $\beta(m)=\alpha(1\otimes m)$.
Its inverse,
$$ \gamma_{MN}^{-1}\colon\hom_{\Mod(B)}(M,\res^A_BN) \to \hom_{\Mod(A)}(A\otimes_BM,N) $$
is given by mapping $\beta\colon M\to\res^A_BN$ to $\alpha\colon A\otimes_BM\to N$ by $\alpha(a\otimes m)=\iota(a)\beta(m)$.
Their inverse-ness is readily checked, as well as their naturality.

Thus we have
$$ A\otimes_B\bul\,\dashv\,\res^A_B $$

Interestinly, $\res^A_B$ has a right adjoint as well.
Namely, consider the functor $\Mod(B)\to\Mod(A)$ $M\mapsto\hom_B(A,M)$ ($A$ is a $B$-module via $\iota$).
$\hom_B(A,M)$ is a $A$-module by $(a\phi)(a')=\phi(a'a)$.
Given a morphism $\alpha\colon M_1\to M_2$ we define $\alpha_*\colon\hom_B(A,M_1)\to\hom_B(A,M_2)$ by post-composition.

Now given $M\in\Mod(B),N\in\Mod(A)$ we define
$$ \gamma_{MN}\colon \hom_{\Mod(B)}(\res^A_BN,M) \to \hom_{\Mod(A)}(N,\hom_B(A,M)) $$
by mapping $\alpha\colon\res^A_BN\to M$ to $\beta\colon N\to\hom_B(A,M)$ defined by $\beta(n)(a)=\alpha(an)$.
Its inverse maps $\beta\colon N\to\hom_B(A,M)$ to $\alpha(n)=\beta(n)(1)$.
All the conditions are readily checked.

Thus we have a chain of adjunctions:
$$ A\otimes_B\bul\,\dashv\,\res^A_B\,\dashv\,\hom_B(A,\bul) $$
when $\iota\colon B\to A$ is a $k$-algebra morphism.

\subsection{Basics of induction}

Fix groups $G,H$, a base field $k$, and a group morphism $\iota\colon H\to G$.

$\iota$ induces a morphism of $k$-algebras $\iota\colon k[H]\to k[G]$.
Since there is an isomorphism of categories $\Rep_k(G)\cong\Mod(k[G])$, we have a chain of adjunctions:

\kern1cm
\centerline{\drawdiagram{
    $\Rep(H)$&$\Rep(G)$
}{
    \diagarrow{from={1,1}, to={1,2}, text=$\ind^H_G$, y off=.5cm, y distance=.75cm}
    \diagarrow{from={1,1}, to={1,2}, text=$\Ind^H_G$, y off=-.5cm, y distance=-.25cm}
    \diagarrow{from={1,2}, to={1,1}, text=$\res^G_H$, y distance=.25cm}
}}
\kern.5cm

Where
$$ \ind^H_G\dashv\res^G_H\dashv\Ind^H_G $$
Specifically $\res^G_H=\res^{k[G]}_{k[H]}$, $\ind^H_G$ is $k[G]\otimes_{k[H]}\bul$, and $\Ind^H_G$ is $\hom_{k[H]}(k[G],\bul)$.

\subsubsection*{Understanding $\ind^H_G$}

We will assume for now that $\iota\colon H\to G$ is an inclusion.
We will examine first $\ind^H_G$.
Let $(g_q)_{q\in G/H}$ be a set of representatives of the cosets in $G/H$ (so $g_q\in q$).
Then we have an isomorphism of right $k[H]$-modules:
$$ k[G]\cong\bigoplus_{q\in G/H}k[H] $$
given by mapping $(d_q)_{q\in G/H}$ on the right to $\sum_{q\in G/H}\delta_{g_q}\cdot d_q$ on the left.
It is readily checked that this is an injective morphism of right $k[H]$-modules, and by considering dimensions we have that it is an isomorphism.
Thus, given a (left) $k[H]$-module $M$:
$$ k[G]\otimes_{k[H]}M \cong \parens{\bigoplus_{q\in G/H}k[H]}\otimes_{k[H]}M \cong \bigoplus_{q\in G/H}(k[H]\otimes_{k[H]}M) \cong \bigoplus_{q\in G/H}M $$
Sending $(m_q)_{q\in G/H}$ on the right to $\sum_{q\in G/H}\delta_{g_q}\otimes m_q$ on the left.

So we can think of $\ind^H_GM$ as $\bigoplus_{q\in G/H}g_qM$ (where $g_q$ ``acts'' on $M$ to give $M$), and we have $g(g_qm)=g_{q'}(hm)$ for $g\in G$ where $gg_q=g_{q'}h$ for $h\in H$.
In particular, we have
$$ \dim(\ind^H_G(V)) = [G:H]\cdot\dim V $$

So suppose we have a $G$-representation $V$, and an $H$-subrepresentation $W\subseteq V$.
Then as the inclusion forms a morphism of $H$-representations $W\to\res^G_HV$, by adjunction we get a $G$-morphism $\ind^H_GW\to V$.
This is an isomorphism if and only if $[G:H]\cdot\dim W=\dim V$.
And it is an injection iff the subspaces $(g_qW)_{q\in G/H}$ are linearly independent (i.e.\ have a direct sum in $V$).

\subsubsection*{Understanding $\Ind^H_G$}

Let us examine now $\Ind^H_G$.
Given a $k[H]$-module $M$, we can think of $\hom_{k[H]}(k[G],M)$ as a subspace of $\hom_k(k[G],M)$, which we can identify with $\St(G,M)$.
In particular, $\hom_{k[H]}(k[G],M)$ can be identified with maps $f\colon G\to M$ such that $f(hx)=hf(x)$ for all $h\in H$ and $x\in G$.
The $G$-action on this space is $(gf)(x)=f(xg)$ (as per the action on $\hom_{k[H]}(k[G],M)$).
So we can consider
$$ \Ind^H_G(M) = \set{f\colon G\to M}[\forall h\in H,x\in G.\,f(hx)=hf(x)],\qquad (gf)(x)=f(xg) $$

\subsubsection*{Back to business}

Note that $k[G]$ can be viewed as $\ind^1_G(k)$, and we can compute $[kG:E]$ using the adjunction $\ind^1_G\dashv\res^G_1$.

We want to compare $\ind^H_G$ and $\Ind^H_G$; we show that we have a natural transformation $\ind^H_G\to\Ind^H_G$, which is an isomorphism when $[G:H]$ is finite.
By the adjunction, a $G$-morphism $\ind^H_G(V)\to\Ind^H_G(V)$ is equivalent to an $H$-morphism $V\to\res^G_H\Ind^H_G(V)$ (we abuse notation and write $\Ind^H_G(V)$ for the right side).
We define such a morphism by mapping $v\in V$ to the function $G\to V$ defined by mapping $h\in H$ to $hv$ and $g\in G-H$ to $0$.

If $[G:H]$ is finite, we define its inverse $\Ind^H_G(V)\to\ind^H_G(V)$ as follows: let $(g_q)_{q\in G/H}$ be representatives of the cosets.
Then map $f\in\Ind^H_G(V)$ to $\sum_{q\in G/H}\delta_{g_q^{-1}}\otimes f(g_q)$.
This indeeds forms an inverse morphism.

So in our case, where $G$ is finite, we have that $\ind^H_G$ and $\Ind^H_G$ are isomorphic.

We can also abstract this argument.
Let $\iota\colon B\to A$ be a morphism of $k$-algebras.
Then we have the following natural morphism (in $M$):
$$ \hom_B(A,B)\otimes_B M \to \hom_B(A,M) $$
$\hom_B(A,B)$ is a right $B$-module since $B$ is a right $B$-module, so the left-hand side is well-defined.
And both sides are left $A$-modules since $A$ is.
This morphism maps $T\otimes m$ to $a\mapsto T(a)m$.

Now suppose $e\colon A\to B$ is a morphism of left $B$-modules.
Then we can construct a morphism of $(A,B)$-bimodules
$$ A\to\hom_B(A,B) $$
by mapping $a$ to $a'\mapsto e(a'a)$.
This gives a morphism of $A$-modules, natural in $M$:
$$ A\otimes_BM\to\hom_B(A,B)\otimes_BM $$

Composing gives us a morphism (natural in $M$):
$$ A\otimes_BM\to\hom_B(A,B)\otimes_BM\to\hom_B(A,M) $$
Note that this is a morphism $\ind^B_A\to\Ind^B_A$.

We claim that the second arrow is an isomorphism when $A$ is a free left $B$-module of finite rank.
Indeed, for a left $B$-module $N$, we can extend our second arrow to
$$ \hom_B(N,B)\otimes_BM\to\hom_B(N,M) $$
This is clearly an isomorphism when $N=B$.
And if it is an isomorphism for $N_1,N_2$ then it is an isomorphism for $N_1\oplus N_2$.
Thus if $N$ is a free left $B$-module of finite rank (so $N=B\oplus\cdots\oplus B$), then this is an isomorphism.

The first arrow is an isomorphism if $A\to\hom_B(A,B)$ is also an isomorphism.

In our case, we have $A=k[G]$, $B=k[H]\subseteq k[G]$ and $e\colon k[G]\to k[H]$ defined by $e(\delta_h)=\delta_h$ for $h\in H$ and $e(\delta_g)=0$ for $g\in G-H$.
Then $A$ is indeed a free left $B$-module of finite rank (when $[G:H]$ is finite): $k[G]=k[H]^{\oplus[G:H]}$.
And $A\mapsto\hom_B(A,B)$ is readily shown to be an isomorphism.

\subsection{Applications to dimensions of irreducible representations}

\bdefn

    Let $\theta\colon G\to G'$ be a group isomorphism, and $(V,\rho)$ a $G$-representation.
    This defines a $G'$-representation $({}^\theta V,{}^\theta\rho)$ via $\theta$'s inverse: ${}^\theta V$ is just $V$ and
    $$ {}^\theta\rho(g') = \rho(\theta^{-1}g') $$
    i.e.\ $g'v=\theta^{-1}(g)v$ for all $g'\in G',v\in V$.

\edefn

Note that ${}^\theta V$ shares pretty much all important properties with $V$ (they are pretty much the same representation).

We can specialize this definition: given an automorphism $\theta$ of $G$, and a subgroup $H$ of $G$, an $H$-representation $V$ can be made into a $\theta(H)$-representation ${}^\theta V$ by
restricting $\theta$ along $H$.
In particular we focus on conjugation by a specific element of $G$, say $g_0$, call it $\theta_{g_0}\colon g\mapsto g_0^{-1}gg_0$.
Then ${}^{g_0}V={}^{\theta_{g_0}}V$ has its action given by $h*v=g_0hg_0^{-1}v$ (this is a $g_0^{-1}Hg_0$-representation).

If $\theta$ is a $G$-automorphism clearly ${}^\theta V$ is isomorphic to $V$.
In particular ${}^gV\cong V$ for all $g\in G$.

\bprop

    Let $H$ be a subgroup of $G$ and $\theta$ a $G$-automorphism, and $V$ an $H$-representation.
    Then $\ind^{\theta(H)}_G({}^\theta V)\cong{}^\theta\ind^H_GV$.
    In particular for all $g\in G$, $\ind^{gHg^{-1}}_G({}^gV)\cong\ind^H_GV$ (since ${}^g\ind^H_GV\cong\ind^H_GH$ as it is a $G$-representation).

\eprop

\bprop

    Let $N$ be a normal subgroup of $G$, $V$ a $G$-representation, and $g\in G$.
    If $F\subseteq V$ is an $N$-isotypic component, $gF$ is an $N$-isotypic component as well.

\eprop

\bdefn

    A representation is {\emphcolor isotypical} if all its irreducible subrepresentations are isomorphic (i.e.\ it has only one isotypic component).

\edefn

\blemm

    Let $N$ be a normal subgroup of $G$, and $E$ an irreducible $G$-representation.
    Let $F\subseteq E$ be an $H$-isotypic component.
    Consider
    $$ K = \set{g\in G}[gF=F] \subseteq G $$
    Then $K$ is a subgroup of $G$ containing $N$, and $F$ is a $K$-subrepresentation of $E$.
    Consider the $G$-morphism $\ind^K_GF\to E$ corresponding to the inclusion ($K$-morphism) $F\to E$; it is an isomorphism.
    $F$ as a $K$-representation is irreducible.

\elemm

\bproof

    Since $E$ is irreducible, it is sufficient to show that $\ind^K_GF\to E$ is injective.
    By the structure of the induced representation, this map is injective if and only if the subspaces $(g_qF)_{q\in G/K}$ (where $g_q\in q$) are linearly independent in $E$.
    We will show this by showing they are distinct $H$-isotypic components in $E$.
    Them being isotypic follows from the previous proposition.
    And if $g_qF=g_{q'}F$ then $g_{q'}^{-1}g_q\in K$ so $q'=q$.

    All that remains is to show that $F$ is an irreducible $K$-representation.
    Suppose not, let $F=F'\oplus F''$ for non-trivial $F',F''$.
    Then
    $$ E \cong \ind^K_G(F) \cong \ind^K_G(F')\oplus\ind^K_G(F'') $$
    These are non-zero, and this contradicts $E$ being irreducible.
    \qed

\eproof

\blemm

    Let $N$ be a normal subgroup of $G$, and let $E$ be an irreducible $G$-representation.
    Then either $\res^G_H(E)$ is isotypical or there exists a subgroup $N\subseteq H\subsetneq G$ such that $E$ is isomorphic to a representation of the form $\ind^K_G(F)$ where $F$
    is an irreducible $H$-representation.

\elemm

\bproof

    Let $F\subseteq\res^G_H(E)$ be an isotypic component, and let $H=\set{g\in G}[gF=F]$ as above.
    Then $N\subseteq H\subseteq G$, and $F$ is a $H$-subrepresentation of $E$.
    By the previous lemma, $F$ is an irreducible $H$-representation and $\ind^H_G(F)\cong E$ as $G$-representations.
    We need to see that if $H=G$ then $\res^G_N(E)$ is isotypical.
    Indeed, if $H=G$ then $F$ is a $G$-subrepresentation of $E$ and as such is equal to it (as $E$ is irreducible).
    So $E=F$ is an isotypical $H$-representation.
    \qed

\eproof

\bthrm

    Let $A\subseteq G$ be an Abelian normal subgroup.
    Then the dimension of any irreducible $G$-representation divides $[G:A]$.

\ethrm

\bproof

    We induct on $\abs G$ (the base case is trivial).
    Apply the above lemma to our situation (with $N=A$), and we must consider two cases.
    In the first case $\res^G_A(E)$ is isotypical.
    Since $A$ is Abelian, an isotypical representation is one on which $A$ acts by scalars.
    Let $\rho\colon G\to\gl(E)$ be the representation, then we have that $\rho(A)$ lies in the center of $\rho(G)$.
    $E$ can be considered an irreducible $\rho(G)$-representation.
    Then we have shown that $\dim E$ divides $[\rho(G):Z(\rho(G))]$, which in turn turn divides $[\rho(G):\rho(A)]$, which divides $[G:A]$ as desired.

    In the second case $E$ is isomorphic to $\ind^K_G(F)$ with $A\subseteq K\subsetneq G$ and $F$ irreducible.
    By induction $\dim F$ divides $[K:A]$.
    Thus
    $$ \dim E = \dim\ind^K_G(F) = [G:K]\dim F $$
    which divides $[G:K]\cdot[K:A]=[G:A]$ as desired.
    \qed

\eproof

\subsection{Irreducible representations of semidirect products $A\rtimes H$, $A$ commutative}

We quickly recall the definition of inner and outer semidirect products:

\bdefn

    Let $G$ be a group, $N,H\subseteq G$ be subgroups with $N$ normal.
    Suppose that $N\cap H=1$ and $NH=G$, then we call $G$ the {\emphcolor inner semidirect product} of $N$ with $H$, writteb $G=N\rtimes H$.

\edefn

Note that every element of $G$ can be written uniquely in the form $nh$ for $n\in N,h\in H$.
Indeed, suppose $n_1h_1=n_2h_2$, then $n_2^{-1}n_1=h_2h_1^{-1}\in N\cap H$ so $n_1=n_2,h_1=h_2$.
Moreover, as $N$ is a normal subgroup of $G$, $H$ acts on $N$ by conjugation.
In fact, multiplication in $G$ can be written in terms of multiplication in $N,H$ and $H$'s action on $N$:
$$ n_1h_1n_2h_2 = n_1(h_1n_2h_1^{-1})h_1h_2 = n_1(h_1.n_2)h_1h_2 $$
This leads to the following definition:

\bdefn

    Let $N,H$ be groups, and $\phi\colon H\to{\rm Aut}(N)$ a group homomorphism (a group action on a group).
    We define their {\emphcolor outer semidirect product} (relative to $\phi$) to be $N\rtimes_\phi H$ as follows.
    The underlying set is the cartesian product $N\times H$, and multiplication is given by
    $$ (n_1,h_1)\cdot(n_2,h_2) = (n_1(\phi(h_1)n_2),h_1h_2) $$

\edefn

Note that the inner direct product $N\rtimes H$ is isomorphic to $N\rtimes_\theta H$ where $\theta$ is the group action given by conjugation.
Furthermore, $N$ is a normal subgroup of the semidirect product and $(N\rtimes H)/N\cong H$.

For this section suppose $G=A\rtimes H$ where $A$ is Abelian, and $k$ is an algebraically closed group whose characteristic does not divide $\abs G$.

Since $H$ acts on $A$, it also acts on the group of characters of $A$, $\ch_k(A)=\hom(A,k^\times)$ by $(h\chi)(a)=\chi(h.a)$.
In particular, since outer semidirect products are inner semidirect products (of themselves), viewing $A,H\subseteq G$ subgroups, $(h\chi)(a)=\chi(h^{-1}ah)$.
This extends to all of $G$: $G$ acts on $\ch_k(A)$ by $(g\chi)(a)=\chi(g^{-1}ag)$.

For a $G$-representation $V$ and a character $\chi\in\ch_k(A)$, define
$$ V_{A,\chi} = \set{v\in V}[\forall a\in A.\,av=\chi(a)v] \subseteq V $$
This is the isotypical component of $\res^G_AV$ corresponding to the irreducible $A$-representation $k_\chi$.
Let us also define $\supp(V)\subseteq\ch_k(A)$ to be all characters $\chi$ for which $V_{A,\chi}\neq0$.
$\supp(V)$ is $G$-invariant: if $v\in V_{A,\chi}$ then $gv\in V_{A,g\chi}$.
Furthermore the support is nonempty if $V$ is finite-dimensional (and nonzero).

For a character $\chi\in\ch_k(A)$ define
$$ G_\chi = \set{g\in G}[g\chi=\chi] \subseteq G,\qquad H_\chi = H\cap G_\chi = \set{h\in H}[h\chi=\chi] $$
then $G_\chi=A\rtimes H_\chi$: $A$ is clearly subgroup of $G_\chi$, $A\cap H_\chi\subseteq A\cap H=1$, $A$ is normal in $G$ and so in $G_\chi$ as well, and $AH_\chi\subseteq G_\chi$ and for $ah\in G_\chi$ since
$a\in G_\chi$ we have $h\in G_\chi$ so in $H_\chi$.

Note that since $gV_{A,\chi}=V_{A,g\chi}$,
$$ G_\chi = \set{g\in G}[gV_{A,\chi}=V_{A,\chi}] $$

\blemm

    Suppose $V$ is finite-dimensional and nonzero, then $V$ is $G$-irreducible iff $\supp(V)$ is $G$-transitive (has one orbit) and for one/all $\chi\in\supp(V)$, $V_{A,\chi}$ is $H_\chi$-irreducible.
    In that case, the inclusion $V_{A,\chi}\to V$ induces a $G$-isomorphism $\ind^{G_\chi}_G(V_{A,\chi})\to V$.

\elemm

\bproof

    If $V$ is irreducible, let $\chi\in\supp(V)$, and define
    $$ V' = \sum_{g\in G}gV_{A,\chi} = \sum_{g\in G}V_{A,g\chi} $$
    This is a nonzero subrepresentation, and as such is equal to $V$.
    These are all $A$-isotypical components, so for $V_{A,\mu}$ which is another $A$-isotypical component, it must be contained in one of them.
    I.e.\ $V_{A,\mu}=V_{A,g\chi}$ for some $g\in G$ for all $\mu\in\supp(V)$.
    Since $\mu$ is in the support this means $\mu=g\chi$, so $\supp(V)=G\chi$, a single orbit, so $\supp(V)$ is indeed $G$-transitive.

    Furthermore $V_{A,\chi}$ is $\set{g\in G}[gV_{A,\chi}=V_{A,\chi}]$-irreducible (by a previous lemma).
    But this is just $G_\chi$, since $V_{A,\chi}=V_{A,g\chi}$ implies equality over all of $\supp(V)$ (by transitivity).
    So $V_{A,\chi}$ is $G_\chi$-irreducible.
    Since $A$ acts on $V_{A,\chi}$ by scalars, $H_\chi$-subrepresentations are the same as $G_\chi$-subrepresentations, so $V_{A,\chi}$ is $H_\chi$-irreducible as well.

    Finally, the same lemma tells us that $\ind^{G_\chi}_GV_{A,\chi}$ is isomorphic to $V$.

    Conversely, suppose $\supp(V)$ is $G$-transitive and $V_{A,\chi}$ is $H_\chi$-irreducible.
    Let $\chi\in\supp(V)$.
    Let $V'\subseteq V$ be a nonzero $G$-subrepresentation, then $\supp(V')\neq\emptyset$.
    Since $\supp(V')\subseteq\supp(V)$ is $G$-invariant and $\supp(V)$ is transitive, $\supp(V')=\supp(V)$.
    Then $V'_{A,\chi}=V'\cap V_{A,\chi}$ is nonzero, but this is a $H_\chi$-subrepresentation of $V_{A,\chi}$ so $V'\cap V_{A,\chi}=V_{A,\chi}$, i.e.\ $V_{A,\chi}\subseteq V'$.
    This is true for all $\chi\in\supp(V)$, and since these are all the $A$-isotypic components of $V$, this means that $V=V'$.
    So $V$ is indeed $G$-irreducible.
    \qed

\eproof

Given a character $\chi\in\ch_k(A)$ and a $H_\chi$-representation $F$, this can be viewed as a $G_\chi$-representation $F_\chi$ over the same vector space $F$ where
$$ (ah).v = \chi(a)(hv) $$
i.e.\ $A$ acts by $\chi$ and $H_\chi$ as it does on $F$ (since $G_\chi=A\rtimes H_\chi$).
From the above lemma, we see that every irreducible $G$-representation is isomorphic to $\ind^{G_\chi}_GF_\chi$ for some $\chi\in\ch_k(A)$ and $H_\chi$-irreducible representation $F$.
Conversely $V=\ind^{G_\chi}_GF_\chi$ is irreducible for all such $\chi,F$.
Indeed, let $(h_q)_{q\in G/G_\chi}$ be representatives in $H$ for cosets in $G/G_\chi$, then
$$ V \cong \oplus_qh_qF_\chi $$
Notice that $F_\chi\subseteq V_{A,\chi}$, but we actually have equality.
Indeed, for $q\neq G_\chi$ we have $h_qF_\chi\subseteq V_{A,h_q\chi}$ and $h_q\chi\neq\chi$, so we must have equality (since $V$ is their direct sum).
So we readily see that $\supp(V)=G\chi$, and $V_{A,\chi}$ is an irreducible $G_\chi$-representation.
So by the lemma $V$ is irreducible.

To summarize: we can parameterize irreducible $G$-representations as follows.
We take one $\chi\in\ch_k(A)$ for each $H$-orbit, and we look at all irreducible $H_\chi$-representations $F$ and consider $\ind^{G_\chi}_GF_\chi$.
This gives all irreducible $G$-representations.

\subsection{Characters of induced representations}

Let $G$ be a finite group, with subgroup $H\subseteq G$ and $k$ an algebraically closed field whose characteristic does not divide $\abs G$.

Our goal is to compute a map which maps the characters of $H$-representations to the characters of their induced $G$-representations.
That is, a map
$$ \find^H_G\colon\St(H,k)^\cl\to\St(G,k)^\cl $$
such that for a finite-dimensional $H$-representation $V$:
$$ \find^H_G(\ch_V) = \ch_{\Ind^H_G(V)} $$
Such a map must exist: since $\Ind^H_G(V\oplus W)\cong\Ind^H_G(V)\oplus\Ind^H_G(W)$ (in general $\hom(\bul,A\oplus B)\cong\hom(\bul,A)\oplus\hom(\bul,B)$), we must have that
$\find^H_G(\ch_V)$ is determined by the irreducible components of $V$.
And since the characters of irreducible $H$-representations form a basis for the $H$-class functions, such a map is well-defined.

Now, notice that clearly
$$ \ch_{\res^G_HV} = (\ch_W)|_H $$
where $(\ch_W)|_H$ is the restriction of $\ch_W$ to $H$.
Recall our non-degenerate bilinear form $(\bul,\bul)$ on $\St(G,k)^\cl$, which satisfies $(\ch_V,\ch_W)=\dim\hom_G(V,W)$.
Thus we have for all $G$-representations $W$:
$$ (\ch_W,\find^H_G(\ch_V)) = (\ch_W,\ch_{\Ind^H_G V}) = \dim\hom_G(W,\Ind^H_GV) = \dim\hom_H(\res^G_HW,V) = (\ch_{\res^G_HW},\ch_V) $$
which is equal to $(\ch_W|_H,\ch_V)$.
So $\find^H_G$ is the adjoint of the restriction operator $\bul|_H\colon\St(G,k)^\cl\to\St(H,k)^\cl$.

As a quick aside, we can give $\St(G,k)$ a $G$-representation: $(gf)(x)=f(gxg^{-1})$.
Then $\St(G,k)^\cl$ are simply the $G$-invariants of this representation: those $f\colon G\to k$ for which $f(g\bul g^{-1})=f(\bul)$ for all $g\in G$.
In particular, we then are given the averaging operator:
$$ \av\colon\St(G,k) \to \St(G,k)^\cl,\qquad \av(f)(x) = \frac1{\abs G}\sum_{g\in G}f(gxg^{-1}) $$

Thus by studying the restriction operator, we can compute $\find^H_G$ explicitly.
We can consider $\bul|_H$ on our class functions as the composition
$$ \St(G,k)^\cl \xvarrightarrow{\rm inclusion} \St(G,k) \xvarrightarrow{\bul|_H} \St(H,k) \xvarrightarrow{\av} \St(H,k)^\cl $$
where $\av$ is the averaging operator just discussed.
This is because $\av$ is simply projection onto $\St(H,k)^\cl$.

The adjoint of $\av$ (with respect to the bilinear form) is simply inclusion (since $(\av f,f')=(f,\av f')$ and $\av$ is the identity on class functions).
The adjoint of $\bul|_H$ is the composition of the adjoints of the above, i.e.\
$$ \St(H,k)^\cl \xvarrightarrow{\rm inclusion} \St(H,k) \xvarrightarrow{\rm adj} \St(G,k) \xvarrightarrow{\av} \St(G,k)^\cl $$
where ${\rm adj}$ is the adjoint to $\bul|_H$ on all functions.
This is given by ${\rm adj}(f)(x)=[G:H]f(x)$ if $x\in H$ and $0$ otherwise, indeed for $\phi\colon H\to k$ and $\psi\colon G\to k$,
$$ ({\rm adj}\phi,\psi) = \frac1{\abs G}\sum_{g\in G}{\rm adj}\phi(g)\psi(g^{-1}) = \frac1{\abs G}\sum_{h\in H}[G:H]\phi(h)\psi(h^{-1}) = \frac1{\abs H}\sum_{h\in H}\phi(h)\psi(h^{-1}) $$
which is just equal to $(\phi,\psi|_H)$, as desired.

Thus we have that
$$ \find^G_H(f)(x) = \av\circ{\rm adj}(f)(x) = \frac1{\abs G}\sum_{g\in G}{\rm adj}(gxg^{-1}) = \frac1{\abs H}\sum_{g\in G,gxg^{-1}\in H}f(gxg^{-1}) $$
It is more convenient to switch $g$ with $g^{-1}$, i.e.\
$$ \find^G_H(f)(x) = \frac1{\abs H}\sum_{g\in G,g^{-1}xg\in H}f(g^{-1}xg) $$
Now notice that if $g\in G$ appears in the sum, then $gh$ appears in the sum for all $h\in H$.
Since $f$ is a class function, $f(g^{-1}xg)=f(h^{-1}g^{-1}xgh)$.
That is to say, we can split the sum over cosets, removing the factor of $1/\abs H$ as each coset has size $\abs H$:
$$ \find^G_H(f)(x) = \sum_{g\in G/H,g^{-1}xg\in H}f(g^{-1}xg) $$
(By $g\in G/H$ we mean that we choose a representative for the coset, the representative does not matter since $f(\bul x\bul^{-1})$ is equal to $f(x)$.)

\subsection{Mackey's theory}

Again, $G$ is a finite field whose order is not divisible by the characterstic of the algebraically closed field $k$.

Given a subgroup $H\subseteq G$, we would like to find irreducible representations of $G$ from the induced representations of irreducible $H$-representations.
Note that every irreducible $G$-representation $F$ is a subrepresentation of $\Ind^H_GE$ for $E$ an irreducible $H$-representation.
Indeed, since $\res^G_HF\neq0$, take a projection onto an irreducible subrepresentation $\res^G_HF\to E$.
The adjunction gives us a non-zero map $F\to\Ind^H_GE$, which must be an injection since $F$ is irreducible.

So if we know the irreducible representations of $H$, we can try and find the irreducible representations of $G$.
To do so, we consider their induced representations and study how they decompose into irreducible components.

Note that if we take $H=G$, then we get nothing of interest since $\Ind^G_G={\rm id}$.
And if we take $H=1$, then since its only irreducible representation is the trivial one we get that all irreducible $G$-representations are components of
$\Ind^1_G{\rm triv}=kG$, which again doesn't tell us anything we don't know.

Let us focus on a simpler question: when is $\Ind^H_GE$ irreducible for an irreducible $H$-representation $E$?
Now, $V$ is irreducible iff $\dim\hom_G(V,V)=1$, and
$$ \dim\hom_G(\Ind^H_GE,\Ind^H_GE) = \dim\hom_H(\res^G_H\Ind^H_GE,E) $$
so we would like to understand $\res^G_H\Ind^H_GE$.
It turns out it is beneficial to study $\res^G_K\circ\Ind^H_G$ for $H,K\subseteq G$ subgroups.

\bdefn

    Let $H,K\subseteq G$ be subgroups of $G$.
    Then a {\emphcolor $(H,K)$-double coset} is a set of the form
    $$ HgK = \set{hgk}[h\in H,k\in K] $$
    for $g\in G$.
    Double cosets are disjoint and cover $G$, and we denote by $H\ls G/K$ the set of all $(H,K)$-double cosets of $G$.

\edefn

It is not hard to see that $HgK$ is the orbit of $g$ under the action of $H\times K$ on $G$ given by $(h,k).g=hgk^{-1}$.
This quickly shows us that indeed double cosets are disjoint and cover $G$.
This action can help study the double cosets.

\bprop[title=Mackey]

    Let $H,K\subseteq G$ be subgroups, and choose representatives of the double cosets $(g_q)_{q\in H\ls G/K}$ (so $g_q\in q$).
    Then there is an isomorphism of functors:
    $$ \res^G_K\circ\Ind^H_G \cong \bigoplus_{q\in H\ls G/K}\Ind^{K\cap g_q^{-1}Hg_q}_K\circ\res^{g_q^{-1}Hg_q}_{K\cap g_q^{-1}Hg_q}\circ{}^{g_q^{-1}}\bul $$

\eprop

Recall that ${}^g\bul$ is a functor between $G$-representations which preserves the base space, and changes the action by conjugation.
That is for $x\in G$ and $v\in V$, $x.v=(g^{-1}xg)v$.

\bproof

    Recall that $\res^G_K\Ind^H_GV$ consists of functions $f\colon G\to V$ where $f(hg)=hf(g)$ for all $g\in G,h\in H$, and its action is $(kf)(g)=f(gk)$ for $k\in K$.

    For $q\in H\ls G/K$, let us consider the subspace $S_q(V)\subseteq\res^G_K\Ind^H_GV$ given by all those functions $f$ whose support lies within $q$, i.e.\ $f(x)=0$ if $x\notin q$.
    This is clearly a $K$-subrepresentation: it is clearly linear, and since $qK=q$ it is closed under $K$'s action.
    Furthermore, we also clearly have that the $S_q(V)$s are linearly independent, and every function $f$ can be written as a sum of functions from the $S_q(V)$s.
    That is,
    $$ \res^G_K\Ind^H_GV = \bigoplus_{q\in H\ls G/K}S_q(V) $$

    So let us now investigate these $S_q(V)$s.
    $S_q(V)$ is clearly isomorphic to the set of functions $f\colon q\to V$ where $f(hg)=hf(g)$ for $g\in G,h\in H$.
    Given $f\in S_q(V)$ consider the function $f'\colon K\to V$ defined by $f'(k)=f(g_qk)$.
    This defines a linear map $S_q(V)\to\St(K,k)$ mapping $f$ to $f'$.

    Now notice that for $k\in K$ and $\ell\in K\cap g_q^{-1}Hg_q$, we have
    $$ f'(\ell k) = f(g_q\ell k) = f(g_q\ell g_q^{-1}\cdot g_qk) = (g_q\ell g_q^{-1})f(g_qk) = (g_q\ell g_q^{-1})f'(k) $$
    That is to say,
    $$ f'\in\Ind^{K\cap g_q^{-1}Hg_q}_K\res^{g_q^{-1}Hg_q}_{K\cap g_q^{-1}Hg_q}{}^{g_q^{-1}}V $$
    We claim that our map $\Phi_q\colon S_q(V)\to\Ind^{K\cap g_q^{-1}Hg_q}_K\res^{g_q^{-1}Hg_q}_{K\cap g_q^{-1}Hg_q}{}^{g_q^{-1}}V$ is $K$-equivariant.
    Indeed, $(kf)'(k')=(kf)(g_qk')=f(g_qk'k)$, and $(k.f')(k')=f'(k'k)=f(g_qk'k)$.

    Now we claim that $\Phi_q$ is an isomorphism.
    We will explicitly describe its inverse.
    For $g\in q$, we have $g=hg_qk$ for $h\in H,k\in K$, so given $f'$ define $f(g)=hf'(k)$.
    Notice that $hf'(k)=hf(g_qk)=f(hg_qk)=f(g)$, so this gives an inverse in one direction.
    And $\Phi_q(f)(k)=f(g_qk)=f'(k)$ (since $g_qk=1g_qk$), so it is an inverse in the other direction as well.

    Thus we have
    $$ \res^G_K\Ind^H_G(V) = \bigoplus_{q\in H\ls G/K}S_q(V) \cong \bigoplus_{q\in H\ls G/K}\Ind^{K\cap g_q^{-1}Hg_q}_K\res^{g_q^{-1}Hg_q}_{K\cap g_q^{-1}Hg_q}{}^{g_q^{-1}}V $$
    as desired.
    (This isomorphism is also natural, which can be verified.)
    \qed

\eproof

Notice that if $H\subseteq G$ is normal, then
$$ S_q(V) \cong \Ind^{K\cap H}_K\res^H_{K\cap H}{}^{g_q^{-1}}V $$
and if $K=H$ then
$$ S_q(V) \cong {}^{g_q^{-1}}V $$

Let us denote by $\irr_k(G)$ the (isomorphism classes of) irreducible representations of $G$ over $k$.
Given a normal subgroup $H\subseteq G$, we have an action of $G$ on $\irr_k(H)$ by mapping $[E]$ to $[{}^gE]$ ($[\bul]$ are the isomorphism classes).
As discussed before, $H$ acts trivially here.
Thus we actually have an action of $G/H$ on $\irr_k(H)$.

\bprop

    Let $H\subseteq G$ be a normal subgroup, $E,F$ irreducible $H$-representations.
    \benum
        \item The dimension of $\hom_G(\Ind^H_GE,\Ind^H_GF)$ is equal to the cardinality of
        $$ {\rm trans}_{G/H}([E],[F]) = \set{q\in G/H}[{{}^q[E]=[F]}] $$
        \item $\Ind^H_GE$ is irreducible iff it is a free point for this action (that is, $\stab_{G/H}([E])=\set1$).
        \item $\Ind^H_GE$ and $\Ind^H_GF$ have common irreducible components iff $[E]$ and $[F]$ lie in the same orbit.
        In this case, $\Ind^H_GE$ and $\Ind^H_GF$ are isomorphic.
    \eenum

\eprop

\bproof

    \benum
        \item Since $H$ is normal in $G$, double cosets are simply one-sided cosets, in particular $H\sl G/H=G/H$.
        So let $(g_q)_{q\in G/H}$ be representatives of cosets in $G/H$.
        Then letting $K=H$ and using Mackey, we have
        $$ \hom_G(\Ind^H_GE,\Ind^H_GF) \cong \hom_H(\res^G_H\Ind^H_GE,F) \cong \bigoplus_{q\in G/H}\hom_H({}^{g_q^{-1}}E,F) $$
        By Schur, the dimensions of the hom-sets on the right-hand side are $0$ or $1$, depending on whether or not ${}^{g_q^{-1}}E$ is isomorphic to $F$.
        Thus the dimension is equal to the number of $q\in G/H$ such that ${}^{g_q^{-1}}[E]=[F]$, which gives the desired result.

        \item This is clear, since $\Ind^H_GE$ is irreducible iff $\dim\hom_G(\Ind^H_GE,\Ind^H_GE)=1$.
        By the above point this is iff ${\rm trans}_{G/H}([E],[E])=\stab_{G/H}([E])$ is a singleton.

        \item Note that $V,W$ have a common irreducible component iff $\hom_G(V,W)\neq0$.
        Thus the first assertion follows: they share an irreducible component iff ${\rm trans}_{G/H}([E],[F])\neq\emptyset$, i.e.\ iff they lie in the same orbit.
        Now, if $[E]$ and $[F]$ lie in the same orbit we must show that $\Ind^H_GE\cong\Ind^H_GF$.
        Suppose ${}^gE\cong F$, then $\Ind^H_G({}^gE)\cong\Ind^H_G(F)$.
        It is not hard to show that $\Ind^H_G({}^gE)\cong{}^g\Ind^H_G(E)\cong\Ind^H_G(E)$ as desired.
        \qed
    \eenum

\eproof

For example, suppose $H\subseteq G$ is normal with an index of $2$.
Then the orbits of $\irr_k(H)$ have a size of either $1$ or $2$ (since $G/H$ has a size of $2$).
If an orbit is a single point $[E]$, then $\Ind^H_GE$ decomposes as the sum of two irreducible representations.
This is because
$$ \dim\endo_G(\Ind^H_GE) = \#\set{q\in G/H}[{{}^q[E]=[E]}] = \abs{G/H} = 2 $$
and this tells us that it has two (non-isomorphic) irreducible subrepresentations.
And if an orbit is two points $[E],[E']$ then $\Ind^H_GE$ (which is isomorphic to $\Ind^H_GE'$ as they lie in the same orbit) is irreducible (since it must be a free point).
This gives us all irreducible $G$-representations.

\subsection{Brauer's induction theorem}

We focus now on representations over $\bC$.
For a subring $A\subseteq\bC$ we denote by $R_A(G)$ the $A$-span of characters of finite-dimensional $G$-representations in $\St(G,\bC)^\cl$.
This is an $A$-algebra, and the characters of irreducible $G$-representations form an $A$-basis for $R_A(G)$ (since $\ch_V$ can be written as the
$\bZ$-combination of irreducible characters).

Given a subgroup $H\subseteq G$, recall that we have a linear map $\find^H_G\colon\St(H,\bC)^\cl\to\St(G,\bC)^\cl$.
This map restricts to $R_A(H)\to R_A(G)$ (since characters are mapped to characters).
We will simplify notation and write $\find^H_G$ simply as $\Ind^H_G$.
In the other direction restriction also restricts to these algebras: $\res^G_H\colon R_A(G)\to R_A(H)$.

In particular we have $R_\bC(G)=\St(G,\bC)^\cl$.
And we denote $R(G)=R_\bZ(G)$, this is called the {\it representation ring} of $G$, elements of which are called {\it virtual characters.}
Notice that a virtual character $f\in R(G)$ is the character (or its minus) of an irreducible representation iff $(f,f)=1$.
Indeed, $f=\sum_in_i\chi_i$ for irreducible characters $\chi_i$, and so $(f,f)=\sum_in_i^2$ which is equal to $1$ iff there is a unique $n_i=\pm1$
with the rest being $0$.

Now suppose we have a family of subgroups $(H_i)_{i\in I}$ such that every element in $G$ is conjugate to an element in some $H_i$.
{\bf Artin's induction theorem} states that the map
$$ \bigoplus_{i\in I}\Ind^{H_i}_G\colon\bigoplus_{i\in I}R_\bQ(H_i)\to R_\bQ(G) $$
is surjective.
Some linear algebra shows that this is surjective iff the map
$$ \bigoplus_{i\in I}\Ind^{H_i}_G\colon\bigoplus_{i\in I}R_\bC(H_i)\to R_\bC(G) $$
is.
We will show that if $f\in R_\bC(G)$ is orthogonal to the image of this map, then it is zero.
Since the orthogonal complement is zero, this means that the image is the entire space.

Indeed, if $f$ is orthogonal to the image then $(f,\Ind^{H_i}_Gf')=0$ for all $f'\in R_\bC(H_i)$, by reciprocity this is equal to
$(\res^G_{H_i}f,f')=0$.
Since the form is nondegenerate, this means $\res^G_{H_i}f=0$ for all $i$.
But then $f=0$ since every $g\in G$ is conjugate to an element in some $H_i$ and $f$ is a class function.

We can take $H_i$ to be the cyclic subgroups of $G$.
And so every function in $R_\bQ(G)$ is the $\bQ$-linear combination of the induced characters of $\ch_\bC(H_i)$ (as these span $R_\bQ(H_i)$).
Thus we get

\bthrm[title=Artin's induction theorem]

    Let $V$ be a finite-dimensional $G$-representation.
    Then there exist cyclic subgroups $H_1,\dots,H_n\subseteq G$ and characters $\chi_i\in\ch_\bC(H_i)$ such that $\ch_V$ is a rational
    linear combination of the induced characters $\Ind^{H_1}_G\chi_1,\dots,\Ind^{H_n}_G\chi_n$.

\ethrm

The goal for the rest of this section is to prove the following:

\bthrm[title=Brauer's induction theorem]

    Let $V$ be a finite-dimensional $G$-representation.
    Then there exist subgroups $H_1,\dots,H_n\subseteq G$ and characters $\chi_i\in\ch_\bC(H_i)$ such that $\ch_V$ is a $\bZ$-combination
    of the induced characters $\Ind^{H_1}_G\chi_1,\dots,\Ind^{H_n}_G\chi_n$.

\ethrm

\bdefn

    A group $H$ is {\emphcolor $0$-nilpotent} if it is trivial, and {\emphcolor $n$-nilpotent} if $H/Z(H)$ is $n-1$-nilpotent.
    $H$ is {\emphcolor nilpotent} if it is $n$-nilpotent for some $n$.

\edefn

Subgroups and quotients of nilpotent groups are nilpotent, as well as the product of nilpotent groups.
Further $p$-groups are nilpotent.

\blemm

    Let $H$ be a non-Abelian nilpotent group, then $H$ contains a normal Abelian subgroup which is not central.

\elemm

\bproof

    Since $H$ is Abelian, $Z(H)\neq H$.
    Consider then the projection $H\to H/Z(H)$, and take an element $h\in H$ which is mapped to a non-trivial element of $Z(H/Z(H))$.
    Let $K$ be the subgroup generated by $Z(H)$ and $h$.
    It is not central, is Abelian, and it is normal: it is the preimage of a central group in $H/Z(H)$ (the group generated by the image of $h$).
    \qed

\eproof

\blemm

    Let $H$ be a finite nilpotent group, and let $E$ be an irreducible $H$-representation.
    Then there exists a subgroup $K\subseteq H$ and a character $\chi\in\ch_\bC(K)$ such that $E\cong\Ind^K_H(\bC_\chi)$.

\elemm

\bproof

    We proceed by induction on the order of $H$.
    First we reduce this to the case when $E$ is a faithful $H$-representation.
    Let $L\subseteq H$ be the kernel of the representation; if $L\neq1$ then viewing $E$ as an $H/L$-representation, by induction there exists a subgroup $\tilde K\subseteq H/L$ and a
    character $\chi\in\ch_\bC(\tilde K)$ such that $E$ is isomorphic to $\Ind^{\tilde K}_{H/L}(\bC_\chi)$.
    $\tilde K$ is of the form $K/L$ for some subgroup $L\subseteq K\subseteq H$, and we can view $\chi$ as a character in $\ch_\bC(K)$ (by precomposing with the projection $K\to\tilde K$).
    It is not hard to show (in fact, it is shown in homework) that $\Ind^{K/L}_{H/L}(\bC_\chi)\cong\Ind^K_L(\bC_\chi)$, as desired.
    So all that remains is to show this for the case when $L=1$, i.e.\ when $E$ is faithful.

    If $H$ is Abelian then this is clear: $E$ is of the form $\bC_\chi$ for $\chi\in\ch_\bC(H)$, so taking $H=K$ gives us the desired result.
    Then by the above lemma, there is a non-central Abelian normal subgroup $L\subseteq H$.
    We showed (a while ago) that either $E$ is either isotypical as an $L$-representation, or there exists a proper subgroup $L\subseteq K\subsetneq H$ such that $E\cong\Ind^K_H(F)$ for an
    irreducible $K$-representation $F$.
    In the latter case, by induction there is a subgroup $K'\subseteq K$ and a character $\chi\in\ch_\bC(K')$ such that $F\cong\Ind^{K'}_K(\bC_\chi)$.
    Then $E\cong\Ind^K_H\Ind^{K'}_K(\bC_\chi)\cong\Ind^{K'}_H(\bC_\chi)$ as desired.

    We will show that the former case is impossible: $E$ cannot be an isotypical $L$-representation.
    Since $L$ is Abelian, $L$ acts on $E$ by scalars.
    Then the center of the image of $L$ in $\gl(E)$ is contained in the center of the image of $H$ in $\gl(E)$.
    Since the representation is faithful, this means that the center of $L$ is contained in the center of $H$, which is impossible since $L$ is Abelian non-central.
    \qed

\eproof

\bdefn

    For a prime $p$, a finite group $H$ is {\emphcolor $p$-elementary} if it is isomorphic to the product of a $p$-group with a cyclic group.
    A group is {\emphcolor elementary} if it is $p$-elementary for some $p$.

\edefn

Since $p$-groups and cyclic groups are nilpotent, elementary groups are nilpotent as well.

We now state {\bf Brauer's induction theorem}:

\bthrm[title=Brauer's induction theorem]

    Let $\c E_G$ be the collection of elementary subgroups of $G$, then the map
    $$ \bigoplus_{H\in\c E_G}\Ind^H_G\colon\bigoplus_{H\in\c E_G}R(H)\to R(G) $$
    is surjective.

\ethrm

This indeed implies our original formulation of Brauer's induction theorem: a character of a $G$-representation is equal to the $\bZ$-combination of induced characters of elementary subgroup representations.
But by the above lemma, for an elementary subgroup $H$ and an $H$-representation $V$, we have that $\ch_V=\Ind^K_H(\chi)$ for some $\chi\in\ch_\bC(K)$.
Then $\Ind^H_G$ maps this to $\Ind^K_H(\chi)$, giving us the original result.

Let us for a second discuss something that probably should've been discussed before.
Recall that there is a canonical isomorphism $G\cong G^\op$ by mapping $g\mapsto g^{-1}$.
A right $G$-set is simply a left $G^\op$-set, which means that in general given a left $G$-set $X$ it can be given a right $G$-set structure by defining $xg=g^{-1}x$, i.e.\ precomposing with the isomorphism $G\to G^\op$.

Recall that a $G$-representation is simply a left $kG$-module.
Equivalently we can view a $G$-representation as a right $kG$-module by setting $vg=g^{-1}v$.

\blemm

    Let $H\subseteq G$ be a subgroup, $V$ a $G$-representation and $W$ an $H$-representation.
    Then
    $$ \ind^H_G(W\otimes\res^G_HV) \cong \ind^H_GW\otimes V $$

\elemm

\bproof

    We compute:
    $$ \ind^H_G(W\otimes\res^G_HV) = kG\otimes_{kH}(W\otimes\res^G_HV),\qquad \ind^H_GW\otimes V = (kG\otimes_{kH}W)\otimes V $$
    While a first glance may seem to indicate a trivial isomorphism via associativity: it is not the case.
    This is because the $kH$-module structure on $W\otimes\res^G_HV$ is not given by the $kH$-module structure on $W$, but by $\res^G_HV$'s as well: $h(w\otimes v)=hw\otimes hv$, not $hw\otimes v$.

    So instead we will compute their characters:
    $$ \ch_{\ind^H_G(W\otimes\res^G_HV)}(x) = \Ind^H_G(\ch_W\cdot\ch_V|_H) = \sum_{g\in G/H,g^{-1}xg\in H}\ch_W(g^{-1}xg)\ch_V(x) = (\Ind^H_G(\ch_W)\cdot\ch_V)(x) $$
    which is just the character of $\Ind^H_G(W)\otimes V$, as desired.
    \qed

\eproof

Notice that this implies that the image of our map is an ideal.
Indeed, for $\ch_V\in R(G)$ and $\ch_W\in R(H)$ we have that $\Ind^H_G(\ch_W)\cdot\ch_V=\Ind^H_G(\ch_{W\otimes\res^G_HV})$ by above, which is in the image of our map.
Let us denote the image of our map by $I$; so to show that our map is surjective, it is sufficient to show that $1\in I$.

For subrings $A,B\subseteq\bC$ let us denote by $R^B_A(G)\subseteq R_A(G)$ the subset of functions whose values lie in $B$.
$R^B_A(G)$ is clearly a subring of $R_A(G)$, and restriction and induction preserve $R^B_A(\bul)$.
Let us denote by $I^B_A$ the image of the map
$$ \bigoplus_{H\in\c E_G}\Ind^H_G\colon\bigoplus_{H\in\c E_G}R^B_A(H)\to R^B_A(G) $$
so that $I=I^\bC_\bZ$.
Define $A$ to be the subring of $\bC$ generated by all $\abs G$th roots of unity.

\blemm

    To show that our original map is surjective, it is sufficient to show $1\in I^\bC_A$.

\elemm

\bproof

    Since $A$ is a finitely-generated $\bZ$-module, all of its elements are integral.
    Thus $A\cap\bQ=\bZ$, and so $A/\bZ$ is a torsion-free finitely-generated $\bZ$-module.
    And so we can write $A=\bZ\oplus A'$ for a $\bZ$-submodule $A'\subset A$.
    Now consider the projection $p\colon A\to\bZ$ which induces a projection $P\colon R_A(G)\to R_\bZ(G)$ given by mapping an $A$-linear combination of irreducible characters to the
    linear combination of the projection of those values (i.e.\ $a_1\chi_1+\cdots+a_n\chi_n\mapsto p(a_1)\chi_1+\cdots+p(a_n)\chi_n$).

    Note that $P(a\chi)=p(a)\chi$ holds for all characters $\chi$ (not just irreducible ones).
    Indeed, $\chi$ is the $\bZ$-linear combination of irreducible characters, $\chi=\sum_in_i\chi_i$.
    Then $P(a\chi)=P(\sum_ian_i\chi_i)=\sum_ip(an_i)\chi_i=p(a)\sum_in_i\chi_i=p(a)\chi$.

    Now if $1\in I^\bC_A$ then we can write $1=\sum_ia_i\ch_{\Ind^{H_i}_GE_i}$ where $a_i\in A$ and $H_i$ are elementary subgroups of $G$.
    Then
    $$ 1 = P(1) = P\parens{\sum_ia_i\ch_{\Ind^{H_i}_GE_i}} = \sum_iP\parens{a_i\ch_{\Ind^{H_i}_GE_i}} = \sum_ip(a_i)\ch_{\Ind^{H_i}_GE_i} \in I $$
    as desired.
    \qed

\eproof

In particular since $I^\bZ_A\subseteq I^\bC_A$, it is sufficient to show that $1\in I^\bZ_A$.

So we consider $R^\bZ_A(G)\subseteq R^\bZ_\bC(G)$.
Notice that $R^\bZ_\bC(G)=\St(G,\bZ)^\cl$, so it is a finitely generated free Abelian group (as it is generated by the characters of $G$-representations).

We have the following lemma:

\blemm

    Let $L$ be a finitely generated Abelian group, $M\subseteq L$ a subgroup and $\ell\in L$.
    If $\ell\in M+p^kL$ for all prime $p$ and $k\geq1$, then $\ell\in M$.

\elemm

Applying this to $L=R^\bZ_\bC(G)$ and $M=I^\bZ_A$ and $\ell=1$, we have the following reduction:

\blemm

    In order to show that our map is surjective, it is enough to show that for every prime $p$ and $k\geq1$, there exists $f_{pk}\in I^\bZ_A$ such that all values
    of $1-f_{pk}$ are divisible by $p^k$.

\elemm

\blemm

    It is enough to show that for every prime $p$, there is a $f_p\in I^\bZ_A$ all of whose values are coprime to $p$.

\elemm

\bproof

    Let $k\geq1$, then by Euler's theorem all values of $1-f_p^{\phi(p^k)}$ are divisible by $p^k$.
    So defining $f_{pk}=f_p^{\phi(p^k)}$ gives us the requires condition for the previous lemma.
    \qed

\eproof

\bdefn

    Let $H$ be finite and $h\in H$.
    $h$ is {\emphcolor $p$-regular} if $h$'s order is coprime to $p$, and {\emphcolor $p$-torsion} if its order is a power of $p$.

\edefn

\blemm

    Let $H$ be finite and $p$ prime, then every $h\in H$ can be uniquely written as $h=h_rh_t$ for $h_r$ $p$-regular and $h_t$ $p$-torsion, where $h_r,h_t$ commute.

\elemm

We now reduce further:

\blemm

    It is sufficient to prove the following: for every $p$ prime and $g\in G$ $p$-regular, there exists $f_g\in I^\bZ_A$ such that $f_g(x)$ is coprime to $p$
    if $x_r$ is conjugate to $g$ and $0$ otherwise.

\elemm

\bproof

    Let $f_p$ be the sum of all $f_g$s where $g$ runs over the conjugacy classes of $p$-regular elements in $G$.
    Clearly $f_p$'s values are all coprime to $p$.
    \qed

\eproof

Let us consider a Sylow $p$-subgroup $S$ of $Z_G(g)=\set{h\in G}[hg=gh]$.
Then $E=S\cdot\gen g$ is a $p$-elementary subgroup of $G$.
Let $o(g)$ be the order of $g$, and define $f=o(g)\cdot\delta_{g}\in R^\bZ_A(\gen g)$.
$f$ does indeed lie in $R^\bZ_A$, since $\delta_g$ can be written (recall the commutative Fourier transform) as $\delta_g=\sum_{\chi\in\ch(\gen g)}\chi(g)e_\chi$ where
$e_\chi\chi(\bul)^{-1}/o(g)$, so $f=\sum_\chi\chi(g)^{-1}\chi$ which is an $\bZ\subseteq A$-span of characters, and $o(g)\delta_g$ clearly has its image lie in $\bZ$, so $f\in R^\bZ_A$ as desired.
Now let $f_1\in R^\bZ_A(E)$ be the pullback of $f$ under the projection $E\to\gen g$, i.e.\ $f_1=o(g)\delta_{Sg}$.
Finally we define $f_g=\Ind^E_G(f_1)\in I^\bZ_A$, and we claim that $f_g$ has the desired trait from the above lemma.

Indeed,
$$ f_g(x) = \sum_{y\in G/E,y^{-1}xy\in E}f_1(y^{-1}xy) = o(g)\cdot\#\set{y\in G/E}[y^{-1}xy\in Sg] = \#\set{y\in G/S}[y^{-1}xy\in Sg] $$
If $x_r$ is not conjugate to $g$, then this is $0$.
And if $x_r$ is conjugate to $g$, then by conjugating we can assume $x_r=g$ and so $x\in Z_G(\gen g)$, and by further conjugating by Sylow's theorem we can assume $x\in E$.
So we write $x=sg$ with $s\in S$.
Then $y\in G$ satisfying $y^{-1}xy\in Sg$ has $y^{-1}sgy=(y^{-1}sy)(y^{-1}gy)\in Sg$.
This requires that $y^{-1}sy\in S,y^{-1}gy=g$ i.e.\ $y\in Z_G(g)$ and $y^{-1}sy\in S$.
So we want to show that $\#\set{y\in Z_G(g)/S}[y^{-1}sy\in S]$ is coprime with $p$.

We can interpret this set as $\#\set{yS\in Z_G(g)/S}[s(yS)=yS]$.
So we want to show that the number of fixed points of $s$ in $S$'s action on $Z_G(g)/S$ is coprime with $p$.
We have the following result:

\blemm

    Let $X$ be a finite set whose size is coprime with $p$.
    Let $\sigma\colon X\to X$ be a permutation whose order is a power of $p$.
    Then the number of fixed points of $\sigma$ is coprime to $p$.

\elemm

\bproof

    $\gen\sigma$ acts on $X$, and the set of fixed points of this action are precisely the fixed points of $\sigma$.
    $\gen\sigma$ is a $p$-group, and in general ${\rm fp}(X)\equiv\#X\pmod p$ for any $p$-group set $X$.
    Since $\#X$ is coprime with $p$, ${\rm fp}(X)$ is as well.

    We show that ${\rm fp}(X)\equiv\#X\pmod p$ for a $P$-set $X$ ($P$ a $p$-group).
    Indeed, we have $\#X=\sum_{\sigma\in X/P}\#\sigma$.
    We know that $\#\sigma$ divides $\#P$ for all orbits $\sigma$.
    For a non-trivial orbit, this means its size is divisible by $p$ and thus equal to $0$ modulo $p$.
    Thus $\#X$ is equal to the number of trivial orbits, i.e.\ fixed points, modulo $p$.
    \qed

\eproof

$s$ is a permutation of $Z_G(g)/S$ whose order divides $s$'s order in $G$, which is a power of $p$ (following the group morphism which is $S$'s action).
Thus we get the desired result.

