\section{Schemes}

\subsection{Affine schemes}

\bdefn

    Let $R$ be a ring, then its {\emphcolor (prime) spectrum} is its set of prime ideals,
    $$ \spec(R) = \set{\p\leq R}[\hbox{$\p$ is prime}] . $$

\edefn

\bnote

    We can similarly define the {\emphcolor maximal spectrum} of a ring to be its set of maximal ideals.
    This is in general less useful than the prime spectrum, but its definition is analogous:
    $$ \specm(R) = \set{\m\leq R}[\hbox{$\m$ is maximal}] . $$

\enote

We can give $\spec(R)$ a topological structure.
Instead of defining the open sets of a spectrum, we will define its closed sets.

\bdefn

    Let $S\subseteq R$ be a subset, then define its {\emphcolor vanishing set} to be
    $$ V(S) = \set{\p\in\spec(R)}[S\subseteq\p] . $$

\edefn

Notice that clearly we have
$$ V(S) = V((S)) , $$
so that every vanishing set is a vanishing set of an ideal.
Note that $V(I)=V(J)$ does not imply $I=J$: for instance $I=2\bZ$ and $J=4\bZ$ in $R=\bZ$; their vanishing sets are both $\set{2\bZ}$.

\blemm

    A proper ideal $\p\leq R$ is prime iff $IJ\leq\p$ implies $I\leq\p$ or $J\leq\p$ for all ideals $I,J\leq R$.

\elemm

\bproof

    The ``if'' part is clear: if $ab\in\p$, then $(a)\cdot(b)=(ab)\in\p$ so either $(a)$ or $(b)$ is a subset of $\p$.
    Conversely, suppose that $\p$ is a prime ideal and $IJ\leq\p$ but $I,J$ are both not subsets of $\p$.
    Then let $x\in I,y\in J$ be elements not in $\p$: we then have $xy\in IJ\leq\p$, a contradiction to $\p$'s primality.
    \qed

\eproof

\bprop

    The collection of vanishing sets forms the collection of closed sets of a topology on $\spec(R)$.

\eprop

\bproof

    We need to show that $\spec(R),\emptyset$ are vanishing sets, the arbitrary intersection of vanishing sets is a vanishing set, and the binary
    union of vanishing sets is a vanishing set.
    Clearly we have
    $$ \spec(R) = V(0),\qquad \emptyset = V(R) $$
    so the first point is true.

    Now consider a collection $\set{S_\alpha}_\alpha$ of subsets.
    Then we have:
    $$ \bigcap_\alpha V(S_\alpha) = V\parens{\bigcup_\alpha S_\alpha} , $$
    since $S_\alpha\leq\p$ for all $\alpha$ iff $\bigcup_\alpha S_\alpha\leq\p$.
    In terms of ideals, we have that
    $$ \bigcap_\alpha V(I_\alpha) = V\parens{\parens{\bigcup_\alpha I_\alpha}} = V\parens{\sum_\alpha I_\alpha} . $$

    For the final point on binary unions, we can restrict our attention to vanishing sets of ideals.
    Due to the above lemma we have that
    $$ V(I) \cup V(J) = V(IJ) . $$
    Thus the collection of vanishing sets forms a topology of closed sets.
    \qed

\eproof

\bdefn

    The topology defined above on $\spec(R)$ is called the {\emphcolor Zariski topology} of $\spec(R)$.

\edefn

\bexam

    Consider $R=\bC[x]$, a PID.
    Its prime ideals are the principal ideals generated by irreducible elements and zero, in this case
    $$ \spec(\bC[x]) = \set{(x-\lambda)}[\lambda\in\bC]\cup\set{(0)} . $$

    For a polynomial $f\in\bC[x]$, we have that $(x-\lambda)\in V(f)$ iff $\lambda$ is a root of $f$.
    Thus
    $$ V(f) = \set{(x-\lambda)}[f(\lambda)=0] . $$
    As we showed above, for $I\leq\bC[x]$, we have that
    $$ V(I) = V\parens{\bigcup_{f\in I}\set f} = \bigcap_{f\in I}V(f) = \set{(x-\lambda)}[\hbox{$\lambda$ is a root of every polynomial in $I$}] . $$
    Thus $\spec(\bC[x])$'s proper closed sets are precisely the finite subsets of $\spec(\bC[x])-(0)$, i.e.\ it induces the cofinite topology on this subspace.
    The ideal $(0)$ is not contained in any proper closed set, so its closure is all of $\spec(\bC[x])$.

\eexam

\bexam

    Consider $R=\bZ$, a PID.
    Its prime ideals are the principal ideals generated by primes and zero, so
    $$ \spec(\bZ) = \set{(p)}[\hbox{$p$ prime}] \cup \set{(0)} . $$
    For $n\in\bZ$, $(p)\in V(n)$ iff $p$ divides $n$ so that
    $$ V(n) = \set{(p)}[\hbox{$p$ is a prime divisor of $n$}] . $$
    Thus $\spec(\bZ)$'s proper closed sets are precisely the finite subsets of $\spec(\bZ)-0$, so it again induces the cofinite topology on this subspace.
    As before, $(0)$'s closure is all of $\spec(\bZ)$.

\eexam

In the two examples above, we saw that the point $(0)$'s closure is all of $\spec(R)$.
If $(0)$ is a prime ideal, then this is always true: $(0)\in V(I)$ iff $I=(0)$ in which case $V(0)=\spec(R)$.
Thus $\spec(R)$ is the only closed set containing the zero ideal, and is thus its closure.

\bexam

    Consider the case that $R=k$ is a field.
    The only prime ideal of $k$ is the zero ideal, so $\spec(k)$ is a single-point space.
    There is only one topology on such a space: the trivial and discrete topology coincide.

\eexam

The map $R\mapsto\spec(R)$ is functorial.

\bdefn

    Let $\phi\colon R\to S$ be a ring morphism, then define $\phi^*\colon\spec(S)\to\spec(R)$ by
    $$ \phi^*(\q) = \phi^{-1}(\q),\qquad \q\in\spec(S) . $$

\edefn

This is well-defined, as we showed before that the preimage of a prime ideal is prime.

\bprop

    $\phi^*$ is continuous, and functorial.

\eprop

\bproof

    To show that $\phi^*$ is continuous, we need to show that its preimage on closed sets is closed.
    So let $I\leq R$ be an ideal, and consider $(\phi^*)^{-1}V(I)$:
    $$ \q \in (\phi^*)^{-1}V(I) \iff \phi^{-1}\q \in V(I) \iff I \leq \phi^{-1}\q \iff \phi(I) \leq \q , $$
    so that we have
    $$ (\phi^*)^{-1}V(I) = V(\phi(I)) , $$
    which is closed as desired.
    Note that while $\phi(I)\subseteq S$ may not be an ideal, all vanishing sets are closed.

    Functorality is clear.
    \qed

\eproof

\bnote

    As $\specm(R)\subseteq\spec(R)$, we endow $\specm(R)$ with the subspace topology.
    In particular its closed sets are
    $$ V_m(S) = \set{\m\in\specm(R)}[S\subseteq\m] . $$
    But note that $R\mapsto\specm(R)$ is not functorial: the above $\phi^*$ does not necessarily restrict to maximal ideals.

\enote

Let us consider more the Zariski topology.
Clearly $V(\bul)$ is contravariant (order reversing).

\bprop

    Let $I,J\leq R$ be ideals, then $V(I)=V(J)$ iff $\sqrt I=\sqrt J$.
    In particular $V(I)=V(\sqrt I)$ and $\spec(R)=V(\nil(R))$.

\eprop

\bproof

    We know that $\sqrt I$ is the intersection of all prime ideals containing $I$, meaning
    $$ \sqrt I = \bigcap_{\p\in V(I)}\p . $$
    So the first implication is clear: if $V(I)=V(J)$ then $\sqrt I=\sqrt J$.
    Now if $\sqrt I=\sqrt J$, let $\p\in V(I)$.
    If $\p\notin V(J)$ then there is an $f\in J$ not in $\p$, but then $f\in\sqrt J$ and $f\notin\p$ so $f\notin\sqrt I$, a contradiction.
    \qed

\eproof

\bnote

    Let the {\emphcolor Jacobson radical} of an ideal $I\leq R$ be the intersection of all maximal ideals containing $I$:
    $$ \jac_R(I) = \bigcap_{I\leq\m}\m = \bigcap_{\m\in V_m(I)}\m . $$
    We can alter the proof above to show that $V_m(I)=V_m(J)$ iff $\jac_R(I)=\jac_R(J)$.
    Thus $V_m(I)=V_m(\jac_R(I))$ and $\specm(R)=V_m(\jac(R))$, where $\jac(R)=\jac_R(R)$ is the usual Jacobson radical of a ring.

\enote

\bdefn

    Let $Z\subseteq\spec(R)$ be a subset.
    Define its {\emphcolor ideal} to be
    $$ \ideal(Z) = \set{f\in R}[Z\subseteq V(f)] . $$

\edefn

We can characterize $\ideal(Z)$ more concretely:
$$ \ideal(Z) = \bigcap_{\p\in Z}\p . $$
Indeed, if $f\in\ideal(Z)$ then $Z\subseteq V(f)$ so for every $\p\in Z$ we have that $\p\in V(f)$, meaning $f\in\p$.
Conversely if $f\in\p$ for every $\p\in Z$, then $\p\in V(f)$ for every $\p$ and so $Z\subseteq V(f)$.

From this characterization it is obvious that $\ideal(\bul)$ is also order reversing.

\bprop

    \benum
        \item Let $Z\subseteq\spec(R)$ be a subset, then $V(\ideal(Z))=\cl Z$, $Z$'s closure.
        \item Let $J\leq R$ be an ideal, then $\ideal(V(J))=\sqrt J$.
    \eenum

\eprop

\bproof

    For the first point we need to show that $V(\ideal(Z))$ is the smallest vanishing set containing $Z$.
    It contains $Z$ since if $\p\in Z$ then $\ideal(Z)\leq\p$ by definition (it is the intersection of all such $\p$), so $\p\in V(\ideal(Z))$.
    Now if $Z\subseteq V(J)$ then for $\p\in Z$ we have that $J\leq\p$, so that $J\subseteq\bigcap_{\p\in Z}\p=\ideal(Z)$.
    Since $V$ is contravariant, we have that $V(\ideal(Z))\subseteq V(J)$, as desired.

    For the second point, we have
    $$ \ideal(V(J)) = \bigcap_{\p\in V(J)}\p = \bigcap_{J\leq\p}\p = \sqrt J, $$
    as desired.
    \qed

\eproof

\bcoro

    Let $\p\in\spec(R)$ be a point, then its closure is $\cl{\set\p}=V(\p)$.
    In particular $\set\p$ is closed iff $\p$ is maximal.

\ecoro

\bproof

    This is because $\ideal\set\p=\p$, so apply the above proposition.
    And $V(\p)=\set\p$ iff $\p$ is maximal.
    \qed

\eproof

\bexam

    This shows that $\spec(R)$ need not be $T_1$.
    Recall that a space $X$ is $T_1$ if for every two distinct $x\neq y\in X$, there is an open set $U$ such that $x\in U$ and $y\notin U$.
    Equivalently a space is $T_1$ iff its singletons are closed: if singletons are closed let $U=\set y^c$, and conversely if for every $y\neq x$ we choose a
    $U_y$ such that $y\in U_y$ and $x\notin U_y$ we see that
    $$ X-\set x = \bigcup_{y\neq x}U_y $$
    is open and thus $\set x$ is closed.

    In any case, clearly if we have prime ideals which are not maximal, then $\spec(R)$ is not $T_1$.
    For instance, take $X=\spec(\bZ)$.
    Since $(0)\in X$ is not maximal, it is not closed and thus the space cannot be $T_1$.

    $\spec(R)$ {\it is\/} $T_0$ though: all distinct points are topologically distinct.
    Indeed, if $\p,\q\in\spec(R)$ have the same neighborhoods they have the same closure and thus $V(\p)=V(\q)$.
    Thus $\p\in V(\q)$ and $\q\in V(\p)$ so that $\p\leq\q\leq\p$ so that they are equal.

\eexam

\bdefn

    Let $P,Q$ be two posets.
    An {\emphcolor antitone Galois connection} are two order-reversing functions $F\colon P\to Q$ and $G\colon Q\to P$ such that for every $p\in P,q\in Q$,
    $$ q \leq Fp \iff p \leq Gq . $$

\edefn

An alternative method of looking at antitone Galois connections is via category theory.
Recall that we can view posets as categories, where there exists a unique morphism $a\to b$ iff $a\leq b$.

\bprop

    An antitone Galois connection between two posets $P,Q$ consists of two functors $F\colon P\to Q^\op\colon G$ such that $F\dashv G$ forms an adjunction.

\eprop

\bproof

    Order reversing functions between posets are just contravariant functors.
    So if $F,G$ form an antitone Galois connection, then they are functors $F\colon P^\op\to Q$ and $G\colon Q^\op\to P$.
    Then $F^\op\colon P\to Q^\op\colon G$ forms an adjunction: we have $q\leq Fp$ iff $F^\op p\leq^\op q$ iff $p\leq Gq$.
    The converse is shown similarly.
    \qed

\eproof

\bcoro

    $F\colon P\to Q\colon G$ forms an antitone Galois connection iff they are order reversing and $p\leq GFp$ and $q\leq FGq$ for all $p\in P,q\in Q$.

\ecoro

\bproof

    This is just the unit-counit formulation of an adjunction (remember contravariance!).
    A natural transformation between functors $X,Y$ between posets amounts to saying $X(\bul)\leq Y(\bul)$.
    \qed

\eproof

\bcoro

    The two functions $V\colon\pow(R)\to\pow(\spec(R))\colon\ideal$ form an antitone Galois connection.

\ecoro

\bproof

    The maps are order reversing and $S\subseteq\sqrt{(S)}=\ideal(V(S))$ and $Z\subseteq\cl Z=V(\ideal(Z))$, so they form an antitone Galois connection.
    Showing directly that $Z\subseteq V(S)$ iff $S\subseteq\ideal(Z)$ is easy to do as well.
    \qed

\eproof

\bdefn

    Let $R$ be a ring, $X=\spec(R)$ its spectrum.
    For $f\in R$, define
    $$ X_f = X - V(f) = \set{\p\in\spec(R)}[f\notin\p] . $$

\edefn

\bprop

    The collection $\set{X_f}_{f\in R}$ forms a basis of open sets for $X=\spec(R)$.

\eprop

\bproof

    This is equivalent to showing that $\set{V(f)}_{f\in R}$ forms a basis of closed sets; that they are closed and every closed set is an intersection of closed basis sets.
    Clearly $V(f)$ are closed, and for $S\subseteq R$ we have
    $$ V(S) = V\parens{\bigcup_{f\in S}\set f} = \bigcap_{f\in S}V(f) $$
    as desired.
    \qed

\eproof

This shows that for an open set $X-V(S)$, we have
$$ X - V(S) = \bigcup_{f\in S}X_f . $$

\bprop

    Let $R$ be a ring and $X=\spec(R)$.
    \benum
        \item Let $\set{f_\alpha}_\alpha$ be a collection of elements from $R$.
        Then $X=\bigcup_\alpha X_{f_\alpha}$ iff $(\set{f_\alpha}_\alpha)=R$.
        \item Let $f\in R$, then $X_f=X$ iff $f\in R^\times$.
        \item Let $f\in R$, then $X_f=\emptyset$ iff $f\in\nil(R)$.
    \eenum

\eprop

\bproof

    We saw that
    $$ \bigcup_\alpha X_{f_\alpha} = X - V(\set{f_\alpha}_\alpha) $$
    which is equal to $X$ iff $V(\set{f_\alpha}_\alpha)=\emptyset$.
    This is iff
    $$ V(\set{f_\alpha}_\alpha) = V((\set{f_\alpha}_\alpha)) = V(R) \iff \sqrt{(\set{f_\alpha}_\alpha} = R . $$
    Now $\sqrt J=R$ iff $J=R$, as $1\in\sqrt J$ iff $1\in J$.
    So this is iff $(\set{f_\alpha}_\alpha)=R$, as desired.

    The second point follows from the first: $X_f=X$ iff $(f)=R$ iff $f\in R^\times$.

    For the third point, $X_f=\emptyset$ iff $V(f)=X=V(0)$ iff $\sqrt{(f)}=\nil(R)$.
    This occurs iff $f\in\nil(R)$: clearly it implies this, and conversely $0\leq(f)\leq\nil(R)$ and radicals are order-preserving so we get $\sqrt{(f)}=\nil(R)$.
    \qed

\eproof

\bcoro

    $X=\spec(R)$ is compact.

\ecoro

\bproof

    We need to show every open cover of $X$ has a finite subcover.
    Let $\U$ be an open cover of $X$, then we can assume that $\U$ consists of basic open sets: $\U=\set{X_{f_\alpha}}_\alpha$.
    By the first point above, since $\U$ is a cover, $(\set{f_\alpha}_\alpha)=R$.
    Thus $1\in(\set{f_\alpha}_\alpha)$ so there are finitely many $f_1,\dots,f_n$ such that $1=x_1f_1+\cdots+x_nf_n$.
    Then $(f_1,\dots,f_n)=R$ and so again by the above proposition, $\set{X_{f_1},\dots,X_{f_n}}$ is a finite subcover of $\U$, as desired.
    \qed

\eproof

\bnote

    This result shows that the usual topological concept of compactness is in general not strong enough for algebraic geometry.
    Thus in algebraic geometry topological compactness is usually called {\emphcolor quasicompactness}, and compactness is reserved for a stronger
    concept.

\enote

\bprop

    Let $f\in R$, and consider $R_f$, the localization of $R$ at $\set{f^n}_{n<\omega}$.
    $\spec(R_f)$ is homeomorphic to $X_f$.

\eprop

\bproof

    Let $\iota\colon R\to R_f$ be the structure map.
    Then $\q\mapsto\iota^{-1}\q$ is a bijective correspondence between prime ideals of $R_f$ and prime ideals of $R$ which do not intersect $\set{f^n}$.
    A prime ideal $\p\leq R$ does not intersect $\set{f^n}$ iff $f\notin\p$.
    Thus $\iota^*\colon Y=\spec(R_f)\to\spec(R)$ is injective and its image are all prime ideals such that $f\notin\p$, i.e.\ $\im\iota^*=X_f$.

    We claim that $\iota^*$ is a homeomorphism onto its image.
    We already know it is continuous and bijective, so we need only show it is open; we will show it sends basic open sets to basic open sets.
    So let $g/f^n\in R_f$, with the basic open set
    $$ Y_{g/f^n} = \set{\q\in Y}[g/f^n\notin\q] . $$

    We claim that $\iota^*(Y_{g/f^n})=X_g\cap X_f$.
    Indeed, if $\q\in Y_{g/f^n}$ then $g/f^n\notin\q$ and so $g\notin\iota^*\q$ (since $g/f^n$ is $\iota(g)$ times a unit, so if $g\in\iota^*\q$ then $\iota(g)\in\q$
    and then $g/f^n\in\q$); thus $\iota^*\q\in X_g$ as desired.
    Conversely if $\p\in X_f\cap X_g$ then $g/f^n\notin \p R_f$: otherwise $g/f^n=h/f^m$ for $h\in\p$ and thus $f^k(f^mg-f^nh)=0$ so that $f^{m+k}g\in\p$, but $f,g\notin\p$
    a contradiction.
    Now, $\p R_f$ is prime and $\iota^*\p R_f=\p$, showing the equality.
    \qed

\eproof

We now focus on the connectedness of topological spaces, in particular spectra of rings.
Recall that a topological space is {\emphcolor disconnected} if it can be written as the disjoint union of two open subsets, equivalently if it has a non-trivial clopen subset.

\blemm

    A surjective ring morphism $\phi\colon R\to S$ induces a closed embedding $\phi^*\colon\spec(S)\to\spec(R)$ (i.e.\ its image is closed, and it is a topological embedding).
    In particular if $I\leq R$ is an ideal, then the projection $\pi\colon R\to R/I$ induces a closed embedding $\pi^*\colon\spec(R/I)\to\spec(R)$ whose image is $V(I)$.

\elemm

\bproof

    $\phi$ induces an isomorphism of rings $R/\ker\phi\to S$, which carries prime ideals of $R/\ker\phi$ bijectively to prime ideals of $S$.
    Thus $\phi^*$'s image is the set of all prime ideals containing $\ker\phi$, i.e.\ $\im\phi^*=V(\ker\phi)$ which is closed by definition.
    Since $\phi$ is surjective, $\phi^*$ is injective and thus a bijection onto its image.
    So to show it is a topological embedding, we need only show that it is a closed map.

    Indeed, for $J\leq S$,
    $$ \phi^*V(J) = \set{\phi^{-1}\q}[J\leq\q] = \set{\phi^{-1}\q}[\phi^{-1}J\leq\phi^{-1}\q] = V(\phi^{-1}J)\cap\im\phi^* , $$
    which is closed.
    (The second equality is due to $\phi$'s surjectivity.)
    \qed

\eproof

\bprop

    If $R=S\times T$ is a product ring, then $\spec(R)$ is homeomorphic to $\spec(S)\sqcup\spec(T)$, and thus disconnected.
    Conversely if $\spec(R)=Y\sqcup Z$ is disconnected, then there exist rings $S,T$ such that $Y,Z$ are homeomorphic to $\spec(S),\spec(T)$ respectively, and $R\cong S\times T$.

\eprop

\bproof

    Let $X=\spec(R)$.
    Consider $(1,0),(0,1)\in R$, so that $R$'s unit is their sum.
    Then we showed above that $X=X_{(1,0)}\cup X_{(0,1)}$, and this is a disjoint union since $V(0,1)\cup V(1,0)=X$ (if $\p\in X$ and $(0,1),(1,0)\notin\p$ then their product $0$ is not in
    $\p$, a contradiction).
    We also showed that $X_{(1,0)}\cong\spec(R_{(1,0)})$.

    So let $\pi\colon R\to S$ be the projection, so $\pi(1,0)=1\in S^\times$ so that by the universal property of localization $\pi$ induces a map $\pi'\colon R_{(1,0)}\to S$ which is
    surjective since $\pi$ is.
    Now if $\pi'((a,b)/(1,0))=\pi'((x,y)/(1,0))$ then $a=x$ so $(a,b)(1,0)=(x,y)(1,0)$ and thus $(a,b)/(1,0)=(x,y)/(1,0)$ showing that $\pi'$ is also injective.
    Thus $S\cong R_{(1,0)}$, and similarly $T\cong R_{(0,1)}$ so that we have desired result.

    Since $Y,Z$ are closed, there are $I,J\leq R$ such that $Y=V(I)$ and $Z=V(J)$.
    Then let $S=R/I$ and $T=R/J$; we showed above that $\pi_1\colon R\to R/I$ induces a homeomorphism $\pi_1^*\colon\spec(R/I)\to V(I)\subseteq\spec(R)$ and similar for $R/J$.
    Now, $Y\cap Z=V(I)\cap V(J)=V(I+J)$ is empty so that $I+J=R$, and thus $I,J$ are coprime.
    By the chinese remainder theorem,
    $$ R \cong R/I \times R/J = S\times T, $$
    as desired.
    \qed

\eproof

\bdefn

    An element $f\in R$ is {\emphcolor idempotent} if $f^2=f$.
    The {\emphcolor trivial idempotents} of $R$ are $0$ and $1$.

\edefn

If $f\in R$ is idempotent, then $fR$ is a ring.
Indeed, it is closed under addition and multiplication (it is the ideal generated by $f$), and $f$ is its multiplicative identity since it is idempotent.

\bprop

    $R\cong S\times T$ for non-trivial rings $S,T$ iff $R$ has nontrivial idempotents.

\eprop

\bproof

    If $R\cong S\times T$ then $(1,0),(0,1)$ are nontrivial idempotents.
    Conversely let $f\in R$ be a nontrivial idempotent, then define $S=fR$ and $T=(1-f)R$.
    Then $R\to S\times T$ given by $g\mapsto(fg,(1-f)g)$ is clearly an injective since if $fg=(1-f)g=0$ then $g=0$, and it is surjective since for $(fg,(1-f)h)\in S\times T$ then we want a $x\in R$ such that
    $fg=fx$ and $(1-f)h=(1-f)x$, so $(1-f)h=x-fg$ and thus $x=(1-f)h+fg$ gives the desired element.
    They are also non-trivial since $f\neq0$.
    \qed

\eproof

\bdefn

    A ring $R$ is {\emphcolor connected} if it has no nontrivial idempotents.

\edefn

\bcoro

    $\spec(R)$ is connected iff $R$ is.

\ecoro

\bdefn

    A topological space $X$ is {\emphcolor reducible} if there exist two proper closed subsets $Z,W\subsetneq X$ such that $X=Z\cup W$.
    Otherwise $X$ is {\emphcolor irreducible}.

\edefn

Note that irreducibility implies connectedness; if $X$ is disconnected then its union $Z\sqcup W$ of open sets are also closed.
The converse is not true in general (take the axes of $\bR^2$).

\bprop

    $X=\spec(R)$ is irreducible iff $\nil(R)$ is prime.

\eprop

\bproof

    If $\nil(R)$ is not prime, then there are $f,g\notin\nil(R)$ such that $fg\in\nil(R)$.
    Let $Z=V(f)$ and $W=V(g)$, so that
    $$ Z \cup W = V(f) \cup V(g) = V(fg) = X , $$
    where the second equality is because $V(f)\cup V(g)=V((f)(g))=V(fg)$, and the final equality is because $V(fg)\supseteq V(\nil(R))=X$.

    Conversely if $X$ is reducible, then $X=V(I)\cup V(J)$ so that $V(IJ)=X$ for $V(I),V(J)\neq X$.
    But then $IJ\leq\nil(R)$ and $I$ and $J$ cannot be contained in $\nil(R)$, so $\nil(R)$ is not prime.
    \qed

\eproof

\bcoro

    The spectrum of a domain is irreducible.

\ecoro

\bexam

    The affine space $\bA^n_\bC=\spec(\bC[x_1,\dots,x_n])$ is irreducible since $\bC[x_1,\dots,x_n]$ is a domain.

    Similarly the spectrum of $R=\bC[x_1,\dots,x_n]/(f)$ is irreducible when $f$ is irreducible (and then $(f)$ is prime and $R$ is a domain).
    But the converse does not hold: let $R=\bC[x]/(x^2)$, $x^2$ is reducible, and $\nil(R)=\sqrt{(x^2)}=(x)$ is prime.

\eexam

