\section{Commutative Algebra}

\subsection{Rings and Ideals}

We recall some basic notions regarding rings.

\bdefn

    A {\emphcolor ring} is a set $R$ equipped with two operations $+$ and $\cdot$, making it into a commutative group and a monoid respectively.
    Furthermore, they satisfy the distributivity properties:
    $$ a\cdot(b+c) = ab + ac,\qquad (a+b)\cdot c = ac + bc $$
    for all $a,b,c\in R$.
    The additive and multiplicative identities are denoted $0$ and $1$ respectively.

\edefn

\bnote

    In this course all rings are assumed to be {\emphcolor commutative}: $\cdot$ forms a commutative monoid.

\enote

\bdefn

    A {\emphcolor ring morphism} is a function between two rings, $f\colon R\to S$, which preserves structure: $f(x+y)=f(x)+f(y)$, $f(xy)=f(x)f(y)$,
    and $f(1)=1$.
    A {\emphcolor ring isomorphism} is a ring morphism with an inverse (equivalently a bijective ring morphism).

    A {\emphcolor subring} of a ring $R$ is a subset $S\subseteq R$ equipped with a ring structure making the inclusion $S\to R$ a ring morphism.

    An {\emphcolor ideal} of a ring $R$ is a nonempty subset $I\subseteq R$ closed under addition and multiplication by $R$: for $r\in R$ and $x\in I$,
    $rx\in I$.
    For an ideal $I\subseteq R$, we define its {\emphcolor quotient ring} $R/I$ to be the set of all cosets $[r]=r+I$ with ring structure given by
    $$ 0 = [0] = I,\qquad 1 = [1],\qquad [x] + [y] = [x+y],\qquad [x]\cdot[y] = [xy] , $$
    equivalently the unique structure making the projection map $\pi\colon R\to R/I,r\mapsto[r]$ a ring morphism.

\edefn

\bprop

    Let $f\colon R\to S$ be a ring morphism, then
    \benum
        \item For every ideal $J\leq S$, $f^{-1}(J)\leq R$ is an ideal.
        \item In particular $f$'s {\emphcolor kernel}, $\ker f=f^{-1}0$, is an ideal of $R$.
        And $f$'s {\emphcolor image}, $\im f=f(R)$ is a subring of $S$.
        $f$ then quotients to give a ring isomorphism $R/\ker f\cong\im f$.
        \item For an ideal $I\leq R$, consider the projection $\pi\colon R\to R/I$.
        Then $J\mapsto\bar J=\pi(J)$ and $\bar J\mapsto\pi^{-1}\bar J$ gives a one-to-one correspondence between ideals of $R$ containing $I$
        and ideals of $R/I$.
    \eenum

\eprop

\bdefn

    For a ring $R$ and a subset $S\subseteq R$, we denote the smallest ideal containing $S$ by $(S)$.
    This exists as the arbitrary intersection of ideals is an ideal.
    When $S=\set x$ is a singleton, we denote $(\set x)=(x)$; such an ideal is called a {\emphcolor principal ideal}.

\edefn

\bprop

    The ideal generated by $S$, $(S)$, is the set of all finite $R$-combinations of elements from $S$:
    $$ (S) = \set{\sum_{s\in S}r_ss}[r_s\in R,\hbox{ all but finitely many being zero}] . $$
    In particular, $(x)=Rx$ is the set of all multiples of $x$.

\eprop

\bproof

    It is not hard to see that this is an ideal, and clearly every ideal containing $S$ must contain it.
    \qed

\eproof

\bdefn

    Let $R$ be a ring, an element $x\in R$ is
    \benum
        \item a {\emphcolor zero divisor} if there is a $y\in R$ such that $xy=0$;
        \item a {\emphcolor unit} if there is a $y\in R$ such that $xy=1$;
        \item {\emphcolor nilpotent} if $x^n=1$ for some $n>0$.
    \eenum
    The set of units of $R$ is denoted $R^\times$ and forms a group.

\edefn

\bdefn

    A ring $R\neq0$ is
    \benum
        \item {\emphcolor reduced} if every $x\neq0$ is not nilpotent;
        \item an {\emphcolor integral domain} if every $x\neq0$ is not a zero divisor;
        \item a {\emphcolor field} if every $x\neq0$ is a unit.
    \eenum

\edefn

It is easy to see that a field is an integral domain is reduced.

\bprop

    For a ring $R$, the following are equivalent:
    \benum
        \item $R$ is a field,
        \item $R$ has no non-trivial ideals,
        \item Every ring morphism out of $R$ to a non-zero ring is injective.
    \eenum

\eprop

\bproof

    If $R$ is a field, then every nonzero element is a unit and thus the ideal generating it must be all of $R$.
    Conversely if $R$ has no non-trivial ideals, then for $x\neq0$ we must have that $1\in(x)$ and thus there is a $y\in R$ such that $xy=1$ so it is a unit.

    If $R$ has no non-trivial ideals then for every $f\colon R\to S\neq0$ we must have $\ker f=0$, as it is an ideal not containing $1$.
    Thus $f$ is injective.
    Conversely if $I\neq R$ is an ideal of $R$, then $\pi\colon R\to R/I$ is injective, but its kernel is $I$, so $I=0$.
    \qed

\eproof

\bprop

    If $\set{I_\alpha}_\alpha$ is a collection of ideals from $R$, then
    \benum
        \item $\bigcap_\alpha I_\alpha$ is an ideal, the largest ideal contained in each $I_\alpha$.
        \item $\sum_\alpha I_\alpha=\set{\sum_\alpha x_\alpha}[x_\alpha\in I_\alpha\hbox{ all but finitely many are zero}]$ is an ideal, the smallest ideal containing each $I_\alpha$.
    \eenum

\eprop

\bproof

    The first point is easily verified, and the second one follows because
    $$ \sum_\alpha I_\alpha = \parens{\bigcup_\alpha I_\alpha} . \qed $$

\eproof

\bdefn

    If $I,J$ are ideals then we define their {\emphcolor product} to be the ideal generated from all products of elements from $I$ and $J$:
    $$ IJ = (\set{xy}[x\in I,y\in J]) . $$

\edefn

Since $I,J$ are ideals and are thus closed under multiplication from $R$, for $x\in I$ and $y\in J$ we have that $xy\in I\cap J$.
Thus $IJ\subseteq I\cap J$.

\bdefn

    For an ideal $I\subseteq R$, we define its {\emphcolor radical} to be
    $$ \sqrt I = \set{r\in R}[r^n\in I\hbox{ for some $n\geq0$}] . $$

\edefn

\blemm

    The radical of an ideal is an ideal.

\elemm

\bproof

    If $a^n\in I$ and $r\in R$ then $(ra)^n=r^na^n\in I$ so that $ra\in\sqrt I$.
    And if $a^n,b^m\in I$ then
    $$ (a+b)^{n+m} = \sum_{i=0}^{n+m}\binom{n+m}i a^ib^{n+m-i} \in I , $$
    since for each $i$, if $i<n$ then $n+m-i>m$, so $a+b\in\sqrt I$.
    \qed

\eproof

\subsection{Prime and maximal ideals}

\bdefn

    A proper ideal $I\subset R$ is
    \benum
        \item {\emphcolor prime} if $ab\in I$ implies $a\in I$ or $b\in I$.
        Equivalently if $a,b\notin I$ then $ab\notin I$.
        \item {\emphcolor maximal} if it is not contained in any other proper ideal.
    \eenum

\edefn

\bprop

    An ideal $I$ is
    \benum
        \item prime iff $R/I$ is an integral domain;
        \item maximal iff $R/I$ is a field.
    \eenum
    Consequently, a maximal ideal is prime.

\eprop

\bproof

    $R/I$ is an integral domain iff $[a][b]\neq[0]$ when $[a],[b]\neq[0]$.
    This means that $ab\notin I$ when $a,b\notin I$, as desired.

    $R/I$ is a field iff it has no nontrivial ideals.
    Since ideals of $R/I$ are in bijection with ideals containing $I$, this is iff $I$ is maximal.
    \qed

\eproof

In particular the zero ideal $0=(0)=\set0$ is prime iff $R\cong R/(0)$ is an integral domain.

\bcoro

    Let $I$ be an ideal, then the correspondence $J\oto\bar J$ between ideals containing $I$ and ideals of $R/I$
    preserves primality and maximality: $J$ is prime/maximal iff $\bar J$ is.

\ecoro

\bproof

    We have that
    $$ (R/I)/\bar J = (R/I)/(J/I) \cong R/J , $$
    so that $(R/I)/\bar J$ is an integral domain/field iff $R/J$ is.
    The result follows.
    \qed

\eproof

\bdefn

    \benum
        \item An element $0\neq x\in R$ is {\emphcolor irreducible} if it is not a unit, and if $x=yz$ then either $y$ or $z$ is a unit.
        \item An integral domain $R$ is a {\emphcolor principal ideal domain} (or a {\emphcolor PID}) if all of its ideals are principal.
    \eenum

\edefn

\bprop

    If $R$ is a PID and $0\neq x$, then $(x)$ is prime iff $x$ is irreducible iff $(x)$ is maximal.

\eprop

\bproof

    If $(x)$ is prime, $x$ is not a unit since it is a proper ideal.
    And if $x=yz$ we have that $y$ or $z$ is in $(x)$, say $y=xt$.
    Then $x=yz=xtz$, and so $x(1-tz)=0$, and since $R$ is an integral domain this means $tz=1$ so $z$ is a unit.

    Now if $x$ is irreducible, then suppose $(x)\subseteq I$ is a proper ideal containing $(x)$.
    Since $R$ is a PID, $I=(y)$ so that $x\in(y)$ meaning $x=yz$.
    $(y)$ is a proper ideal so $y$ is not a unit, thus $z$ is, and then $y=xz^{-1}$ so that $y\in(x)$ and thus $(x)=(y)$.
    So $(x)$ is maximal, as desired.
    \qed

\eproof

\bexam

    The ring of integers $\bZ$ is a PID.
    Its irreducible elements are precisely its prime elements, so that its prime ideals are $(0)$ and $(p)$ for $p\in\bZ$ prime.

\eexam

\bexam

    If $k$ is a field, then its ring of polynomials $k[x]$ is a PID.
    Indeed, if $0\neq I\leq k[x]$ is an ideal, then let $f\in I$ have minimal degree.
    By polynomial division we see that $I=(f)$: for $g\in I$ we have that $g=fq+r$, but since $r=g-fq\in I$ has smaller degree than $f$,
    it must be zero.

    Its prime ideals are $(0)$ and $(f)$ for an irreducible monic polynomial $f$.

\eexam

\blemm

    If $f\colon R\to S$ is a ring morphism, and $\q\leq S$ is prime, then $f^{-1}\q\leq R$ is prime.

\elemm

\bproof

    A direct proof is easy, instead consider the composition:
    $$ \bar f\colon R\xtos fS\xtos\pi S/\q . $$
    By definition $f^{-1}\q=\ker\bar f$, and so $\bar f$ induces an embedding $R/f^{-1}\q\inj S/\q$.
    Subrings of integral domains are integral domains, so $R/f^{-1}\q$ is an integral domain, thus $f^{-1}\q$ is prime.
    \qed

\eproof

Subrings of fields are not fields, so the above proof does not show that the preimage of a maximal ideal is maximal.
Indeed, this is not true: consider the unique map $f\colon\bZ\to\bQ$.
Then $0\leq\bQ$ is maximal, but $f^{-1}0=\ker f=0$ is not maximal in $\bZ$.

\bthrm

    For every ring $R$ and every proper ideal $I<R$, there is a maximal ideal $\m$ containing $I$.

\ethrm

\bproof

    Consider the set of all proper ideals containing $I$.
    This is nonm-empty, as it contains $I$.
    Furthermore, the union of every chain of proper ideals is a proper ideal; it being an ideal is easy to see, and it is proper
    since $1$ is not in any element of the chain.
    Thus by Zorn, this set has a maximal element, which must be a maximal ideal.
    \qed

\eproof

\bcoro

    Let $x\in R$, then $x$ is a unit iff $x$ is not in any maximal ideal.
    That is,
    $$ R^\times = \bigcap_\m\m^c , $$
    where the intersection is over all maximal ideals of $R$.

\ecoro

\bproof

    An element is a unit iff it is not contained in any proper ideal.
    This is clear enough: if $x$ is not a unit then $(x)$ is proper, and proper ideals don't contain units.
    Every proper ideal can be extended to a maximal ideal, so the result follows.
    \qed

\eproof

\bdefn

    Let $R$ be a ring.
    Then define its {\emphcolor nilradical} to be $\nil(R)=\sqrt0$, i.e.\ the ideal of all nilpotent elements.

\edefn

Clearly $R$ is reduced iff $\nil(R)=0$.

\bprop

    Let $I\leq R$ be an ideal, and let $\pi\colon R\to R/I$ be the projection.
    Then $\sqrt I=\pi^{-1}\nil(R/I)$.

\eprop

\bproof

    We note that if $x^n\in I$ then $\pi(x)^n\in I$ so that $\pi(x)\in\nil(R/I)$ as desired.
    Conversely, if $\pi(x)\in\nil(R/I)$ then $\pi(x^n)=I$ so that $x^n\in I$ as desired.
    \qed

\eproof

\bprop

    For a ring $R$,
    \benum
        \item $R/\nil(R)$ is reduced;
        \item The nilradical is the intersection of all prime ideals:
        $$ \nil(R) = \bigcap_\p\p . $$
    \eenum

\eprop

\bproof

    We noted above that $\nil R=\sqrt{\nil R}=\pi^{-1}\nil(R/\nil(R))$.
    But we have also that $\nil R=\pi^{-1}0$, so $\pi^{-1}0=\pi^{-1}\nil(R/\nil(R))$ and since $\pi$ is surjective this means $\nil(R/\nil(R))=0$
    so that $R/\nil(R)$ is reduced.

    Clearly if $x\in\nil(R)$ then $x\in\p$ for every prime $\p$ since $x^n=0\in\p$.
    Now if $x\notin\nil(R)$ then we claim there exists a prime ideal $\p$ such that $x\notin\p$.
    We prove this using Zorn.
    Let $T=\set{1,x,x^2,\dots}$ and let $S$ be the poset of all ideals $I$ such that $I\cap T=\emptyset$; note that all ideals in $S$ are proper.
    As explained before the union of a chain of ideals is an ideal, and clearly if all of the ideals in the union don't intersect $T$, neither
    does the union.
    Thus by Zorn, $S$ has a maximal element $\p$.

    We claim that $\p$ is prime.
    Indeed, suppose that $ab\in\p$ but $a,b\notin\p$.
    Then by maximality we have that $x^n\in(a,\p)$ and $x^m\in(b,\p)$ for some $n,m\geq0$.
    But then $x^{n+m}\in(a,\p)(b,\p)\subseteq\p$, a contradiction.

    Thus there is a prime ideal $\p$ such that $\p\cap T=\emptyset$, in particular $x\notin\p$.
    So if $x\notin\nil(R)$ then $x\notin\bigcap_\p\p$, as desired.
    \qed

\eproof

\bcoro

    Let $I$ be an ideal of $R$, then
    $$ \sqrt I = \bigcap_{I\leq\p}\p , $$
    where the intersection is over all prime ideals containing $I$.

\ecoro

\bproof

    We showed above that $\sqrt I=\pi^{-1}\nil(R/I)$.
    By above $\nil(R/I)$ is the intersection over all prime ideals of $R/I$, which are all prime ideals of $R$ containing $I$:
    $$ \sqrt I = \pi^{-1}\bigcap_{I\leq\p}\bar\p = \bigcap_{I\leq\p}\pi^{-1}\bar\p = \bigcap_{I\leq\p}\p . \qed $$

\eproof

\bdefn

    Let $R$ be a ring, then define its {\emphcolor Jacobson radical} to be
    $$ J(R) = \bigcap_\m\m , $$
    the intersection of all maximal ideals.

\edefn

\blemm

    Let $\m\leq R$ be maximal.
    Then $x\notin\m$ iff there is a $y\in R$ such that $1-xy\in\m$.

\elemm

\bproof

    $x\notin\m$ iff $\bar x\in R/\m$ is nonzero, or equivalently is a unit.
    This is iff there is a $y\in R$ such that $\bar x\bar y=\bar 1$.
    This just means $1-xy\in\m$, as desired.
    \qed

\eproof

\bcoro

    For a ring $R$, its Jacobson radical $J(R)$ is the set of all elements $x\in R$ such that $1-yx\in R^\times$ for all $y\in R$.

\ecoro

\bproof

    Now if $x\notin J(R)$ then $x\notin\m$ for some $\m$, and then $1-yx\in\m$ for some $y$ and thus is not a unit.
    And if $1-yx$ is not a unit, then $1-yx\in\m$ for some maximal $\m$ and thus $x\notin\m$ so that $x\notin J(R)$.
    \qed

\eproof

\bdefn

    A ring $R$ is {\emphcolor local} if it has a unique maximal ideal $\m$.
    The field $k=R/\m$ is called its {\emphcolor residue field}.
    By abuse of notation we write $(R,\m)$ for a local field.

\edefn

\bprop

    \benum
        \item If $(R,\m)$ is a local ring then $R-\m\subseteq R^\times$.
        \item Conversely if $I<R$ is a proper ideal with $R-I\subseteq R^\times$ then $(R,I)$ is a local ring.
        \item If $\m\leq R$ is maximal such that $1+\m\subseteq R^\times$ then $(R,\m)$ is local.
    \eenum

\eprop

\bproof

    If $x\notin\m$ then $(x)$ is not contained in $\m$, and since $\m$ is the unique maximal ideal this means that $(x)=R$ so that $x$ is a unit.
    Also we noted that $R^\times$ is the intersection of all complements of maximal ideals, so this follows from that as well.

    Conversely, we have that
    $$ R-I \subseteq R^\times \bigcap_\m\m^c \implies I \supseteq \bigcup_\m\m , $$
    which means that $I$ contains all maximal ideals, which can only happen if it is the unique maximal ideal.

    We want to show that $R-\m\subseteq R^\times$.
    Indeed, if $x\notin\m$ then there is a $y$ such that $1-yx\in\m$ so that $yx\in1+\m$ and is thus a unit.
    So $x$ is a unit, as desired.
    \qed

\eproof

\bexam

    For $p\in\bZ$ prime, define
    $$ \bZ_{(p)} = \set{\frac ab}[a,b\in\bZ,b\notin(p)] \subseteq \bQ ,\qquad I = \set{\frac ab\in\bZ_{(p)}}[a\in(p)] . $$
    Then $(\bZ_{(p)},I)$ is local.
    Indeed, if $\frac ab\notin I$ then it is invertible (its inverse is $\frac ba$).
    We can generalize this, as we will in the next section.

\eexam

\subsection{Localization of rings}

Consider the following problem: for a ring $A$ and a subset $S\subseteq A$, we want a ring $S^{-1}A$ and a morphism $\iota\colon A\to S^{-1}A$
such that any morphism $f\colon A\to B$ such that $f(S)\subseteq B^\times$, $f$ factors uniquely through $\iota$.
That is, there is a unique $\hat f\colon S^{-1}A\to B$ such that $f=\hat f\circ\iota$:

\centerline{\drawdiagram{
    $A$&$S^{-1}A$\cr
    &$B$
}{
    \diagarrow{from={1,1}, to={1,2}, text=$\iota$, y distance=-.25cm}
    \diagarrow{from={1,1}, to={2,2}, text=$f$, y distance=-.25cm}
    \diagarrow{from={1,2}, to={2,2}, text=$\exists!$, x distance=.25cm, dashed}
}}

This is a universal property, and as such any solution is unique up to unique isomorphism.
In this section we will investigate the construction of such a solution.

\bdefn

    A subset $S\subseteq A$ is {\emphcolor multiplicatively closed} if $1\in S$ and for $a,b\in S$ we have $ab\in S$.

\edefn

It is sufficient to consider only the case that $S$ is multiplicatively closed.
Otherwise, let $\hat S$ be $S$'s multiplicative closure: the set of all finite products of elements from $S$.
Then $\hat S^{-1}R$ is a solution to the problem for $S$.
Indeed, if $f(S)\subseteq B^\times$, then $f(\hat S)\subseteq B^\times$ as well, as the product of units.
Then there is a unique factoring through $\hat S$'s $\iota$.

\bdefn

    For $S\subseteq A$ multiplicatively closed, define an equivalence on $A\times S$ by $(a,s)\sim(b,t)$ iff there is a
    $u\in S$ such that $u(at-bs)=0$, in which case we write $(a,s)\sim_u(b,t)$.

\edefn

\blemm

    This is indeed an equivalence relation.

\elemm

\bproof

    The subrelations $\sim_u$ are all clearly reflexive and symmetric, thus their union $\sim$ is as well.
    Now if $(a,s)\sim_\alpha(b,t)\sim_\beta(c,v)$ then $\alpha at=\alpha sb$ and $\beta bv=\beta ct$.
    So then $\alpha t\beta av=\alpha sb\beta v=\alpha\beta tcs$ so that $(a,s)\sim_{\alpha t\beta}(c,v)$.
    \qed

\eproof

\bexam

    A more naive equivalence would be $\sim_1$, i.e.\ $(a,s)\sim(b,t)$ iff $at-bs=0$.
    This corresponds to $\sim$ when $A$ is an integral domain, but is not generally an equivalence.

\eexam

\bdefn

    For $S\subseteq A$ multiplicatively closed, define $S^{-1}A$ to be the quotient of $A\times S$ by $\sim$.
    Write the equivalence class of $(a,s)$ by $a/s$.
    We give this a ring structure by
    $$ 0 = 0/1,\qquad 1 = 1/1,\qquad a/s + b/t = (at+bs)/st,\qquad a/s \cdot b/t = (ab)/(st) . $$
    Define $\iota\colon A\to S^{-1}A$ by $\iota(a)=a/1$.

\edefn

\bprop

    $S^{-1}A$ is indeed a ring, and $\iota$ is a ring morphism.

\eprop

\bproof

    We need to show that addition and multiplication are well-defined.
    Indeed, if $(a,s)\sim_u(a',s')$ and $(b,t)\sim_v(b',t')$ then $(at+bs,st)\sim_{uv}(a't'+b's',s't')$ which can be verified.
    And similarly $(ab,st)\sim_{uv}(a'b',s't')$.
    From well-definedness it is easy to see that addition and multiplication define a ring structure and $\iota$ is a ring morphism.
    \qed

\eproof

A few properties are immediate:
\benum
    \item For $s\in S$, $\iota(s)=1/s$ has an inverse $s/1$, so that $\iota(S)\subseteq(S^{-1}R)^\times$;
    \item We have that $a/s=0/1$ iff there is a $t\in S$ such that $at=0$.
    In particular, $\iota(a)=0$ iff there is a $t\in S$ such that $ta=0$.
    \item Every element of $S^{-1}A$ is of the form
    $$ a/s = a/1\cdot1/s = \iota(a)\cdot\iota(s)^{-1} . $$
\eenum

\bthrm

    This construction solves the above universal property.

\ethrm

\bproof

    Let $f\colon A\to B$ be such that $f(S)\subseteq B^\times$: we need to show that there is a unique map $g\colon S^{-1}A\to B$
    such that $f=g\circ\iota$.
    Indeed, we must define
    $$ g(a/s) = g(\iota(a)\iota(s)^{-1}) = (g\circ\iota(a))\cdot(g\circ\iota(s))^{-1} = f(a)f(s)^{-1} . $$
    This shows that $g$ is unique, if it exists.
    We need to show that this map is well-defined and a ring morphism.

    If $a/s=b/t$ then we must have that $f(a)f(s)^{-1}=f(b)f(t)^{-1}$ so that $g$ is well-defined.
    Indeed, if $u(at-bv)=0$ for $u\in S$, we have
    $$ f(u)\cdot(f(at)-f(bv)) = 0 $$
    and since $f(S)\subseteq B^\times$ we have that $f(u)$ is a unit so that $f(at)-f(bv)$, meaning this is well-defined.
    And this is a ring morphism, as can be checked easily.
    \qed

\eproof

\bcoro

    Let $f\colon A\to B$ be a ring morphism such that
    \benum
        \item $f(S)\subseteq B^\times$,
        \item if $f(a)=0$ then $ta=0$ for some $t\in S$,
        \item every element of $B$ is of the form $f(a)f(s)^{-1}$ for $a\in A$ and $s\in S$.
    \eenum
    Then there is a unique isomorphism $g\colon S^{-1}A\to B$ such that $g\circ\iota=f$.

\ecoro

\bproof

    $f$ solves the universal problem above for $S$ as well.
    If $h\colon A\to C$ is a map where $h(S)\subseteq C^\times$, then we need to define $\hat h\colon B\to C$ by
    $$ \hat h(f(a)f(s)^{-1}) = h(a)h(s)^{-1} , $$
    which is well-defined and a ring morphism for similar reasons as above.
    Since universal properties are unique up to unique isomorphism, the result follows.

    Alternatively we can prove this directly: let $g\colon S^{-1}A\to B$ be the unique map where $g\circ\iota=f$ by the universal property.
    By definition $g(a/s)=f(a)f(s^{-1})$.
    By the third point $g$ is surjective, and it is injective since if $g(a/s)=0$ we have $f(a)=0$ so that $ta=0$, which means that $a/s=0$.
    \qed

\eproof

\bexam

    If $A$ is an integral domain, then $S=A-0$ is multiplicatively closed, and $S^{-1}A$ is the field of fractions of $A$.

\eexam

\bprop

    For $A\neq0$, then $S^{-1}A=0$ iff $0\in S$.

\eprop

\bproof

    We have that $S^{-1}A=0$ iff $1/1=0/1$ iff there is a $s\in S$ such that $s\cdot1=0$.
    \qed

\eproof

\bdefn

    Let $\p\subseteq A$ be prime.
    Then $S=A-\p$ is multiplicatively closed, and we denote its localization by $A_\p=S^{-1}A$.

    For $a\in A$, $S=\set{1,a,a^2,\dots}$ is multiplicatively closed, and we denote its localization by $A_a=S^{-1}A$.

\edefn

\bnote

    Note that we have two distinct localizations here, which have similar names in some cases.
    If $A$ is a PID and $p\in A$ is irreducible, then $A_{(p)}$ is localization by $A-(p)$, which $A_p$ is localization by $\set{p^n}_{n<\omega}$.
    Elements of the former are of the form $a/s$ where $p$ does not divide $s$, while elements of the former are of the form $a/s$ where $s=p^n$.
    Quite different!

\enote

\bdefn

    For a ring morphism $f\colon A\to B$ and an ideal $I\subseteq A$, define the ideal $IB=(f(I))$ of $B$.
    Explicitly, $IB$ is the set of all finite sums $\sum_if(a_i)b_i$ for $a_i\in I$.

\edefn

In general we have $f^{-1}(f(I))\supseteq I$ and $f(f^{-1}(J))\subseteq J$.
These imply $f^{-1}(IB)\supseteq I$ and $f^{-1}(J)B\subseteq J$.

\bdefn

    For two ideals $I,J$, define $(I:J)=\set{x\in A}[xJ\subseteq I]$.
    This is an ideal containing $I$.
    For $a\in A$, define $(I:a)=(I:(a))=\set{x\in A}[xa\subseteq I]$.

    For $S\subseteq R$ multiplicatively closed, define the ideal $S^{-1}I=IS^{-1}A\subseteq S^{-1}A$.
    This is the ideal generated by $\iota(I)$.

\edefn

So we have two maps: $S^{-1}\bul$ mapping ideals of $A$ to ideals of $S^{-1}A$, and $\iota^{-1}$ mapping ideals of $S^{-1}A$ to $A$:
$$ \set{\hbox{Ideals of $A$}} \xtof{\quad S^{-1}\quad}[\quad\iota^{-1}\quad] \set{\hbox{Ideals of $S^{-1}A$}} . $$
We wish to relate the two.

\bprop

    \benum
        \item For an ideal $I\subseteq A$, $S^{-1}I=\set{a/s}[a\in I,s\in S]$.
        \item For an ideal $J\subseteq S^{-1}A$, we have $S^{-1}\iota^{-1}J=J$.
        \item For an ideal $I\subseteq A$, we have $\iota^{-1}S^{-1}I=\bigcup_{s\in S}(I:s)$.
        \item We have $S^{-1}I=S^{-1}A$ iff $I\cap S\neq\emptyset$.
        \item For $\p\subseteq A$ prime, if $I\cap S\neq\emptyset$ then $S^{-1}\p\subseteq S^{-1}A$ is prime (otherwise it is all of $S^{-1}A$).
        \item The correspondence $\q\mapsto\iota^{-1}\q$ is a bijection between prime ideals of $S^{-1}A$ and prime ideals of $A$ not intersecting $S$.
        Its inverse is $S^{-1}$.
        That is, we have a one-to-one correspondence:
        $$ \set{\hbox{Prime ideals of $A$ not intersecting $S$}} \xtof{\quad S^{-1}\quad}[\quad\iota^{-1}\quad] \set{\hbox{Prime ideals of $S^{-1}A$}} . $$
    \eenum

\eprop

\bproof

    \benum
        \item $\set{a/s}[a\in I,s\in S]$ is an ideal and contains $\iota(I)$.
        It is clearly closed under multiplication by $S^{-1}A$ and for $a/s+b/t=(at+bs)/(st)$, notice that $at+bs\in I$.
        By definition $S^{-1}I$ is the smallest ideal containing $\iota(I)$, and it consists of finite sums $\sum_i\iota(a_i)x_i/s_i$.
        This clearly contains $a/s$, and thus they must be equal.

        \item As explained before, in general we have $S^{-1}\iota^{-1}J=\iota^{-1}(J)S^{-1}A\subseteq J$, we must prove the opposite inclusion.
        As above, we have $S^{-1}\iota^{-1}J=\set{a/s}[\iota(a)\in J,s\in S]$.
        For $x/s\in J$, we have that $\iota(x)=x/1\in J$ and so $x/s\in S^{-1}\iota^{-1}J$ as desired.

        \item If $a\in\iota^{-1}S^{-1}I$ then $\iota(a)\in S^{-1}I$ so by the first point $a/1=b/s$ for $b\in I$.
        Then $asu=bu$ for some $u\in S$ and since $bu\in J$ this means that $a\in(I:su)$.
        Conversely if $a\in(I:s)$ for $s\in S$ then $as\in I$ and so $\iota(a)=a/1=(as)/s\in S^{-1}I$.

        \item If $S^{-1}I=S^{-1}A$ then $1/s=a/t$ for $a\in I$, so $as=t$ meaning $as\in I\cap S$.
        Conversely if $a\in I\cap S$ then $1/1=a/a\in S^{-1}I$.

        \item If $S\cap\p=\emptyset$ then we need to show that $S^{-1}\p$ is prime.
        If $a_1/s_1,a_2/s_2\notin S^{-1}\p$ we want to show $(a_1a_2)/(s_1s_2)\notin S^{-1}\p$.
        By assumption $a_1,a_2\notin\p$.

        Now we claim that if $a\notin\p$, $a/s\notin S^{-1}\p$, which would complete the proof.
        Assume the contrary: $a/s=b/t$ for $b\in\p$ then there is a $u\in S$ such that $atu=bsu$.
        Then $atu\in\p$, but $a\notin\p$ and $t,u\in S$ and thus also are not in $\p$, a contradiction to $\p$'s primality.

        \item Preimages of prime ideals are prime, so $\iota^{-1}$ maps prime ideals to prime ideals.
        By the second point $S^{-1}\iota^{-1}\q=\q$ for all prime ideals $\q\subseteq S^{-1}A$.
        It remains to show that $\iota^{-1}S^{-1}\p=\p$ for all prime $\p\subseteq A$ which don't intersect $S$.
        We know that $\p\subseteq\iota^{-1}S^{-1}\p$ as is true in general, so we want to show that $\iota^{-1}S^{-1}\p=\bigcup_{s\in S}(\p:s)$ is contained in $\p$.
        Indeed, if $a\in(\p:s)$ then $as\in\p$ and since $s\notin\p$ we have that $a\in\p$ as desired.
        \qed
    \eenum

\eproof

\bcoro

    Let $\p\subseteq A$ be prime, and consider its localization $A_\p$.
    Then $A_\p$ is a local ring with maximal ideal $\p A_\p$.
    Moreover, the correspondence $\q\to\q A_\p$ is a bijection between prime ideals of $A_\p$ and prime ideals $\q\subseteq\p$.

\ecoro

\bproof

    Let $S=A-\p$, so that $A_\p=S^{-1}A$, and $\p A_\p=S^{-1}\p$.
    The last point is clear: $\q\cap S=\emptyset$ iff $\q\subseteq\p$.
    $A_\p$ being local is also clear: $S^{-1}\bul$ is order-preserving, so $S^{-1}\q\subseteq S^{-1}\p$ for all prime ideals $\q$.
    Thus $S^{-1}\p$ is maximal.
    \qed

\eproof

\bcoro

    If $a\in A$ is not nilpotent, then there is a prime ideal $\p\subseteq A$ such that $a\notin\p$.

\ecoro

\bproof

    As before, consider the multiplicative set $S=\set{a^n}_{n<\omega}$, and $A_a=S^{-1}A$.
    Since $a$ is not nilpotent, $0\notin S$, so that $A_a\neq0$.
    So take a prime ideal $\p\subseteq A_a$, and then $\iota^{-1}\p$ is a prime ideal of $A$ not intersecting $S$, thus not containing $a$.
    \qed

\eproof

\bnote

    This completes the proof before that
    $$ \nil(A) = \bigcap_\p\p $$
    but without appealing to Zorn, and thus not relying on the Axiom of Choice.

\enote

\subsection{Modules}

\bdefn

    Let $A$ be a ring, then a {\emphcolor module} over $A$ is an Abelian group $M$ along with a ring morphism
    $\phi\colon A\to\endo(M)$.
    We define {\emphcolor multiplication} to be a function $\mu\colon A\times M\to M$, written by juxtaposition, by
    $$ \mu(a,m) = \phi(a)m . $$

\edefn

An $A$-module can also be defined via multiplication: require that multiplication distributes both ways, and $1\in A$ is an identity:
$$ (a+b)m = am + bm,\qquad a(m+n) = am + an,\qquad a(bm) = (ab)m,\qquad 1m = m . $$

\bdefn

    For two $A$-modules $M$ and $N$, an {\emphcolor $A$-module morphism} is a map $f\colon M\to N$ which is an Abelian group morphism and also
    multiplicative:
    $$ f(am) = af(m) $$
    for all $a\in A,m\in M$.

\edefn

\bnote

    I'm going to skip the rest of the standard definitions, for Chrissake.

\enote

\bdefn

    If $M$ is an $A$-module, we define $\ann_A(M)\leq A$ the {\emphcolor annihilator} of $M$ by
    $$ \ann_A(M) = \set{a\in A}[aM=0] = \ker\phi , $$
    where $\phi$ is the module map $\phi\colon A\to\endo(M)$.
    This is then an ideal.
    For $S\subseteq M$ a subset we define $\ann_A(S)$ similarly.

    A module $M$ is {\emphcolor faithful} if $\ann_A(M)=0$.

\edefn

$\ann_A(S)$ is equal to $\ann_A(\gen S)$, the annihilator of the submodule generated by $S$ (all $R$-combinations of elements from $S$),
and is thus also an ideal.

\bdefn

    For a ring $A$, an {\emphcolor $A$-algebra} is a ring $B$ equipped with a ring morphism $\phi\colon A\to B$.

\edefn

Note that $B$ can be embedded into $\endo(B)$ mapping $b\in B$ to multiplication by $b$.
Thus an $A$-algebra is an $A$-module with multiplication given by $a\cdot b=\phi(a)b$.

\bdefn

    For two $A$-algebras $B,C$, an {\emphcolor $A$-algebra morphism} is a map $f\colon B\to C$ which is both a ring morphism and an
    $A$-module morphism.

\edefn

If $\phi\colon A\to B,\psi\colon A\to C$ are the algebra maps, then $f\colon B\to C$ is an algebra morphism iff it is a ring morphism and
$\psi=f\circ\phi$:
$$ \psi(a) = a1_C = af(1_B) = f(a1_B) = f\circ\phi(a) . $$

\bdefn

    For an $A$-module $M$ and a submodule $N$, its quotient module $M/N$ is the set of cosets $m+N$ with the usual $M$-module structure.
    The usual isomorphism results hold.

\edefn

\bprop

    Let $I\leq A$ be a module, then every $A/I$-module is an $A$-module with $\ann_A(M)\supseteq I$ and vice-versa.
    In particular an $A$-module $M$ can be viewed as a faithful $A/\ann_A(M)$-module.

\eprop

\bproof

    Consider the projection $\pi\colon A\to A/I$.
    Composing with $\phi\colon A/I\to\endo(M)$ gives an $A$-module structure to an $A/I$-module and $\ker(\phi\circ\pi)\supseteq\ker\phi=I$.
    Conversely if $\phi\colon A\to\endo(M)$ has kernel containing $I$ we can quotient it by $I$.
    \qed

\eproof

\bdefn

    For a collection of $A$-modules $\set{M_\alpha}_\alpha$ we define their {\emphcolor direct product} and {\emphcolor direct sum}
    by
    $$ \prod_\alpha M_\alpha = \set{(m_\alpha)}[m_\alpha\in M_\alpha],\qquad
    \bigoplus_\alpha M_\alpha = \set{(m_\alpha)\in\prod_\alpha M_\alpha}[\hbox{All but finitely many of $m_\alpha$ are zero}] . $$
    $\prod_\alpha M_\alpha$'s module structure is given by pointwise operations, and $\bigoplus_\alpha M_\alpha$ is viewed as a submodule.
    These form the products and coproducts in the category of $A$-modules.

\edefn

\bdefn

    An $A$-module $M$ is {\emphcolor free} if
    $$ M \cong A^{\oplus S} = \bigoplus_{\alpha\in S}A $$
    for some set $S$.
    Note that $\set{e_\alpha}_{\alpha\in S}$ (the usual standard basis vectors) forms a basis of $A^{\oplus S}$ in the sense that every
    element of $A^{\oplus S}$ can be written as a unique combination of standard basis vectors.
    A collection of elements $\set{m_\alpha}_{\alpha\in S}$ is a {\emphcolor basis} of $M$ if mapping $m_\alpha\mapsto e_\alpha$ extends
    to an isomorphism $M\cong A^{\oplus S}$.

\edefn

A basis of $M$ is equivalently a set of elements which generate $M$ and are also linearly independent in the usual way.

A free module is called such because they are free in the usual sense with respect to the obvious forgetful functor $\Mod_A\to\Set$.
That is, $\hom_A(A^{\oplus S},B)\cong\hom_\Set(S,B)$.

\bprop

    If $A^{\oplus S}\cong A^{\oplus S'}$ then $S$ and $S'$ have the same cardinality.

\eprop

\bproof

    Let $\m\leq A$ be maximal, then the isomorphism $A^{\oplus S}\cong A^{\oplus S'}$ induces an isomorphism
    $(A/\m)^{\oplus S}\cong(A/\m)^{\oplus S'}$.
    $A/\m$ is a field, and so this is an isomorphism of $A/\m$-modules (vector spaces) of dimensions $\abs S,\abs{S'}$.
    Since the dimension is invariant, their cardinalities must be equal.
    \qed

\eproof

\bdefn

    If $M$ is a free $A$-module, then its {\emphcolor rank} is the cardinality of $S$ where $M\cong A^{\oplus S}$.

\edefn

\bdefn

    An $A$-module $M$ is {\emphcolor finitely generated} ({\emphcolor fg}) if there is a finite set $S\subseteq M$ such that $M=\gen S$.

\edefn

\bprop

    An $A$-module is finitely generated iff it is a quotient of $A^n$ for some $n$ (up to isomorphism).

\eprop

\bproof

    If $\phi\colon A^n\to M$ is surjective, then $\phi(e_i)$ generate $M$ clearly and so it is finitely-generated.
    Thus if $M\cong A^n/N$ is a quotient of $A^n$ then the projection $\pi\colon A^n\to A^n/N\cong M$ is surjective and thus $M$
    is finitely-generated.
    Conversely let $m_i$ generate $M$, and define $\phi\colon A^n\to M$ by $\phi(e_i)=m_i$ which extends by linearity to all of $A^n$ since
    it is free.
    This is surjective and then $A^n/\ker\phi\cong M$ as desired.
    \qed

\eproof

\bprop

    Let $M$ be a fg $A$-module, and $I\leq A$ an ideal such that $M=IM$.
    Then there is an $x\in I$ such that $(1-x)M=0$ ($xm=m$ for all $m\in M$).

\eprop

\bproof

    Let $m_1,\dots,m_n$ generate $M$, we induct on $n$.
    For $n=0$ we have $M=0$ so this is clear.
    Consider $N=M/(m_n)$, then $IN=(IM+(m_n))/(m_n)=M/(m_n)=N$ and $N$ is generated by $n-1$ elements, so by induction there is an $x\in I$
    such that $(1-x)N=0$.
    Meaning $(1-x)M\subseteq(m_n)$, and since $m_n\in M=IM$ we have
    $$ (1-x)m_n \in (1-x)IM = I(1-x)M \subseteq I(m_n), $$
    so $(1-x)m_n=ym_n$ for some $y\in I$ and thus $(1-x-y)m_n=0$.
    Now $(1-x-y)(1-x)\in 1+I$ and satisfies
    $$ (1-x-y)(1-x)M \subseteq (1-x-y)(m_n) = 0 $$
    as desired.
    \qed

\eproof

\bcoro[title=Nakayama's lemma]

    Let $M$ be a fg $A$-module, such that $J(A)M=M$, then $M=0$.

\ecoro

\bproof

    By above there is an $x\in J(A)$ such that $(1-x)M=0$.
    Since $x\in J(A)$, $1-x\in A^\times$ so that $(1-x)M=M=0$ as desired.
    \qed

\eproof

\bcoro

    Let $M$ be a fg $A$-module and $N\subseteq M$ a submodule such that $M=IM+N$ for some $I\subseteq J(A)$, then $M=N$.

\ecoro

\bproof

    Consider $M/N$, then $I(M/N)=(IM+N)/N=M/N$, so by above $M/N=0$ meaning $M=N$.
    \qed

\eproof

\bcoro

    Let $(A,\m)$ be local, $M$ a fg $A$-module.
    Then $x_1,\dots,x_n\in M$ generate $M$ iff $\bar x_1,\dots,\bar x_n\in M/\m M$ generate $M/\m M$ over $A/\m$.

\ecoro

\bproof

    If $x_i$ generate $M$ then $\bar x_i$ clearly generate $M$.
    Conversely, let $N=(x_1,\dots,x_n)$ be the submodule generated by $M$, and consider the composition $f\colon N\to M\to M/\m M$.
    $f$'s image is the span of $(\bar x_i)_i$, so all of $M/\m$.
    Thus $f$ is surjective, so $M=N+\m M$, since for every $m\in M$ there is an $n\in N$ such that $f(n)-m\in\m M$ by surjectivity,
    and by above since $J(A)=\m$ this means $M=N$ as desired.
    \qed

\eproof

\bdefn

    Recall the definition of an exact sequence:
    $$ M \xtos fN\xtos gP $$
    is {\emphcolor exact} at $N$ if $\im f=\ker g$.

\edefn

\subsection{Localization of modules}

\bdefn

    Let $S\subseteq A$ be multiplicatively closed and $M$ an $A$-module.
    Define an equivalence on $M\times S$ by $(m,s)\sim(m',s')$ iff there is a $t\in S$ such that $t(s'm-sm')=0$.
    This is an equivalence relation and we denote the equivalence class of $(m,s)$ by $m/s$.
    The quotient $S^{-1}M$ forms an $A$-module by
    $$ m/s + m'/s' = (s'm+sm')/(ss'),\qquad a\cdot(m/s) = (am)/s . $$

\edefn

Note that localization at $S$ forms an additive functor $S^{-1}\colon\Mod_A\to\Mod_{S^{-1}A}$.
Indeed, $S^{-1}M$ is an $S^{-1}A$ module by
$$ a/s\cdot m/s' = (am)/(ss') . $$
And if $f\colon M\to N$ is an $A$-module morphism, then define
$$ S^{-1}f\colon S^{-1}M\to S^{-1}N,\qquad S^{-1}f(m/s) = f(m)/s . $$
It is not hard to see that this is functorial and additive.

There is the canonical morphism $\iota\colon A\to S^{-1}A$ and as such it lifts to a ``forgetful'' functor $\Mod_{S^{-1}A}\to\Mod_A$,
where for $M\in\Mod_{S^{-1}A}$ we view it as an $A$-module by $a\cdot M=\iota(a)\cdot M$.

\bprop

    The functor $S^{-1}\colon\Mod_A\to\Mod_{S^{-1}A}$ is left adjoint to the forgetful functor $\Mod_{S^{-1}A}\to\Mod_A$.

\eprop

\bproof

    We need to show that there is a canonical bijection
    $$ \hom_{S^{-1}A}(S^{-1}M,N) \cong \hom_A(M,N) , $$
    for all $A$-modules $M$ and $S^{-1}A$-modules $N$.
    For $f\colon M\to N$ map it to $\hat f\colon S^{-1}M\to N$ by $\hat f(m/s)=1/s\cdot f(m)$.
    And for $g\colon S^{-1}M\to N$ map it to $\tilde g\colon M\to N$ by $\tilde g(m)=g(m/1)$.
    \qed

\eproof

\bdefn

    An additive functor $F\colon\Mod_A\to\Mod_B$ is
    \benum
        \item {\emphcolor left exact} if it preserves left-exact sequences: if
        $$ 0 \to M \xtos f N \xtos g P $$
        is exact, then so is
        $$ 0 \to FM \xtos{Ff} FN \xtos{Fg} FP . $$
        \item {\emphcolor right exact} if it preserves right-exact sequences: if
        $$ M \xtos f N \xtos g P \to 0 $$
        is exact, then so is
        $$ FM \xtos{Ff} FN \xtos{Fg} FP\to 0. $$
        \item {\emphcolor exact} if it is both left and right exact.
    \eenum

\edefn

\bprop

    An additive functor is
    \benum
        \item left exact iff it preserves kernels iff it preserves finite limits iff it preserves the left side of a short exact sequence.
        \item right exact iff it preserves cokernels iff it preserves finite colimits iff it preserves the right side of a short exact sequence.
        \item exact iff it preserves finite limits and colimits iff it preserves short exact sequences iff it preserves all exact sequences.
    \eenum

\eprop

\bproof

    Left-exact sequences are kernel diagrams, and right-exact sequences are cokernel diagrams.
    So by definition, left exact functors are those which preserve kernels, and right exact functors are those which preserve cokernels.
    Since additive functors preserve finite biproducts, left/right exact functors preserve limits/colimits.

    Clearly left-exact functors preserve the left side of a short exact sequence (similarly for right-exact functors).
    To show the converse, if left-exact sequences can be quotiented to give an appropriate short exact sequence, similar for right-exact
    sequences.
    \qed

\eproof

\bcoro

    Left adjoints are right exact, and right adjoints are left exact.

\ecoro

\bproof

    Left adjoints preserve all colimits, and right adjoints all limits.
    \qed

\eproof

\bcoro

    $S^{-1}$ is an exact functor.

\ecoro

\bproof

    By the above corollary, it is right-exact.
    To show that it is exact, it is sufficient to show that it preserves injections.
    If $f\colon M\to N$ is injective then $m/s\in\ker S^{-1}f$ iff $f(m)/s=0/1$, so that $tf(m)=0$ for some $t\in S$.
    But then $tf(m)=f(tm)$ so $tm\in\ker\phi$, so $tm=0$, which means $m/s=0/1$ as desired.
    \qed

\eproof

In general if $U\colon\Mod_A\to\Mod_B$ is a restriction functor corresponding to a morphism $\iota\colon B\to A$, then it has a left adjoint
via the tensor product.
In particular, viewing $A$ as a $(A,B)$-module, then $A\otimes_BM$ has an $A$-module structure (given by $a(a'\otimes m)=(aa')\otimes m$).
$U$'s left adjoint is then $A\otimes_BM$.

Since left adjoints are unique up to natural isomorphism,
$$ S^{-1}M \cong S^{-1}A\otimes_AM . $$
Thus $S^{-1}A\otimes_A\bul$ is an exact functor, which means that $S^{-1}A$ is a {\emphcolor flat} $A$-module.
This can also be shown directly: $S^{-1}A$ is the filtered colimit of free $A$-modules, and is thus flat.

\bcoro

    $S^{-1}$ preserves arbitrary coproducts:
    $$ S^{-1}\bigoplus_\alpha M_\alpha \cong \bigoplus_\alpha S^{-1}M_\alpha $$
    naturally.

\ecoro

\bproof

    Left adjoints preserve arbitrary colimits.
    \qed

\eproof

\bnote

    While $S^{-1}$ is exact, it is not {\emphcolor continuous}: it does not preserve arbitrary limits (equivalently it does not preserve
    arbitrary products).
    That is, we do {\it not\/} have that
    $$ S^{-1}\prod_\alpha M_\alpha \cong \prod_\alpha S^{-1}M_\alpha . $$
    This is of course true when the index set is finite, since then it is just a biproduct.

\enote

\bcoro

    \benum
        \item If $N\subseteq M$ is a submodule, $S^{-1}N$ can be naturally embedded in $S^{-1}M$.
        \item There is a natural isomorphism $S^{-1}(M/N)\cong (S^{-1}M)/(S^{-1}N)$.
        \item For $f\colon M\to N$ of $A$-modules, we have natural isomorphisms
        $$ \ker S^{-1}f \cong S^{-1}\ker f,\qquad \im S^{-1}f \cong S^{-1}\im f,\qquad S^{-1}\coker f\cong\coker S^{-1}f . $$
        \item For every two submodules $M_1,M_2\subseteq M$ under the natural embeddings,
        $$ S^{-1}(M_1\cap M_2) = S^{-1}M_1\cap S^{-1}M_2 \subseteq S^{-1}M . $$
    \eenum

\ecoro

\bproof

    The first point is due to $S^{-1}$ being exact and thus preserving injections.
    The second point is because $S^{-1}$ preserves the short exact sequence $0\to N\to M\to M/N\to0$.
    The third point is clear as well, with only $\im$ not being immediately obvious.
    $\im f$ can be characterized as $\im f=\ker\coker f$, so it follows.

    For the last point, note that $M_1\cap M_2$ is the kernel of $M_1\oplus M_2\to M$ by $(x,y)\mapsto x-y$
    (via the diagonal embedding), and so this follows from $S^{-1}$'s exactness.
    \qed

\eproof

\bdefn

    For a prime ideal $\p\leq A$ and an $A$-module $M$, denote by $M_\p$ the localization of $M$ by $S=A-\p$.

\edefn

\blemm

    For an $A$-module $M$, consider the natural map
    $$ \phi\colon M\to\prod_\m M_\m,\qquad m\mapsto (m/1)_\m , $$
    where the product is over all maximal ideals of $A$.
    Then $\phi$ is injective.

\elemm

\bproof

    Assume there is an $0\neq x\in M$ such that $\phi(x)=0$.
    Then define
    $$ I = \ann_A(x) = \set{a\in A}[ax=0] . $$
    $I$ is a proper ideal of $A$, and thus is contained in a maximal ideal $\m$.
    Then for every $s\in A-\m$ we have that $s\notin I$ and so $sx\neq0$, and thus $m/1$ is not zero in $M_\m$, a contradiction.
    \qed

\eproof

\bcoro

    For an $A$-module $M$, the following are equivalent:
    \benum
        \item $M=0$,
        \item $S^{-1}M=0$ for every multiplicative set $S\subseteq A$,
        \item $M_\p=0$ for every prime ideal $\p\leq A$,
        \item $M_\m=0$ for every maximal ideal $\m\leq A$.
    \eenum

\ecoro

\bproof

    Clearly each point implies the next.
    By above $M$ embeds into $\prod_\m M_\m$, so if $M_\m=0$ for all $\m$ then $M=0$; the last point implies the first.
    \qed

\eproof

\bcoro

    For a morphism of $A$-modules $f\colon M\to N$, the following are equivalent:
    \benum
        \item $f$ is injective/surjective/an isomorphism,
        \item $S^{-1}f$ is injective/surjective/an isomorphism for every multiplicative $S\subseteq A$,
        \item $f_\p$ is injective/surjective/an isomorphism for every prime $\p\leq A$,
        \item $f_\m$ is injective/surjective/an isomorphism for every maximal $\m\leq A$.
    \eenum

\ecoro

\bproof

    By $S^{-1}$'s exactness, the first point implies the second, and clearly all the other points imply the next.
    To show the last implies the first, let $K=\ker f$ so that $0\to K\to M\xto fN$ is exact.
    Then $0\to K_\m\to M_\m\xto{f_\m}N_\m$ is exact, and so $K_\m=\ker f_\m=0$ for all $\m$.
    By above this means $K=0$ so that $f$ is injective.
    Similar for surjectivity.
    \qed

\eproof

\subsection{Finitely generated, integral, and finite algebras}

\bdefn

    For an $A$-algebra $B$, and $S\subseteq B$ denote by $A[S]$ the smallest $A$-subalgebra containing $S$ in $B$.

\edefn

Let $\phi\colon A\to B$ be the algebra map, then $A[S]$ is the subring of $B$ generated by $\phi(A)\cup S$.
Alternatively, for every $s\in S$ define an indeterminant $x_s$, and let $X=\set{x_s}_{s\in S}$ be the set of all these indeterminants.
Consider the polynomial algebra $A[X]$, and define $f\colon A[X]\to B$ by $f(x_s)=s$.
Then $A[S]$ is just the image of $f$.
Explicitly, for a function $\alpha\colon S\to\bZ_{\geq0}$ with finite support, define $S^\alpha=\prod_{\alpha(s)\neq0}s^{\alpha(s)}$.
Then
$$ A[S] = \set{\hbox{Finite sums }\sum_\alpha\phi(a_\alpha)S^\alpha}[a_\alpha\in A] . $$

Note that we can consider $B$ as an $A$-module.
Then for $S\subseteq B$, we can consider the {\it submodule\/} generated by $S$: $\gen S$.
Since $A[S]$ is itself a module, $\gen S\subseteq A[S]$ automatically, and we do not have equality in general.

\bdefn

    An $A$-algebra $B$ is {\emphcolor finitely generated} if there is a finite $S\subseteq B$ such that $B=A[S]$.
    $B$ is a {\emphcolor finite algebra} if it is finitely generated as an $A$-module.

\edefn

A finite $A$-algebra is also finitely generated as an $A$-algebra, but not vice versa in general.

\bexam

    Suppose $S=\set s$ is a singleton, then $A[s]$ is the set of polynomials $\sum_k\phi(a_k)s^k$.

\eexam

\bexam

    Clearly the polynomial algebra $A[x_1,\dots,x_n]$ is a finitely generated $A$-algebra.
    But it is not a finite $A$-algebra: we can reduce this to the case of fields, in which $x_i^k$ are all linearly independent and thus
    there is no finite basis.

\eexam

\blemm

    An $A$-algebra $B$ is fg iff it is the quotient of some finite polynomial algebra $A[x_1,\dots,x_n]$ (up to isomorphism of $A$-algebras).

\elemm

\bproof

    This is equivalent to saying $B$ is fg iff there is a surjection $A[x_1,\dots,x_n]\to B$.
    This is clearly true: if $B$ is fg then for $s_1,\dots,s_n$ generating $B$ map $x_i\mapsto s_i$ and extend by freeness.
    Conversely, quotients of fg algebras are fg.
    \qed

\eproof

\bdefn

    Let $B$ be an $A$-algebra.
    An element $b\in B$ is {\emphcolor integral} over $A$ if $b$ is the root of a monic polynomial over $A$.
    That is, there exist $a_1,\dots,a_n\in A$ such that $b^n+a_1b^{n_1}+\cdots+a_n\cdot1=0$.

    $b\in B$ is {\emphcolor algebraic} over $A$ if $b$ is the root of any polynomial over $A$.

\edefn

Clearly integral elements are algebraic.
For fields the converse holds as well, as we can divide a polynomial by its leading coefficient.

\bexam

    The set of integral elements in the $\bZ$-algebra $\bQ$ is $\bZ$.
    Clearly an integer is integral, as $n\in\bZ$ is the root of the monic $x-n$.
    Conversely if $b=p/q\in\bQ$ is integral with $p,q$ coprime then
    $$ b^n + a_1b^{n-1} + \cdots + a_n1 = 0,\qquad a_i\in\bZ . $$
    Then we have that
    $$ p^n + a_1p^{n-1}q + \cdots + a_nq^n = 0 $$
    and so $q$ must divide $p^n$, but since $p,q$ are coprime this can only happen if $q=1$ and $b\in\bZ$.

\eexam

\blemm

    If $M$ is a fg $A$-module, then for every $\phi\in\endo_A(M)$ there exist $a_1,\dots,a_n\in A$ such that
    $$ \phi^n + a_1\phi^{n-1} + \cdots + a_n = 0 $$
    in $\endo_A(M)$.

\elemm

\bproof

    The proof is by a generalization of Cayley-Hamilton.
    Let $n\geq1$ and consider $R=\bZ[X_{ij}]$, the polynomial ring over $n^2$ indeterminants $X_{ij}$, $1\leq i,j\leq n$.
    Let $X=(X_{ij})\in M_n(R)$, and define its {\emphcolor characteristic polynomial}: $\Phi_X(T)\in R[T]$ given by
    $$ \Phi_X(T) = \det(T\id - X) . $$
    By Cayley-Hamilton over $R$'s field of fractions, $\Phi_X(X)=0$.

    Now if $A$ is any ring and $Y\in M_n(A)$, there is a unique ring morphism $\phi\colon R\to A$ which maps $X$ to $Y$ (coordinate-wise).
    Indeed, such a map must map $X_{ij}$ to $Y_{ij}$ which uniquely determines $\phi$.
    Then $\phi$ lifts to the ring morphism $R[T]\to A[T]$ (map $\sum_na_nT^n$ to $\sum_n\phi(a_n)T^n$), which maps $\Phi_X(T)$ to $\Phi_Y(T)$ (which is also given by the determinant $\det(T\id-Y)$ since the determinant
    is functorial: $\det(\phi(B))=\phi(\det(B))$).
    Thus $\Phi_X(X)=0$ implies that $\Phi_Y(Y)=\phi(\Phi_X(\phi(X)))=\phi(\Phi_X(X))=0$, as desired.
    So Cayley-Hamilton applies to any commutative ring.

    Now for our lemma, let $m_1,\dots,m_n\in M$ be generators.
    Then for all $j$ we have
    $$ \phi(m_j) = \sum_i a_{ij}m_i ,\qquad a_{ij}\in A . $$
    We associate to $\phi$ the representative matrix $[\phi]=(a_{ij})\in M_n(A)$ (which is not unique, unlike over fields).
    This has the usual properties that $[\phi^n]=[\phi]^n$ and $[\phi+\psi]=[\phi]+[\psi]$ (meaning that the composition or sum of representatives is a representative matrix).
    Now Cayley-Hamilton gives us that
    $$ \sum_n b_n[\phi]^n = 0 $$
    for some $b_n\in A$, and thus
    $$ \bracks{\sum_nb_n\phi^n} = 0 , $$
    and since only the zero endomorphism has the zero representative, this means $\sum_nb_n\phi^n=0$ as desired.
    \qed

\eproof

\bprop

    For an $A$-algebra $B$ and $b\in B$, the following are equivalent:
    \benum
        \item $b$ is integral over $A$,
        \item $A[b]\subseteq B$ is a finite algebra (fg module) over $A$,
        \item $A[b]$ is contained in a finite algebra over $A$,
        \item There exists a faithful $A[b]$-module $M$ such that $M$ is finitely generated as an $A$-module.
    \eenum

\eprop

\bproof

    Clearly $(2)$ implies $(3)$ implies $(4)$: take $M=C$, and since $A[b]\subseteq M$ it is faithful (the algebra map is inclusion, which is
    injective).
    $(1)$ implies $(2)$ since if $b^n+a_1b^{n-1}+\cdots+a_n1=0$ then $b^n$ is in $\gen{1,\dots,b^{n-1}}\subseteq B$ and then by induction so is
    every higher power of $b$, so that $A[b]=\gen{1,\dots,b^{n-1}}$ is finite.

    $(4)$ implies $(1)$: we have a ring morphism $\phi\colon A[b]\to\endo_A(M)$, so that every $f\in A[b]$ gives us an endomorphism
    $\phi(f)\in\endo_A(M)$.
    In particular by the above lemma there are $a_1,\dots,a_n\in A$ such that
    $$ \rho(b)^n + a_1\rho(b)^{n-1} + \cdots + a_n = 0 \in \endo_A(M) . $$
    For $f(b)=b^n+a_1b^{n-1}+\cdots+a_n\in A[b]$, we see that $\rho(f(b))=0$.
    But $M$ is a faithful $A[b]$-module, so $f(b)=0$ and $b$ is integral as desired.
    \qed

\eproof

\blemm

    \benum
        \item If $B$ is a finite $A$-algebra, and $M$ is a fg $B$-module, then $M$ is a fg $A$-module.
        \item If $B$ is a finite $A$-algebra, and $C$ is a finite $B$-algebra, then $C$ is a finite $A$-algebra.
    \eenum

\elemm

\bproof

    Clearly the second point follows from the first.
    Now if $M$ is generated by $m_1,\dots,m_n$ over $B$, and $B$ is generated by $b_1,\dots,b_k$ over $A$ then $b_im_j$ generates $M$ over $A$.
    Indeed, for $m\in M$ if $m=\sum_ib'_im_i$, then $b'_i=\sum_ja_{ij}b_j$ so
    $$ m = \sum_{i,j}a_{ij}(b_jm_i) , $$
    as desired.
    \qed

\eproof

\bcoro

    Let $B$ be an $A$-algebra, and suppose that $b_1,\dots,b_n\in B$ are integral over $A$.
    Then $A[b_1,\dots,b_n]$ is a finite $A$-algebra.

\ecoro

\bproof

    By induction: for $n=1$ this is our above proposition.
    Then if $b_{n+1}$ is integral over $A$, it is integral over $A[b_1,\dots,b_n]$ so that $A[b_1,\dots,b_{n+1}]$ is a finite $A[b_1,\dots,b_n]$-algebra, and thus a finite $A$-algebra
    by transitivity which we just showed.
    \qed

\eproof

\bcoro

    Let $B$ be an $A$-algebra, then
    $$ B_0 = \set{b\in B}[\hbox{$b$ is integral over $A$}] $$
    is an $A$-subalgebra of $B$

\ecoro

\bproof

    Clearly $0,1\in B_0$.
    Now if $b_1,b_2\in B_0$ then $A[b_1+b_2],A[b_1b_2]\subseteq A[b_1,b_2]$ which is a finite $A$-algebra and thus $b_1+b_2,b_1b_2$ are integral over $A$; $B_0$ is a subring of $B$.
    And if $a\in A,b\in B_0$ then $A[ab]\subseteq A[b]$ is finite, so that $ab\in B_0$.
    \qed

\eproof

\bdefn

    \benum
        \item For an $A$-algebra $B$, we define $A$'s {\emphcolor integral closure} in $B$ to be the set of all integral elements over $A$ in $B$; which we showed above is an $A$-subalgebra.
        \item An $A$-algebra $B$ is {\emphcolor integral} if every element of $B$ is integral over $A$.
        \item A morphism of rings $f\colon A\to B$ is {\emphcolor integral} if $B$ is an integral $A$-algebra via $f$.
        \item For a morphism of rings $f\colon A\to B$, $f(A)$ is {\emphcolor integrally closed} in $B$ if $B_0=f(A)$ (i.e.\ no $b\in B-f(A)$ is integral over $A$).
        \item A domain $A$ is {\emphcolor normal} if it is integrally closed in its field of fractions.
    \eenum

\edefn

\bexam

    We saw that $\bZ$ is normal (its field of fractions is $\bQ$); the same proof shows that any UFD is normal.
    In particular $k[x_1,\dots,x_n]$ and $\bZ[x_1,\dots,x_n]$ are normal.

    Consider the $k$-subalgebra $A=k[x^2,x^3]\subseteq k[x]$.
    Then $x\notin A$ is integral over $A$ since it is a root of $t^6-x^6\in A[t]$, and $x=x^3/x^2$ is in its field of fractions.
    Thus $A$ is not normal.

\eexam

\bnote

    There is similarity between integral morphisms and algebraic extensions of fields.
    But while algebraically closed fields exist, there are no ``integrally closed ring'': every ring can be embedded non-integrally into another ring.

\enote

\bcoro

    An $A$-alegbra $B$ is finite over $A$ iff it is integral and finitely generated.

\ecoro

\bproof

    If $B$ if a finite $A$-algebra, then for every $b\in B$, $A[b]\subseteq B$ is finite so $b$ is integral over $A$; $B$ is integral and finitely generated.
    Conversely if $B$ is integral and finitely generated, suppose $B=A[b_1,\dots,b_n]$; since each $b_i$ is integral we have that $B$ is finite.
    \qed

\eproof

\bcoro

    Let $B$ be an integral $A$-algebra, and $C$ be a $B$-algebra.
    \benum
        \item If $c\in C$ is integral over $B$, then it is integral over $A$.
        \item If $C$ is integral over $B$, then it is integral over $A$.
    \eenum

\ecoro

\bproof

    If $c$ is integral over $B$, then
    $$ c^n + b_1c^{n-1} + \cdots + b_n = 0 $$
    for some $b_i\in B$, which are integral over $A$.
    Then $B'=A[b_1,\dots,b_n]\subseteq B$ is a finite $A$-algebra, and $c$ is integral over $B'$ and so $B'[c]$ is a finite $B'$-algebra.
    Since $B'$ is a finite $A$-algebra, by transitivity $B'[c]$ is a finite $A$-algebra as well containing $A[c]$, so $c$ is integral over $A$.

    The second point follows from the first.
    \qed

\eproof

\bcoro

    Let $B$ be an $A$-algebra, and $B_0$ be the integral closure of $A$ inside $B$.
    Then $B_0$ is integrally closed in $B$.

\ecoro

\bproof

    If $b\in B$ is integral over $B_0$, then by the above corollary, since $B_0$ is an integral $A$-algebra, $b$ is integral over $A$ and thus $b\in B_0$.
    \qed

\eproof

Let $B$ be an $A$-algebra under $f\colon A\to B$.
\benum
    \item For an ideal $J\leq B$, define $J_A=f^{-1}(J)\leq A$.
    Then $f$ induces an injective homomorphism $A/J_A\to B/J$, so that $B/J$ is an $A/J_A$-algebra.
    \item For a multiplicative set $S\subseteq A$, $f(S)\subseteq B$ is multiplicative.
    Then we have a natural morphism $S^{-1}A\to f(S)^{-1}B$ mapping $a/s$ to $f(a)/f(s)$.
    Thus $f(S)^{-1}B$ is an $S^{-1}A$-algebra.
    \item Viewing $B$ as an $A$-module, we have already defined $S^{-1}B$, and this clearly coincides with $f(S)^{-1}B$.
    Thus $S^{-1}B$ is naturally an $S^{-1}A$-algebra.
\eenum

\bprop

    Let $B$ be an integral $A$-algebra.
    \benum
        \item Let $J\leq B$ be an ideal, then $B/J$ is an integral $A/J_A$-algebra.
        \item Let $S\subseteq B$ be multiplicative, then $S^{-1}B$ is an integral $S^{-1}A$-algebra.
    \eenum

\eprop

\bproof

    For $\bar x\in B/J$, lift it to $x\in B$, which is integral over $A$.
    Thus $x^n+a_1x^{n-1}+\cdots+a_n=0$ for $a_i\in A$, and then
    $$ \bar x^n + \bar a_1\bar x^{n-1} + \cdots + \bar a_n = 0 , $$
    where $\bar a_i=a_i+J_A\in A/J_A$, so that $\bar x$ is integral over $A/J_A$.

    An element of $S^{-1}B$ is of the form $\tilde x=x/f(s)$ with $x\in B,s\in S$.
    By assumption, $x^n+a_1x^{n-1}+\cdots+a_n=0$ and thus
    $$ \parens{\frac x{f(s)}}^n + \frac{a_1}s\parens{\frac x{f(s)}}^{n-1} + \cdots + \frac{a_n}{s^n} = \frac{x^n+a_1x^{n-1}+\cdots+a_0}{f(s)^n} = 0 , $$
    so that $\tilde x$ is integral over $S^{-1}A$ as desired.
    \qed

\eproof

\bprop

    Let $A\subseteq B$ be integral domains so that $B$ is integral over $A$.
    Then $A$ is a field iff $B$ is a field.

\eprop

\bproof

    If $A$ is a field, then let $0\neq b\in B$.
    It is integral over $A$ so $b^n+a_1b^{n-1}+\cdots+a_n=0$ for $a_i\in A$, and assume $n$ is minimal.
    Then
    $$ b(b^{n-1} + a_1b^{n-2} + \cdots + a_{n-1}) = -a_n . $$
    Since by minimality $b^{n-1}+a_1b^{n-1}+\cdots+a_{n-1}\neq0$, we have $a_n\neq0$.
    Thus $-a_n$ is a unit, and thus $b$ must be a unit as well (if $xy=u$ is a unit, then $x$ is a unit since $x(yu^{-1})=1$).
    So $B$ is a field.

    Conversely if $B$ is a field, then let $0\neq x\in A$.
    Then $x^{-1}\in B$ since it is a field.
    Since $B$ is integral over $A$, there are $a_i\in A$ such that
    $$ x^{-n} + a_1x^{-(n-1)} + \cdots + a_n = 0 , $$
    and multiplying by $x^{n-1}$ we get
    $$ x^{-1} + a_1 + \cdots + a_nx^{n-1} , $$
    so that $x^{-1}$ is the sum of elements from $A$, and thus is in $A$.
    \qed

\eproof

\bcoro

    Let $B$ be an integral $A$-algebra, and $\q\leq B$ a prime ideal.
    Let $\p=\q_A\leq A$ be its preimage, then $\q$ is maximal iff $\p$ is maximal.

\ecoro

\bproof

    Both $\p,\q$ are prime so that $B/\q$ and $A/\p$ are integral domains.
    As we showed, $B/\q$ is integral over $A/\p$.
    Thus $B/\q$ is a field iff $A/\p$ is a field; so $\q$ is maximal iff $\p$ is maximal.
    \qed

\eproof

\bcoro

    Let $B$ be an integral $A$-algebra, $\q\leq\q'\leq B$ be prime ideals such that $\q_A=\q'_A$.
    Then $\q=\q'$.

\ecoro

\bproof

    Let $\p=\q_A=\q'_A$; it is prime over $A$.
    Consider the following commutative square

    \commsq{A&B\cr A_\p&B_\p}
        {f}{f_\p}{\iota_A}{\iota_B}

    which commutes since the structure morphism $A\to S^{-1}A$ is a natural transformation (literally what the commutative square says).

    Since $f^{-1}\q=\p$ we see that $\q\cap f(A-\p)=\emptyset$ and similarly for $\q'$.
    Thus $\q B_\p\subseteq\q' B_\p\subseteq B_\p$ are prime ideals (since they don't intersect the localizing set $f(A-\p)$).

    We are going to show that $\q B_\p\subseteq B_\p$ is a maximal ideal so that $\q B_\p=\q' B_\p$.
    And then since $\q\mapsto\q B_\p$ is bijective with inverse $\iota_B^{-1}$, we will have that $\q=\q'$.

    Since $B$ is integral over $A$, $B_\p$ is integral over $A_\p$.
    Thus, by the above corollary, it is sufficient to show that $\tilde\p=f_\p^{-1}(\q B_\p)\subseteq A_\p$ is maximal.
    By assumption,
    $$ \iota_A^{-1}(\tilde\p) = \iota_A^{-1}(f_\p^{-1}(\q B_\p)) = f^{-1}\circ\iota_B^{-1}(\q B_\p) = f^{-1}(\q) = \p . $$
    Thus we have that $\tilde\p=\p A_\p$ (since $\iota_A^{-1}$ is bijective) which is maximal, as desired.
    \qed

\eproof

The main result we want is the following.

\bthrm

    Let $f\colon A\to B$ be an injective integral morphism.
    Then for every prime $\p\leq A$, there exists a prime ideal $\q\leq B$ such that $\p=\q_A$.

\ethrm

\bproof

    Since $B_\p=f(A-\p)^{-1}B$, and $0\notin f(A-\p)$ because $f$ is injective, $B_\p\neq0$.
    As before, $B_\p$ is integral over $A_\p$ and we have the same commutative square

    \commsq{A&B\cr A_\p&B_\p}
        {f}{f_\p}{\iota_A}{\iota_B}

    Let $\m\subseteq B_\p$ be maximal, so that $f_\p^{-1}(\m)\subseteq A_\p$ is maximal (as we showed $\q_A$ is maximal iff $\q$ is).
    Thus $f_\p^{-1}(\m)=\p A_\p$.
    Let $\q=\iota_B^{-1}(\m)$, which is prime in $B$, and
    $$ f^{-1}(\q) = f^{-1}\circ\iota_B^{-1}(\m) = \iota_A^{-1}\circ f_\p^{-1}(\m) = \iota_A^{-1}(\p A_\p) = \p , $$
    as desired.
    \qed

\eproof

\bnote

    Equivalently, this theorem says that if $f\colon A\to B$ is integral injective, then $f^*\colon\spec(B)\to\spec(A)$ is surjective.

    Note that injectivity is necessary here.
    This is because $\pi\colon A\to A/I$ is integral for any ideal $I\leq A$, but $\pi^*\colon\spec(A/I)\to\spec(A)$ is not surjective: its image is $V(I)$, the set of prime ideals containing $I$.

\enote

\bcoro

    Let $B$ be an integral $A$-algebra, then for every prime $\q\leq B$ and every prime ideal $\p'\supseteq\p=\q_A$, there exists a prime ideal $\q'\supseteq\q$ of $B$ such that $\q'_A=\p'$.

\ecoro

\bproof

    Consider the commutative square

    \commsq{A&B\cr A/\p&B/\q}
        {f}{\bar f}{\pi_A}{\pi_B} 

    Then $\p'\supseteq\p$ gives rise to a prime ideal $\bar\p'=\p'/\p\subseteq A/\p$.
    Since $B/\q$ is an integral $A/\p$-algebra, there is a prime ideal $\bar\q'\subseteq B/\q$ such that $\bar f^{-1}(\bar\q')=\bar\p'$.
    Then $\q'=\pi^{-1}_B(\bar\q')$ is a prime ideal of $B$ containing $\q$ and
    $$ f^{-1}(\q') = f^{-1}\circ\pi^{-1}_B(\bar\q') = \pi^{-1}_A\circ\bar f^{-1}(\bar\q') = \pi^{-1}_A(\bar\p') = \p' , $$
    as desired.
    \qed

\eproof

\bprop

    Let $B$ be an $A$-algebra, $C\subseteq B$ be the integral closure of $A$ inside of $B$.
    Then for any multiplicative $S\subseteq A$, $S^{-1}C$ is the integral closure of $S^{-1}A$ inside of $S^{-1}B$.

\eprop

\bproof

    Since $C$ is integral over $A$, $S^{-1}C$ is integral over $S^{-1}C$.
    Now if $b/s$ is integral over $S^{-1}A$, then $b/1=s\cdot b/s$ is integral over $S^{-1}A$ as well, and thus
    $$ \parens{\frac b1}^n + \frac{a_1}{s_1}\parens{\frac b1}^{n-1} + \cdots + \frac{a_n}{s_n} = 0 $$
    for $a_i/s_i\in S^{-1}A$.
    Let $t=s_1\cdots s_n$ and $u_i=t/s_i\in S$, and multiplying this equation by $t^n$ gives us
    $$ \parens{\frac{tb}1}^n + a_1u_1\parens{\frac{bt}1}^{n-1} + \cdots + a_nu_nt^{n-1} = 0 . $$
    Thus there is a $v\in S$ such that
    $$ v((bt)^n + a_1u_1(bt)^{n-1} + a_2u_2t(bt)^{n-2} + \cdots + a_nu_nt^{n-1}) = 0 . $$
    Then $btv$ is integral over $A$ (multiply by $v^{n-1}$), so that $btv\in C$.
    Then $b/s=(btv)/(stv)\in S^{-1}C$, as desired.
    \qed

\eproof

\bcoro

    Let $A$ be a domain, then the following are equivalent:
    \benum
        \item $A$ is normal,
        \item $S^{-1}A$ is normal for every multiplicative $S\subseteq A$ not containing $0$,
        \item $A_\p$ is normal for every prime $\p$,
        \item $A_\m$ is normal for every maximal $\m$.
    \eenum

\ecoro

\bproof

    Let $K$ be $A$'s field of fractions, and let $C\subseteq K$ be $A$'s integral closure in $K$.
    Then $S^{-1}C$ is $S^{-1}A$'s integral closure in $S^{-1}K=K$ for all multiplicative $S$ not containing $0$.
    If $A$ is normal, then $C=A$ so that $S^{-1}A$ is integrally closed in $K$, and thus normal.
    So (1) implies (2), which implies (3) which implies (4).

    Now if $A_\m$ is normal, then $A_\m\to C_\m$ is an isomorphism for all $\m$.
    We showed that this implies $A\to C$ is an isomorphism, so that $A$ is normal.
    \qed

\eproof

\subsection{Noetherian rings and modules}

\blemm

    Let $\Sigma$ be a poset, then the following conditions are equivalent:
    \benum
        \item every nonempty subset of $\Sigma$ has a maximal element,
        \item every increasing chain $x_1\leq x_2\leq\cdots$ in $\Sigma$ eventually stabilizes: at some point $x_n=x_{n+1}$ (equivalently, there are no infinite strictly
        increasing chains).
    \eenum
    A poset satisfying these equivalent conditions is said to have the {\emphcolor ascending chain condition} ({\emphcolor ACC}).

\elemm

\bproof

    The first implies the second as we can consider the subset $\set{x_i}$.
    The second implies the first since if $\Sigma'\subseteq\Sigma$ doesn't have a maximal element we can form a strictly increasing chain inductively.
    \qed

\eproof

\bdefn

    An $A$-module $M$ is {\emphcolor Noetherian} if its poset of $A$-submodules has the ACC.
    A ring $A$ is {\emphcolor Noetherian} if it is Noetherian as a module over itself (equivalently its ideals satisfy the ACC).

\edefn

\bexam

    Any field is a Noetherian ring, since it only has two ideals.

    $\bQ$ is not a Noetherian $\bZ$-module, since $\frac1{n!}\bZ$ does not stabilize (the reason for the factorial is so that it forms an increasing chain).

    The ring $k[x_1,x_2,\dots]$ with infinitely many indeterminants is not Noetherian, since $(x_1)\subsetneq(x_1,x_2)\subsetneq(x_1,x_2,x_3)\subsetneq\cdots$ forms a strictly increasing chain.

\eexam

\bprop

    An $A$-module is Noetherian iff every submodule is finitely generated.
    Thus a ring is Noetherian iff every ideal is finitely generated.

\eprop

\bproof

    If $M$ is Noetherian and $N\subseteq M$ is a submodule, suppose it is not finitely generated.
    Then inductively we will form an infinitely increasing sequence of submodules.
    Let $x_1\in N$ arbitrary, then since $N$ is not finitely generated $x_1,\dots,x_n$ does not generate $N$ so that there is an $x_{n+1}\in N-\gen{x_1,\dots,x_n}$.
    Then $N_n=\gen{x_1,\dots,x_n}$ is a strictly increasing infinite chain of submodules of $M$, a contradiction.

    Conversely if every submodule of $M$ is finitely generated, then $M$ is Noetherian.
    Otherwise, let $M_1\subsetneq M_2\subseteq\cdots$ be an increasing chain of submodules.
    Let $N=\bigcup_nM_n$ be their union, which is also a submodule of $M$, and thus is finitely generated.
    Suppose $x_1,\dots,x_n\in N$ generate $N$, and since $M_i$ form a chain there is an $m$ such that $x_1,\dots,x_n\in M_m$ and thus $M_m=N$ so that $M_i$ eventually stabilizes to $M_m$.
    \qed

\eproof

\bcoro

    Every PID is Noetherian.

\ecoro

\blemm

    Let $0\to M'\xto fM\xto gM''\to0$ be a short exact sequence of $A$-modules.
    Then $M$ is Noetherian iff $M',M''$ are.

    Equivalently, if $M$ is an $A$-module and $N\subseteq M$ a submodule, $M$ is Noetherian iff $N$ and $M/N$ are.

\elemm

\bproof

    If $M$ is Notherian, let $M'_n\subseteq M',M''_n\subseteq M''$ be increasing sequences of submodules.
    Then $f(M'_n)$ and $g^{-1}(M''_n)$ are submodules of $M$ and thus stabilize: $f(M'_n)=f(M'_{n+1})$ and $g^{-1}(M'_n)=g^{-1}(M'_{n+1})$ eventually.
    By the injectivity of $f$ and surjectivity of $g$, this means $M'_n$ and $M''_n$ both eventually stabilize.

    Conversely if $M',M''$ are Noetherian, let $M_n\subseteq M$ be an increasing sequence of submodules.
    Then $f^{-1}(M_n)=M_n\cap M'$ and $g(M_n)\subseteq M''_n$ are increasing sequences of submodules of $M',M''$ respectively.
    Thus eventually they stabilize: $f^{-1}(M_n)=f^{-1}(M_{n+1})$ and $g(M_n)=g(M_{n+1})$ eventually.
    We claim that this means $M_n=M_{n+1}$.

    Indeed, for $x\in M_{n+1}$ consider $g(x)\in g(M_{n+1})$ so there is a $y\in M_n$ such that $g(x)=g(y)$.
    Then $x-y\in\ker g=\im f$ so that $x-y=f(z)$ for $z\in M'$, and $x-y\in M_{n+1}$ so that $z\in f^{-1}(M_{n+1})$ and thus $z\in f^{-1}(M_n)$.
    This means that $x-y\in M_n$ and thus $x\in M_n$ as desired.
    \qed

\eproof

\bcoro

    \benum
        \item The finite direct sums of Noetherian modules is Noetherian.
        \item If $A$ is a Noetherian ring and $M$ is a fg $A$-module, then $M$ is a Notherian module.
        \item If $A$ is a Notherian ring and $B$ a finite $A$-algebra, then $B$ is a Noetherian ring.
        In particular, $A/I$ is Notherian for any ideal $I\leq A$.
    \eenum

\ecoro

\bproof

    For the first point it is sufficient to show this for binary direct sums.
    If $M,N$ are Noetherian, then $0\to M\to M\oplus N\to N\to0$ is a short exact sequence and thus $M\oplus N$ is Notherian by above.

    Since $M$ is fg, $M\cong A^n/N$ for some submodule $N\leq A^n$ (consider the surjection $A^n\to M$ mapping to generators).
    By the first point, $A^n$ is Noetherian and thus $M$ is as well.

    $B$ is a fg $A$-module, and thus a Noetherian $A$-module.
    Since ideals of $B$ are $A$-submodules, the ACC on $A$-submodules implies ACC on $B$'s ideals.
    \qed

\eproof

\blemm

    Let $A$ be a Noetherian ring, then all of its localizations are Noetherian rings as well.

\elemm

\bproof

    Let $S\subseteq A$ be multiplicative, and consider an increasing sequence of ideals $\tilde I_1\subseteq\tilde I_2\subseteq\cdots$ in $S^{-1}A$.
    Then $I_n=\iota^{-1}(\tilde I_n)$ forms an increasing sequence of ideals of $A$, and thus stabilizes.
    Since $\tilde I_n=S^{-1}I_n$, we get that $\tilde I_n$ stabilizes.
    \qed

\eproof

\bthrm[title=Hilbert basis theorem]

    If $A$ is Noetherian, then $A[x]$ is a Noetherian ring.

\ethrm

\bproof

    Let $I\subseteq A[x]$ be an ideal, we wish to show it is finitely generated.
    Let us adopt the notation that for $f\in A[x]$, $a_f\in A$ denotes the leading term of $f$ (and $a_f=0$ only when $f=0$).

    We claim that $J=\set{a_f}[f\in I]$ is an ideal in $A$.
    If $f\in I$ and $b\in A$ then either $ba_f=0$ or $ba_f=a_{bf}$; in both cases $ba_f\in J$.
    Now if $f,g\neq0$ are in $I$ then $a_f+a_g$ is either zero or the leading term of $x^{\deg g}f+x^{\deg f}g\in J$ and thus $a_f+a_g\in J$.

    Since $J$ is Noetherian, there are $f_1,\dots,f_n\in I$ such that $J=(a_{f_1},\dots,a_{f_n})$.
    Let $d_i=\deg f_i$, and $d$ be the maximum $d_i$.
    Consider the fg $A$-submodule
    $$ M = \sum_{i=0}^{d-1}Ax^i \subseteq A[x] . $$
    Then $M\cap I$ is an $A$-submodule of $M$, which is Noetherian (since $M$ is fg and $A$ is Noetherian so $M$ is Noetherian, and thus its submodules are Noetherian).

    Now define $I'=(f_1,\dots,f_n)\subseteq I$.
    We claim that
    $$ I = I' + (M\cap I)\hbox{ as $A$-submodules} . $$
    Let $f\in I$, then we claim there is an $f'\in I'$ such that $\deg(f-f')<d$ and thus $f-f'\in M\cap I$.
    If $\deg f<d$ then we are done, so assume $\deg f\geq d$.
    By induction, it is sufficient to show that there is an $f'\in I'$ such that $\deg(f-f')<\deg f$.

    Now, $a_f\in J=(a_{f_1},\dots,a_{f_n})$ so that $a_f=\sum_ib_ia_{f_i}$ for $b_i\in A$.
    Then define
    $$ f' = \sum_{i=1}^n b_ix^{(\deg f)-d_i}f_i \in I' . $$
    By assumption, $\deg f\geq d\geq d_i$ so that $\deg(x^{(\deg f)-d_i}f_i)=d_i$, and so $\deg f'\leq\deg f$.
    Moreover, the leading coefficient of $f'$ is
    $$ a_{f'} = \sum_{i=1}^n b_ia_{f_i} = a_f, $$
    so that $\deg(f-f')<\deg f$ as desired.

    Now since $I=I'+(M\cap I)$, $I'$ is a fg ideal and $M\cap I$ is a fg $A$-module (and thus fg by $A[x]$), we have that $I$ is a fg ideal, as desired.
    \qed

\eproof

\bcoro

    \benum
        \item If $A$ is a Noetherian ring, so is $A[x_1,\dots,x_n]$.
        \item If $A$ is a Noetherian ring, and $B$ is a fg $A$-algebra, $B$ is Noetherian.
        \item Any fg algebra over a field is Noetherian.
        \item Any fg algebra over a PID (in particular $\bZ$) is Noetherian.
    \eenum

\ecoro

\bproof

    The first point follows by induction and Hilbert's basis theorem.
    The second point is because if $B$ is a fg $A$-algebra, it is isomorphic to $A[x_1,\dots,x_n]/I$ for some ideal $I\leq A[x_1,\dots,x_n]$, and is thus Noetherian.
    The last two points are due to the second (fields and PIDs are Noetherian).
    \qed

\eproof

We now utilize these results to reprove a lemma we showed above.

\blemm

    Let $A$ be a ring and $T\in M_n(A)$ a matrix.
    Then there are $a_1,\dots,a_n\in A$ such that
    $$ T^n + a_1T^{n-1} + \cdots + a_m = 0 . $$

\elemm

\bproof

    Denote $T=(a_{ij})$, and consider the subalgebra $A'=\bZ[\set{a_{ij}}]\subseteq A$.
    This is a fg $\bZ$-algebra and thus Noetherian.
    Since $T\in M_n(A')$, we can assume from the getgo that $A$ is Noetherian.

    Since $M_n(A)\cong A^{n^2}$ as $A$-modules, $M_n(A)$ is a Noetherian $A$-module.
    Thus the increasing sequence of $A$-modules
    $$ N_m = \gen{1,T,\dots,T^m} \subseteq M_n(A) $$
    stabilizes, so eventually $N_n=N_{n-1}$ so that $T^n\in N_{n-1}$ and we get the desired result.
    \qed

\eproof

\blemm

    If $M$ is a fg $A$-module, then for $\phi\in\endo_A(M)$, there are $a_1,\dots,a_n\in A$ such that
    $$ \phi^n + a_1\phi^{n-1} + \cdots + a_n = 0 . $$

\elemm

\bproof

    When $M=A^n$ there is a natural isomorphism of $A$-algebras $M_n(A)\cong\endo_A(M)$, so the result follow from the above lemma.
    Otherwise, consider a surjective map $\pi\colon A^n\to M$, then we claim there is a $\tilde\phi\colon A^n\to A^n$ such that the below diagram commutes

    \commsq{A^n&A^n\cr M&M}
        {\tilde\phi}{\phi}{\pi}{\pi}

    Meaning $\pi\circ\tilde\phi=\phi\circ\pi$.
    For every $i$ choose a preimage $V_i\in\pi^{-1}(\phi(\pi(e_i)))\in A^n$ and define
    $$ \tilde\phi(x_1,\dots,x_n) = \sum_{i=1}^n x_iv_i . $$
    This satisfies the condition.

    Now there is a monic polynomial $f\in A[x]$ such that $f(\tilde\phi)=0$ by the $M=A^n$ case.
    Then
    $$ f(\phi)\circ\pi = \pi\circ f(\tilde\phi) = 0, $$
    which implies $f(\phi)=0$ because $\pi$ is surjective.
    \qed

\eproof

\subsection{Simple modules and length}

\bdefn

    A poset $\Sigma$ has the {\emphcolor descending chain condition} ({\emphcolor DCC}) if it satisfies the dual of the ACC (equivalently, $\Sigma^\op$ has ACC).
    This means that every subset has a minimal element, or every descending chain stabilizes.

\edefn

\bdefn

    An $A$-module is {\emphcolor Artinian} if it has DCC on its submodules.
    A ring $A$ is {\emphcolor Artinian} if it is Artinian as a module over itself, equivalently it has DCC on its ideals.

\edefn

Note that the short exact sequence result on Noetherian rings holds for Artinian rings as well.
Thus finite direct sums of Artinian modules are Artinian, fg modules over Artinian rings are Artinian, and thus finite algebras over Artinian rings are Artinian.

Note that it is not the case that a module is Artinian iff its submodules are all fg.
That is, Artinian is not equivalent to Noetherian (and for modules they are distinct conditions, for rings we will see that Artinian implies Noetherian).

\bexam

    $\bZ$ is a Noetherian but not an Artinian ring.
    It is not Artinian because $\bZ\supsetneq(n)\supsetneq(n^2)\supsetneq(n^3)\supsetneq\cdots$ is an infinite decreasing sequence of ideals.

    Similarly $k[x]$ is Noetherian but not Artinian: $(x^k)$ forms a descending sequence of ideals.

\eexam

\bdefn

    A nonzero module $M$ is {\emphcolor simple} if it has no non-trivial submodules.

\edefn

\blemm

    An $A$-module $M$ is simple iff it is isomorphic (as $A$-modules) to $A/\m$ for a maximal ideal $\m\leq A$.

\elemm

\bproof

    If $M$ is simple, then for $0\neq m\in M$ the submodule $\gen m\subseteq M$ is isomorphic to $A/\ann(m)$ (the map $a\mapsto am$ has kernel $\ann(m)$).
    Since $M$ is simple, $\gen m=M$ and so $A/\ann(m)\cong M$.
    There is then a bijection between submodules of $M$ and ideals of $A$ containing $\ann(m)$.
    Since $M$ is simple, there cannot be any nontrivial such ideals, so $\ann(m)$ is maximal.

    And if $M\cong A/\m$ then again there is a bijection between submodules of $M$ and ideals containing $\m$.
    So $M$ is simple, as $\m$ is maximal.
    \qed

\eproof

\bdefn

    A {\emphcolor composition series} of $M$ is a chain of submodules
    $$ 0 = M_0 \subsetneq M_1 \subsetneq M_2 \subsetneq \cdots \subsetneq M_n = M $$
    which is maximal.
    This is equivalent to $M_{i+1}/M_i$ being simple for all $i$ (no modules between $M_i$ and $M_{i+1}$).
    The {\emphcolor length} of such a series is $n$ (not $n+1$).

    Denote by $\ell_A(M)$ the minimal length of a composition series of $M$ (or infinity if no composition series exists).

\edefn

\blemm

    If $M$ has finite length $\ell(M)=n$, then all of its composition series have the same length $n$, and every chain of submodules can be extended to a composition series.

\elemm

\bproof

    First we claim that if $N\subsetneq M$ is a submodule, then $\ell(N)<\ell(M)$.
    Indeed, if $\set{M_i}$ is a composition series for $N$, then $N_i=M_i\cap N$ is a sequence of submodules of $N$.
    Then $N_i=N_{i+1}\cap M_i$ so
    $$ N_{i+1}/N_i = N_{i+1}/(N_{i+1}\cap M_i) \cong (N_{i+1}+M_i)/M_i \subseteq M_{i+1}/M_i $$
    which is either simple or zero.
    Thus removing repeated values, we have formed a composition series for $N$ of shorter length, so $\ell(N)\leq\ell(M)$.

    Now if $\ell(N)=\ell(M)$ then $N_{i+1}+M_i=M_{i+1}$ for all $i$, and thus we claim $N_i=M_i$ for all $i$ so that $N=M$.
    This is by induction, since $M_0=N_0=0$, and $N_{i+1}=N_{i+1}+N_i=N_{i+1}+M_i=M_{i+1}$, as desired.

    Now if $\set{M_i}_0^n$ is any chain in $M$, then
    $$ \ell(M) = \ell(M_n) > \ell(M_{n-1}) > \cdots > \ell(M_0) = 0, $$
    so $\ell(M)\geq n$.

    Now if $\set{M_i}$ is any composition series of $M$ of length $n$, then $\ell(M)\geq n$ by what we just showed.
    But by definition $\ell(M)\leq n$ so that $\ell(M)=n$ and thus all composition series have length $\ell(M)$.

    And for any general chain of submodules of $M$, if it is not a composition series then new terms can be inserted.
    This process can be repeated, and must terminate (since once it reaches a length of $\ell(M)$ it must be a composition series).
    \qed

\eproof

\bnote

    Note that we can prove something stronger, the {\emphcolor Jordan-Holder theorem}: for any two composition series $M_i,N_j$ there is a permutation $\sigma$ between $i$ and $j$
    such that $M_{i+1}/M_i\cong N_{\sigma(i)+1}/N_{\sigma(i)}$.
    A proof of this can be found in my representation theory homework somewhere.

\enote

\blemm

    An $A$-module $M$ has finite length iff it is both Noetherian and Artinian.

\elemm

\bproof

    If $M$ has finite length, then any increasing/decreasing sequence has length bound by $\ell(M)$ so that $M$ is both Noetherian and Artinian.

    Conversley if $M$ is Artinian and Noetherian, then consider the set of submodules $N\subseteq M$ of finite length.
    It is nonempty since it contains $0$, and because $M$ is Noetherian contains a maximal element $M'$.
    If $M=M'$ then we are done.
    Otherwise, the set of submodules $M'\subsetneq N\subseteq M$ is nonempty (containing $M$) and has a minimal element $N$ since $M$ is Artinian.
    But then $N/M'$ is simple, and so composing a composition series of $M'$ with $\subsetneq N$ forms a composition series of $N$, so $N$ has finite length, contradicting $M'$'s maximality.
    \qed

\eproof

\bprop

    Let $M$ be fg over a Notherian ring $A$.
    \benum
        \item If $M\neq0$, then $M$ has a submodules $N\subseteq M$ such that $N\cong A/\p$ for some prime $\p\leq A$.
        \item There exists a chain $0=M_0\subsetneq M_1\subsetneq\cdots\subsetneq M_n=M$ of submodules such that for each $i$, there is an isomorphism of $A$-modules
        $M_{i+1}/M_i\cong A/\p_i$ for prime ideals $\p_i\leq A$.
        \item The chain in the previous point is not necessarily unique, but
        \benum
            \item for a prime ideal $\p\leq A$ we have $\ann M\subseteq\p$ iff $\p_i\subseteq\p$ for some $i$.
            In particular, $\ann M\subseteq\p_i$ for all $i$, and the minimal ideals between $\p_1,\dots,\p_n$ are precisely minimal prime ideals $\ann(M)\subseteq\p$.
            \item for every minimal prime ideal $\ann(M)\subseteq\p$, the $\A_\p$ module $M_\p$ has finite length equal to the number of times $\p$ occurs in $\set{\p_1,\dots,\p_n}$.
        \eenum
    \eenum

\eprop

\bproof

    Let $\Theta$ be the set of all ideal $I\leq A$ of the form $I=\ann(m)$ for $0\neq m\in M$.
    Since $A$ is Noetherian, there is an $m$ such that $\ann(m)$ is maximal.
    Since $\gen m\subseteq M$ is isomorphic to $A/\ann(m)$, it is sufficient to show $\ann(m)$ is prime.
    So suppose $a,b\notin\ann(m)$, then $bm\neq0$ and $\ann(m)\subseteq\ann(bm)$.
    Thus by maximality $\ann(m)=\ann(bm)$ and then $a\notin\ann(bm)$ so $abm\neq0$ meaning $ab\notin\ann(m)$, showing it is prime.
    This proves the first point.

    Let $\Sigma$ be the set of all submodules of $M$ which possess such a chain.
    It is nonempty since the zero module is in it.
    Since $M$ is Noetherian, let $M'$ be a maximal element of $\Sigma$.

    Now consider $M''=M/M'$; if $M''=0$ we are finished.
    Otherwise by the first point it has a submodule $N''=N'/M'$ isomorphic to $A/\p$.
    But then composing the chain of $M'$ with $\subsetneq N'$ gives such a chain, contradicting $M'$'s maximality.

    For the final point, notice that for $M'\subsetneq M$,
    $$ \ann(M')\cdot\ann(M/M') \subseteq \ann(M) \subseteq \ann(M')\cap\ann(M/M') . $$
    The second inclusion is clear, the first is because for $a\in\ann(M')$ and $b\in\ann(M/M')$ and $m\in M$ then $bm\in M'$ and $abm\in aM'=0$.
    Thus for a prime ideal $\p\leq A$, $\ann(M)\subseteq\p$ iff $\ann(M')\subseteq\p$ or $\ann(M/M')\subseteq\p$.
    Thus by induction $\ann(M)\subseteq\p$ iff $\ann(M_{i+1}/M_i)\subseteq\p$ for some $i$.
    Since $\ann(A/\p_i)=\p_i$, the first point follows.

    For the second point, the chain $\set{M_i}$ gives the chain $\set{S^{-1}M_i}$ and
    $$ S^{-1}M_{i+1}/S^{-1}M_i \cong S^{-1}(M_{i+1}/M_i) \cong S^{-1}(A/\p_i) \cong S^{-1}A/S^{-1}\p_i . $$
    Moreover $S^{-1}\p_i=S^{-1}A$ iff $S\cap\p_i\neq\emptyset$, and otherwise is a prime ideal.
    For $S=A-\p$ for a minimal prime ideal $\ann(M)\subseteq\p$, then $S^{-1}\p_i=A_\p$ if $\p\neq\p_i$ and otherwise $S^{-1}\p_i=\p A_\p$ is a maximal ideal.
    Thus the sequence $(M_i)_\p$ forms a composition series after deleting the repeated entries (when $\p\neq\p_i$), and we are left with a composition series (since the quotients are
    quotients by a maximal ideal) whose length is equal to the number of $\p_i$ equal to $\p$.
    \qed

\eproof

\bdefn

    A minimal prime ideal is a prime ideal not containing any other prime ideal.

\edefn

\blemm

    A Noetherian ring $A$ has finitely many minimal prime ideals.

\elemm

\bproof

    The minimal prime ideals of $A$ are the minimal primes containing $\ann(A)=0$.
    Then for any chain as in the above lemma (whose components are $A/\p$ for some prime), the minimal prime ideals are the minimal primes from $\set{\p_1,\dots,\p_n}$ which is finite.
    \qed

\eproof

\bprop

    A ring is Artinian iff it is Noetherian and every prime ideal is maximal.

\eprop

\bproof

    If $A$ is Noetherian and every prime ideal is maximal, then every prime ideal is also minimal.
    Thus $A$ has finitely many prime/maximal ideals, $\m_1,\dots,\m_k$.
    Then their product $I=\m_1\cdots\m_k$ is contained in the intersection of all prime ideals, $\nil(A)$.
    Because $A$ is Noetherian $I$ is finitely generated, so $I^n=0$ for some $n$ (since every element of $I$ is nilpotent, and it is finitely generated).

    We have a chain of submodules
    $$ 0 = M_0 \subseteq M_1 \subseteq \cdots \subseteq M_k = A , $$
    where $M_{i+1}/M_i\cong A/\m_i$ for some $i$ (by the above proposition).
    As a subquotient, $M_{i+1}/M_i$ is a Noetherian $A$-module, and thus a Notherian module over the field $K=A/\m_i$.
    A Noetherian vector space is finite-dimensional (it is fg), so that $M_{i+1}/M_i$ is of finite length over $K$ (its length is equal to its dimension).
    Then $M_{i+1}/M_i$ has finite length over $A$ as well, and then inductively we see that each $M_i$ has finite length, and in particular $A$ is Artinian.

    Conversely if $A$ is Artinian, then all of its prime ideals are maximal (homework), so we need only show it is Noetherian.
    If $\m_1,\dots,\m_{n+1}$ are distinct maximal ideals then $\m_1\cdots\m_n\supset\m_1\cdots\m_{n+1}$ is strict, so by Artinian-ness there must be finitely many maximal ideals.
    If the inclusion were not strict then $\m_1\cdots\m_n\subseteq\m_{n+1}$, and then by $\m_{n+1}$'s primality, $\m_i\subseteq\m_{n+1}$ for some $i$.
    But then $\m_i=\m_{n+1}$ by maximality, a contradiction to distinctness.

    So let $I=\m_1\cdots\m_n$ be the product of all $A$'s prime/maximal ideals.
    Consider the descending chain $I\supseteq I^2\supseteq I^3\supseteq\cdots$ which must terminate, so $I^k=I^{k+1}$ eventually.
    We claim that $I^k=0$.

    If not, let $x_1,\dots,x_{k+1}\in I$ such that $x_1\cdots x_{k+1}\neq 0$.
    Because $I^k=I^{k+1}$, $\prod_{i=2}^{k+1}x_i=\sum_\alpha y_{\alpha,1}\cdots y_{\alpha,k+1}$ where the sum is over a finite set and each $y_{\alpha,i}\in I$.
    Thus $x_1y_{\alpha_0,1}\cdots y_{\alpha_0,k+1}\neq0$ for some $\alpha_0$.
    Then from the original tuple $x_1,\dots,x_{k+1}\in I$ we obtain a new tuple $x'_1=x_1,x'_2,\dots,x'_{k+2}$ ($x'_i=y_{\alpha_0,i-1}$) with nonzero product, and the first element unchanged.

    We similarly may write the product of the final $k$ elements $x'_3\cdots x'_{k+2}$ as an element of $I^{k+1}$ and obtain $x_1,x'_2,x'_3,x''_4,\dots,x''_{k+3}$ with nonzero product.
    Continuing this inductively, we produce an infinite sequence $y_1,y_2,\dots$ such that $y_1\cdots y_m\neq0$ for all $m>0$.
    Now consider the descending chain $(y_1)\supseteq(y_1y_2)\supseteq\cdots$, which stabilizes since $A$ is Artinian.
    Thus for some $m>0$ and $a\in A$,
    $$ \prod_{i=1}^my_i = a\prod_{i=1}^{m+1}y_i \implies \parens{\prod_{i=1}^my_i}(1-ay_{m+1}) = 0 . $$
    Since $y_{m+1}\in I$ and thus $ay_{m+1}\in I$ is in the Jacobson radical ($I$ is the product of all maximal ideals, and thus contained in their intersection), we have that $1-ay_{m+1}\in A^\times$.
    Thus $\prod_{i=1}^my_i=0$, a contradiction.

    So $I^k=0$, and so there is a chain (taking $M_j=\m_1\cdots\m_{n-j}$),
    $$ 0 = M_0 \subseteq M_1 \subseteq \cdots \subseteq M_m = A $$
    where $M_{i+1}/M_i$ is annihlated by some $\m_i$, and thus is Artinian over the field $K=A/\m_i$ so finite-dimensional over $A$.
    Thus $M_{i+1}/M_i$ has finite length over $A/\m_i$, and thus over $A$, and then by induction we see that each $M_i$ has finite length, in particular so does $A$ and thus is Noetherian.
    \qed

\eproof

\bnote

    Recall that $\bZ$ is Notherian but not Artinian.
    This is because while all of its non-zero prime ideals are maximal, $(0)$ is a prime ideal which is not maximal.
    We see that in general an integral domain which is not a field cannot be Artinian.

\enote

\subsection{Tensor products}

I assume some familiarity with the tensor product (see e.g.\ my summary on representation theory otherwise).
The tensor product is a bifunctor
$$ \bul\otimes_B\bul\colon {}_A\Mod_B\times{}_B\Mod_C \to {}_A\Mod_C , $$
with the universal property that it represents the balanced map functor:
$$ \hom(M\otimes_BN,P) \cong {\rm Bal}(M,N;\,P) . $$
Balanced maps are maps $M\times N\to P$ which are additive and multiplicative in each coordinate, and $f(mb,n)=f(m,bn)$ for all $b\in B$.

The tensor product also has a right-adjoint.
For an $(A,B)$-bimodule $M$, $\bul\otimes_AM$ is a functor $\Mod_A\to\Mod_B$ with right adjoint $\hom_B(M,\bul)$.

In the case of commutative rings (this course), all modules are bimodules, so the tensor product is a bifunctor
$$ \bul\otimes_A\bul\colon \Mod_A\to\Mod_A, $$
which represents the bilinear map functor (bilinear in the usual sense).

Note still that we use tensor products between module categories.
In particular if $B$ is an $A$-algebra, then $B\otimes_AM$ is a $B$-module ($b(c\otimes m)=(bc)\otimes m$).
This defines a map $\Mod_A\to\Mod_B$, and its right-adjoint is the restriction map (restrict a $B$-module to an $A$-module).
Note that this is a special case of the tensor-hom adjunction: $\bul\otimes_AB$ is left-adjoint to $\hom_B(B,\bul)\cong B$.

So we have a natural isomorphism
$$ \hom_B(B\otimes_AM,P) \cong \hom_A(M,P|_A),\qquad M\in\Mod_A, P\in\Mod_B . $$
The correspondence is given by mapping $f\colon B\otimes_AM\to P$ to $\tilde f\colon M\to P$ by $\tilde f(m)=f(1\otimes m)$.
And $g\colon M\to P|_A$ to $\hat g\colon B\otimes_AM\to P$ by $\hat g(1\otimes m)=g(m)$.

\bprop

    Let $M$ be an $A$-module.
    \benum
        \item For a multiplicative set $S\subseteq A$, $S^{-1}A\otimes_AM$ is naturally isomorphic to $S^{-1}M$.
        \item For an ideal $I\leq A$, $A/I\otimes_AM$ is naturally isomorphic to $M/IM$.
    \eenum

\eprop

\bproof

    The canonical map $M\to S^{-1}M$ mapping $m\mapsto m/1$ defines a map $f\colon S^{-1}A\otimes_AM\to S^{-1}M$ by
    $f(1\otimes m)=m/1$.
    Then
    $$ f(a/s\otimes m) = a/sf(1\otimes m) = am/s . $$
    We define its inverse $g\colon S^{-1}M\to S^{-1}A\otimes_AM$ by
    $$ g(m/s) = 1/s\otimes m . $$
    We claim this is well-defined: if $m/s=m'/s'$ then $1/s\otimes m=1/s'\otimes m'$.
    If $t(ms'-m's)$ then
    $$ \frac1s\otimes m = \frac{ts'}{ts's}\otimes m = \frac1{ts's}\otimes ts'm = \frac1{ts's}\otimes tsm' = \frac1{s'}\otimes m' $$
    as desired, and clearly $g$ is a module morphism.
    $f$ and $g$ are also clearly morphisms.

    The proof for the second point is similar: projection $M\to M/IM$ is an $A$-module morphism.
    Thus this gives rise to an $A/IA$-module morphism $f\colon(A/IA)\otimes_AM\to M/IM$ by $f(1\otimes m)=\bar m$.
    Then
    $$ f(\bar a\otimes m) = af(1\otimes m) = a\bar m = \bar{am} . $$
    Its inverse is $g\colon M/IM\to(A/IA)\otimes_AM$ given by $g(\bar m)=1\otimes m$; this is easily seen to be well-defined and an inverse.
    \qed

\eproof

\bcoro

    Let $B$ be an $A$-algebra and $M$ a $B$-module.
    Then there is a natural map
    $$ B\otimes_A M|_A \to M,\qquad b\otimes m\to bm . $$
    In particular this is a natural map $m_B\colon B\otimes_AB\to B$ by $b\otimes b'\mapsto bb'$.

\ecoro

\bproof

    This map is the map corresponding to the identity of $\hom_A(M|_A,M|_A)$.
    The second point is for $M=B$.
    \qed

\eproof

\blemm

    If $B,C$ are $A$-algebras, then their tensor product $B\otimes_AC$ is naturally an $A$-algebra and their coproduct in the category of
    $A$-algebras.

\elemm

\bproof

    Let $D=B\otimes_AC$.
    The natural algebra structure is given by $(b\otimes c)(b'\otimes c')=(bb')\otimes(cc')$, and $A\to D$ by $a\mapsto a(1\otimes 1)$.
    This is the composition of
    $$ D\times D\to D\otimes_AD,\qquad (d,d')\mapsto d\otimes d', $$
    and
    $$ D\otimes_AD = (B\otimes_AC)\otimes_A(B\otimes_AC) \cong (B\otimes_AB)\otimes_A(C\otimes_AC) \to B\otimes_AC $$
    where the final map is the tensor of $m_B\otimes m_C$.

    The inclusion maps $i_B\colon B\to D$ and $i_C\colon C\to D$ are given by $b\mapsto b\otimes1$ and $c\mapsto1\otimes c$.

    Now if $E$ is another $A$-algebra and there are maps $f\colon B\to E,g\colon C\to E$, then if $h\colon D\to E$ commutes with the
    rest of the diagram, i.e.
    $$ g = h\circ i_C,\qquad f = h\circ i_B, $$
    then we must have
    $$ h(b\otimes c) = h((b\otimes1)(1\otimes c)) = h(b\otimes1)h(1\otimes c) = hi_B(b)hi_C(c) = f(b)g(c) . $$
    Thus $h$ is unique, and this is well-defined (it is bilinear).
    \qed

\eproof

Recall the notion of left- and right-exact functors (preserving finite limits and colimits respectively).
Tensor products are left-adjoints and thus right-exact:

\bprop

    If $M'\to M\to M''\to0$ is an exact sequence of $A$-modules, then for any $A$-module $N$, the induced sequence
    $$ M'\otimes_AN \to M\otimes_AN \to M''\otimes_AN \to 0 $$
    is exact.

\eprop

\bnote

    Tensor products are not exact; they do not preserve injections.
    Indeed, if $f\colon A\to A$ is the map $f(x)=ax$ for $a\in A$, then $f\otimes M=f\otimes\id_M$ is the endomorphism of
    $A\otimes_AM\cong M$ which maps $m\mapsto am$.
    When $a$ is not a zero divisor (e.g.\ $A$ is a domain), $f$ is injective, but $f\otimes M$ need not be (e.g.\ $M=A/(a)$).

\enote

Since $\hom_A(M,\bul)$ is right-adjoint, it is left exact so it preserves finite limits.
It is not hard to show that $\hom_A(\bul,M)$ is left-exact as well (but contravariantly).
Furthermore the Yoneda embedding reflects exactness:

\bprop

    A sequence of $A$-modules $M_\bul=M'\to M\to M''\to0$ is exact iff
    $$ \hom(M_\bul,N) = 0 \to \hom(M'',N) \to \hom(M,N) \to \hom(M',N) $$
    is exact for all $N\in\Mod_A$.
    Similarly $N_\bul=0\to N'\to N\to N''$ is exact iff
    $$ \hom(M,N_\bul) = 0 \to \hom(M,N') \to \hom(M,N) \to \hom(M,N'') $$
    is exact for all $M$.

\eprop

\bproof

    The ``only if'' part is due to left exactness.
    The converse (second point) is done by choosing $M=N'$ and looking at where $\id$ goes (this shows that the composition of maps
    is zero), then $M$ to be the kernel of the second map, and showing that this implies exactness at $M$, and then you just need to
    show the first map is injective.
    \qed

\eproof

While not all tensor products are exact, some are.
We give this case a special term.

\bdefn

    An $A$-module $M$ is {\emphcolor flat} if the functor $\bul\otimes_AM$ is exact.

\edefn

\bnote

    Recall that a functor is exact iff it preserves finite limits and colimits iff it preserves short exact sequences iff it preserves
    arbitrary exact sequences.

    Since $\bul\otimes_AM$ is right-exact for any module $M$, $M$ is flat iff $\bul\otimes_AM$ preserves injections.

\enote

\bexam

    If $S\subseteq A$ is multiplicative, then $\bul\otimes_AS^{-1}A$ is naturally isomorphic to $S^{-1}\bul$, which is exact.
    Thus $S^{-1}A$ is a flat $A$-module.

    If $M$ is free, so $M\cong A^{\oplus I}$, then
    $$ \bul\otimes_AM \cong \bul\otimes_AA^{\oplus I} (\bul\otimes_AA)^{\oplus I} \cong \bul^{\oplus I} . $$
    The second isomorphism is due to left adjoints preserving colimits.
    And $\bul^{\oplus I}$ is exact (it maps injective maps to injective maps, more generally direct sums in module categories are exact,
    satisfying Grothendieck's axiom (AB4)).

\eexam

\blemm

    Let $M$ be an $A$-module and $S\subseteq A$ a multiplicative set.
    \benum
        \item If $N$ is an $S^{-1}A$-module, it is an $A$-module via the canonical map $A\to S^{-1}A$, and there is a natural isomorphism
        $$ M\otimes_AN \to S^{-1}M\otimes_{S^{-1}A}N,\qquad m\otimes n\mapsto m/1\otimes n . $$
        \item For an $A$-module $N$, there is a natural isomorphism
        $$ S^{-1}M\otimes_{S^{-1}A}S^{-1}N \to S^{-1}(M\otimes_AN),\qquad \frac ms\otimes\frac nt \mapsto \frac{m\otimes n}{st} . $$
        In particular for every prime $\p\leq A$, there is a natural isomorphism
        $$ M_\p \otimes_{A_\p} N_\p \cong (M\otimes_AN)_\p . $$
    \eenum

\elemm

\bproof

    There is a natural isomorphism (by associativity of the tensor)
    $$ S^{-1}M\otimes_{S^{-1}A}N \cong (M\otimes_AS^{-1}A)\otimes_{S^{-1}A}N \cong M\otimes_AN $$
    which maps $m/s\otimes n$ to $m\otimes 1/s\otimes n$ to $m\otimes n/s$.
    So in reverse it maps $m\otimes n$ to $m/1\otimes n$ as desired.

    There is a natural isomorphism
    $$ S^{-1}M\otimes_{S^{-1}A}S^{-1}N \cong M\otimes_AS^{-1}N \cong M\otimes_A(N\otimes_AS^{-1}A) \cong (M\otimes_AN)\otimes_A S^{-1}A , $$
    and this is naturally isomorphic to $S^{-1}(M\otimes_AN)$.
    \qed

\eproof

\bcoro

    Flatnes is local, i.e.\ the following are equivalent:
    \benum
        \item $M$ is a flat $A$-module,
        \item $S^{-1}M$ is a flat $S^{-1}A$-module for all multiplicative $S\subseteq A$,
        \item $M_\p$ is a flat $A_\p$-module for all prime $\p\leq A$,
        \item $M_\m$ is a flat $A_\m$-module for all maximal $\m\leq A$.
    \eenum

\ecoro

\bproof

    If $M$ is flat then by the above lemma
    $$ \bul\otimes_{S^{-1}A}S^{-1}M \cong S^{-1}(\bul\otimes_AM) , $$
    which is a composition of exact functors and thus exact.

    Now if $M_\m$ is flat for all maximal $\m\leq A$, we show $M$ is flat.
    Recall we need only show that $\bul\otimes_AM$ preserves injectives.
    Recall that $f$ is injective iff $f_\m$ is injective for all maximal $\m\leq A$.
    Now, $f_\m$ is isomorphic to $f\otimes M_\m$, which is injective since $M_\m$ is flat, we see $f_\m$ is injective for all $\m$ as desired.
    \qed

\eproof

