\input pdfToolbox
\input preamble
\input pdfmsym

\pdfmsymsetscalefactor{10}

\setlayout{horizontal margin=2cm, vertical margin=2cm}
\parindent=0pt
\parskip=3pt plus 2pt minus 2pt

\input preamble

\footline={}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\headline={\pageborder{rgb{1 1 .7}}{rgb{.6 .4 0}}{5}}

%\color rgb{1 .5 .1}

{\def\boxshadowcolor{rgb{.7 .7 .5}}
\bppbox{rgb{1 1 .7}}{rgb{.6 .4 0}}{rgb{.2 .1 0}}

    \leftline{\setfontandscale{bf}{20pt}\quad Commutative Algebra and}
    \rightline{\setfontandscale{bf}{20pt}Algebraic Geometry\quad}
    \smallskip
    \centerline{\setfont{it}Summary by Ari Feiglin \setfont{rm}({\tt ari.feiglin@gmail.com})}

\eppbox

\bigskip

\bppbox{rgb{1 1 .7}}{rgb{.6 .4 0}}{rgb{.2 .1 0}}
    \section*{Contents}
    
    \tableofcontents
\eppbox

\vfill\break

\null\vfill

\bppbox{rgb{1 1 .7}}{rgb{.6 .4 0}}{rgb{.2 .1 0}}

    This summary is based on the course ``Fundamentals of Commutative Algebra and Algebraic Geometry'' taught in the Hebrew University of Jerusalem in the Spring semester
    of 2026 by Prof.\ Zev Rosengarten.

    While the summary covers the entirety of the material taught, it does not do so in the same order.
    In particular, while the course taught elements of algebraic geometry and commutative algebra simultaneously, these two topics were split into distinct parts in this summary.
    Within each part, the sections are kept in-order.
    This was done for ease of reference, but not ease of reading.
    Thus this summary should be used as a reference, and not read in-order.

    Issues may arise when reading this summary in-order, for example topics on algebraic geometry may be referenced before they are defined.
    Furthermore, a more full picture of the topic is apparent when you read about geometric aspects of algebra concurrently with reading about the algebra itself.
    Thus if you would like to use this summary to learn about the material yourself, it is recommended you follow the syllabus of the course.

    My personal recommendation for the order to read this summary is as follows:
    \benum
        \item Sections 1 and 2 of Commutative Algebra (Rings and Modules; Localization).
        \item Sections 1.1 and 1.2 of Algebraic Geometry (Affine Schemes, Basic definitions; Topological properties of affine schemes).
        \item Section 3 of Commutative Algebra (Properties of Modules).
        \item Section 2.1 of Algebraic Geometry (Varities; Affine varieties); this is essentially just more commutative algebra, there is little geometry here.
        \item Section 4 of Commutative Algebra (Tensoring and Flatness).
        \item Second 1.3 of Algebraic Geometry (Affine Schemes, Categorical constructions); this is a short section which requires tensor products.
    \eenum

\eppbox}

\vfill\break

\color{black}

\pageno=1
\newif\ifpageodd
\pageoddtrue
\headline={%
    \hbox to \hsize{\color{black}%
        \ifpageodd\hfil{\it\currsubsection\quad\bf\folio}\global\pageoddfalse%
        \else{\bf\folio\quad\it\currsubsection}\hfil\global\pageoddtrue\fi%
    }%
}


\input commutative-algebra

\vfill\break
\input algebraic-geometry

\bye

