\section{Connections}

\subsection{Connections}

\bdefn

    Let $E\to M$ be a smooth vector bundle over $M$, and denote its space of sections by $\Gamma(E)$.
    A {\emphcolor connection} in $E$ is a map
    $$ \nabla\colon\X(M)\times\Gamma(E)\to\Gamma(E) $$
    written $(X,Y)\mapsto\nabla_XY$ which has the following properties:
    \benum
        \item $\nabla_XY$ is $C^\infty(M)$-linear in $X$:
        $$ \nabla_{f_1X_2+f_2X_2}Y = f_1\nabla_{X_1}Y + f_2\nabla_{X_2},\qquad f_1,f_2\in C^\infty(M). $$
        \item $\nabla_XY$ is $\bR$-linear in $Y$:
        $$ \nabla_X(a_1Y_1+a_2Y_2) = a_1\nabla_XY_1 + a_2\nabla_XY_2,\qquad a_1,a_2\in\bR . $$
        \item $\nabla$ satisfies the product rule: for $f\in C^\infty(M)$:
        $$ \nabla_X(fY) = f\nabla_XY + (Xf)Y . $$
    \eenum

\edefn

\blemm

    If $\nabla$ is a connection over $E$, and $X,\tilde X\in\X(M)$ and $Y,\tilde Y\in\Gamma(E)$ are sections which agree
    on a neighborhood of $p\in M$, then
    $$ \nabla_XY|_p = \nabla_{\tilde X}\tilde Y|_p . $$

\elemm

\bproof

    By linearity it is sufficient to assume $X=0$ or $Y=0$ in a neighborhood of $p$ and show that in each case
    $\nabla_XY=0$.
    For $Y$, suppose $Y$ is zero in a neighborhood $U$ of $p$.
    Let $\psi$ a bump function which is $1$ at $p$ and supported in $U$.
    Thus $\phi Y=0$, and so by the axioms
    $$ 0 = \nabla_X(\phi Y) = \phi\nabla_XY + (X\phi)Y . $$
    Since $Y$ is zero on $\phi$'s support, the second term is zero.
    So evaluating the first term at $p$ gives
    $$ 0 = \phi(p)\nabla_XY|_p = \nabla_XY|_p , $$
    as desired.

    Choosing a similar bump function for $X$ gives
    $$ 0 = \nabla_{\phi X}Y = \phi\nabla_XY $$
    and evaluating at $p$ gives the desired result again.
    \qed

\eproof

\blemm

    Let $\nabla$ be a connection in $E\to M$.
    For an open subset $U\subseteq M$ there is a unique connection $\nabla^U$ on the restricted bundle $E|_U$ such that
    for all $X\in\X(M)$ and $Y\in\Gamma(E)$,
    $$ \nabla_{X|_U}^U(Y|_U) = (\nabla_XY)|_U . $$

\elemm

\bproof

    First we show uniqueness.
    For $X\in\X(U)$ and $Y\in\Gamma(E|_U)$ arbitrary and $p\in U$ we use a bump function to define sections
    $\tilde X\in\X(U)$ and $\tilde Y\in\Gamma(E|_U)$ such that $\tilde X$ and $\tilde Y$ agree with $X,Y$ on a neighborhood
    of $p$ so that
    $$ \nabla_X^UY|_p = \nabla_{\tilde X|_U}^U(\tilde Y|_U)|_p = (\nabla_{\tilde X}\tilde Y)|_p , $$
    showing uniqueness.
    This also demonstrates how to construct $\nabla^U$.
    By the above lemma $\nabla_{\tilde X}\tilde Y|_p$ is independent on choice of $\tilde X,\tilde Y$, and it is easy to see
    that the above formula defines a connection.
    \qed

\eproof

\bnote

    The restricted connection is usually denoted $\nabla$ as well instead of $\nabla^U$.
    This lemma shows there can be no confusion.

\enote

\bprop

    $\nabla_XY$ depends only on $X$'s value at $p$ and $Y$'s values in any neighborhood of $p$.

\eprop

\bproof

    The lemma before this showed that the connection only depends on $X$ and $Y$ in any neighborhood of $p$.
    To show that it depends only on $X_p$, it is sufficient to assume $X_p=0$.
    Let $U$ be a coordinate neighborhood of $p$ and write $X=X^i\ipdv{x^i}$ so that $X^i(p)=0$.
    Then at $p$ by the above lemma on restricting the connection,
    $$ \nabla_XY|_p = \nabla_{X^i\ipdv{x^i}}Y|_p = (X^i\nabla_{\ipdv{x^i}}Y)|_p = 0 $$
    since $X^i(p)=0$.
    \qed

\eproof

Our main concern are connections in the tangent bundle, called just {\emphcolor connections on $M$}.
These are connections
$$ \nabla\colon\X(M)\times\X(M)\to\X(M) . $$

Let $(E_i)$ be a smooth local frame for $TM$ on an open subset $U\subseteq M$.
Then for every $i,j$ we can represent $\nabla_{E_i}E_j$ in $U$ over this coordinate frame.
So there are smooth functions $\Gamma^k_{ij}\in C^\infty(U)$ such that
$$ \nabla_{E_i}E_j = \Gamma^k_{ij}E_k . $$
These are $n^3$ smooth functions ($n=\dim M$) called the {\emphcolor connection coefficients} of $\nabla$ in the given frame.

\bprop

    Let $\nabla$ be a connection on $M$, $E_i$ a local frame over $U$, and $\Gamma^k_{ij}$ be $\nabla$'s connection coefficients
    in this frame.
    Then for $X,Y\in\X(U)$ with representations $X=X^iE_i,Y=Y^jE_j$, one has
    $$ \nabla_XY = (X(Y^k) + X^iY^j\Gamma^k_{ij})E_k . $$

\eprop

\bproof

    Shut up and compute.
    $$ \eqalign{
        \nabla_XY &= \nabla_X(Y^jE_j) = X(Y^j)E_j + Y^j\nabla_XE_j = X(Y^j)E_j + Y^j\nabla_{X^iE_i}E_j\cr
            &= X(Y^j)E_j + X^iY^j\nabla_{E_i}E_j\cr
            &= X(Y^j)E_j + X^iY^j\Gamma^k_{ij}E_k .
    } $$
    \qed

\eproof

\bprop

    Let $\nabla$ be a connection on $M$, and $(E_i)$ and $(\tilde E_j)$ are two local frames over $U$ related by
    $\tilde E_i=A^j_iE_j$ for a matrix $A=(A^j_i)$.
    Let $\Gamma^k_{ij},\tilde\Gamma^k_{ij}$ be the connection coefficients relative to these frames, then
    $$ \tilde\Gamma^k_{ij} = (A^{-1})^k_pA^q_iA^r_j\Gamma^p_{qr} + (A^{-1})^k_pA^q_iE_q(E^p_j) . $$

\eprop

\bproof

    Tedious computations.
    \qed

\eproof

\bprop

    If $TM$ has a global frame $(E_i)$ then any choice of smooth functions $\Gamma^k_{ij}\in C^\infty(M)$ defines a
    connection on $M$ with these as the connection coefficients relative to $(E_i)$.

\eprop

\bproof

    Write $X=X^iE_i$ and $Y=Y^jE_j$, and show that the formula above
    $$ \nabla_XY = (X(Y^k) + X^iY^j\Gamma^k_{ij})E_k $$
    defines a connection on $M$.
    This is clear, or an easy set of computations.
    \qed

\eproof

\bexam

    $T\bR^n$ has a global frame $\ipdv{x^i}$.
    So define $\Gamma^k_{ij}=0$ for all $i,j,k$, this gives us the {\emphcolor Euclidean connection}:
    $$ \bar\nabla_XY = X(Y^i)\pdv{x^i} . $$

\eexam

\bnote

    Linear combinations of connections need not be connections.
    They are obviously still linear, but need not satisfy the product rule.
    For example if $\nabla$ is a connection then
    $$ 2\nabla_X(fY) = 2(XfY + f\nabla_XY) . $$
    But the product rule requires
    $$ (2\nabla_X(fY)) = XfY + f(2\nabla_X)Y = XfY + 2f\nabla_XY , $$
    which are clearly not equal.
    Similarly for sums.

\enote

\bprop

    Every manifold admits a connection on it.

\eprop

\bproof

    Let $M$ be a smooth manifold and cover it with coordinate charts $\set{U_\alpha}$.
    These are all diffeomorphic to Euclidean space and thus can be given the (restriction of the) Euclidean connection
    $\nabla^\alpha$.
    Let $\set{\phi_\alpha}$ be a partition of unity subordinate to $\set{U_\alpha}$ and define
    $$ \nabla_XY = \sum_\alpha\phi_\alpha\nabla^\alpha_XY . $$
    Since partitions of unity are locally finite, this defines a smooth vector field on $M$.
    This is clearly still linear, but we need to show the product rule, as explained in the note above.

    We have
    $$ \nabla_X(fY) = \sum_\alpha\phi_\alpha(XfY + f\nabla^\alpha_XY) = \parens{\sum_\alpha\phi_\alpha}XfY
    + f\sum_\alpha\phi_\alpha\nabla^\alpha_XY = XfY + f\nabla_XY $$
    because the sum of a partition of unity is one.
    \qed

\eproof

\bprop

    Let $\nabla^0,\nabla^1$ be two connections on $M$, then the map $D\colon\X(M)\times\X(M)\to\X(M)$ given by
    $$ D(X,Y) = \nabla^1_XY - \nabla^0_XY $$
    is bilinear over $C^\infty(M)$ and thus defines a $(1,2)$-tensor field called the {\emphcolor difference tensor}.

\eprop

\bproof

    Additivity is clear.
    Multiplicity is by multiplicity in the lower input and the product rule.
    \qed

\eproof

\bthrm

    Let $\nabla^0$ be a connection on $M$, then the set $\A(TM)$ of all connections on $M$ is equal to
    $$ \A(TM) = \set{\nabla^0 + D}[D\in\T^{(1,2)}TM] , $$
    where $\nabla^0+D$ is the connection given by $(\nabla^0+D)_XY=\nabla^0_XY+D(X,Y)$.

\ethrm

\bproof

    The proposition above shows that $\A(TM)$ is contained in this affine space, and it is routine to show that elements of
    the affine space are all connections.
    \qed

\eproof

\subsection{Covariant derivatives of tensor fields}

Recall that a $(1,1)$-tensor over $V$ is just a linear map $V\to V$.
Thus we can naturally define the trace of such a tensor to be the trace of its corresponding linear map.
In general we define the {\emphcolor trace operator}
$$ \tr\colon T^{(k+1,\ell+1)}(V) \to T^{(k,\ell)}(V) $$
by mapping $F\in T^{(k+1,\ell+1)}(V)$ to $\tr(F)\in T^{(k,\ell)}(V)$ where $\tr(F)(\omega^1,\dots,\omega^k,v_1,\dots,v_\ell)$
is defined to be the trace of the $(1,1)$-tensor
$$ F(\omega^1,\dots,\omega^k,\bul,v_1,\dots,v_\ell,\bul) \in T^{(1,1)}(V) . $$
This is clearly a tensor by linearity of the trace.
In terms of a basis
$$ (\tr F)^I_J = F^{I,m}_{J,m} $$
summing over $m$.

Note that we can take the trace of a tensor on different indices, not just the last ones, as long as one is covariant and
the other contravariant.

\bprop

    Let $M$ be a smooth manifold with connection $\nabla$.
    Then $\nabla$ uniquely determines a connection on the tensor bundle $T^{(k,\ell)}TM$ also denoted $\nabla$,
    such that
    \benum
        \item In $T^{(1,0)}TM=TM$, $\nabla$ agrees with the given connection.
        \item In $T^{(0,0)}TM=M\times\bR$ (whose sections are identified with real-valued functions),
        $\nabla$ is ordinary differentiatio of functions,
        $$ \nabla_Xf = Xf . $$
        \item $\nabla$ satisfies the product rule with respect to tensor products,
        $$ \nabla_X(F\otimes G) = (\nabla_XF)\otimes G + F\otimes(\nabla_XG) . $$
        \item $\nabla$ commutes with trace (contractions): if $\tr$ is the trace of a tensor on any pair of indices, one
        covariant and one contravariant, then
        $$ \nabla_X(\tr F) = \tr(\nabla_XF) . $$
    \eenum
    Furthermore the connection satisfies
    \benum
        \item For a vector field $Y$ and covector field $\omega$,
        $$ \nabla_X\iprod{\omega,Y} = \iprod{\nabla_X\omega,Y} + \iprod{\omega,\nabla_XY} $$
        where $\iprod{\omega,Y}$ is just $\omega(Y)$.
        \item For $F\in\T^{(k,\ell)}TM$, covector fields $\omega^1,\dots,\omega^k$, and vector fields $Y^1,\dots,Y^\ell$,
        $$ \eqalign{
            (\nabla_XF)(\omega^1,\dots,\omega^k,Y_1,\dots,Y_\ell) &= X(F(\bar\omega,\bar Y))\cr
            &\hphantom{{}={}}- \sum_{i=1}^k F(\omega^1,\dots,\nabla_X\omega^i,\dots,\omega^k,\bar Y)
             - \sum_{j=1}^\ell F(\bar\omega,Y_1,\dots,\nabla_XY_j,\dots,Y_\ell) .
        } $$
    \eenum

\eprop

\bproof

    First we show that the conditions imply the two properties at the end.
    Note that $\iprod{\omega,Y}=\omega_iY^i=\tr(\omega\otimes Y)$.
    Thus
    $$ \nabla_X\iprod{\omega,Y} = \nabla_X(\tr(\omega\otimes Y)) = \tr\nabla_X(\omega\otimes Y)
    = \tr(\nabla_X\omega\otimes Y + \omega\otimes\nabla_XY) = \iprod{\nabla_X\omega,Y} + \iprod{\omega,\nabla_XY} . $$
    The second property is proven by induction, noting that
    $$ F(\bar\omega,\bar Y) = \tr\cdots\tr(F\otimes\bar\omega\otimes\bar Y) $$
    where we take $k+\ell$ traces, each operating on an upper index of $F$ and the lower index of the corresponding
    covector field, or a lower index of $F$ and the upper index of the corresponding vector field.

    For uniqueness, notice that by the conditions and properties,
    $$ (\nabla_X\omega)(Y) = \iprod{\nabla_X\omega,Y} = \nabla_X(\omega(Y)) - \omega(\nabla_XY) . $$
    Thus we have a unique method of computing the covariant derivatives of covector fields, and by the second property
    this gives us an expression for $\nabla_XF$ for any tensor, showing uniqueness.

    Now to show existence, define $\nabla_X\omega$ for covector fields as in the previous equation, and extend this to
    all tensors.
    Then $\nabla_XF$ is $C^\infty(M)$-multilinear (as can be checked), thus defining a smooth tensor field.
    Checking that $\nabla_X$ actually defines a connection is similarly routine.
    \qed

\eproof

The expressions for the covariant derivative given in the proof/proposition above are not useful for computations.

\bprop

    Let $\nabla$ be a connection on $M$, $(E_i)$ a local frame, $(\epsilon^j)$ its dual coframe, and $\Gamma^k_{ij}$
    the connection coefficients of $\nabla$ with respect to $(E_i)$.
    Let $X=X^iE_i$ be a local vector field.
    \benum
        \item The covariant derivative of $\omega=\omega_i\epsilon^i$ is given locally by
        $$ \nabla_X\omega = (X(\omega_k)-X^j\omega_i\Gamma^i_{jk})\epsilon^k . $$
        \item If $F\in\Gamma(T^{(k,\ell)}TM)$ is a smooth tensor field with local representation
        $$ F = F^I_JE_I\otimes\epsilon^J , $$
        then its covariant derivative is
        $$ \nabla_XF = \parens{X(F^I_J) + \sum_{s=1}^kX^mF^{I^p_{\hat s}}_J\Gamma^{i_s}_{mp} - \sum_{s=1}^\ell X^mF^I_{J^p_{\hat s}}\Gamma^p_{mj_s}}
        E_I\otimes\epsilon^J , $$
        where $I^p_{\hat s}=(i_1,\dots,i_{s-1},p,i_{s+1},\dots,i_k)$.
    \eenum

\eprop

\bproof

    We know that
    $$ (\nabla_X\omega)(Y) = \nabla_X(\omega(Y)) - \omega(\nabla_XY) . $$
    And
    $$ \nabla_X\omega = (\nabla_X\omega)(E_i)\epsilon^i  = \parens{\nabla_X(\omega(E_i)) - \omega(\nabla_XE_i)}\epsilon^i . $$
    Since $\omega(E_i)=\omega_i$ and $\nabla_X$ of a map is just its derivative relative to $X$, and by expanding $\nabla_X$ by linearity,
    we get that this is equal to
    $$ = \parens{\nabla_X(\omega_i) - \omega(X^j\Gamma^k_{ij}E_k)}\epsilon^i = \parens{X(\omega_i) - X^j\omega_k\Gamma^k_{ij}}\epsilon^i . $$

    The second formula is by considering $(\nabla_XF)(\epsilon^I,E_J)$ via the formula in the theorem above.
    \qed

\eproof

Note that $\nabla_\bul F$ is $C^\infty(M)$-linear for any $F$, and thus defines a tensor:

\bprop

    Let $\nabla$ be a connection on $M$, and $F\in\Gamma(T^{(k,\ell)}TM)$ a tensor field.
    Define
    $$ \nabla F\colon\X^*(M)^k \times \X(M)^\ell \times \X(M) \to C^\infty(M) $$
    by
    $$ \nabla F(\omega^I,Y_J,X) = (\nabla_X F)(\omega^I,Y_J) . $$
    This defines a smooth $(k,\ell+1)$-tensor field on $M$ called the {\emphcolor total covariant derivative} of $F$.

\eprop

\bproof

    $\nabla_XF$ is a tensor field and thus is $C^\infty(M)$-multilinear.
    Since $\nabla$ is a connection, $\nabla_\bul F$ is $C^\infty(M)$-linear.
    Thus in total $\nabla F$ is $C^\infty(M)$-multilinear and thus defines a tensor field.
    \qed

\eproof

When representing the total covariant derivative in a local coordinate frame, we have
$$ \nabla F = \nabla F(\epsilon^I,E_J,E_j)E_I\otimes\epsilon^J\otimes\epsilon^j =
(\nabla_{E_j}F)(\epsilon^I,E_J)E_I\otimes\epsilon^J\otimes\epsilon_j . $$
It is customary to denote the components by
$$ F^I_{J;j} = \nabla F(\epsilon^I,E_J,E_j) , $$
since the final index ($j$) has different significance than the rest.

So for example, if $Y$ is a vector field $Y=Y^iE_i$ then
$$ \nabla Y = Y^i_{;j}E_i\otimes\epsilon^j,\qquad Y^i_{;j} = \nabla_{E_j} Y(\epsilon^i) = E_jY^i + Y^k\Gamma^i_{jk} . $$
And similarly for covector field $\omega=\omega_i\epsilon^i$,
$$ \nabla\omega = \omega_{i;j}\epsilon^i\otimes\epsilon^j,\qquad \omega_{i;j} = E_j\omega_i - \omega_k\Gamma^k_{ij} . $$

In general we have
$$ F^I_{J;m} = \nabla_{E_m}F(\epsilon^I,E_J) $$
and we know the coordinate representation of $\nabla_{E_j}F$, giving us
$$ F^I_{J;j} = E_m(E^I_J) + \sum_{s=1}^k F^{I^p_{\hat s}}_J\Gamma^{i_s}_{mp} - \sum_{s=1}^\ell F^I_{J^p_{\hat s}}\Gamma^p_{mj_s} . $$

Since $\nabla F$ is itself a tensor field, we can take its own total covariant derivative: $\nabla^2 F=\nabla\nabla F$,
which is a $(k,\ell+2)$-tensor field.
Define for $X,Y\in\X(M)$,
$$ \nabla^2_{X,Y}F(\bul) = \nabla^2 F(\bul,Y,X) . $$
The reason for swapping order of $X,Y$ is because $\nabla^2_{X,Y}$ should be read inside-out: first we differentiate with respect to $Y$
and then $X$.

Note that $\nabla^2_{X,Y}F$ does not equal $\nabla_X(\nabla_YF)$.
This can be seen because the former is $C^\infty(M)$-linear in $Y$ but the latter is not.

\bprop

    $$ \nabla^2_{X,Y}F = \nabla_X(\nabla_YF) - \nabla_{\nabla_XY}F . $$

\eprop

\bproof

    Note that
    $$ \nabla_YF = \tr(\nabla F\otimes Y) , $$
    where the trace is on its final two indices.
    This can be checked directly by considering the representations of each by the coordinate bases (both give $F^I_{J;m}Y^m$).
    Similarly one can check that
    $$ \nabla^2_{X,Y}F = \tr(\tr(\nabla^2F\otimes X)\otimes Y) , $$
    the inner trace being on the last index of $\nabla^2F$ with $X$'s, and the outer trace the last free index (the second-to-last
    in $\nabla^2F$) with $Y$'s.

    Thus
    $$ \nabla_X\nabla_YF = \nabla_X(\tr(\nabla F\otimes Y)) = \tr\nabla_X(\nabla F\otimes Y) =
    \tr(\nabla_X(\nabla F)\otimes Y + \nabla F\otimes\nabla_XY), $$
    applying the same idea again gives
    $$ = \tr(\tr(\nabla^2F\otimes X))+\tr(\nabla F\otimes\nabla_XY)) . $$
    The first term is $\nabla^2_{X,Y}F$, the second is $\nabla_{\nabla_XY}F$, as desired.
    \qed

\eproof

\bexam

    Let $u\in C^\infty(M)$ be a smooth function.
    Then $\nabla u\in\X^*(M)$ is just $du$; they both act on vector fields the same way: $\nabla_Xu=Xu=du(X)$.

    The $2$-tensor $\nabla^2u=\nabla(du)$ is called the {\emphcolor covariant Hessian} of $u$.
    By the above proposition:
    $$ \nabla^2 u(Y,X) = \nabla^2_{X,Y}u = \nabla_X(\nabla_Y u) - \nabla_{\nabla_XY}u = X(Yu) - (\nabla_XY)u . $$

\eexam

\subsection{Pullbacks of vector bundles and connections}

\bdefn

    If $F\colon M\to N$ is a smooth map, and $\pi\colon E\to N$ a vector bundle on $N$, we define its {\emphcolor pullback}
    to be the vector bundle $\pi^*\colon F^*E\to M$ defined by
    $$ F^*E = \set{(p,e)\in M\times E}[F(p)=\pi(e)] \subseteq M\times E , $$
    where $\pi^*$ is projection onto the first component.

\edefn

Projecting onto the second component gives a map $f\colon F^*E\to E$ so that the following diagram commutes

\commsq{F^*E&M\cr E&N}
    {\pi^*}{\pi}{f}{F}

To show that this is indeed a vector bundle we need to give local trivializations.
Let $(U,\Phi)$ be a local trivialization of $E$, where $\Phi\colon\pi^{-1}U\to U\times\bR^k$, then $(F^{-1}U,\Phi^*)$ is a local trivialization of $F^*E$ where
$$ \Phi^*\colon(\pi^*)^{-1}(F^{-1}U) \to F^{-1}U\times\bR^k,\qquad \Phi^*(p,e) = (p,\proj_2(\Phi(e))) . $$
Note that this is well-defined, since $(\pi^*)^{-1}(F^{-1}U)=f^{-1}(\pi^{-1}U)$ so if $(p,e)$ is in it then $e\in\pi^{-1}U$ and $Fp=\pi e$.
Thus $\Phi(e)$ is defined, and $p\in F^{-1}U$.

Now $\Phi^*$ is also bijective.
It is injective since if $\Phi^*(p,e)=\Phi^*(q,e')$ then clearly $p=q$, and so $Fp=\pi e=\pi e'$ so $e,e'\in E_{Fp}$.
Then $\proj_2\Phi|_{E_{Fp}}$ is a bijection $E_{Fp}\to\bR^k$ so $e=e'$.
It is surjective since if $(p,v)\in F^{-1}U\times\bR^k$ then consider the bijection $\Phi|_{E_{Fp}}\colon E_{Fp}\to\bR^k$, there is an $e\in E_{Fp}$ such that
$\Phi(e)=v$.
And $e\in E_{Fp}$ means that $\pi e=Fp$ so $(p,e)$ is the preimage of $(p,v)$.

And $\Phi^*$'s restriction to $(\pi^*)^{-1}\set p=\set{(p,e)}[\pi e=Fp]$ is
$$ \Phi^*(p,e) = (p,\proj_2\Phi(e)) $$
whose second component is $\proj_2\Phi|_{E_{Fp}}$ which is a linear isomorphism to $\bR^k$.

If $\Psi$ is another local trivialization, then
$$ \Phi^*\circ(\Psi^*)^{-1}(p,v) = \Phi^*(p,\Psi^{-1}(Fp,v)) = (p,\proj_2\Phi\Psi^{-1}(Fp,v)) , $$
and the second coordinate is linear, it is equal to $\tau(Fp)v$ where $\tau$ is the transition function between $\Phi,\Psi$.

Thus $\pi^*\colon F^*E\to M$ is a vector bundle over $M$.
Moreover, $f$ is a bundle morphism over $F$.
Indeed,
$$ f|_{F^*E_p}\colon F^*E_p \to E_{Fp} $$
is the map from $F^*E_p=\set{(p,e)\in F^*E}$ to $E_{Fp}$ by mapping $(p,e)\mapsto e$.
This is clearly linear.

The pullback bundle is precisely a categorical pullback.
That is, it satisfies a certain universal property.
Specifically, if $\tau\colon X\to M$ is any other vector bundle over $M$ and $g\colon X\to E$ is a vector bundle morphism over $F$
then there is a unique vector bundle morphism $h\colon X\to F^*E$ over $M$ commuting everything.
This is summarized in the following diagram:

\bigskip
\centerline{\drawdiagram{
    $X$\cr
    &$F^*E$&$E$\cr
    &$M$&$N$
}{
    \diagarrow{from={1,1}, to={2,3}, curve=1cm, text=$g$, y distance=.75cm}
    \diagarrow{from={1,1}, to={3,2}, curve=-1cm, text=$\tau$, x distance=-.75cm}
    \diagarrow{from={1,1}, to={2,2}, dashed, text=$\exists!h$, y distance=.25cm}
    \diagarrow{from={2,2}, to={2,3}, text=$f$, y distance=.25cm}
    \diagarrow{from={3,2}, to={3,3}, text=$F$, y distance=-.25cm}
    \diagarrow{from={2,2}, to={3,2}, text=$\pi^*$, x distance=-.25cm}
    \diagarrow{from={2,3}, to={3,3}, text=$\pi$, x distance=.25cm}
}}

The map $h$ is given explicitly by $h(x) = (\tau(x),g(x))$.

Taking the pullback is functorial.
To formalize this we construct a category whose objects are pairs $(F\colon M\to N,\pi\colon E\to N)$ where $F$ is smooth and $\pi$ is a
vector bundle.
A morphism $(F\colon M\to N,\pi\colon E\to N)\to(F'\colon M'\to N',\pi'\colon E'\to N')$ is then a triplet of maps
$(f_M\colon M\to M',f_N\colon N\to N',f_E\colon E\to E')$ such that the obvious diagram commutes:
$$ f_N\circ F = F'\circ f_M,\qquad \pi'\circ f_E = f_N\circ\pi , $$
and furthermore that $f_E$ forms a vector bundle morphism, so is linear over $\pi$'s fibers.
This is just a subcategory of a functor category (of shape $\bul\to\bul\ot\bul$).
Then the pullback is then just the restriction of the pullback (as a limit) to this subcategory.

Since taking limits is functorial, given such a morphism in this category there is a unique map $G\colon F^*E\to(F')^*E'$ which commutes
with everything.
Explicitly it is given by
$$ G(p,e) = (f_Mp,f_Ee) . $$
Then $G$ forms a bundle morphism over $f_M$.

More useful is another form of functorality.
Suppose we have the following diagram:

\centerline{\drawdiagram{
    &&$E$\cr
    $M$&$N$&$P$
}{
    \diagarrow{from={1,3}, to={2,3}, text=$\pi$, x distance=.25cm}
    \diagarrow{from={2,1}, to={2,2}, text=$F$, y distance=.25cm}
    \diagarrow{from={2,2}, to={2,3}, text=$G$, y distance=.25cm}
}}

We can fill this in by taking the pullback twice:

\centerline{\drawdiagram{
    $F^*(G^*E)$&$G^*E$&$E$\cr
    $M$&$N$&$P$
}{
    \diagarrow{from={1,1}, to={2,1}, text=$\pi^{**}$, x distance=.25cm}
    \diagarrow{from={1,2}, to={2,2}, text=$\pi^*$, x distance=.25cm}
    \diagarrow{from={1,3}, to={2,3}, text=$\pi$, x distance=.25cm}
    \diagarrow{from={1,1}, to={1,2}, text=$f$, y distance=-.25cm}
    \diagarrow{from={1,2}, to={1,3}, text=$g$, y distance=-.25cm}
    \diagarrow{from={2,1}, to={2,2}, text=$F$, y distance=.25cm}
    \diagarrow{from={2,2}, to={2,3}, text=$G$, y distance=.25cm}
}}

But we can also pull back $E$ through $G\circ F$ directly:

\bigskip
\centerline{\drawdiagram{
    $(G\circ F)^*E$\cr
    &$F^*(G^*E)$&$G^*E$&$E$\cr
    &$M$&$N$&$P$
}{
    \diagarrow{from={1,1}, to={3,2}, text=$\tau$, curve=-1cm, x distance=-.75cm}
    \diagarrow{from={1,1}, to={2,4}, text=$h$, curve=1cm, y distance=.75cm}
    \diagarrow{from={2,2}, to={3,2}, text=$\pi^{**}$, x distance=.25cm}
    \diagarrow{from={2,3}, to={3,3}, text=$\pi^*$, x distance=.25cm}
    \diagarrow{from={2,4}, to={3,4}, text=$\pi$, x distance=.25cm}
    \diagarrow{from={2,2}, to={2,3}, text=$f$, y distance=-.25cm}
    \diagarrow{from={2,3}, to={2,4}, text=$g$, y distance=-.25cm}
    \diagarrow{from={3,2}, to={3,3}, text=$F$, y distance=.25cm}
    \diagarrow{from={3,3}, to={3,4}, text=$G$, y distance=.25cm}
}}

By the universal property of the pullback, we see that there are unique maps between $(G\circ F)^*E$ and $F^*(G^*E)$, and their composition must
then by the identity.
So we see that $(G\circ F)^*E$ and $F^*G^*E$ are naturally isomorphic.

Since $\id^*E$ is naturally isomorphic to $E$ (showing this explicitly is easy, but we can also appeal to the universal property).
Thus the pullback is functorial in the sense that
$$ (G\circ F)^*E = F^*G^*E,\qquad \id^*E = E $$
up to natural isomorphism.

In particular we see that if $F$ is a diffeomorphism, $F^*E$ is naturally isomorphic to $E$.

\bprop

    If $f\colon E\to E'$ is a vector bundle isomorphism over $M$, then we can push forward a connection on $E$ to $E'$.
    Explicitly, let $\nabla$ be a connection on $E$, then there is a unique connection $\tilde\nabla$ on $E'$ such that
    $$ \tilde\nabla_XY = f\circ\nabla_Xf^{-1}Y $$
    for all $X\in\X(M)$ and $Y\in\Gamma(E')$.

\eprop

\bproof

    Since $f$ is a bundle isomorphism, $f^{-1}Y$ is a section of $E$ for $Y\in\Gamma(E')$.
    Then $\nabla_Xf^{-1}Y$ is a section of $E$, and $f\circ\nabla_Xf^{-1}Y$ is a section of $E'$.
    Seeing that this is a connection is simple.
    \qed

\eproof

\bprop

    Sections of the pullback bundle $F^*E$ are equivalent to maps $\sigma\colon M\to E$ where $\sigma(p)\in E_{Fp}$.
    Viewing these as $C^\infty(M)$-modules, they are isomorphic.

\eprop

\bproof

    We want to give a natural correspondence
    $$ \Gamma(F^*E) = \set{\sigma\colon M\to F^*E}[\pi^*\circ\sigma=\id] \oto \set{\sigma\colon M\to E}[\sigma(p)\in E_{Fp}] . $$
    If $\sigma\in\Gamma(F^*E)$ is a section, then define $\tilde\sigma=f\circ\sigma$.
    This is a smooth map and $f\circ\sigma(p)\in E_{Fp}$ since $\pi\circ f\circ\sigma(p)=F\circ\pi^*\circ\sigma(p)=F(p)$.
    Conversely if $\sigma\colon M\to E$ is smooth with $\sigma(p)\in E_{Fp}$ then define $\sigma^*(p)=(p,\sigma(p))$, which is well-defined since $\pi\sigma(p)=Fp$
    by assumption.
    Then $\pi^*\circ\sigma^*(p)=\pi^*(p,\sigma(p))=p$ so $\sigma^*$ is a section.

    These are inverses, since
    $$ (\tilde\sigma)^*(p) = (p,\tilde\sigma(p)) = (p,f\circ\sigma(p)) , $$
    and this is equal to $\sigma(p)$ by definition ($\sigma(p)=(p,e)$ and $e=f(p,e)$).
    And
    $$ \widetilde{\sigma^*}(p) = f\circ\sigma^*(p) = f(p,\sigma(p)) = \sigma(p) . \qed $$

\eproof

Note that this gives us a map $\Gamma(E)\to\Gamma(F^*E)$ mapping $\sigma\in\Gamma(E)$ to $\sigma\circ F$.
In particular if $E_i\colon U\to E$ is a local frame for $E$, then $E_i^*=E_i\circ F$ is a local frame for $F^*E$ in $F^{-1}U$.
This is because $E_i^*|_p=E_i|_{Fp}$ are linearly independent and the rank of $E$ and $F^*E$ are the same.

Now consider connections on pullback bundles, so maps
$$ \nabla\colon \X(M) \times \Gamma(F^*E) \to \Gamma(F^*E) $$
satisfying the conditions of a connection.

\bprop

    Let $F\colon M\to N$ be smooth, $\pi\colon E\to N$ a vector bundle on $N$, and $\pi^*\colon F^*E\to M$ its pullback bundle.
    Further let $\nabla$ be a connection in $E$, then there is a unique connection $\nabla^*$ in $F^*E$ such that
    $$ \nabla_X^*(Y\circ F)|_p = (\nabla_{dF_p(X_p)}Y)|_{Fp} $$
    for all $X\in\X(M)$, $Y\in\Gamma(E)$, and $p\in M$.

\eprop

\bnote

    Vector fields along $F$ of the form $Y\circ F$ where $Y$ is a vector field on a neighborhood of $\im F$ are called
    {\emphcolor extendible}.

\enote

\bproof

    To show uniqueness, let $p\in M$ and let $(E_i)$ be a local frame of $F(p)\in U\subseteq N$ and consider its pullback frame $(E_i^*)$ of
    $F^*E$.
    Then $\nabla_X^*$ is determined by its values on $E_i^*=E_i\circ F$, which are given by the above formula.
    Thus it is unique, and it is easy to check that $\nabla^*$ is a connection.
    \qed

\eproof

If $F$ is not clear from context, we also may denote the pullback of a connection $\nabla$ by $\nabla^F$.

While in general we cannot push a connection {\it forward}, we can pull back a connection via a diffeomorphism.
In particular if $F\colon M\to N$ is a diffeomorphism and $\nabla$ is a connection on $E\to M$, then we can pull back $\nabla$
via $F^{-1}\colon N\to M$.
That is, we have a connection $\nabla^{F^{-1}}$ on $(F^{-1})^*E$ given by the above proposition.

A specific case of interest are pullbacks of the tangent bundle, i.e.\ $F^*TN$.

\bdefn

    Let $F\colon M\to N$ be smooth, then sections of the pullback $F^*TN$ are called {\emphcolor vector fields along $F$}, and the space of sections
    is denoted
    $$ \Gamma(F^*TN) = \X(F) . $$
    More generally the sections of the pullback $F^*T^{(k,\ell)}TN$ are called {\emphcolor tensor fields along $F$}.

\edefn

As we saw, a vector field along $F$ amounts to a map $X\colon M\to TN$ such that $X_p\in T_{Fp}N$ for all $p\in M$.
Similarly a tensor field along $F$ amounts to a map $X\colon M\to T^{(k,\ell)}TN$ such that $X_p\in T^{(k,\ell)}T_{Fp}N$ for all $p\in M$.
The case of most importance is when $F=\gamma$ is a curve.

When $F\colon M\to N$ is a diffeomorphism, we know that $TM$ and $TN$ are isomorphic vector bundles, and thus $TM$ is isomorphic to the pullback
$F^*TN$.
Explicitly the isomorphism is $h\colon F^*TN\to TM$ by
$$ h(p,u) = (p,(dF^{-1})_{Fp}u),\qquad p\in M,u\in T_{Fp}N . $$
If $\nabla$ is a connection on $N$, then pulling it back gives a connection on $F^*TN$.
Since $F^*TN$ and $TM$ are isomorphic vector bundles over $M$, this defines a connection on $TM$ via $h$.
Explicitly this connection is given by
$$ \tilde\nabla_XY = h\circ\nabla^F_Xh^{-1}Y . $$
Evaluating at $p$ gives
$$ \tilde\nabla_XY|_p = h\nabla_{dF_p(X_p)}(h^{-1}\circ Y\circ F^{-1}) . $$
Note that $h^{-1}YF^{-1}$ is just $dF_{F^{-1}p}Y_{F^{-1}p}$, i.e.\ $(F_*Y)_p$.
So this is equal to
$$ \tilde\nabla_XY|_p = (dF^{-1})_{Fp}\nabla_{dF_p(X_p)}(F_*Y)_p , $$
and thus
$$ \tilde\nabla_XY = (F^{-1})_*\nabla_{F_*X}(F_*Y) . $$

\bprop

    If $F\colon M\to N$ is a diffeomorphism, and $\nabla$ a connection on $N$, then we define a connection $F^*\nabla$ on $M$ by
    $$ (F^*\nabla)_XY = (F^{-1})_*\nabla_{F_*X}(F_*Y) . $$
    This is called the {\emphcolor pullback} of $\nabla$.

\eprop

Note then that there are two notions of a pullback.
The one just introduced is a special case of the one obtained by pulling back to a pullback bundle.

\subsection{Parallel transport}

\bdefn

    Let $\nabla$ be a connection on $M$, and $F\colon A\to M$ a smooth map from an open subset $A\subseteq\bR^n$ to $M$.
    We define the operator
    $$ \frac D{\partial t_i} \colon \Gamma(F^*T^{(k,\ell)}TM) \to \Gamma(F^*T^{(k,\ell)}TM) $$
    by
    $$ \frac D{\partial t_i}X = \nabla^F_{\ipdv{t_i}}X , $$
    the {\emphcolor covariant derivative along $F$}.

\edefn

The case we are most interested is when $n=1$ and $F=\gamma$ is a curve.
In such a case we write $D_t$ for $D/{\partial t}$.

\bthrm[title=Existence and uniqueness of parallel transport]

    Let $\gamma\colon(a,b)=I\to M$ be a curve, $t_0\in(a,b)$, and $v\in T^{(k,\ell)}T_{\gamma(t_0)}M$.
    Then there is a unique tensor field $X$ along $\gamma$ such that $X(t_0)=v$ and $D_tX=0$.

\ethrm

\bproof

    We prove this for when $X$ is a vector field; the proof for tensor fields is similar but more index-heavy.

    We assume that $\alpha(a,b)$ is contained in a coordinate domain $(U,\phi)$.
    Otherwise can cover $(a,b)$ with subintervals which $\gamma$ maps into coordinate domains of $M$.
    By uniqueness, the vector fields in the subintervals agree on their intersections, and thus we can glue them together
    to get the desired vector field $X$.

    For $X\in\X(\gamma)$, consider its representation relative to local the local frame $\partial_i\circ\gamma$
    ($\partial_i=\ipdv{x^i}$), $X=X^i(\partial_i\circ\gamma)$.
    Then
    $$ D_tX = \nabla^\gamma_{dt}(X^i(\partial_i\circ\gamma)) = dX^i\parens{\drv t}\partial_i +
    X^i\nabla_{d\gamma(d/dt)}\partial_i . $$
    If $\gamma$'s local representation is $(\gamma^1(t),\dots,\gamma^n(t))$ then
    $$ d\gamma(d/dt) = \dot\gamma(t) = \dot\gamma^i(t)(\partial_i\circ\gamma) . $$
    Thus
    $$ D_tX = \pdvof{X^i}t\partial_i + X^i\dot\gamma^j\nabla_{\partial_j}\partial_i =
    \pdvof{X^i}t\partial_i + X^i\dot\gamma^j\Gamma_{ij}^k(\alpha(t))\partial_k . $$
    So $D_tX=0$ iff
    $$ \pdvof{X^i}t(t) + X^i(t)\dot\gamma^j(t)\Gamma_{ij}^k(\alpha(t)) = 0,\qquad \hbox{for $k=1,\dots,n$} . $$
    This is a linear ODE in $X$, and thus has a unique solution given any initial value.
    \qed

\eproof

\bdefn

    A vector field $X\in\X(\gamma)$ is called {\emphcolor parallel along $\gamma$} if $D_tX=0$.
    By the above theorem, for any initial value $v$, there is a unique vector field parallel along $\gamma$ starting
    at $v$.

    For $t_0,t_1\in I$ define
    $$ P^\gamma_{t_0,t_1}\colon T_{\gamma(t_0)}M\to T_{\gamma(t_1)}M $$
    to map $v\in T_{\gamma(t_0)}M$ to $X(t_1)$ where $X$ is the unique vector field parallel along $\gamma$
    with $X(t_0)=v$.
    This is called the {\emphcolor parallel transport map}, $X$ is also called the {\emphcolor parallel transport} of $v$
    along $\gamma$.

\edefn

\bprop

    Parallel transport is a linear isomorphism $P^\gamma_{t_0,t_1}\colon T_{\gamma(t_0)}M\to T_{\gamma(t_1)}M$.

\eprop

\bproof

    It is easy to verify that if $X,Y$ are the parallel transports of $v,u\in T_{\gamma(t_0)}M$ along $\gamma$,
    then $X+Y$ is the parallel transport of $v+u$.
    Similarly $cX$ is the parallel transport of $cv$, proving linearity.

    Furthermore if $X$ is the parallel transport of $v$, then $X$ is also the parallel transport of $u=X(t_1)$.
    So the parallel transport map maps $u$ to $X(t_0)=v$, showing that $P^\gamma_{t_0,t_1}$ and $P^\gamma_{t_1,t_0}$
    are inverses.
    \qed

\eproof

\bcoro

    For any curve $\gamma\colon I\to M$ there exists a global frame of $\gamma^*TM$ consisting of parallel vector fields
    along $\gamma$.

\ecoro

\bproof

    Choose any $t_0\in I$ and a basis $v_1,\dots,v_n$ of $T_{\gamma(t_0)}M$.
    Then the above proposition shows that parallel transporting $v_i$ keeps it a basis (as it is an isomorphism).
    That is if we let $X_i$ be the parallel transport of $v_i$, then $X_i$ forms a global frame (in the pullback),
    and by definition each $X_i$ is parallel.
    \qed

\eproof

\subsection{Geodesics}

\bdefn

    Let $\nabla$ be a connection on $M$, and $\gamma\colon I\to M$ a smooth curve.
    Its {\emphcolor acceleration} is defined to be the vector field $D_t\gamma'(t)$.
    $\gamma$ is a {\emphcolor geodesic} if its acceleration is zero: $D_t\gamma'(t)=0$.

\edefn

Note that $\gamma$ is a geodesic iff its derivative is parallel along $\gamma$.

In local coordinates if $\gamma(t)=(x^1(t),\dots,x^n(t))$ then
$$ D_t\gamma' = D_t\parens{\dot x^i(t)\partial_i} = \ddot x^i\partial_i + \dot x^i\dot x^j\Gamma^k_{ij}(\gamma(t))\partial_k , $$
so $\gamma$ is a geodesic iff this is zero.

\bthrm[title=Existence and uniqueness of geodesics]

    Let $\nabla$ be a connection on $M$.
    Then for every $p\in M$, $w\in T_pM$, and $t_0\in\bR$, there is an open interval $t_0\in I\subseteq\bR$ and a geodesic $\gamma\colon I\to M$
    where $\gamma(t_0)=p$ and $\gamma'(t_0)=w$.
    Any two such geodesics are equal on their common domain.

\ethrm

\bproof

    Our ODE above is second-order, so we introduce auxillary variables $v^i=\dot x^i$.
    Then our geodesic equation becomes
    $$ \dot x^k(t) = v^k(t),\qquad \dot v^k(t) = -v^i(t)v^j(t)\Gamma^k_{ij}(x(t)) . $$
    Consider $(x^1,\dots,x^n,v^1,\dots,v^n)$ as coordinates on $U\times\bR^n$, then the equation above is just the flow equation
    of the vector field $G\in\X(U\times\bR^n)$ given by
    $$ G_{(x,v)} = v^k\evpdv{x^k}_{(x,v)} - v^iv^j\Gamma^k_{ij}\evpdv{v^k}_{(x,v)} . $$
    By the fundamental theorem on flows, for every $(p,w)\in U\times\bR^n$, $G$ has a unique integral curve $\zeta\colon I\to U\times\bR^n$ where
    $\zeta(t_0)=(p,w)$.
    If we write $\zeta(t)=(x^1(t),\dots,x^n(t),v^1(t),\dots,v^n(t))$ then we see $\gamma(t)=(x^1(t),\dots,x^n(t))$ satisfies the existence.

    Uniqueness follows because any other geodesic $\tilde\gamma$ satisfying the properties defines an integral curve of $G$ by
    $\tilde\zeta(t)=(\gamma(t),\dot\gamma(t))$, and thus must be equal to $\zeta$ on their common domain by uniqueness of flows.
    \qed

\eproof

\bdefn

    A {\emphcolor maximal} geodesic is one whose domain cannot be extended.

\edefn

\bcoro

    For any initial condition $\gamma(0)=p,\gamma'(0)=w\in T_pM$, there is a unique maximal geodesic satisfying these conditions.

\ecoro

\bproof

    Let $I$ be the union of all the open intervals containing $0$ on which the initial conditions are satisfied (there is a geodesic).
    All the geodesics on these intervals agree on their intersections, and thus gluing them together yields a geodesic from $I$ which is
    clearly maximal.
    \qed

\eproof

\bdefn

    The unique maximal geodesic $\gamma$ with $\gamma(0)=p,\gamma'(0)=v$ is called simpy the {\emphcolor geodesic with initial point $p$
    and initial velocity $v$}, and is denoted simply by $\gamma_v$ ($p$ can be recovered by $p=\pi(v)$).

\edefn

Note our vector field $G$ defined in our proof above.
We claim that this vector field actually extends to a global vector field on the tangent bundle: $G\in\X(TM)$.
For any coordinate domain $U\subseteq M$ with coordinates $(x^i)$ then let $(x^i,v^i)$ denote the local coordinates for $\pi^{-1}U\subseteq TM$
so that in these coordinates $G$ is given by the formula above:
$$ G_{(x,v)} = v^k\evpdv{x^k}_{(x,v)} - v^iv^j\Gamma^k_{ij}\evpdv{v^k}_{(x,v)} . $$
To show that $G$ defines a global vector field, we need to show that the definition above agrees on the intersection of coordinate domains.
Instead of doing this forcefully, we instead will show that $G$ acts on a smooth function $f\in C^\infty(TM)$ by
$$ Gf(p,v) = \evdrv t_{t=0}f(\gamma_v(t),\gamma'_v(t)) . $$
Since this is independent on the coordinates, we see that $G$ is well-defined on all of $TM$.

Let $\gamma_v=(x^1,\dots,x^n)$ and $\gamma'_v=(v^1,\dots,v^n)$, so $\dot x^i=v^i$.
Then the right-hand side is
$$ \evdrv t_{t=0}f(x^1(t),\dots,x^n(t),v^1(t),\dots,v^n(t)) =
\parens{\pdvof f{x^k}(x(t),v(t))\dot x^k(t) + \pdvof f{v^k}(x(t),v(t))\dot v^k}_{t=0} . $$
By the geodesic equations we know $\dot x^k=v^k$ and $\dot v^k=-v^iv^j\Gamma^k_{ij}(x)$, and $x(0)=p,v(0)=v$, so this is equal to
$$ = \pdvof f{x^k}(p,v)v^k - \pdvof f{v^k}(p,v)v^iv^j\Gamma^k_{ij}(p) . $$
Which is precisely $Gf(p,v)$, as desired.

\bdefn

    The vector field $G\in\X(TM)$ defined above is called the {\emphcolor geodesic vector field}.
    $\gamma$ is a geodesic iff $(\gamma,\gamma')$ is an integral curve of $G$ iff there is an integral curve $\zeta$ of $G$ whose projection
    onto $M$ is $\gamma$.

\edefn

\bprop

    Let $M,\tilde M$ be smooth manifolds, $\tilde\nabla$ a connection on $\tilde M$, and $\phi\colon M\to\tilde M$
    a diffeomorphism.
    Let $\nabla=\phi^*\tilde\nabla$ be the pullback connection on $M$, and let $\gamma\colon I\to M$ be a smooth curve.
    \benum
        \item $\phi$ commutes with covariant derivatives:
        $$ d\phi\circ D_tV = \tilde D_t(d\phi\circ V), $$
        where $D_t$ is the covariant derivative along $\gamma$ with respect to $\nabla$, $\tilde D_t$ along $\phi\circ\gamma$
        with respect to $\tilde\nabla$, and $V\in\X(\gamma)$.
        \item $\phi$ maps geodesics to geodesics: if $\gamma$ is a $\nabla$-geodesic, then $\phi\circ\gamma$ is a
        $\tilde\nabla$-geodesic.
        \item $\phi$ commutes with parallel transport:
        $$ d\phi_{\gamma(t_1)}\circ P^\gamma_{t_0t_1} = P^{\phi\circ\gamma}_{t_0t_1}\circ d\phi_{\gamma(t_0)} . $$
    \eenum

\eprop

\bproof

    Recall that
    $$ \nabla_XY = \phi^{-1}_*\tilde\nabla_{\phi_*X}(\phi_*Y) . $$
    Thus for extendible $V\circ\gamma$,
    $$ D_t(V\circ\gamma) = \nabla_{\ipdv t}V = \phi^{-1}_*\tilde\nabla_{\phi_*\ipdv t}(\phi_*V) . $$
    And from this we get the desired result, recalling that
    $$ \phi_*X|_p = d\phi_{\phi^{-1}p}(X_{\phi^{-1}p}) . $$

    The second point follows from the first, since if $D_t\gamma'=0$ then
    $$ \tilde D_t((\phi\circ\gamma)') = \tilde D_t(d\phi\circ\gamma') = d\phi\circ D_t\gamma' = 0 . $$
    So $\phi\circ\gamma$ is a geodesic.

    And if $X$ is the parallel transport of $v\in T_{\gamma(t_0)}M$ then $d\phi\circ X$ is the parallel transport
    of $d\phi_{\gamma(t_0)}v$ since $\tilde D_t(d\phi\circ X)=0$ by the first point.
    The result follows.
    \qed

\eproof

\subsection{The exponential map}

\blemm[title=Rescaling lemma]

    For every $p\in M,v\in T_pM,c,t\in\bR$, 
    $$ \gamma_{cv}(t) = \gamma_v(ct) , $$
    whenever either side is defined (i.e.\ they are defined on the same interval and are equal).

\elemm

\bproof

    If $c=0$ then both sides are just $p$ for all $t\in\bR$.
    Assume $c\neq0$, then it is sufficient to assume $\gamma_v(ct)$ is defined and to show $\gamma_{cv}(t)$ is and is equal to it.
    The other direction is due to setting $v=v/c$ and $t=t/c$.

    Suppose $\gamma=\gamma_v$'s domain is $I\subseteq\bR$, and define $\tilde\gamma\colon c^{-1}I\to M$ by $\tilde\gamma(t)=\gamma(ct)$.
    We show that $\tilde\gamma$ is a geodesic with initial point $p$ and initial velocity $cv$, then by uniqueness and maximality
    $\tilde\gamma=\gamma_{cv}$.
    It is immediate that $\tilde\gamma(0)=\gamma(0)=p$.
    Furthermore,
    $$ \tilde\gamma'(t) = c\gamma'(ct) \implies \tilde\gamma'(0) = c\gamma'(0) = cv $$
    as desired.
    Now we need to show that $\tilde\gamma$ is a geodesic.

    Let $D_t$ and $\tilde D_t$ denote the differentiation operators along $\gamma,\tilde\gamma$ respectively.
    Then in local coordinates
    $$ \tilde D_t\tilde\gamma'(t) = \parens{\drv t\dtilde\gamma^k + \Gamma^k_{ij}(\tilde\gamma(t))\dtilde\gamma^i\dtilde\gamma^j}\partial_k , $$
    by the chain rule,
    $$ = \parens{c^2\ddot\gamma^k(ct) + c^2\Gamma^k_{ij}(\gamma(ct))\dot\gamma^i(ct)\dot\gamma^j(ct)}\partial_k = c^2D_t\gamma'(ct) = 0 , $$
    so $\tilde\gamma$ is indeed a geodesic, as desired.
    \qed

\eproof


Let $\E\subseteq TM$ be the subset
$$ \E  = \set{v\in TM}[\hbox{$\gamma_v$ is defined on an interval containing $[0,1]$}] . $$
Then define the {\emphcolor exponential map}
$$ \exp\colon\E\to M,\qquad \exp(v) = \gamma_v(1) . $$
We also define, for $p\in M$, $\E_p=\E\cap T_pM$ and the {\emphcolor restricted exponential map} $\exp_p\colon\E_p\to M$ to be just the
restriction of the exponential to $\E_p$.

\bnote

    The exponential map here is not in general related to the exponential of a Lie group.
    They are related closely for bi-invariant metrics, but otherwise they may be different.
    We will thus make it clear which exponential map we are dealing with.

\enote

\bdefn

    A subset $S$ of a vector space $V$ is {\emphcolor star-shaped} with respect to $x\in S$ if for every $y\in S$, the line connecting $x$ and
    $y$ is contained in $S$.

\edefn

\bprop

    \benum
        \item For every $p\in M$, $\E$ is an open subset of $TM$ containing every zero vector (i.e.\ for $0\in\E_p$ for all $p$),
        and each $\E_p$ is star-shaped with respect to $0$.
        \item The exponential map is smooth.
        \item For each $v\in TM$ the geodesic $\gamma_v$ is given by
        $$ \gamma_v(t) = \exp(tv) , $$
        when either side is defined.
        \item For each $p\in M$, $d(\exp_p)_0\colon T_pM\cong T_0(T_pM)\to T_pM$ is the identity map of $T_pM$, under the usual identification of
        $T_0(T_pM)$ with $T_pM$.
        Thus there is a neighborhood $V$ of $0\in T_pM$ and $U$ of $p$ such that $\exp_p\colon V\to U$ is a diffeomorphism.
    \eenum

\eprop

\bproof

    Let $n=\dim M$.
    The rescaling lemma says that
    $$ \exp(cv) = \gamma_{cv}(1) = \gamma_v(c) $$
    whenever either side is defined, giving us (3).

    When $v\in\E_p$ then $\gamma_v$ is defined on at least $[0,1]$, thus for $t\in[0,1]$ the rescaling lemma gives us
    $$ \exp_p(tv) = \gamma_{tv}(1) = \gamma_v(t) , $$
    so $\gamma_{tv}$ is defined on at least $[0,1]$ as well; $tv\in\E_p$, so $\E_p$ is star shaped with respect to $0$.

    Now we show that $\E$ is open and $\exp$ is smooth.
    Let $G$ be the geodesic vector field, so by the fundamental theorem of flows there is an open subset $\D\subseteq\bR\times TM$
    containing $\set 0\times TM$ and a smooth map $\theta\colon\D\to TM$ such that $\theta^{(p,v)}(t)=\theta(t,(p,v))$ is the unique
    maximal integral curve of $G$ starting at $(p,v)$.

    If $(p,v)\in\E$ then $\gamma_v$ is defined on at least $[0,1]$, and thus so is the integral curve of $G$ starting at $(p,v)\in TM$.
    So $(1,(p,v))\in\D$, and thus there is a neighborhood of $(1,(p,v))$ in $\bR\times TM$ contained in $\D$.
    In particular there is an open neighborhood $U$ of $(p,v)$ such that $\set1\times U\subseteq\D$.
    This means that $U\subseteq\E$ and so $\E$ is open.

    Geodesics are the projections of integral curves of $G$, so
    $$ \exp(v) = \gamma_v(1) = \pi\circ\theta(1,(p,v)) , $$
    which is smooth.

    To compute $d(\exp_p)_0(v)$, let $\tau$ be a curve in $T_pM$ starting at $0$ whose initial velocity is $v$.
    Take $\tau(t)=tv$, and thus
    $$ d(\exp_p)_0(v) = \evdrv t_{t=0}(\exp_p\circ\tau)(t) = \evdrv t_{t=0}\exp_p(tv) = \evdrv t_{t=0}\gamma_v(t) = v , $$
    as desired.
    \qed

\eproof

