\section{Orientations}

\bdefn

    An atlas $\A$ on a smooth manifold $M$ is {\emphcolor oriented} if for every two charts $(U,\phi),(V,\psi)$ in $\A$
    and $p\in U\cap V$,
    $$ \det(d(\psi\circ\phi^{-1})_{\phi p}) > 0 $$
    where the differential of $\psi\circ\phi^{-1}\colon\bR^n\to\bR^n$ is taken with respect to the standard basis, i.e.\ the Jacobian.

    A pair $(M,\A)$ where $\A$ is an oriented atlas on the smooth manifold $M$ is called an {\emphcolor oriented smooth manifold}.

\edefn

\bdefn

    Let $(M,\A),(N,\B)$ be oriented manifolds.
    A diffeomorphism $F\colon M\to N$ {\emphcolor preserves orientation} if for every two charts $(U,\phi)\in\A,(V,\psi)\in\B$:
    $$ \det(d(\psi\circ F\circ\phi^{-1})_{\phi p}) > 0 $$
    for all $p\in U\cap F^{-1}V$.

\edefn

Note that $\det(d(\phi\circ F^{-1}\circ\psi^{-1}))$ is the multiplicative inverse of $\det(d(\psi\circ F\circ\phi^{-1}))$ and so $F$ preserves
orientation $(M,\A)\to(N,\B)$ iff $F^{-1}\colon(N,\B)\to(M,\A)$ does.
Furthermore, by multiplicativety of the determinant and functorality of the differential, the composition of orientation-preserving maps
is also orientation-preserving.

\bdefn

    Two oriented atlases on $M$ are equivalent if $\id_M$ preserves orientation.
    An {\emphcolor orientation} on $M$ is a choice of one of these equivalence classes.

    A smooth manifold $M$ is {\emphcolor orientable} if it has an orientation.

\edefn

By the above discussion, this is indeed an equivalence and thus well-defined.
Not every smooth manifold is orientable!

\blemm

    If $M$ has an atlast $\A$ consisting of just two charts $(U_1,\phi_1),(U_2,\phi_2)$ such that $\phi_1(U_2)$ is connected, then $M$
    is orientable.

\elemm

\bproof

    If $\A$ is already an oriented atlas, then we are finished.
    Otherwise $\det(d\phi_2\circ\phi_1^{-1})$ is not positive for all $p\in\phi_1(U_2)$.
    Since $\phi_1(U_2)$ is connected and the determinant is continuous and never zero, it must always be negative.
    So define $\phi_2'=r\circ\phi_2$ and $U_2'=\phi_2^{-1}(r^{-1}\circ\phi_2(U_2))$ where $r$ is the map
    $$ r\colon\bR^n\to\bR^n,\ (x_1,\dots,x_n) \mapsto (-x_1,x_2,\dots,x_n) . $$
    $r$'s Jacobian has determinant $-1$, and so
    $$ \det(d\phi_2'\circ\phi_1^{-1}) = \det(dr)\cdot\det(d\phi_2\circ\phi_1^{-1}) = -\det(d\phi_2\circ\phi_1^{-1}) > 0 . $$
    Thus this new atlas is oriented, as desired.
    \qed

\eproof

\bexam

    $S^n$ is orientable: recall that we have atlases $(U_i^\pm,\phi_i^\pm)$ where $U_i^\pm=\set{\bar x\in S^n\subseteq\bR^{n+1}}[\pm x^i>0]$
    and $\phi_i^\pm$ removes the $i$th coordinate.
    We see that all the $\phi_i^+$ and $\phi_i^-$ are compatible orientation-wise amongst themselves but not together.
    Thus we can swap $\phi_i^-$ with $r\phi_i^-$ to get an orientation.

\eexam

