\section{Riemannian Manifolds}

\subsection{Riemannian manifolds}

A symmetric $2$-tensor $\alpha\in\sym^2(V^*)$ is {\emphcolor positive definite} if $\alpha(v,v)>0$ for all $v\neq0$.
A symmetric covariant $2$-tensor field is positive definite if it is positive definite at every point.

\bdefn

    Let $M$ be a smooth manifold, a {\emphcolor Riemannian metric} on $M$ is a smooth symmetric covariant $2$-tensor field on $M$.
    A {\emphcolor Riemannian manifold} is a smooth manifold $M$ equipped with a Riemannian metric $g$.

\edefn

Since $g_p$ is an positive semidefinite and symmetric, it forms an inner product on $T_pM$.
We denote this by $\iprod{\bul,\bul}_g$.

\bnote

    Equivalently, we can characterize a Riemannian metric to be an association to each $p\in M$ an inner product $\iprod{\bul,\bul}_g$ on $T_pM$
    which is smooth in the sense that $\iprod{X,Y}_g$ is in $C^\infty(M)$ for every $X,Y\in\X(M)$.

\enote

In any smooth coordinate chart $(x^i)$, a Riemannian metric has components
$$ g = g_{ij}dx^i\otimes dx^j . $$
Note then that by symmetry, $g_{ij}=g_{ji}$ and so
$$ g = \frac12(g_{ij}dx^i\otimes dx^j + g_{ij}dx^i\otimes dx^j) = \frac12(g_{ij}dx^i\otimes dx^j + g_{ji}dx^i\otimes dx^j) =
\frac12(g_{ij}dx^i\otimes dx^j + g_{ij}dx^j\otimes dx^i) = g_{ij}dx^idx^j , $$
where $dx^idx^j$ is the symmetric product.

\bexam

    The {\emphcolor Euclidean metric} on $\bR^n$ is $\bar g$ given by gloabl coordinates
    $$ \bar g = \delta_{ij}dx^idx^j . $$
    Writing $\alpha^2$ for the symmetric product of $\alpha$ with itself, we have that
    $$ \bar g = (dx^1)^2 + \cdots + (dx^n)^2 . $$
    Then for $v,w\in T_p\bR^n$ we have
    $$ \bar g_p(v,w) = \delta_{ij}v^iw^j = \sum_i v^iw^i = v\cdot w, $$
    so that the Euclidean metric is the Euclidean inner product on $T_p\bR^n$.

\eexam

\bexam

    If $(M_1,g^1)$ and $(M_2,g^2)$ are Riemannian manifolds, we can define a Riemannian metric $g=g^1\oplus g^2$ on $M_1\times M_2$ by
    $$ g((v_1,w_1),(v_2,w_2)) = g^1(v_1,v_2) + g^2(w_1,w_2) , $$
    for $(v_i,w_i)\in T_pM_1\oplus T_qM_2\cong T_{(p,q)}(M_1\times M_2)$.
    This is the sum of two smooth tensor fields, and thus smooth itself, and symmetric.
    It is positive definite at each point because $g^1,g^2$ are.

    Given local coordinates $(x^i)$ for $M_1$ and $(y^j)$ for $M_2$, if $g^1$ has components $g^1_{ij}dx^idx^j$ and $g^2$ has components $g^2_{ij}dy^idy^j$, then
    $g$'s coordinate representation is
    $$ (g_{ij}) = \pmatrix{g^1_{ij}\cr&g^2_{ij}} . $$

\eexam

\bprop

    Every smooth manifold admits a Riemannian metric.

\eprop

\bproof

    Let $(U_\alpha,\phi_\alpha)$ cover $M$ by countably many coordinate charts.
    Then $\phi_\alpha\colon U_\alpha\to\bR^m$ is smooth and thus we can pull back the Euclidean metric to $U_\alpha$ by $g_\alpha=\phi_\alpha^*\bar g$.
    Its coordinate representation is $\delta_{ij}dx^idx^j$ relative to $\phi_\alpha$.
    Now let $\set{\psi_\alpha}$ be a partition of unity subordinate to $U_\alpha$, and define
    $$ g = \sum_\alpha\psi_\alpha g_\alpha . $$
    This is smooth, by local finiteness.
    It is also obviously symmetric, so only positivity needs to be checked.
    All the functions in question are nonnegative, so we need to show that $g(v,v)\neq0$ when $v\neq0$.
    Indeed,
    $$ g(v,v) = \sum_\alpha\psi_\alpha(p)g_\alpha(v,v) , $$
    and if $0\neq v\in T_pM$ then some $\psi_\alpha(p)$ must be positive (since they sum to $1$), and thus $g(v,v)\geq\psi_\alpha(p)g_\alpha(v,v)>0$.
    \qed

\eproof

Note that such a choice of Riemannian metric is not natural.

\bdefn

    If $(M,g)$ is a Riemannian metric, then the {\emphcolor norm} of $v\in T_pM$ is defined to be $\norm v=\iprod{v,v}_g^{1/2}=g_p(v,v)^{1/2}$.
    $v,w\in T_pM$ are {\emphcolor orthogonal} if $\iprod{v,w}_g=0$.

\edefn

\bprop

    Let $(M,g)$ be a Riemannian manifold, and $(X_j)$ a smooth local frame for $M$ over an open subset $U\subseteq M$.
    Then there is a smooth local frame $(E_j)$ over $U$ which is orthonormal at every point, and $\lspan(E_1|_p,\dots,E_j|_p)=\lspan(X_1|_p,\dots,X_j|_p)$
    for all $j=1,\dots,n$.

    In particular for every $p\in M$ there is an orthonormal frame in a neighborhood around $p$.

\eprop

\bproof

    Apply Gram-Schmidt, and note that the resulting fields are smooth.
    \qed

\eproof

\bdefn

    If $F\colon M\to N$ is smooth, and $g$ is a Riemannian metric on $N$, then the pullback $F^*g$ is a smooth $2$-tensor field on $M$.
    If it is positive definite, it is a Riemannian metric on $M$, called the {\emphcolor pullback metric}.

\edefn


\bprop

    The pullback $F^*g$ is Riemannian iff $F$ is a smooth immersion.

\eprop

\bproof

    The pullback is defined by
    $$ F^*g_p(v,w) = g(dF_pv,dF_pw) , $$
    and in general $\iprod{F\bul,F\bul}$ is an inner product only when $F$ is injective.
    \qed

\eproof

\bdefn

    If $(M,g)$ and $(N,\tilde g)$ are Riemannian manifolds, a smooth map $F\colon M\to N$ is a {\emphcolor (Riemannian) isometry} if it is a diffeomorphism
    which satisfies $F^*\tilde g=g$.
    A {\emphcolor local isometry} is one where for every $p\in M$ there is a neighborhood $p\in U$ such that $F|_U\colon U\to N$ is an isometry onto an open subset of $N$.
    Equivalently, if it satisfies $F^*\tilde g=g$ and $F$ is a local diffeomorphism.

    If there exists an isometry between Riemannian manifolds, we say they are {\emphcolor isometric}.
    A Riemmanian $n$-manifold is a {\emphcolor flat Riemannian manifold} (and the metric is a {\emphcolor flat metric}) if it is locally isometric to $(\bR^n,\bar g)$.

\edefn

\bprop

    Let $(M,g)$ be a Riemannian manifold.
    The following are equivalent:
    \benum
        \item $g$ is flat;
        \item Each point of $M$ is contained in the domain of a smooth coordinate chart in which $g$ has the coordinate representation $g=\delta_{ij}dx^idx^j$;
        \item Each point of $M$ is contained in the domain of a smooth coordinate chart in which the coordinate frame is orthonormal;
        \item Each point of $M$ is contained in the domain of a commuting orthnormal frame.
    \eenum

\eprop

\bproof

    The downwards implications are clear.

    \TODO{Revisit after doing flows}
    \qed

\eproof

\bdefn

    If $S\subseteq M$ is an immersed submanifold with $(M,g)$ Riemannian, then $S$ inherits the pullback metric $\iota^*g$ where $\iota\colon S\inj M$ is the inclusion.
    This is called the {\emphcolor induced metric}.

\edefn

The induced metric is indeed a metric since $\iota$ is by definition an immersion.
Under this metric $S$ is called a {\emphcolor Riemannian submanifold}.

\bexam

    The induced metric $\iota^*g$ induced on $S^n$ by the inclusion $S^n\inj\bR^{n+1}$ is called the {\emphcolor round} or {\emphcolor standard metric} on the sphere.

\eexam

\bdefn

    Let $(M,g)$ be a Riemannian $n$-manifold, and $S\subseteq M$ a $k$-dimensional submanifold.
    We say that for $p\in S$, a vector $v\in T_pM$ is {\emphcolor normal to $S$} if it is orthogonal to $T_pS$ with respect to $\iprod{\bul,\bul}_g$.
    The {\emphcolor normal space to $S$} at $p$ is the subspace $N_pS\subseteq T_pM$ of all vectors orthogonal to $T_pS$ (i.e.\ $N_pS=(T_pS)^\perp$).

    Let $NS=\coprod_{p\in S}N_pS$, and $\pi_{NS}\colon NS\to S$ to be the obvious projection.

\edefn

\bprop

    $NS\to S$ forms a smooth $(n-k)$-rank vector bundle over $S$, with a unique smooth structure where the linear structures on the fibers are given by $N_pS$.
    For each $p\in S$ there is a smooth frame for $NS$ on a neighborhood of $p$ that is orthonormal with respect to $g$.

\eprop

\bproof

    Take a coordinate frame $(X_i)$ for $T_pS$, and then we can extend it and apply Gram-Schmidt to get a coordinate frame $(X_i,Y_j)$ where $Y_j$ are in $N_pS$.
    These form a basis, and thus satisfy the local frame criterion for subbundles.
    \qed

\eproof

\subsection{The Riemannian distance function}

\bdefn

    Suppose $(M,g)$ is a Riemannian manifold, and $\gamma\colon[a,b]\to M$ a piecewise smooth curve segment.
    $\gamma$'s {\emphcolor length} is
    $$ L_g(\gamma) = \int_a^b\norm{\gamma'(t)}_g\,dt . $$
    This is well-defined since $\gamma'(t)$ is defined at all but finitely many values of $t$.

\edefn

Note that if $F\colon M\to N$ is a local isometry, then since $F^*g_N=g_M$ we have that $F$ preserves lengths: if $\gamma\colon[a,b]\to M$ is piecewise smooth,
$$ L_{g_N}(F\circ\gamma) = \int_a^b\norm{(F\circ\gamma)'(t)} = \int_a^b\norm{dF_{\gamma(t)}\circ\gamma'(t)} = \int_a^b\norm{\gamma'(t)} = L_{g_M}(\gamma). $$

\bdefn

    A {\emphcolor reparametrization} of a curve $\gamma\colon[a,b]\to M$ is a curve of the form $\tilde\gamma=\gamma\circ\phi$ where $\phi\colon[c,d]\to[a,b]$ is
    a diffeomorphism.

\edefn

\bprop

    Curve length is invariant under reparametrization.

\eprop

\bproof

    Let $\phi\colon[c,d]\to[a,b]$ be a diffeomorphism and $\gamma\colon[a,b]\to M$ a piecewise smooth curve.
    Since $\phi$ is a diffeomorphism, either $\phi'>0$ or $\phi'<0$ everywhere.
    Assume $\phi'$ is always positive, then
    $$ L_g(\gamma\circ\phi) = \int_c^d\norm{(\gamma\circ\phi)'(t)} = \int_c^d\norm{\phi'(t)\gamma'(\phi(t))} = \int_c^d\norm{\gamma'(\phi(t))}\phi'(t) = \int_a^b\norm{\gamma'(t)} , $$
    where the final equality is due to change of variables.
    If $\phi'<0$ then we'd have to multiply the second to last term by $-1$, then change of variables swaps the sign again.
    \qed

\eproof

\bdefn

    Let $(M,g)$ be a connected Riemannian manifold.
    For $p,q\in M$ define their {\emphcolor Riemannian distance}, denoted $d_g(p,q)$, to be the infimum over $L_g(\gamma)$ over all piecewise smooth curves $\gamma$
    from $p$ to $q$.

\edefn

\bexam

    Consider $\bR^n$ with the Euclidean metric $g$.
    Then $d_g(p,q)=\norm{p-q}$, i.e.\ the Riemannian distance is the same as Euclidean distance.

    Indeed, $\gamma\colon[0,1]\to\bR^n$ with $\gamma(t)=tp+(1-t)q$ has $\gamma'(t)=p-q$ and thus
    $$ L_g(\gamma) = \int_0^1\norm{p-q} = \norm{p-q} . $$
    So $d_g(p,q)\leq\norm{p-q}$.

    Now if $\gamma$ is any other piecewise smooth curve, we need to show that $L_g(\gamma)\geq\norm{p-q}$.
    If $\gamma\colon[a,b]\to\bR^n$ is a curve from $p$ to $q$, then if it diverses from the line connecting $p$ and $q$, say at $s$, we have
    $$ \eqalign{
        \int_a^b\norm{\gamma'(t)} &= \int_a^s\norm{\gamma'(t)} + \int_s^b\norm{\gamma'(t)}\cr
            &\geq \norm{\int_a^s\gamma'(t)} + \norm{\int_s^b\gamma'(t)}\cr
            &= \norm{\gamma(s)-\gamma(a)} + \norm{\gamma(b)-\gamma(s)} > \norm{\gamma(b)-\gamma(a)}
    } $$
    where the final strict inequality is due to the triangle inequality.

    Thus $\gamma(t)=p+s(t)(q-p)$ where $s\colon[a,b]\to[0,1]$ is a smooth map (since $\gamma$ is), with $s(a)=0,s(b)=1$.
    Then $\gamma'(t)=s'(t)(q-p)$ and so
    $$ \int_a^b\norm{\gamma'(t)} = \norm{q-p}\int_a^b\abs{s'(t)} \geq \norm{q-p}\abs{\int_a^bs'(t)} = \norm{q-p}\abs{s(b)-s(a)} = \norm{q-p} , $$
    as desired.

\eexam

\blemm

    Let $g$ be a Riemannian metric on an open subset $U\subseteq\bR^n$.
    Given a compact subset $K\subseteq U$, there exist positive constants $c,C$ such that for all $x\in K$ and $v\in T_x\bR^n$,
    $$ c\norm v_{\bar g} \leq \norm v_g \leq C\norm v_{\bar g} . $$

\elemm

\bproof

    Define $L\subseteq T\bR^n$ by
    $$ L = \set{(x,v)\in T\bR^n}[x\in K,\norm v_{\bar g}=1] . $$
    Under the canonical identification of $T\bR^n$ with $\bR^n\times\bR^n$, this is just $L=K\times S^{n-1}$, which is compact.
    Since $\norm v_g$ is continuous and strictly positive on $L$, there are positive constants $c,C$ such that $c\leq\norm v_g\leq C$ ($\norm v_g$
    obtains its maximum and minimum on a compact set) for all $(x,v)\in L$.
    If $x\in K$ and $v\in T_x\bR^n$ is nonzero, let $\lambda=\norm v_{\bar g}$ so that $(x,\lambda^{-1}v)\in L$ so
    $$ \norm v_g = \lambda\norm{\lambda^{-1}v}_g \leq \lambda C = C\norm v_{\bar g}, $$
    and similar for $c$.
    \qed

\eproof

\bthrm

    The Riemannian distance function on a connected Riemannian manifold $(M,g)$ is a metric in the usual sense.
    It induces the same topology on $M$ as its manifold topology.

\ethrm

\bproof

    All of the conditions for $d_g$ being a metric are clear, other than positivity.
    Note that the triangle inequality is because if $\gamma$ connects $a$ to $b$, and $\tilde\gamma$ connects $b$ to $c$, then we can concatenate the curves
    to get a curve from $a$ to $c$ whose length is the sum of the curves.

    Now we need to show that $d_g(p,q)\neq0$ when $p\neq q$.
    Let $U$ be a chart containing $p$ but not $q$.
    We use the coordinate map to identify $U$ with a an open subset of $\bR^n$, and let $\bar g$ be the Euclidean metric in these coordinates.
    Let $V$ be a regular coordinate ball of radius $\epsilon$ centered at $p$ such that $\cl V\subseteq U$.
    Then the above lemma shows that there are constants $c,C>0$ such that $c\norm v_{\bar g}\leq\norm v_g\leq C\norm v_{\bar g}$ for all $v\in T_xM$ when $x\in\cl V$.
    Then for any smooth curve $\gamma$ lying in $\cl V$, $cL_{\bar g}(\gamma)\leq L_g(\gamma)\leq CL_{\bar g}(\gamma)$.

    Now suppose $\gamma\colon[a,b]\to M$ is piecewise smooth from $p$ to $q$.
    Let $t_0$ be the infimum of all $t$ such that $\gamma(t)\notin\cl V$.
    By continuity, $\gamma(t_0)\in\partial V$ and so $\gamma(t)\in\cl V$ for $a\leq t\leq t_0$.
    Thus
    $$ L_g(\gamma) \geq L_g(\gamma|_{[a,t_0]}) \geq cL_{\bar g}(\gamma|_{[a,t_0]}) \geq cd_{\bar g}(p,\gamma(t_0)) = c\epsilon . $$
    The final inequality is because $d_{\bar g}$ agrees with normal Euclidean distance, and $\norm{p-\gamma(t_0)}=\epsilon$ since $\gamma(t_0)\in\partial V$.
    Thus we conclude that $d_g(p,q)\geq c\epsilon>0$ as desired.

    Finally to show that the two topologies coincide, we show that open sets in one are open in the other and vice versa.
    Suppose that $U\subseteq M$ is open, and let $p\in U$, and let $V$ be a regular coordinate ball of radius $\epsilon$ around $p$ such that $\cl V\subseteq U$.
    We have that $d_g(p,q)\geq c\epsilon$ when $q\notin\cl V$, and thus if $d_g(p,q)<c\epsilon$ then $q\in\cl V\subseteq U$.
    So the open ball of radius $c\epsilon$ centered around $p$ is contained in $U$, since $p$ was arbitrary this means $U$ is open.

    Conversely if $W$ is open in the metric topology, let $p\in W$.
    Let $V$ be a regular coordinate ball around $p$ of radius $r$, and let $\bar g$ be the Euclidean metric on $\cl V$.
    Then we have positive constants $c,C>0$ bounding the metric of $d_g$ by $d_{\bar g}$ as before.
    Let $\epsilon<r$ be small enough such that the metric ball around $p$ of radius $C\epsilon$ is contained in $W$, and let $V_\epsilon$ be the set of points of $\cl V$
    whose Euclidean distance from $p$ is less than $\epsilon$.
    If $q\in V_\epsilon$ let $\gamma$ be the straight line segment from $p$ to $q$, then
    $$ d_g(p,q) \leq L_g(\gamma) \leq CL_{\bar g}(\gamma) < C\epsilon . $$
    Thus $V_\epsilon$ is contained in the metric ball of radius $C\epsilon$ around $p$, and thus in $W$.
    $V_\epsilon$ is open in the manifold topology (since it is an open set under the coordinate chart of $\cl V$), and $p\in V_\epsilon$, we see that $W$ is open in the
    manifold topology.
    \qed

\eproof

\bdefn

    A Riemannian manifold $(M,g)$ is {\emphcolor complete}, and $g$ is a {\emphcolor complete Riemannian metric}, if $(M,d_g)$ is a complete metric space.
    A subset $B\subseteq M$ is {\emphcolor bounded} if it is bounded in the metric space.

\edefn

\bcoro

    Every smooth manifold is metrizable.

\ecoro

\bproof

    If $M$ is a manifold without boundary, then choose a Riemannian metric $g$ on $M$.
    If $M$ is connected, then the above proposition shows that it is metrizable by $d_g$.
    Otherwise, let $\set{M_i}$ be the components of $M$, each of which is metrizable by $d_g|_{M_i}$.
    The disjoint union of metrizable spaces is metrizable: let $p_i\in M_i$ and define
    $$ d_g(x,y) = d_g(x,p_i) + 1 + d_g(p_j,y), $$
    when $x,y$ are not in the same component.

    If $M$ has boundary, it can be embedded into a manifold without boundary (its double), which is metrizable.
    Subspaces of metrizable spaces are metrizable.
    \qed

\eproof

\subsection{The tangent-cotangent isomorphism}

If $V$ is a finite-dimensional inner product space, there is a natural isomorphism $V\to V^*$ given by $v\mapsto\iprod{v,\bul}$.
This is clearly linear and injective, since if $\iprod{v,\bul}=0$ then $\iprod{v,v}=0$ so $v=0$.
We can also compute its inverse: let $e_i$ be an orthonormal basis of $V$, then for $\phi\in V^*$ define $v=\sum_i\phi(e_i)e_i$.

In our case, if $(M,g)$ is a Riemannian manifold, we define $\hat g\colon TM\to T^*M$ where for $v\in T_pM$ we define $\hat g(v)\in T^*_pM$ by
$$ \hat g(v)(u) = g_p(v,u) . $$
Its action on smooth vector fields is then
$$ \hat g(X)(Y) = g(X,Y) ,\qquad X,Y\in\X(M) . $$
Now, $\hat g(X)(\bul)$ is $C^\infty(M)$-linear, and thus $\hat g(X)$ is a smooth covector field.
And $\hat g(\bul)$ is $C^\infty(M)$-linear and thus a bundle morphism as desired.
As we explained, $\hat g$ is bijective, and thus a bundle isomorphism.

In smooth coordinates $(x^i)$, we have $g=g_{ij}dx^idx^j$.
Thus if $X,Y$ are smooth vector fields,
$$ \hat g(X)(Y) = g_{ij}X^iXY^j, $$
so that the covector field $\hat g(X)$ has coordinates
$$ \hat g(X) = g_{ij}X^idx^j . $$
So $\hat g$ is the bundle homomorphism whose matrix with respect to coordinate frames for $TM$ and $T^*M$ is the same as $g$'s matrix.

It is customary to write
$$ \hat g(X) = X_jdx^j,\qquad \hbox{where } X_j=g_{ij}X^j , $$
so we say that $\hat g(X)$ is obtained by ``lowering an index''.
The notation $X^\flat$ is often used for $\hat g(X)$ because of this; $\flat$ is musical notation for lowering a tone.
The inverse map $\hat g^{-1}\colon T^*M\to TM$'s matrix is the inverse of $(g_{ij})$, which we denote $(g^{ij})$.
That is,
$$ g_{ij}g^{jk} = g^{ij}g_{jk} = \delta_{ik}. $$
The inverse of a symmetric matrix is symmetric, and so for $\omega\in\X^*(M)$, the vector field $\hat g^{-1}(\omega)$ has coordinate representation
$$ \hat g^{-1}(\omega) = \omega^i\pdv{x^i},\qquad \hbox{where } \omega^i=g^{ij}\omega_j . $$
And so $\hat g^{-1}(\omega)$ is obtained by ``raising an index'', and so we denote $\hat g^{-1}(\omega)$ also by $\omega^\sharp$ (raising a tone).

$\hat g$ and its inverse $\hat g^{-1}$ are thus called {\emphcolor musical isomorphisms}.
Note that by definition, $\omega^\sharp=\hat g^{-1}(\omega)$ is the unique vector field such that for $X\in\X(M)$,
$$ \omega(X) = \iprod{\omega^\sharp,X}_g . $$

Under the musical isomorphisms, we can define the {\emphcolor gradient} of a function $f\in C^\infty(M)$ by
$$ \grad f = (df)^\sharp = \hat g^{-1}(df) . $$
So $\grad f$ is the unique vector field such that
$$ \iprod{\grad f,X}_g = df(X) = Xf . $$
Equivalently,
$$ \iprod{\grad f,\bul} = df . $$
Since $df=\ipdvof{f}{x^i}dx^i$, we get that
$$ \grad f = g^{ij}\pdvof f{x^i}\pdv{x^j} . $$
In particular, with the Euclidean metric on $\bR^n$ we see that
$$ \grad f = \delta^{ij}\pdvof f{x^i}\pdv{x^j} = \sum_i \pdvof f{x^i}\pdv{x^i} , $$
which is just the usual gradient of a real-valued function.

