\section{Tensors}

This section will be rather short, assuming a basic familiarity with tensor products.
The following definition is a brief overview of notation.

\bdefn

    We denote the tensor product of modules by $A\otimes B$.
    We denote the dual of an $R$-module $A$ by $A^*=\hom_R(A,R)$.

\edefn

\bnote

    Primitive elements of the tensor product $A\otimes B$ are written $a\otimes b$; all other elements are sums of these.

    The tensor product is associative (in a natural way), and we denote the tensor of $n$ modules without parentheses.
    In an $n$-fold tensor product, we write the primitive elements without parentheses.

    If $B_i=(b^i_1,\dots,b^i_{n_i})$ is a basis for each $V_1$, then
    $$ B = \set{b^1_{j_1}\otimes b^n_{j_n}}[b^i_{j_i}\in B_i] \subseteq V_1\otimes\cdots\otimes V_n $$
    is a basis for the tensor product.
    In particular the dimension of a tensor product of vector spaces is the product of the dimensions of the vector spaces.

    If $A$ is an $R$-module, then $A\otimes R\cong A$ naturally.
    This natural isomorphism is given by mapping $a\otimes r\mapsto ra$.

    The tensor product is functorial: if $F_i\colon A_i\to B_i$ are maps for $i=1,2$ then define
    $F_1\otimes F_2\colon A_1\otimes A_2\to B_1\otimes B_2$ by
    $$ (F_1\otimes F_2)(a_1\otimes a_2) = (F_1(a_1))\otimes(F_2(a_2)) $$
    and extending by linearity.
    This is well-defined.

    In the case of linear functionals $\phi_i\colon A_i\to\bR$ we have that $\bR\otimes\bR\cong\bR$, and this map is naturally isomorphic to
    $$ (\phi_1\otimes\phi_2)(a_1\otimes a_2) = \phi_1(a_1)\phi_2(a_2) . $$

    The universal property of the tensor product is that
    $$ \hom(V_1\otimes\cdots\otimes V_n,A) \cong L(V_1,\dots,V_n;A) $$
    naturally, where the right side is the space of multilinear maps $V_1\otimes\cdots\otimes V_n\to A$.

\enote

\bnote

    For a field $k$ and a finite-dimensional $k$-linear space $V$, we know that $\dim V^*=\dim V$.
    For a basis $B=(v_1,\dots,v_n)$ we define its {\emphcolor dual basis} $B^*=(v_1^*,\dots,v_n^*)$ in $V^*$ as the linear functionals
    defined by
    $$ v_i^*(v_j) = \delta_{ij} , $$
    and extending by linearity.
    These are linearly independent, and thus must be a basis considering dimensions.

    There is a natural map from a space to its double dual, $V\to V^{**}$.
    For $v\in V$ define $\hat v\in V^{**}$ by $\hat v(\phi)=\phi(v)$ for all $\phi\in V^*$.
    This is injective, and considering dimensions is an isomorphism.
    It is natural, and thus for finite-dimensional spaces $V$ and $V^{**}$ are naturally isomorphic.

    For infinite-dimensional spaces, $\dim V^*$ is strictly greater than $\dim V$, so in particular $V$ is naturally a proper subspace of
    $V^{**}$, but they are not isomorphic.

\enote

\bprop

    Let $V_1,\dots,V_n$ be vector spaces, then there is a natural isomorphism
    $$ (V_1\otimes\cdots\otimes V_n)^* \cong V_1^*\otimes\cdots\otimes V_n^* . $$

\eprop

\bproof

    They have the same dimension, and thus we need only find an injective map.
    We define a map $V_1^*\otimes\cdots\otimes V_n^*$ to the dual space on the left by mapping $\phi_1\otimes\cdots\otimes\phi_n$ to the
    literal tensor product of maps:
    $$ (\phi_1\otimes\cdots\otimes\phi_n)(a_1,\dots,a_n) = \phi_1(a_1)\cdots\phi_n(a_n) . $$
    If we choose a dual basis for each $V_i^*$, we see that it maps to a linearly independent set, and thus this is injective.
    \qed

\eproof

In particular, we then have the natural isomorphisms
$$ V_1\otimes\cdots\otimes V_n \cong (V_1\otimes\cdots\otimes V_n)^{**} \cong (V_1^*\otimes\cdots\otimes V_n^*)^* . $$
Thus,
$$ V_1\otimes\cdots\otimes V_n \cong \hom(V_1^*\otimes\cdots\otimes V_n,\bR) \cong L(V_1^*,\dots,V_n^*,\bR) . $$
So we can identify the tensor product $V_1\otimes\cdots\otimes V_n$ with multilinear maps $V_1^*\times\cdots\times V_n^*\to\bR$ naturally.
Similarly, $V_1^*\otimes\cdots\otimes V_n^*$ can naturally be identified with multilinear maps $V_1\times\cdots\times V_n\to\bR$.

\bdefn

    For a space $A$, denote by $T^k(A)$ the $k$-fold tensor product of $A$ with itself:
    $$ T^k(A) = A\otimes\cdots\otimes A . $$
    The empty tensor product is taken to be the ambient ring, so $T^0(A)=R$.

\edefn

\bdefn

    Let $V$ be a finite-dimensional real vector space.
    A {\emphcolor covariant $k$-tensor} for $k>0$ is an element of $V^*\otimes\cdots\otimes V^*$.
    Equivalently, it is a multilinear map $V\times\cdots\times V\to\bR$.
    If $\alpha$ is a covariant $k$-tensor for some $k$, it is called a {\emphcolor covariant tensor} and its {\emphcolor rank} is $k$.
    The space of covariant $k$-tensors is $T^k(V^*)$.

    A {\emphcolor contravariant $k$-tensor} for $k>0$ is an element of $V\otimes\cdots\otimes V=T^k(V)$.
    Equivalently, a multilinear map $V^*\times\cdots\times V^*\to\bR$.
    The rank is defined similarly.

    For $k,\ell\neq0$ nonnegative, we define a {\emphcolor tensor of type $(k,\ell)$} to be an element of
    $$ T^{(k,\ell)}(V) = V\otimes\cdots\otimes V\otimes V^*\otimes\cdots\otimes V^* = T^k(V) \otimes T^\ell(V^*) . $$

\edefn

Note that
$$ T^{(0,k)}(V) = T^k(V^*),\qquad T^{(\ell,0)}(V) = T^\ell(V) . $$

\bdefn

    We define an action of $S_k$ on $T^k(V)$, viewed as multilinear maps, by
    $$ {}^\sigma\alpha(v_1,\dots,v_n) = \alpha(v_{\sigma(1)},\dots,v_{\sigma(n)}) $$
    for $\sigma\in S_k,\alpha\in T^k(V)$.
    This is indeed a group action, moreover it preserves the linear structure of $T^k(V)$ in that
    $$ {}^\sigma(\alpha+\beta) = {}^\sigma\alpha + {}^\sigma\beta,\qquad {}^\sigma(c\alpha) = c{}^\sigma\alpha . $$

\edefn

Viewing $T^k(V)$ as the $k$-fold tensor product of $V^*$, ${}^\sigma$ acts on primitive elements by
$$ \sigma(v_1\otimes\cdots\otimes v_k) = v_{\sigma^{-1}(1)}\otimes\cdots\otimes v_{\sigma^{-1}(k)} , $$
and extending by linearity.

\bdefn

    A covariant $k$-tensor $\alpha\in T^k(V^*)$ is {\emphcolor symmetric} if swapping its inputs does not change its value.
    Equivalently it has a trivial orbit under the above action: ${}^\sigma\alpha=\alpha$ for all $\sigma\in S_k$ (since transpositions
    generate $S_k$).
    Denote the space of symmetric covariant $k$-tensors by $\Sigma^k(V^*)\subseteq T^k(V^*)$.

    Similarly a covariant $k$-tensor $\alpha\in T^k(V^*)$ is {\emphcolor antisymmetric} if swapping its inputs swaps the sign of its value.
    Equivalently if ${}^\sigma\alpha=\sgn\sigma\alpha$ for all $\sigma\in S_k$.
    Denote the space of alternating covariant $k$-tensors by $\Lambda^k(V^*)\subseteq T^k(V^*)$.
    Antisymmetric tensors are also called {\emphcolor exterior forms} or {\emphcolor multi-covectors}.

\edefn

Define the map $\sym\colon T^k(V^*)\to\Sigma^k(V^*)$ by
$$ \sym\alpha = \frac1{k!}\sum_{\sigma\in S_k}{}^\sigma\alpha . $$
This map is called {\emphcolor symmetrization}, and it is a projection onto $\Sigma^k(V^*)$.
Indeed,
$$ {}^\tau\sym\alpha = \frac1{k!}\sum_{\sigma\in S_k}{}^{\tau\sigma}\alpha = \sym\alpha $$
so that $\sym$ is well-defined, and $\sym\alpha=\alpha$ iff $\alpha$ is symmetric.
Clearly the equality implies $\alpha$'s symmetry since $\sym\alpha$ is symmetric, and if $\alpha$ is symmetric then
$$ \sym\alpha = \frac1{k!}\sum_{\sigma\in S_k}\alpha = \alpha . $$

\bdefn

    If $\alpha\in\Sigma^k(V^*)$ and $\beta\in\Sigma^\ell(V^*)$ are symmetric, then their tensor $\alpha\otimes\beta$ is not symmetric in general.
    Define their {\emphcolor symmetric product} to be the symmetrification of their tensor product:
    $$ \alpha\beta = \sym(\alpha\otimes\beta) . $$
    Explicitly,
    $$ \alpha\beta(v_1,\dots,v_{k+\ell}) =
    \frac1{(k+\ell)!}\sum_{\sigma\in S_{k+\ell}}\alpha(v_{\sigma1},\dots,v_{\sigma k})\beta(v_{\sigma(k+1)},\dots,v_{\sigma\ell}) . $$

\edefn

\bprop

    The symmetric product satisfies the following:
    \benum
        \item it is bilinear,
        \item it is commutative,
        \item it is associative.
    \eenum

\eprop

We will not prove this.

We can also define the {\emphcolor alternation} of a tensor by
$$ \alt\alpha = \frac1{k!}\sum_{\sigma\in S_k}\sgn\sigma{}^\sigma\alpha . $$
This is indeed antisymmetric,
$$ {}^\tau\alt\alpha = \frac1{k!}\sum_\sigma\sgn\sigma{}^{\tau\sigma}\alpha = \sgn\tau\alt\alpha . $$
If $\alpha$ is already antisymmetric then
$$ \alt\alpha = \frac1{k!}\sum_\sigma\sgn\sigma{}^\sigma\alpha = \frac1{k!}\sum_\sigma(\sgn\sigma)^2\alpha = \alpha $$
so this is also a projection.

Notice that $\sym\alpha=0$ for antisymmetric tensors of dimension $>1$ and similarly $\alt\alpha=0$ for symmetric tensors of dimension $>1$.

\subsection{Tensor fields}

\bdefn

    Let $M$ be a smooth manifold.
    Define its {\emphcolor bundle of covariant $k$-tensors} by
    $$ T^kT^*M = \coprod_{p\in M}T^k(T^*_pM) $$
    where $T_p^*M$ is the dual space of $T_pM$.
    Similarly the {\emphcolor bundle of contravariant $k$-tensors} by
    $$ T^kTM = \coprod_{p\in M}T^k(T_pM) , $$
    and in general the {\emphcolor bundle of $(k,\ell)$-tensors} by
    $$ T^{(k,\ell)}TM = \coprod_{p\in M}T^{(k,\ell)}(T_pM) . $$

\edefn

To show that $T^{(k,\ell)}TM$ has the unique structure of a vector bundle over $M$, we prove that in general the tensor of two vector bundles is a vector bundle.

\bdefn

    Let $\pi'\colon E'\to M$ and $\pi''\colon E''\to M$ be vector bundles over $M$.
    Define their {\emphcolor tensor} to be $E=\coprod_{p\in M}E'_p\otimes E''_p$ with the obvious projection $\pi\colon E\to M$.

\edefn

Let $(\sigma_i)$ and $(\tilde\sigma_i)$ be local frames defined in an open set $U$, and let $\Phi,\tilde\Phi$ be their local trivializations.
Meaning
$$ \Phi(a^i\sigma_i|_p) = (p,\bar a), $$
and by assumption there is a smooth transition function $\tau\colon U\to\gl_n(\bR)$ such that
$$ \tilde\Phi\circ\Phi^{-1}(p,\bar a) = (p,\tau(p)\bar a) . $$
Then if $X=a^i\sigma_i=\tilde a^i\tilde\sigma_i$, we have
$$ \tilde\Phi\circ\Phi^{-1}(p,\bar a) = \tilde\Phi^{-1}(X_p) = (p,\tilde a), $$
so that $\tilde a=\tau(p)\bar a$, i.e.\ $\tilde a^i=\tau(p)^i\bar a$ where $\bul^i$ is the $i$th row.

In particular, $\sigma_i=\tau(p)^j_i\tilde\sigma_j$.

\bprop

    The tensor of two vector bundles has a unique smooth structure making it a vector bundle over $M$ whose fibers are the tensors of the fibers of the original
    vector bundles.

\eprop

\bproof

    Suppose $E',E''$ have rank $k',k''$ respectively.
    Let $(\sigma_i)$ and $(\tau_j)$ be local frames of $E',E''$ respectively over an open subset $U\subseteq M$.
    Then define $\Phi\colon\pi^{-1}U\to U\times\bR^{k'k''}$ by
    $$ \Phi(v^{ij}\sigma_i|_p\otimes\tau_j|_p) = (p,v^{ij}) . $$
    Then for another trivialization $\tilde\Phi$ associated with $(\tilde\sigma_i)$ and $(\tilde\tau_j)$ we have that ($T',T''$ being the transition functions),
    $$ \sigma_i\otimes\tau_j = ((T')^k_i\tilde\sigma_k)\otimes((T'')^\ell_j\tilde\tau_\ell) = (T')^k_i(T'')^\ell_j(\tilde\sigma_k\otimes\tilde\tau_\ell) . $$
    So the transition function is $T^i_j=\sum_{k,\ell}(T')^k_i\cdot(T'')^\ell_j$, which is smooth (as the sum and product of smooth functions) and invertible
    (since $\sigma_i\otimes\tau_j$ and $\tilde\sigma_i\otimes\tilde\tau_j$ form bases).
    \qed

\eproof

We have that
$$ T^{(k,\ell)}TM = \coprod_{p\in M}T^{(k,\ell)}(T_pM) $$
is just the tensor of $TM$ tensored with itself $k$ times and $T^*M$ with itself $\ell$ times, and is thus a vector bundle.

The local frames of $E\otimes E'$ are of the form $(\sigma_i\otimes\tau_j)$ for local frames $\sigma_i$ of $E$ and $\tau_j$ of $E'$.
Thus the local frames of $T^{(k,\ell)}TM$ give representations of the form
$$ A^{i_1\dots i_k}_{j_1\dots j_\ell}\pdv{x^{i_1}}\otimes\pdv{x^{i_k}}\otimes dx^{j_1}\otimes dx^{j_\ell} , $$
and $A^{i_1\dots i_k}_{j_1\dots j_\ell}$ are called the {\emphcolor components} of the local frame, we denote them also by $A^I_J$ if the tuples $I,J$ are understood.

We adopt the notation for the sections of the tensor bundles:
$$ \T^k(M) = \Gamma(T^kT^*M),\qquad \T^{(k,\ell)}(M) = \Gamma(T^{(k,\ell)}TM) . $$
We do not give special notation to $\Gamma(T^kTM)$, as it is not as important to us.
Sections of $T^kT^*M$ are called {\emphcolor covariant $k$-tensor fields}.
Rough sections of $T^{(k,\ell)}TM$ are maps $A\colon M\to T^{(k,\ell)}TM$ such that for every $p\in M$,
$$ A_p\in T^{(k,\ell)}T_pM = L(T_p^*M,\dots,T_p^*M,T_pM,\dots,T_pM;\ \bR) , $$
so $A_p$ is a multlilinear functional from $k$-times $T_p^*M$ and $\ell$-times $T_pM$.

\bprop

    Let $M$ be a manifold, and $A\colon M\to T^{(k,\ell)}TM$ be a rough section.
    Then the following are equivalent:
    \benum
        \item $A$ is smooth,
        \item In every coordinate chart $A$'s components are smooth,
        \item Each point in $M$ is contained in a chart where $A$'s components are smooth,
        \item If $X_1,\dots,X_\ell\in\X(M)$ and $\omega_1,\dots,\omega_k\in\X^*(M)$, the function $A(\bar\omega,\bar X)\colon M\to\bR$ defined by
        $$ A(\bar\omega,\bar X)(p) = A_p(\bar\omega|_p,\bar X|_p) $$
        is smooth.
        \item Whenever $\bar X,\bar\omega$ are smooth vector and covector fields defined in some open subset $U\subseteq M$, $A(\bar\omega,\bar X)$ is smooth on $U$.
    \eenum

\eprop

\bproof
    
    The first three points are clearly equivalent, as they are for general vector bundles.
    We note that if $A=A^I_J\ipdv{x^I}\otimes dx^J$, then
    $$ A\parens{dx^{j_1},\dots,dx^{j_k},\pdv{x^{i_1}},\dots,\pdv{x^{i_\ell}}} = A^I_J , $$
    so that the last point implies the first.
    And if $A$ is smooth, then in terms of local coordinates we have that if $X_i=X_i^j\ipdv{x^j}$ and $\omega^i=\omega^i_jdx^j$ then
    $$ A(\bar\omega,\bar X) = A^I_J\ipdv{x^I}\otimes dx^J(\bar\omega,\bar X) = A^I_J\omega_IX^J, $$
    where $X^J=X_1^{j_1}\cdots X_k^{j_k}$ and similar for $\omega_I$.
    This is smooth, as desired.
    \qed

\eproof

\bprop

    Let $M$ be a manifold, $A\in\T^{(k,\ell)}(M)$, $B\in\T^{(k',\ell')}(M)$, and $f\in C^\infty(M)$.
    Then $A\otimes B$ is a smooth tensor field in $\T^{(k+k',\ell+\ell')}$, and $fA\in\T^{(k,\ell)}$.

\eprop

\bproof

    It is not hard to see that if $A$ has local coordinates $A^I_J$, $B$ has $B^{I'}_{J'}$ then $A\otimes B$ has coordinates
    $$ (A\otimes B)^{II'}_{JJ'} = A^I_JB^{I'}_{J'} , $$
    where we split the tuples $II',JJ'$.
    And similarly
    $$ (fA)^I_J = f\cdot A^I_J. $$
    These are all smooth, as desired.
    \qed

\eproof

Our above proposition showed that a tensor field $A\in\T^{(k,\ell)}(M)$ induces a map
$$ \X^*(M)\times\cdots\times\X^*(M)\times\X(M)\times\cdots\times\X(M)\to C^\infty(M) . $$
This is a multilinear $C^\infty(M)$-module morphism, which can be verified coordinate-wise.

\blemm

    Multilinear $C^\infty(M)$ maps $\X^*(M)^k\times\X(M)^\ell\to C^\infty(M)$ are precisely those induced by $(k,\ell)$-tensor fields.

\elemm

\bproof

    We prove this for $(0,k)$-tensor fields, so $C^\infty(M)$-maps $\X(M)^k\to C^\infty(M)$; the general case is similar, but messier.
    Let $\c A$ be a $C^\infty(M)$-multilinear map.
    We first show that $\c A$ acts locally, so if $X_i$ vanishes in a neighborhood $U$ of $p$ then $0=\c A(\bar X)$.
    Again, let $\psi$ be a bump function supported in $U$ such that $\psi(p)=1$, so that $\psi X_i=0$.
    Then
    $$ 0 = \c A(X_1,\dots,\psi X_i,\dots,X_n)(p) = \psi(p)\c A(\bar X)(p) = \c A(\bar X)(p) $$
    meaning $\c A(\bar X)$ vanishes on $U$.

    Now we show that if $X_i|_p=0$ then $\c A(\bar X)(p)=0$ so that $\c A(\bar X)(p)$ is dependent only on $\bar X$'s values at $p$.
    Let $X_i=X^j_i\ipdv{x^j}$ in a neighborhood of $p$, so the two vector fields agree in a neighborhood and thus we can replace $X$ with this coordinate
    representation and not change $\c A$'s value.
    Then
    $$ \c A(X_1,\dots,X_i,\dots,X_n)(p) = X^j_i(p)\c A(X_1,\dots,\ipdv{x^j},\dots,X_n) = 0 , $$
    as desired.
    Note that we need to extend $X^j_i\ipdv{x^j}$ to a global tensor field, which can be done by the extension lemma.

    Now define $A\colon M\to T^kT^*M$ by
    $$ A_p(v_1,\dots,v_k)(p) = \c A(V_1,\dots,V_k)(p) $$
    where $V_i$ are any vector fields with $V_i(p)=v_i$.
    This is independent of choice of $V_i$, as we showed, and gives the desired tensor field (it is smooth since it induces $\A$ which is smooth).
    \qed

\eproof

Thus we can identify tensor fields with multilinear $C^\infty(M)$ morphisms.

\bdefn

    A {\emphcolor symmetric tensor field} is a covariant tensor field whose image consists of symmetric tensors.

\edefn

\bdefn

    Let $F\colon M\to N$ be smooth, then at any $p\in M$ and any $k$-tensor $\alpha\in T^k(T^*_{Fp}N)$ we define $dF^*_p(\alpha)\in T^k(T^*_pM)$ by
    $$ dF^*_p(\alpha)(v_1,\dots,v_n) = \alpha(dF_p(v_1),\dots,dF_p(v_n)) . $$
    This is the {\emphcolor pullback of $F$ at $p$}.
    Then for a covariant $k$-tensor field $A$ on $N$, we define the {\emphcolor pullback of $A$ by $F$} to be the rough $k$-tensor field on $M$ given by
    $$ (F^*A)_p = dF^*_p(A_{Fp}) . $$
    This acts on vectors $v_1,\dots,v_k\in T_pM$ by
    $$ (F^*A)_p(v_1,\dots,v_k) = A_{Fp}(dF_p(v_1),\dots,dF_p(v_k)) . $$

\edefn

\bprop

    For $F\colon M\to N$ and $G\colon N\to P$ smooth, $A,B$ covariant $k$-tensor fields on $N$, and $f\in C^\infty(N)$,
    \benum
        \item $F^*(fB)=(f\circ F)F^*B$,
        \item $F^*(df)=d(f\circ F)$,
        \item $F^*(A\otimes B)=F^*A\otimes F^*B$,
        \item $F^*(A+B)=F^*A+F^*B$,
        \item $F^*B$ is continuous, and smooth if $B$ is smooth,
        \item $(G\circ F)^*B=F^*(G^*B)$,
        \item $(\id_N)^*B=B$.
    \eenum

\eprop

\bproof

    Similar to before, by definitions.
    \qed

\eproof

\bcoro

    Let $F\colon M\to N$ be smooth, and $B\in\T^k(N)$.
    If $p\in M$ and $(y^i)$ are smooth coordinates for $N$ in a neighborhood of $F(p)$, then $F^*B$ has the representation in a neighborhood of $p$:
    $$ F^*(B_Idy^I) = (B_I\circ F)d(y^{i_1}\circ F)\otimes\cdots\otimes d(y^{i_k}\circ F) . $$

\ecoro

\bproof

    Immediate from the above proposition.
    \qed

\eproof

